### 2013-12-12 paper/paper.tex
paper template for TCIAIG

%% bare_jrnl.tex
%% V1.3
%% 2007/01/11
%% by Michael Shell
%% see http://www.michaelshell.org/
%% for current contact information.
%%
%% This is a skeleton file demonstrating the use of IEEEtran.cls
%% (requires IEEEtran.cls version 1.7 or later) with an IEEE journal paper.
%%
%% Support sites:
%% http://www.michaelshell.org/tex/ieeetran/
%% http://www.ctan.org/tex-archive/macros/latex/contrib/IEEEtran/
%% and
%% http://www.ieee.org/

% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
% *** with production work. IEEE's font choices can trigger bugs that do  ***
% *** not appear when using other class files.                            ***
% The testflow support page is at:
% http://www.michaelshell.org/tex/testflow/

%%*************************************************************************
%% Legal Notice:
%% This code is offered as-is without any warranty either expressed or
%% implied; without even the implied warranty of MERCHANTABILITY or
%% FITNESS FOR A PARTICULAR PURPOSE! 
%% User assumes all risk.
%% In no event shall IEEE or any contributor to this code be liable for
%% any damages or losses, including, but not limited to, incidental,
%% consequential, or any other damages, resulting from the use or misuse
%% of any information contained here.
%%
%% All comments are the opinions of their respective authors and are not
%% necessarily endorsed by the IEEE.
%%
%% This work is distributed under the LaTeX Project Public License (LPPL)
%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
%% distributed and modified. A copy of the LPPL, version 1.3, is included
%% in the base LaTeX documentation of all distributions of LaTeX released
%% 2003/12/01 or later.
%% Retain all contribution notices and credits.
%% ** Modified files should be clearly indicated as such, including  **
%% ** renaming them and changing author support contact information. **
%%
%% File list of work: IEEEtran.cls, IEEEtran_HOWTO.pdf, bare_adv.tex,
%%                    bare_conf.tex, bare_jrnl.tex, bare_jrnl_compsoc.tex
%%*************************************************************************

% Note that the a4paper option is mainly intended so that authors in
% countries using A4 can easily print to A4 and see how their papers will
% look in print - the typesetting of the document will not typically be
% affected with changes in paper size (but the bottom and side margins will).
% Use the testflow package mentioned above to verify correct handling of
% both paper sizes by the user's LaTeX system.
%
% Also note that the "draftcls" or "draftclsnofoot", not "draft", option
% should be used if it is desired that the figures are to be displayed in
% draft mode.
%
\documentclass[journal]{IEEEtran}
%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[journal]{../sty/IEEEtran}

% Some very useful LaTeX packages include:
% (uncomment the ones you want to load)

% *** MISC UTILITY PACKAGES ***
%
%\usepackage{ifpdf}
% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
% compilation based on whether the output is pdf or dvi.
% usage:
% \ifpdf
%   % pdf code
% \else
%   % dvi code
% \fi
% The latest version of ifpdf.sty can be obtained from:
% http://www.ctan.org/tex-archive/macros/latex/contrib/oberdiek/
% Also, note that IEEEtran.cls V1.7 and later provides a builtin
% \ifCLASSINFOpdf conditional that works the same way.
% When switching from latex to pdflatex and vice-versa, the compiler may
% have to be run twice to clear warning/error messages.

% *** CITATION PACKAGES ***
%
%\usepackage{cite}
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
% \cite{} output to follow that of IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
% \cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 4.0 (2003-05-27) and later if using hyperref.sty. cite.sty does
% not currently provide for hyperlinked citations.
% The latest version can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/cite/
% The documentation is contained in the cite.sty file itself.

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
  % \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  % \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation can
% be obtained at: 
% http://www.ctan.org/tex-archive/macros/latex/required/graphics/
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found as epslatex.ps or
% epslatex.pdf at: http://www.ctan.org/tex-archive/info/
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

% *** MATH PACKAGES ***
%
%\usepackage[cmex10]{amsmath}
% A popular package from the American Mathematical Society that provides
% many useful and powerful commands for dealing with mathematics. If using
% it, be sure to load this package with the cmex10 option to ensure that
% only type 1 fonts will utilized at all point sizes. Without this option,
% it is possible that some math symbols, particularly those within
% footnotes, will be rendered in bitmap form which will result in a
% document that can not be IEEE Xplore compliant!
%
% Also, note that the amsmath package sets \interdisplaylinepenalty to 10000
% thus preventing page breaks from occurring within multiline equations. Use:
%\interdisplaylinepenalty=2500
% after loading amsmath to restore such page breaks as IEEEtran.cls normally
% does. amsmath.sty is already installed on most LaTeX systems. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/required/amslatex/math/

% *** SPECIALIZED LIST PACKAGES ***
%
%\usepackage{algorithmic}
% algorithmic.sty was written by Peter Williams and Rogerio Brito.
% This package provides an algorithmic environment fo describing algorithms.
% You can use the algorithmic environment in-text or within a figure
% environment to provide for a floating algorithm. Do NOT use the algorithm
% floating environment provided by algorithm.sty (by the same authors) or
% algorithm2e.sty (by Christophe Fiorio) as IEEE does not use dedicated
% algorithm float types and packages that provide these will not provide
% correct IEEE style captions. The latest version and documentation of
% algorithmic.sty can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithms/
% There is also a support site at:
% http://algorithms.berlios.de/index.html
% Also of interest may be the (relatively newer and more customizable)
% algorithmicx.sty package by Szasz Janos:
% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithmicx/

% *** ALIGNMENT PACKAGES ***
%
%\usepackage{array}
% Frank Mittelbach's and David Carlisle's array.sty patches and improves
% the standard LaTeX2e array and tabular environments to provide better
% appearance and additional user controls. As the default LaTeX2e table
% generation code is lacking to the point of almost being broken with
% respect to the quality of the end results, all users are strongly
% advised to use an enhanced (at the very least that provided by array.sty)
% set of table tools. array.sty is already installed on most systems. The
% latest version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/required/tools/

%\usepackage{mdwmath}
%\usepackage{mdwtab}
% Also highly recommended is Mark Wooding's extremely powerful MDW tools,
% especially mdwmath.sty and mdwtab.sty which are used to format equations
% and tables, respectively. The MDWtools set is already installed on most
% LaTeX systems. The lastest version and documentation is available at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/mdwtools/

% IEEEtran contains the IEEEeqnarray family of commands that can be used to
% generate multiline equations as well as matrices, tables, etc., of high
% quality.

%\usepackage{eqparbox}
% Also of notable interest is Scott Pakin's eqparbox package for creating
% (automatically sized) equal width boxes - aka "natural width parboxes".
% Available at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/eqparbox/

% *** SUBFIGURE PACKAGES ***
%\usepackage[tight,footnotesize]{subfigure}
% subfigure.sty was written by Steven Douglas Cochran. This package makes it
% easy to put subfigures in your figures. e.g., "Figure 1a and 1b". For IEEE
% work, it is a good idea to load it with the tight package option to reduce
% the amount of white space around the subfigures. subfigure.sty is already
% installed on most LaTeX systems. The latest version and documentation can
% be obtained at:
% http://www.ctan.org/tex-archive/obsolete/macros/latex/contrib/subfigure/
% subfigure.sty has been superceeded by subfig.sty.

%\usepackage[caption=false]{caption}
%\usepackage[font=footnotesize]{subfig}
% subfig.sty, also written by Steven Douglas Cochran, is the modern
% replacement for subfigure.sty. However, subfig.sty requires and
% automatically loads Axel Sommerfeldt's caption.sty which will override
% IEEEtran.cls handling of captions and this will result in nonIEEE style
% figure/table captions. To prevent this problem, be sure and preload
% caption.sty with its "caption=false" package option. This is will preserve
% IEEEtran.cls handing of captions. Version 1.3 (2005/06/28) and later 
% (recommended due to many improvements over 1.2) of subfig.sty supports
% the caption=false option directly:
%\usepackage[caption=false,font=footnotesize]{subfig}
%
% The latest version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/subfig/
% The latest version and documentation of caption.sty can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/caption/

% *** FLOAT PACKAGES ***
%
%\usepackage{fixltx2e}
% fixltx2e, the successor to the earlier fix2col.sty, was written by
% Frank Mittelbach and David Carlisle. This package corrects a few problems
% in the LaTeX2e kernel, the most notable of which is that in current
% LaTeX2e releases, the ordering of single and double column floats is not
% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
% single column figure to be placed prior to an earlier double column
% figure. The latest version and documentation can be found at:
% http://www.ctan.org/tex-archive/macros/latex/base/

%\usepackage{stfloats}
% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
% the ability to do double column floats at the bottom of the page as well
% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
% LaTeX2e). It also provides a command:
%\fnbelowfloat
% to enable the placement of footnotes below bottom floats (the standard
% LaTeX2e kernel puts them above bottom floats). This is an invasive package
% which rewrites many portions of the LaTeX2e float routines. It may not work
% with other packages that modify the LaTeX2e float routines. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/sttools/
% Documentation is contained in the stfloats.sty comments as well as in the
% presfull.pdf file. Do not use the stfloats baselinefloat ability as IEEE
% does not allow \baselineskip to stretch. Authors submitting work to the
% IEEE should note that IEEE rarely uses double column equations and
% that authors should try to avoid such use. Do not be tempted to use the
% cuted.sty or midfloat.sty packages (also by Sigitas Tolusis) as IEEE does
% not format its papers in such ways.

%\ifCLASSOPTIONcaptionsoff
%  \usepackage[nomarkers]{endfloat}
% \let\MYoriglatexcaption\caption
% \renewcommand{\caption}[2][\relax]{\MYoriglatexcaption[#2]{#2}}
%\fi
% endfloat.sty was written by James Darrell McCauley and Jeff Goldberg.
% This package may be useful when used in conjunction with IEEEtran.cls'
% captionsoff option. Some IEEE journals/societies require that submissions
% have lists of figures/tables at the end of the paper and that
% figures/tables without any captions are placed on a page by themselves at
% the end of the document. If needed, the draftcls IEEEtran class option or
% \CLASSINPUTbaselinestretch interface can be used to increase the line
% spacing as well. Be sure and use the nomarkers option of endfloat to
% prevent endfloat from "marking" where the figures would have been placed
% in the text. The two hack lines of code above are a slight modification of
% that suggested by in the endfloat docs (section 8.3.1) to ensure that
% the full captions always appear in the list of figures/tables - even if
% the user used the short optional argument of \caption[]{}.
% IEEE papers do not typically make use of \caption[]'s optional argument,
% so this should not be an issue. A similar trick can be used to disable
% captions of packages such as subfig.sty that lack options to turn off
% the subcaptions:
% For subfig.sty:
% \let\MYorigsubfloat\subfloat
% \renewcommand{\subfloat}[2][\relax]{\MYorigsubfloat[]{#2}}
% For subfigure.sty:
% \let\MYorigsubfigure\subfigure
% \renewcommand{\subfigure}[2][\relax]{\MYorigsubfigure[]{#2}}
% However, the above trick will not work if both optional arguments of
% the \subfloat/subfig command are used. Furthermore, there needs to be a
% description of each subfigure *somewhere* and endfloat does not add
% subfigure captions to its list of figures. Thus, the best approach is to
% avoid the use of subfigure captions (many IEEE journals avoid them anyway)
% and instead reference/explain all the subfigures within the main caption.
% The latest version of endfloat.sty and its documentation can obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/endfloat/
%
% The IEEEtran \ifCLASSOPTIONcaptionsoff conditional can also be used
% later in the document, say, to conditionally put the References on a 
% page by themselves.

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
%\usepackage{url}
% url.sty was written by Donald Arseneau. It provides better support for
% handling and breaking URLs. url.sty is already installed on most LaTeX
% systems. The latest version can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/misc/
% Read the url.sty source comments for usage information. Basically,
% \url{my_url_here}.

% *** Do not adjust lengths that control margins, column widths, etc. ***
% *** Do not use packages that alter fonts (such as pslatex).         ***
% There should be no need to do such things with IEEEtran.cls V1.6 and later.
% (Unless specifically asked to do so by the journal or conference you plan
% to submit to, of course. )

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor ordering algorithms}

\usepackage{graphicx,times,amsmath, fontenc, todonotes, url, array, multirow}
\usepackage{algorithm, mathtools}
\usepackage[noend]{algpseudocode}
\usepackage[bookmarks=false]{hyperref}
\hyphenation{op-tical net-works semi-conduc-tor IEEEtran instead}

\newcommand{\degree}{^{\circ}}
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\setb}[2]{\set{#1 \ : \ #2}}
\newcommand{\argmax}{\operatorname*{arg\,max}}

\newcommand{\twoIndent}{\hskip\algorithmicindent\hskip\algorithmicindent}
\newcommand{\threeIndent}{\hskip\algorithmicindent\hskip\algorithmicindent\hskip\algorithmicindent}

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired

\title{\ \\ \LARGE\bf Multi-Objective Monte Carlo Tree Search for Real-Time Games \thanks{Diego~Perez, Spyridon~Samothrakis, Simon~M.~Lucas (School of Computer Science and Electronic Engineering, University of Essex, Colchester CO4 3SQ, UK; email: {\tt \{dperez,ssamot,sml\}@essex.ac.uk}); \newline
Sanaz~Mostaghim (Department of Knowledge and Language Engineering, Otto-von-Guericke-Universität Magdeburg, Magdeburg, Germany; email: {\tt sanaz.mostaghim@ovgu.de})}}

%
%
% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs
%

\author{Diego Perez, {\it Student Member, IEEE}, Sanaz Mostaghim, Spyridon Samothrakis, {\it Student Member, IEEE}, Simon M. Lucas, {\it Senior Member, IEEE}}

% note the % following the last \IEEEmembership and also \thanks - 
% these prevent an unwanted space from occurring between the last author name
% and the end of the author line. i.e., if you had this:
% 
% \author{....lastname \thanks{...} \thanks{...} }
%                     ^------------^------------^----Do not want these spaces!
%
% a space would be appended to the last name and could cause every name on that
% line to be shifted left slightly. This is one of those "LaTeX things". For
% instance, "\textbf{A} \textbf{B}" will typeset as "A B" not "AB". To get
% "AB" then you have to do: "\textbf{A}\textbf{B}"
% \thanks is no different in this regard, so shield the last } of each \thanks
% that ends a line with a % and do not let a space in before the next \thanks.
% Spaces after \IEEEmembership other than the last one are OK (and needed) as
% you are supposed to have spaces between the names. For what it is worth,
% this is a minor point as most people would not even notice if the said evil
% space somehow managed to creep in.

% The paper headers
\markboth{IEEE Transactions on Computational Intelligence and AI in Games}%
{Shell \MakeLowercase{\textit{et al.}}: Bare Demo of IEEEtran.cls for Journals}

% The only time the second header will appear is for the odd numbered pages
% after the title page when using the twoside option.
% 
% *** Note that you probably will NOT want to include the author's ***
% *** name in the headers of peer review papers.                   ***
% You can use \ifCLASSOPTIONpeerreview for conditional compilation here if
% you desire.

% If you want to put a publisher's ID mark on the page you can do it like
% this:
%\IEEEpubid{0000--0000/00\$00.00~\copyright~2007 IEEE}
% Remember, if you use this you must call \IEEEpubidadjcol in the second
% column for its text to clear the IEEEpubid mark.

% use for special paper notices
%\IEEEspecialpapernotice{(Invited Paper)}

\maketitle

\begin{abstract}
Abstract... 
\end{abstract}

% IEEEtran.cls defaults to using nonbold math in the Abstract.
% This preserves the distinction between vectors and scalars. However,
% if the journal you are submitting to favors bold math in the abstract,
% then you can use LaTeX's standard command \boldmath at the very start
% of the abstract to achieve this. Many IEEE journals frown on math
% in the abstract anyway.

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

%\IEEEoverridecommandlockouts
%
%\textwidth 178mm
%\textheight 239mm
%\oddsidemargin -7mm
%\evensidemargin -7mm
%\topmargin -6mm
%\columnsep 5mm

\section{Introduction} \label{sec:intro}

Here it goes, the introduction.

\section{Monte Carlo Tree Search} \label{sec:mcts}

\section{Multi-Objective Optimization} \label{sec:moo}

\section{Multi-Objective Monte Carlo Tree Search} \label{sec:momcts}

\section{Benchmarks} \label{sec:bench}

\section{Experimentation} \label{sec:exp}

\section{Conclusions and Future Work} \label{sec:conc}

\section*{Acknowledgments}
This work was supported by EPSRC grants EP/H048588/1, under the project entitled ``UCT for Games and Beyond''.

% Can use something like this to put references on a page
% by themselves when using endfloat and the captionsoff option.
\ifCLASSOPTIONcaptionsoff
  \newpage
\fi

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% references section

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
%\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
%\bibliography{IEEEabrv,../bib/paper}
%
% <OR> manually copy in the resultant .bbl file
% set second argument of \begin to the number of references
% (used to reserve space for the reference number labels box)

\bibliographystyle{IEEEtran} 
%\bibliography{bibliography,mcts_group_mendeley,mctsbradford_group_mendeley}
\bibliography{biblio}

\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{photos/perez.jpg}}]
{Diego Perez}
received a B.Sc. and a M.Sc. in Computer Science from University Carlos III, Madrid, in 2007. He is currently pursuing a Ph.D. in Artificial Intelligence applied to games at the University of Essex, Colchester. He has published in the domain of Game AI, participated in several Game AI competitions and organized the Physical Travelling Salesman Problem competition, held in IEEE conferences during 2012. He also has programming experience in the videogames industry with titles published for game consoles and PC.
\end{IEEEbiography}

\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{photos/sanaz.jpg}}]
{Sanaz Mostaghim}
BIO HERE.
\end{IEEEbiography}

\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{./photos/spyros.jpg}}]{Spyridon Samothrakis}
 is currently pursuing a PhD in Computational Intelligence and Games at the University of Essex. His interests include game theory, computational neuroscience, evolutionary algorithms and consciousness. 
\end{IEEEbiography}

\begin{biography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{photos/lucas.png}}]{Simon Lucas}
(SMIEEE) is a professor of Computer Science at the University of Essex (UK) where he leads the Game Intelligence Group. His main research interests are games, evolutionary computation, and machine learning, and he has published widely in these fields with over 160 peer-reviewed papers. He is the inventor of the scanning n-tuple classifier, and is the founding Editor-in-Chief of the IEEE Transactions on Computational Intelligence and AI in Games.
\end{biography}

% You can push biographies down or up by placing
% a \vfill before or after them. The appropriate
% use of \vfill depends on what kind of text is
% on the last page and whether or not the columns
% are being equalized.

%\vfill

% Can be used to pull up biographies so that the bottom of the last one
% is flush with the other column.
%\enlargethispage{-5in}

% that's all folks
\end{document}

### 2013-12-12 code/src/framework/Population.java
not re-evaluting the best

@@ -121,24 +121,34 @@ public class Population
     {
         for(int i = 0; i < numIndividuals; ++i)
         {
-            int genes[] = population[i];
-            for(int k = 0; k < genes.length; ++k)
+            if(i != indexBest)
             {
-                for(int j = 0; j < 3; ++j)
+                int genes[] = population[i];
+                for(int k = 0; k < genes.length; ++k)
                 {
-                    EvoExec.currentWeights[k][j] = EvoExec.genes[genes[k]][j];
+                    for(int j = 0; j < 3; ++j)
+                    {
+                        EvoExec.currentWeights[k][j] = EvoExec.genes[genes[k]][j];
+                    }
                 }
-            }
 
-            double res[] = EvoExec.evaluate(trials, genes, i);
-            //double res[] = EvoExec.evaluateVisual();
-            //double res[] = new double[]{rnd.nextDouble(), rnd.nextDouble(), rnd.nextDouble()};
+                double res[] = EvoExec.evaluate(trials, genes, i);
+                //double res[] = EvoExec.evaluateVisual();
+                //double res[] = new double[]{rnd.nextDouble(), rnd.nextDouble(), rnd.nextDouble()};
 
-            int dominance = Utils.dominates(res, bestResult);
-            if(dominance == 1) //bestResult dominates res... but we are MINIMIZING!
-            {
-                System.arraycopy(res, 0, bestResult, 0, bestResult.length);
-                indexBest = i;
+                int dominance = Utils.dominates(res, bestResult);
+                if(dominance == 1) //bestResult dominates res... but we are MINIMIZING!
+                {
+                    System.arraycopy(res, 0, bestResult, 0, bestResult.length);
+                    indexBest = i;
+                }
+            }else{
+                System.err.print( EvoExec.generation + " " + indexBest + " ");
+                for(int idxBest = 0; idxBest < population[indexBest].length; ++idxBest)
+                {
+                    System.err.print(population[indexBest][idxBest]);
+                }
+                System.err.format(". Fitness: %.3f, %.3f, %.3f\n", bestResult[0],bestResult[1],bestResult[2]);
             }
         }
     }

### 2013-12-13 code/src/framework/EvoExec.java
main evo with params

@@ -284,7 +284,21 @@ public class EvoExec extends Exec
      */
     public static void main(String[] args)
     {
-        m_mapNames = new String[]{"maps/ptsp_map61.map"}; //Set here the name of the map to play in.
+        String[] allMaps = new String[]{"maps/ptsp_map01.map","maps/ptsp_map02.map","maps/ptsp_map08.map",
+                        "maps/ptsp_map19.map","maps/ptsp_map24.map","maps/ptsp_map35.map","maps/ptsp_map40.map",
+                        "maps/ptsp_map45.map","maps/ptsp_map56.map","maps/ptsp_map61.map"}; //In an array, to play in mutiple maps with runGames().
+
+        if(args.length > 0)
+        {
+            int mapIdx = Integer.parseInt(args[0]);
+            m_mapNames = new String[]{allMaps[mapIdx]};
+        }else
+        {
+            m_mapNames = new String[]{"maps/ptsp_map01.map"}; //Set here the name of the map to play in.
+        }
+
+        System.err.println("Running in map " + m_mapNames[0] + "...");
+
         //m_mapNames = new String[]{"maps/ptsp_map01.map","maps/ptsp_map02.map","maps/ptsp_map08.map",
         //        "maps/ptsp_map19.map","maps/ptsp_map24.map","maps/ptsp_map35.map","maps/ptsp_map40.map",
         //        "maps/ptsp_map45.map","maps/ptsp_map56.map","maps/ptsp_map61.map"}; //In an array, to play in mutiple maps with runGames().

### 2013-12-17 paper/paper.tex
MCTS algorithm and literature

@@
 
+Monte Carlo Tree Search (MCTS) is a tree search algorithm that was originally applied to board games, concretely to the two-players game of Go.
+This game is played in a square grid board, with a size of $19 \times 19$ in the original game, and $9 \times 9$ in its reduced version.
+The game is played in turns, and the objective is to surround the opponent's stones by placing stones in any available position in the board.
+Due to the very large branching factor of Go, this game is considered the drosophila of Game AI, and MCTS players have reached professional level play in the reduced board size version~\cite{Lee2009}.
+After its success in Go, MCTS has been used extensively by many researchers in this and different domains.
+An extensive survey of MCTS methods, variations and applications, has been written by Browne et al.~\cite{Browne2012}.
+
+MCTS is considered to be an \textit{anytime} algorithm, as it is able to provide a a valid next move to choose at any moment in time.
+This is true independently from how many iterations the algorithm was able to make (although, in general, more iterations usually produce better results).
+This differs from other algorithms (such A*) that normally provide the next ply only after they have finished.
+This makes MCTS a suitable candidate for real-time domains, where the decision time budget is limited, affecting the number of iterations that can be performed.
+
+MCTS is an algorithm that builds a tree in memory.
+Each node in the tree maintains statistics that indicate how often a move is played from a given state ($N(s,a)$), how many times each move is played from there ($N(s)$) and the average reward ($Q(s,a)$) obtained after applying move $a$ in state $s$.
+The tree is built iteratively by simulating actions in the game, making move choices based on the statistics store in the nodes.
+
+Each iteration of MCTS can be divided into several steps~\cite{Gelly2006}: \textit{Tree selection}, \textit{Expansion}, \textit{Monte Carlo simulation} and \textit{Back-propagation} (all summarized in Figure~\ref{fig:mcts}).
+When the algorithm starts, the tree is formed only by the root node, which holds the current state of the game.
+During the \textit{selection} step, the tree is navigated from the root until a maximum depth or the end of the game has been reached.
+
+\begin{figure} [!t]
+	\begin{center}
+	\includegraphics[scale=0.4,natwidth=708,natheight=484]{img/mcts.png}
+	\caption{MCTS algorithm steps.}
+	\label{fig:mcts}
+	\end{center}
+\end{figure}
+
+In every one of this action decisions, MCTS balances between exploitation and exploration.
+In other words, this chooses between taking an action that leads to states with the best outcome found so far, and performing a move to go to less explored game states, respectively.
+In order to achieve this, MCTS uses Upper Confidence Bound (UCB1, see Equation~\ref{eq:ucb1}) as a \textit{Tree Policy}.
+
+\begin{equation}	\label{eq:ucb1}
+a^* = \argmax_{a \in A(s)} \left\{Q(s,a) + C \sqrt{\frac{ \ln N(s) }{ N(s,a) }}\right\}
+\end{equation}
+
+The balance between exploration and exploitation is achieved by setting the value of $C$.
+Higher values of $C$ weight more the second term of the UCB1 Equation~\ref{eq:ucb1}, giving preference to those actions that have been explored less, at the expense of taking actions with the highest average reward $Q(s,a)$.
+A commonly used value is $\sqrt{2}$, as it balances both facets of the search when the rewards are normalized between $0$ and $1$.
+It is worth noting that MCTS, when combined with UCB1 reaches asymptotically logarithmic regret~\cite{coquelin2007bandit}.
+
+If, during the \textit{tree selection} phase, a node has less children than the available number of actions from a given position, a new node is added as a child of the current one (\textit{expansion} phase) and the \textit{simulation} step starts.
+At this point, MCTS executes a Monte Carlo simulation (or roll-out; \textit{default policy}) from the expanded node.
+This is performed by choosing random (either uniformly random, or biased) actions until the game end or a pre-defined depth is reached, where the state of the game is evaluated.
+
+Finally, during the \textit{back-propagation} step, the statistics $N(s)$, $N(s,a)$ and $Q(s,a)$ are updated for each node visited, using the reward obtained in the evaluation of the state.
+These steps are executed in a loop until a termination criteria is met (such as number of iterations).
+
+MCTS has been employed extensively in real-time games in the literature.
+A clear example of this is the popular real-time game \textit{Ms.
+PacMan}.
+The objective of this game is to control Ms.
+PacMan to clear the maze by eating all pills, without being captured by the ghosts.
+An important feature of this game is that it is \textit{open-ended}, as an end game situation is, most of the time, far ahead in the future and can not be devised by the algorithm during its iterations.
+The consequence of this is that MCTS, in its vanilla form, it is not able to know if a given ply will lead to a win or a loss game end state.
+Robles et al~\cite{Robles2009} solved this problem by including hand-coded heuristics that guided MCTS simulations towards more promising portions of the search space.
+Other authors also included domain knowledge to bias the search in MCTS, such as in~\cite{Samothrakis, Ikehata2010}.
+
+MCTS has also been applied to single-player games, like SameGame~\cite{Matsumoto2010}, where the player's goal is to destroy contiguous tiles of the same colour, distributed in a rectangular grid.
+Another use of MCTS is in the popular puzzle Morpion Solitaire~\cite{Edelkamp2010}, a connection game where the goal is to link nodes of a graph with straight lines that must contain at least five vertices.
+Finally, the PTSP has also been addressed with MCTS, both in the single-objective~\cite{Perez2013, Powley2012} and the multi-objective versions~\cite{Powley2013}.
+These papers describe the entries that won both editions of the PTSP Competition.
+
+It is worthwhile mentioning that in most cases found in the literature, MCTS techniques have been used with some kind of heuristic that guides the Monte Carlo simulations or the tree selection policy.
+In the algorithm proposed in this paper, simulations are purely random, as the objective is to compare the search abilities of the different algorithms.
+The intention is therefore to keep the heuristics to a minimum, and the existing pieces of domain knowledge are shared by all the algorithms presented (as in the case of the score function for MO-PTSP, described later).
+
 \section{Multi-Objective Optimization} \label{sec:moo}

### 2013-12-17 paper/paper.tex
MOO problems added, and literature review of those

@@
 
-\usepackage{graphicx,times,amsmath, fontenc, todonotes, url, array, multirow}
+\usepackage{subfigure,graphicx,url,times,multirow,amsmath,amssymb,algorithm,xspace,epsfig,todonotes,array,caption,color}
 \usepackage{algorithm, mathtools}
@@
 
+A multi-objective optimization problem (MOOP) represents a scenario where two or more objective functions are to be optimized (either maximized or minimized).
+The general form of a MOOP is formally described as a maximization function $F_m(x)$, that transforms points in the decision space ($X$) to points in the solution space ($F$).
+The elements of the decision space are vectors of $n$ variables of the form $x = (x_1, x_2, \dots, x_n)$, while elements in the solution space are vectors with a dimension $m$: $F_m(x) = (f_1(x), f_2(x), \dots, f_m(x))$.
+Therefore, each solution provides $m$ different scores (or rewards, or fitness) that are meant to be optimized.
+Without loss of generality, it is assumed from now on that all objectives must be maximized.
+
+It is said that a solution $F_m(x)$ \textit{dominates} another solution $F_m(y)$ if:
+
+\begin{enumerate}
+\item $F_m(x)$ is not worse than $F_m(y)$ in all objectives for all $i = 1, 2, \dots, m$.
+\item At least one objective of $F_m(x)$ is better than its analogous counterpart in $F_m(y)$.
+\end{enumerate}
+
+When this two conditions apply, it is said that $F_m(x) \preceq F_m(y)$ ($F_m(x)$ dominates $F_m(y)$, and $F_m(x)$ is non-dominated by $F_m(y)$).
+The \textit{dominance} condition provides a partial ordering between points in the solution space: if $F_m(x) \preceq F_m(y)$, then $F_m(x)$ is considered to be better than $F_m(y)$.
+
+However, there are some cases where it cannot be said that $F_m(x) \preceq F_m(y)$ or $F_m(y) \preceq F_m(x)$.
+This situation occurs when one of the objectives is better in $F_m(x)$ but a different objective is better in $F_m(y)$ (for instance, when $F_1(x) < F_1(y)$ but $F_2(x) > F_2(y)$).
+In this case, it is said that these solutions are non-dominated with respect to each other.
+Solutions that are not dominated by each other are grouped in a \textit{non-dominated set}.
+Given a non-dominated set $P$, it is said that $P$ is the \textit{optimal Pareto front} if there is no other solution in the solution space that dominates any member of $P$.
+The relation between decision and objective space, dominance and Pareto fronts is depicted in Figure~\ref{fig:moop}.
+
+\begin{figure} [!t]
+	\begin{center}
+	\includegraphics[scale=0.235]{img/moopDec.png}
+	\caption{Decision and Solution spaces in a MOOP with two variables ($x_1$ and $x_2$) and two objectives ($f_1$ and $f_2$).
+In the objective space, yellow dots are non-optimal objective vectors, while blue dots form the Pareto-optimal front.}
+	\label{fig:moop}
+	\end{center}
+\end{figure}
+
+As many Pareto front can be formed by several points in the solution space, it is important to devise a mechanism to assess the quality of a given front.
+A possibility is to use the Hypervolume Indicator (HV): given a Pareto front $P$, $HV(P)$ is defined as the volume of the objective space dominated by $P$.
+More formally, $HV(P) = \mu (x \in \mathbb{R}^{d} : \exists r \in P \hspace{0.25cm} s.t \hspace{0.25cm} r \preceq x)$, where $\mu$ is the de Lebesgue measure on $\mathbb{R}^{d}$.
+If the objectives are to be maximized, the higher the $HV(P)$, the better the front calculated.
+Figure~\ref{fig:hv} shows an example of $HV(P)$ where the objective dimension space is $2$.
+
+\begin{figure}[!t]
+\begin{center}
+\includegraphics[width=0.75\columnwidth]{img/hv}
+\end{center}
+\caption{$HV(P)$ of a given pareto front $P$.}
+\label{fig:hv}
+\end{figure}
+
+For more extensive descriptions, definitions, properties and multi-objective optimization in general, the reader can consult the work by K. Deb~\cite{Deb2001}.
+
+Many different algorithms have been proposed to tackle multi-objective optimization problems in the literature.
+One of the most widely known and used methods is the weighted-sum approach.
+The procedure consists of giving a weight to each one of the objective and produce a single result as the linear combination of objectives and weights.
+By varying the weights provided, it is possible to converge to different solutions of the optimal Pareto front, if this is convex.
+However, K.
+Deb~\cite{Deb2001} explains how linear scalarization approaches fail in those scenarios where the optimal Pareto front is non-convex.
+
+A popular choice for multi-objective optimization problems are evolutionary multi-objective optimization (EMOA) algorithms~\cite{Coello1999, Zhou2011}.
+One of the most well known algorithms in the literature is the Non-dominated Sorting Evolutionary Algorithm 2 (NSGA-2), which pseudocode is shown in Algorithm~\ref{alg:nsga2}.
+As in any evolutionary algorithm, NSGA-2 evolves a set of individuals or solutions to the problem, with the difference that here they are ranked according to dominance criteria and crowding distance (distances between members of the pareto fronts).
+A full description of the algorithm can be found in~\cite{Deb2000}.
+
+\begin{algorithm}[!t]
+\begin{algorithmic}[1]
+\Function{NSGA-2}{}
+	\State $P = NewRandomPopulation$
+	\While{Termination criteria not met}
+		\State $R = P \cup Q$ 
+		\State $F = \Call{fastNonDominatedSort}{R}$
+		\While{$|P| < N$}
+			\State $\Call{crowdingDistanceAssignment}{F_i}$	
+			\State $P = P \cup F_i$
+		\EndWhile
+		\State $\Call{sort}{P}$
+		\State $P = P[0:N]$
+		\State $Q = breed(P)$	
+	\EndWhile
+\EndFunction
+\end{algorithmic}
+\caption{NSGA-2 Algorithm.}
+\label{alg:nsga2}
+\end{algorithm}
+
+The three main pillars of the NSGA-2 algorithm are:
+
+\begin{itemize}
+\item A \textit{fast non-dominated sorting} algorithm, that ranks the individuals of the population and groups them in Pareto fronts.
+\item A \textit{crowding distance}, assigned to each one of the individuals, that measures how close it is to its neighbours.
+The selection genetic operator chooses individuals based on the ranks of the individuals and their crowding distance.
+\item \textit{Elitism}, implemented so the algorithm automatically promotes the best $N$ individuals to the next generation.
+\end{itemize}
+
+A more recent approach, developed by Q.
+Zhang, is the Multi-Objective Evolutionary Algorithm based on Decomposition (MOEA/D)~\cite{Zhang2007}, that decomposes the problem into several single optimization sub-problems and an evolutionary algorithm optimizes them all simultaneously.
+Information is shared between neighbouring sub-problems in order to guide evolution.
+The authors show that MOEA/D performs similarly than, and sometimes even outperforms, NSGA-II in the scenarios tested.
+
+Reinforcement Learning (RL) algorithms have also tackled Multi-objective optimization in some scenarios.
+RL~\cite{Sutton1998} is a broad field in Machine Learning that studies real-time planning and control situations where  an agent has to find out the actions (or sequences of actions) that should be applied in order to maximize the reward from the environment.
+
+An RL problem can be defined as a tuple ($S$, $A$, $T$, $R$, $\pi$).
+$S$ is the set of possible states in the problem (or game), and $s_0$ is the initial state.
+$A$ is the set of available actions the agent can make at any given time, and the transition model $T(s_i, a_i, s_{i+1})$ determines the probability of reaching the state $s_{i+1}$ when action $a_i$ is applied in state $s_i$.
+The reward function $R(s_i)$ provides a single value (\textit{reward}) that the agent must optimize, representing the desirability of the state $s_i$ reached.
+Finally, a decision policy $\pi(s_i) = a_i$ determines which actions $a_i$ must be chosen from each state $s_i \in S$.
+One of the most important challenges in RL, as shown in Section~\ref{sec:mcts}, is the trade-off between exploration and exploitation.
+The decision policy can choose between following actions that provided good rewards in the past and exploring new parts of the search space by selecting new actions.
+
+Multi-objective Reinforcement Learning (MORL)~\cite{Vamplew2010} changes this definition by using instead a vector $R = {r_0, r_1, \dots, r_n}$ as rewards of the problem.
+Thus, MORL problems differ from RL in having more than one objective objectives ($n$) that must be maximized.
+If the objectives are independent or they do not oppose each other, scalarization technique approaches, as described above, could be suitable to tackle the problem.
+Essentially, this would mean to use a conventional RL algorithm on a single objective obtained by a weighted-sum of the multiple rewards.
+However, this is not always the case, as it is usual that the objectives are in conflict and the policy ($\pi$) must balance among them.
+
+According to how $\pi$ approaches this problem, Vamplew et al.~\cite{Vamplew2010} proposed the following distinction for MORL: \textit{single-policy} algorithms are those that provide a preference order in the objectives available (given by the user or by the nature of the problem).
+An example of this type of algorithm can be found at~\cite{Gabor1998}, where the authors introduce an order of preference in the objectives treated and constraint the value of the rewards desired.
+Scalarization approaches would also fit in this category, as the work performed by S.
+Natarajan et al.~\cite{Natarajan2005}.
+
+The second type of algorithms, \textit{multiple-policy}, target to approximate the optimal Pareto front of the problem.
+An example of this type of algorithm is the one given by L.
+Barrett~\cite{Barrett2008}, who propose the Convex Hull Iteration Algorithm.
+This algorithm provides the optimal policy for any linear preference function, by learning all policies that define the convex hull of the Pareto front.
+
 \section{Multi-Objective Monte Carlo Tree Search} \label{sec:momcts}

### 2013-12-17 paper/paper.tex
MO-MCTS algorithm description

@@
 
+Adapting MCTS into Multi-Objective Monte Carlo Tree Search (MO-MCTS) requires the obvious modification of dealing with multiple rewards instead of just one.
+As these are collected at the end of a Monte Carlo simulation, the reward value $r$ now becomes a vector $R = {r_0, r_1, \dots, r_n}$, where $n$ is the number of objectives to optimize.
+Derived from this change, the average value $Q(s,a)$ becomes a vector that stores the average reward of each objective.
+Note that the other statistics ($N(s,a)$ and $N(s)$) do not need to change, as this are just node and action counters.
+The important question to answer next is how to adapt the vector $Q(s,a)$ to use it in the UCB1 formula (Equation~\ref{eq:ucb1}).
+
+An initial attempt on Multi-Objective MCTS was addressed by W.
+Wang and Michele Sebag~\cite{Weijia2012, Weijia2013}.
+In their work, the authors employ a mechanism, based on the HV calculation, to replace the UCB1 equation.
+The algorithm keeps a Pareto archive (P) with the best solutions found in game end states.
+Every node in the tree defines $\overline{r}_{sa}$ as a vector of UCB1 values, in which each $\overline{r}_{sa, i}$ is the result of calculating UCB1 for each particular objective $i$.
+
+The next step is to define the value for each pair of state and action, $W(s,a)$, as in Equation~\ref{eq:wang}.
+$\overline{r}^p_{sa}$ is the projection of $\overline{r}_{sa}$ into the piecewise linear surface defined by the Pareto archive $P$ (see Figure~\ref{fig:wangHV}.
+Then, $HV(P \cup \overline{r}_{sa})$ is declared as the HV of $P$ plus the point $\overline{r}_{sa}$.
+If $\overline{r}_{sa}$ is dominated by P, the distance between $\overline{r}_{sa}$ and $\overline{r}^p_{sa}$ is subtracted from the HV calculation.
+The tree policy selects actions based on a maximization of the value of $W(s,a)$.
+
+\begin{equation} \label{eq:wang}
+	W(s,a) = \left\{
+\begin{array}{l l}      
+    HV(P \cup \overline{r}_{sa}) - dist(\overline{r}^p_{sa}, \overline{r}_{sa} )\hspace{0.5cm} Otherwise \\
+    HV(P \cup \overline{r}_{sa}) \hspace{0.5cm} if \hspace{0.25cm} \overline{r}_{sa} \preceq P\\
+\end{array}\right.
+\end{equation}
+
+\begin{figure}[!t]
+\begin{center}
+\includegraphics[width=0.65\columnwidth]{img/wangHV}
+\end{center}
+\caption{$r^p_x$ is the projection of the $r_x$ value on the piecewise linear surface (discontinuous line). The shadowed surface represents $HV(P)$. From~\cite{Weijia2012}.}
+\label{fig:wangHV}
+\end{figure}
+
+The proposed algorithm was employed successfully in two domains: the DST and the Grid Scheduling problem, matching state of the art results in both domains, at the expense of a high computational cost.
+
+As mentioned before, the objective of this paper is to propose an MO-MCTS algorithm that is suitable for real-time domains.
+Although the work discussed here is influenced by Wang's approach, some modifications need to be done in order to overcome the high computational cost involved by their approach.
+
+In the algorithm proposed in this paper, the vector $\overline{r}$ of rewards that was obtained at the end of an Monte Carlo simulation is back-propagated through the nodes visited in the last iteration until the root is reached.
+In the vanilla algorithm, each node would use this vector $\overline{r}$ to update its own accumulated reward vector $\overline{R}$.
+Instead of doing this, each node in the MO-MCTS algorithm keeps a local Pareto front ($P$), updated at each iteration with the reward vector $\overline{r}$ obtained at the end of the simulation.
+Algorithm~\ref{alg:updPareto} describes how the node statistics are updated.
+
+\begin{algorithm}[!t]
+\begin{algorithmic}[1]
+\Function{Update}{$node, \overline{r}, dominated=false$}
+
+	\State $node.Visits = node.Visits + 1$
+	\State $node.\overline{R} = node.\overline{R} + \overline{r}$
+	\If{$!dominated$} 
+		\If{$node.P \preceq \overline{r}$}	
+			\State $dominated = true$
+		\Else
+			\State $node.P = node.P \cup \overline{r}$
+		\EndIf
+	\EndIf
+
+	\State $\Call{Update}{node.parent,\overline{r},dominated}$
+
+\EndFunction
+\end{algorithmic}
+\caption{Pareto MO-MCTS node update.}
+\label{alg:updPareto}
+\end{algorithm}
+
+In this algorithm, if $\overline{r}$ is not dominated by the local Pareto front, it is added to the front and $\overline{r}$ is propagated to its parent.
+In case $\overline{r}$ is dominated by the node's Pareto front, the local Pareto front and there is no need to keep this propagation up the tree.
+
+Three observations can be made about the mechanism described here:
+
+\begin{itemize}
+\item Each node in the tree has an estimate of the quality of the solutions reachable from there, both as an average (as in the baseline MCTS) and as the best case scenario (by keeping the non-dominated front $P$).
+\item By construction, if a reward $\overline{r}$ is dominated by the local front of a node, it is a given that it will be dominated by the nodes above in the tree, so there is no need to update the fronts of the upper nodes, producing little cost in the efficiency of the algorithm.
+\item It is easy to infer, from the last point, that the Pareto front of a node cannot be worse than the front of its children (in other words, the front of a child will never dominate that of its parent).
+Therefore, the root of the tree contains the best non-dominated front ever found during the search.
+\end{itemize}
+
+This last detail is important for two main reasons.
+First of all, the algorithm allows the root to store information indicating which is the action to take in order to converge to any specific solution in the front discovered.
+This information can be used, when all iterations have been performed, to select the move to perform next.
+If weights for the different objectives are provided, this weights can be used to select the desired solution in the Pareto front of the root node, and hence select the action that leads to that point.
+Secondly, the root's Pareto front can be used to measure the global quality of the search using the hypervolume calculation.
+
+Finally, the information stored at each node regarding the local Pareto front can be used to substitute $Q(s,a)$ in the UCB1 equation.
+The quality of a given pair $(s,a)$ can be obtained by measuring the HV of the Pareto front stored in the node reached from state $s$ after applying action $a$.
+This can be defined as  $Q(s,a) = HV(P)/N(s)$, and the Upper Confidence Bound equation, referred to here as $MO-UCB$, is described as in Equation~\ref{eq:moucb}).
+
+\begin{equation}	\label{eq:moucb}
+a^* = \argmax_{a \in A(s)} \left\{HV(P)/N(s) + C \sqrt{\frac{ \ln N(s) }{ N(s,a) }}\right\}
+\end{equation}
+
 \section{Benchmarks} \label{sec:bench}

### 2013-12-18 paper/paper.tex
benchmarks added, starting experimentation

@@
 
+Two different are used in this research to analyze the performance of the algorithm proposed.
+
+\subsection{Deep Sea Treasure} \label{ssec:dst}
+
+The Deep Sea Treasure (DST) is a well known multi-objective problem introduced by Vamplew et al.~\cite{Vamplew2010}.
+In this single-player puzzle, the agent moves a submarine with the objective of finding a treasure located at the bottom of the ship.
+The world is divided into a grid of $10$ rows and $11$ columns, and the vessel starts at the top left board position.
+There are three types of cells: empty cells (or water), that the submarine can traverse; ground cells that, as the edges of the grid, cannot be traversed; and treasure cells, that provide different rewards and finish the game.
+Figure~\ref{fig:dstBase} (on the left) shows the environment of the DST.
+
+\begin{figure}[!t]
+\begin{center}
+\includegraphics[width=0.75\columnwidth]{img/dst}
+\end{center}
+\caption{Environment of the Deep Sea Treasure (from~\cite{Vamplew2010}): grey squares represent the treasure (with their different values) available in the map.
+The black cells are the sea floor and the white ones are the positions that the vessel can occupy freely.
+The game ends when the submarine picks one of the treasures.}
+\label{fig:dstBase}
+\end{figure}
+
+The ship can perform four different moves: \textit{up}, \textit{down}, \textit{right}  and \textit{left}.
+If the action applied takes the ship off the grid or into the sea floor, the vessel's position does not change.
+There are two objectives in this game: the number of moves performed by the ship, that must be minimized, and the value of the treasure found, that should be maximized.
+As can be seen in Figure~\ref{fig:dstBase}, the most valuable treasures are at a greater distance from the initial position, making these objectives orthogonal.
+
+Additionally, the agent can only make up to $100$ moves.
+This allows the problem to be defined as the maximization of two rewards: $(M,T) = (100 - moves, treasureValue)$.
+Should the ship perform all moves without reaching a treasure, the result would be ($0$,$0$).
+At each step, the score of a location with no treasure is ($-1$, $0$).
+
+The optimal Pareto front of the DST is shown in Figure~\ref{fig:optDst}.
+There are $10$ non-dominated solutions in this front, one per each treasure in the board.
+The front is globally concave, with local concavities at the second $(83,74)$, fourth $(87,24)$ and sixth $(92,8)$ points from the left.
+The HV value of the optimal front is $10455$.
+
+\begin{figure}[!t]
+\begin{center}
+\includegraphics[width=0.9\columnwidth]{img/optDst.png}
+\end{center}
+\caption{Optimal Pareto Front of the Deep Sea Treasure, with both objectives to be maximized. From~\cite{Weijia2012}.}
+\label{fig:optDst}
+\end{figure}
+
+Section~\ref{sec:moo} introduced the problems of linear scalarization approaches when facing non-convex optimal Pareto front.
+The concave shape of the DST's optimal front makes those approximations to converge to the non dominated solutions located at the edges of the Pareto front ($(-19,1)$,$(-1,124)$).
+Note that this happens independently from the weights chosen for the linear approximation: some solutions of the front just can't be reached with these approaches.
+Thus, successful approaches should be able to find all elements of the optimal Pareto front and converge to any of the non dominated solutions.
+
+\subsubsection{Transposition Tables in DST} \label{ssec:transTables}
+
+The DST is a problem specially suited for the use of Transposition Tables (TT)~\cite{Childs2008} within MCTS.
+TT is a technique used to optimize three search based algorithms when two or more states can be found at different locations in the tree.
+It consists of sharing information between these equivalent states in a centralized manner, in order to avoid managing these positions as completely different states.
+Figure~\ref{fig:dstTable} shows an example of this situation in the DST.
+
+\begin{figure}[!t]
+\begin{center}
+\includegraphics[width=0.9\columnwidth]{img/dstTransTables}
+\end{center}
+\caption{Example of two different sequences of actions (R: Right, D: Down) that lie in the same position in the map, but different node in the tree.}
+\label{fig:dstTable}
+\end{figure}
+
+In this example, the submarine starts in the initial position ($P1$, root of the tree) and makes a sequence of moves that places it in $P2$.
+From this location, if the vessel wishes to move to $P3$, two optimal trajectories would be route \textit{a} and \textit{b}.
+In a tree that does not use TT, there would be two different nodes to represent $P3$, although the location and number of moves performed up to this point are the same (thus, the states are equivalent).
+It is worthwhile to highlight that the coordinates of the vessel are not enough to identify two equivalent states, but also the number of moves is needed: imagine a third route from $P2$ to $P3$ with the moves: \textit{Up}, \textit{Right}, \textit{Right}, \textit{Down}, \textit{Down}.
+As the submarine performs $5$ moves, the states are not the same, and the node where the ship is in $P3$ now would be two levels deeper in the tree.
+
+TT tables are implemented in the MCTS algorithms tested in this benchmark by using hash tables that stores a \textit{representative} node for each pair $(position, moves)$ found.
+The key of the hash map needs then to be obtained from three values: position coordinates $x$ and $y$ of the ship in the board, and number of moves, indicated by the depth of the node in the tree.
+Hence, transpositions can only happen at the same depth within the tree, a feature that has been successfully tried before in the literature~\cite{Kozelek2009}.
+
+\subsection{Multi-Objective PTSP} \label{ssec:moptsp}
+
+The Multi-Objective Physical Travelling Salesman Problem (MO-PTSP) is a game employed in a competition held in the IEEE Conference on Computational Intelligence in Games (CIG) in 2013, a modification of the Physical Travelling Salesman Problem (PTSP), previously introduced by Perez et al.~\cite{PerezCEC2012}.
+The MO-PTSP is a real-time game where the agent navigates a ship and must visit $10$ waypoints scattered around the maze.
+
+This needs to be done while optimizing three different objectives:  1) \textbf{Time}: the player must collect all waypoints scattered around the maze in as few time steps as possible; 2) \textbf{Fuel}: the fuel consumption by the end of the game must be minimized; and 3) \textbf{Damage}: the ship should end the game with as little damage as possible.
+
+In the game, the agent must provide an action every $40$ milliseconds.
+The available actions are combinations of two different inputs: \textit{throttle} (that could be \textit{on} or \textit{off}) and \textit{steering} (that could be \textit{straight}, \textit{left} or \textit{right}).
+This allows for $6$ different actions that modify the ship's position, velocity and orientation.
+These vectors are kept from one step to the next, keeping the inertia of the ship, and making the navigation task not trivial.
+
+The ship starts with $5000$ units of fuel, and one unit is spent every time an action supplied has the throttle input \textit{on}.
+There are, however, two ways of collecting fuel: each waypoint visited grants the ship with $50$ more units of fuel.
+Also, fuel canisters are scattered around the maze and their collection provides with $250$ more units.
+
+Regarding the third objective, the ship can suffer damage in two different ways: by colliding with obstacles and driving through lava.
+In the former case, the ship can collide with normal obstacles (that subtract $10$ units of damage) and specially damaging obstacles ($30$ units).
+In the latter, lava lakes are abundant in the MO-PTSP levels and, in contrast with normal surfaces, they deal one unit of damage for each time step the ship spends over this type of surface.
+All this subtractions are deduced from an initial counter of $5000$ points of damage.
+
+Figure~\ref{fig:sampleMap} shows an example of an MO-PTSP map, as drawn by the framework.
+Waypoints yet not visited are painted as blue circles, while those already collected are depicted as empty circles.
+The green ellipses represent fuel canisters, normal surfaces are drawn in brown and lava lakes are printed as red-dotted yellow surfaces.
+Normal obstacles black, damaging obstacles are drawn in red.
+Blue obstacles are elastic walls that produce no damage to the ship.
+The vessel is drawn as a blue polygon and its trajectory is traced with a black line.
+
+\begin{figure} [!t]
+	\begin{center}
+	\includegraphics[width=0.7\columnwidth]{img/moptspmap}
+	\caption{Sample MO-PTSP map.}
+	\label{fig:sampleMap}
+	\end{center}
+\end{figure}
+
+\subsubsection{Macro-actions for MO-PTSP} \label{sssec:macro}
+
+All algorithms tested in the MO-PTSP in this research employ macro-actions.
+Macro-actions is a concept that can be used for coarsening the action space by applying the action chosen by the control algorithm in several consecutive cycles, instead of just in the next one.
+
+Previous research in PTSP~\cite{Perez2013, Perez2012} suggests that using macro-actions for real-time navigation domains increases the performance of the algorithms, and it has been used in the previous PTSP competitions by the winner and other entries.
+
+A macro-action of length $L$ is defined as a repetition of a given action during $L$ consecutive time steps.
+The main advantage of macro-actions is that the control algorithms can see further in the future, and then reduce the implication of open-endedness of real-time games (see Section~\ref{sec:mcts}).
+As this reduces the search space significantly, better algorithm performance is allowed.
+A consequence of using macro-actions is also that the algorithm can employ $L$ consecutive steps to plan the next move (the next macro-action) to make.
+Therefore, instead of spending $40$ milliseconds, as in MO-PTSP, to define the next action, the algorithm can employ $40 \times L$ milliseconds for this task, and hence perform a more extensive search.
+
+In MO-PTSP, the macro-action size is $L = 15$, a value that has shown its proficiency before in PTSP.
+For more information about macro-actions and their application to real-time navigation problems such as the PTSP, the reader is referred to~\cite{Perez2013}.
+
+\subsubsection{Heuristics for MO-PTSP} \label{sssec:heur}
+
+\todo[inline]{Explain heuristics used here to evaluate the states.}
+
 \section{Experimentation} \label{sec:exp}
+
+The experiments performed in this research compare three different algorithms in the two benchmarks presented in Section~\ref{sec:bench}: a single objective MCTS (referred to here as \textit{MCTS}), the Multi-Objective MCTS (\textit{MO-MCTS}) and a rolling horizon version of the NSGA-II algorithm described in Section~\ref{sec:moo} (\textit{NSGA-II}).
+This NSGA-II version evolves a population where the individuals are sequences of actions (macro-actions in the MO-PTSP case), obtaining the fitness from the state of the game after applying such sequence.
+
+All algorithms have a limited number of evaluations before providing an action to perform.
+In order for these games to be real-time, the time budget allowed is close to $40$ milliseconds.
+With the objective of avoiding congestion peaks at the machine where the experiments are run, an average of the number of evaluations doable in $40$ ms.
+is calculated and employed in the tests.
+This leaded to $4500$ evaluations in the DST and $500$ evaluations for MO-PTSP, in the same server where the PTSP and MO-PTSP competitions were run\footnote{Intel Core i5, 230GHz, 4GB of RAM}.
+
+\subsection{Results in DST} \label{ssec:resDST}
+
+\subsection{Results in MO-PTSP} \label{ssec:resMOPTSP}
+
+\subsection{A step further in MO-PTSP: dynamically varying weights} \label{ssec:resVarW}
 

### 2013-12-18 code/src/framework/ExecSync.java

@@ -353,10 +353,10 @@ public class ExecSync extends Exec
      */
     public static void main(String[] args)
     {
-        //m_mapNames = new String[]{"maps/ptsp_map19.map"}; //Set here the name of the map to play in.
-        m_mapNames = new String[]{"maps/ptsp_map01.map","maps/ptsp_map02.map","maps/ptsp_map08.map",
-                "maps/ptsp_map19.map","maps/ptsp_map24.map","maps/ptsp_map35.map","maps/ptsp_map40.map",
-                "maps/ptsp_map45.map","maps/ptsp_map56.map","maps/ptsp_map61.map"}; //In an array, to play in mutiple maps with runGames().
+        m_mapNames = new String[]{"maps/ptsp_map02.map"}; //Set here the name of the map to play in.
+        //m_mapNames = new String[]{"maps/ptsp_map01.map","maps/ptsp_map02.map","maps/ptsp_map08.map",
+        //        "maps/ptsp_map19.map","maps/ptsp_map24.map","maps/ptsp_map35.map","maps/ptsp_map40.map",
+        //        "maps/ptsp_map45.map","maps/ptsp_map56.map","maps/ptsp_map61.map"}; //In an array, to play in mutiple maps with runGames().
 
         m_controllerName = "controllers.greedy.GreedyController"; //Set here the controller name.
         m_controllerName = "controllers.MacroRandomSearch.MacroRSController"; //Set here the controller name.
@@ -375,8 +375,8 @@ public class ExecSync extends Exec
 
 
         /////// 1. To play the game with the key controller.
-        //int delay = PTSPConstants.DELAY;  //PTSPConstants.DELAY: best human play speed
-        //playGame(delay);
+        int delay = PTSPConstants.DELAY;  //PTSPConstants.DELAY: best human play speed
+        playGame(delay);
 
         /////// 2. Executes one game.
         //ParetoMCTSController.EXPLORATION_VIEW_ON = true;
@@ -392,10 +392,11 @@ public class ExecSync extends Exec
 
         ////// 3. Executes N games (numMaps x numTrials), graphics disabled.
         //m_writeOutput = true;
-        m_verbose = false; //hides additional output. runGamesStats prints anyway.
-        int numTrials=30;
-        runGamesStats(numTrials);
+        //m_verbose = false; //hides additional output. runGamesStats prints anyway.
+        //int numTrials=30;
+        //runGamesStats(numTrials);
 
     }
 
 }
+

### 2013-12-18 paper/paper.tex
benchmark heuristics, dst pics pre-nsga2, moptsp results table

@@
 
-\usepackage{subfigure,graphicx,url,times,multirow,amsmath,amssymb,algorithm,xspace,epsfig,todonotes,array,caption,color}
+\usepackage{bm,subfigure,graphicx,url,times,multirow,amsmath,amssymb,algorithm,xspace,epsfig,todonotes,array,caption,color}
 \usepackage{algorithm, mathtools}
@@
 
+\subsubsection{Heuristics for DST} \label{sssec:heurDST}
+
+As DST has two different objectives, number of moves and value of the treasure, the quality of a state can be assessed by two rewards, $r_m$ and $r_v$, for each objective respectively.
+$r_m$ is adjusted to be maximized using the maximum number of moves in the game, while $r_v$ is just the value in the cell with the treasure.
+Equation~\ref{eq:dst} summarizes these rewards:
+
+\begin{equation}	\label{eq:dst}
+\begin{split}
+r_m = 100 - moves \\
+r_v = treasureValue
+\end{split}
+\end{equation}
+
 \subsection{Multi-Objective PTSP} \label{ssec:moptsp}
@@
 The MO-PTSP is a real-time game where the agent navigates a ship and must visit $10$ waypoints scattered around the maze.
+All waypoints must be visited to consider a game as complete, and a game ticks timer is reset every time a waypoint is collected, finishing the game prematurely (and unsuccessfully) if it reaches $800$ game steps before visiting another waypoint.
 
@@
 
-\subsubsection{Heuristics for MO-PTSP} \label{sssec:heur}
-
-\todo[inline]{Explain heuristics used here to evaluate the states.}
+\subsubsection{Heuristics for MO-PTSP} \label{sssec:heurMOPTSP}
+
+In order to evaluate a game state in MO-PTSP, three different measures or rewards are taken, $r_t, r_f, r_d$, one for each one of the objectives time, fuel and damage, respectively.
+All these rewards are defined so they have to be maximized.
+The first reward uses a measure of distance for the time objective, as indicated in Equation~\ref{eq:dist}:
+
+\begin{equation}	\label{eq:dist}
+r_t = 1 - d/D
+\end{equation}
+
+The MO-PTSP framework includes a path-finding library that allows controllers to query the shortest distance of a route of waypoints.
+$d$ indicates the distance from the current position until the last waypoint, following the desired route, and $D$ is the distance of the whole route from the starting position.
+Minimizing the distance to the last waypoint will lead to the end of the game.
+
+Equation~\ref{eq:fuel} shows how the value of the fuel objective, $r_f$, is obtained:
+
+%f_p = fuelPoints
+%f_c = fuelConsumed
+%f_0 = initialFuel
+
+\begin{equation}	\label{eq:fuel}
+r_f = (1 - (f_c/f_0)) \times \alpha + r_t \times (1 - \alpha)
+\end{equation}
+
+$f_c$ is the fuel consumed so far, and $f_0$ is the initial fuel at the start of the game.
+$\alpha$ is a value that balances between the fuel component and the time objective from Equation~\ref{eq:dist}.
+Note that, as waypoints needs to keep being visited, it is necessary to include a distance measure for this reward.
+Otherwise, an approach that prioritizes this objective would not minimize distance to waypoints at all (the ship could just stand still: no fuel consumed is optimal), and therefore could not complete the game.
+The value of $\alpha$ has been determined empirically, in order to provide a good balance between these two components, and it is set to $0.66$.
+
+Finally, Equation~\ref{eq:dam} gives the method to calculate the damage objective, $r_d$:
+
+%d_p = damagePoints
+%d_s = damageSuffered
+%d_m = maximumDamage
+%s = shipSpeed
+
+\begin{equation}	\label{eq:dam}
+r_d = 
+\begin{cases}
+	(1 - (d_s/d_m)) \times \beta_{1} + r_t \times (1 - \beta_{1}), & s > \gamma\\
+	(1 - (d_s/d_m)) \times \beta_{2} + r_t \times (1 - \beta_{2}), & s \leq \gamma
+\end{cases}
+\end{equation}
+
+$d_s$ is the damage suffered so far in the game, and $d_m$ is the maximum damage the ship can take.
+In this case, three different variables are used to regulate the bahaviour of this objective: $\gamma$, $\beta_{1}$ and $\beta_{2}$.
+Both $\beta_{1}$ and $\beta_{2}$ have the same role as $\alpha$ in Equation~ref{eq:fuel}: they balance between the time objective and the damage measure.
+The difference is that $\beta_{1}$ is used in high speeds, while $\beta_{2}$ is employed with low velocities.
+This is distinguished by the parameter $\gamma$, that can be seen as a threshold value for the ship's speed ($s$).
+This differentiation is made in order to avoid low speeds in lava lakes, as this increase the damage suffered importantly.
+The values for these variables have been determined also empirically, and they are set to $\gamma = 0.8$, $\beta_{1} = 0.75$ and $\beta_{2} = 0.25$.
 
@@
 
-The experiments performed in this research compare three different algorithms in the two benchmarks presented in Section~\ref{sec:bench}: a single objective MCTS (referred to here as \textit{MCTS}), the Multi-Objective MCTS (\textit{MO-MCTS}) and a rolling horizon version of the NSGA-II algorithm described in Section~\ref{sec:moo} (\textit{NSGA-II}).
+The experiments performed in this research compare three different algorithms in the two benchmarks presented in Section~\ref{sec:bench}: a single objective MCTS (referred to here simply as \textit{MCTS}), the Multi-Objective MCTS (\textit{MO-MCTS}) and a rolling horizon version of the NSGA-II algorithm described in Section~\ref{sec:moo} (\textit{NSGA-II}).
 This NSGA-II version evolves a population where the individuals are sequences of actions (macro-actions in the MO-PTSP case), obtaining the fitness from the state of the game after applying such sequence.
@@
 
+Figure~\ref{fig:extCase}
+
+\begin{figure*}[!t]
+	\centering
+	\includegraphics[width=2.1\columnwidth]{img/dstAll}
+	\caption{Results DST}
+	\label{fig:extCase}
+\end{figure*}
+
 \subsection{Results in MO-PTSP} \label{ssec:resMOPTSP}
+
+Table~\ref{tab:resMOPTSP}
+
+\begin{table*}[!t]
+\begin{center}
+\begin{tabular}{|c|c|c|c|c|c|}
+\hline
+  & \textbf{$W: (w_t, w_f, w_d)$} & \textbf{MO-MCTS} \bm{$(D, \O, d)$} &  \textbf{MCTS} \bm{$(D, \O, d)$} & \textbf{NSGA-II} \bm{$(D, \O, d)$} & \textbf{PurofMovio} \bm{$(D, \O, d)$} \\ 
+\hline
+ \multirow{3}{*}{\textbf{MO-MCTS}} & $W_0: (0.33,0.33,0.33)$ & \multirow{3}{*}{$-$} & $(8,2,0)$ & $(8,2,0)$ & $(0,5,5)$\\
+ & $W_1: (0.1,0.3,0.6)$ & & $(10,0,0)$ & $(4,6,0)$ & $(0,6,4)$\\
+ & $W_2: (0.1,0.6,0.3)$ & & $(8,2,0)$ & $(7,3,0)$ & $(0,5,5)$\\
+\hline
+ \multirow{3}{*}{\textbf{MCTS}} & $W_0: (0.33,0.33,0.33)$& $(0,8,2)$ & \multirow{3}{*}{$-$} & $(0,2,8)$ & $(0,2,8)$\\
+ & $W_1: (0.1,0.3,0.6)$ & $(0,0,10)$ & & $(4,2,4)$ & $(0,3,7)$\\
+ & $W_2: (0.1,0.6,0.3)$ & $(0,2,8)$ & & $(0,1,9)$ & $(0,6,4)$\\
+\hline
+ \multirow{3}{*}{\textbf{NSGA-II}} & $W_0: (0.33,0.33,0.33)$& $(0,2,8)$ & $(8,2,0)$ & \multirow{3}{*}{$-$} & $(0,4,6)$\\
+ & $W_1: (0.1,0.3,0.6)$ & $(0,6,4)$ & $(4,2,4)$ & & $(0,4,6)$\\
+ & $W_2: (0.1,0.6,0.3)$ & $(0,3,7)$ & $(9,1,0)$ & & $(0,5,5)$\\
+\hline
+ \multirow{3}{*}{\textbf{PurofMovio}} & $W_0: (0.33,0.33,0.33)$& $(5,5,0)$ & $(8,2,0)$ & $(6,4,0)$ & \multirow{3}{*}{$-$}\\
+ & $W_1: (0.1,0.6,0.3)$ & $(4,6,0)$ & $(7,3,0)$ & $(6,4,0)$ & \\
+ & $W_2: (0.1,0.3,0.6)$ & $(5,5,0)$ & $(6,4,0)$ & $(5,5,0)$ & \\
+\hline
+\end{tabular}
+\caption{Results MO-PTSP }
+\label{tab:resMOPTSP}
+\end{center}
+\end{table*}
 

### 2013-12-20 paper/paper.tex
first results in

@@
 
+This algorithm, same as NSGA-II due to their multi-objective nature, provide a non-dominated front as a solution.
+However, in planning and control scenarios like the games analyzed in this research, an action must be provided to perform a move in the next step.
+The question that arises is how to choose the move to make with the information available.
+
+As shown before, it is straightforward to obtain which actions lead to what points in the front given as a solution: first gene in an NSGA-II individual, a root's child in MO-MCTS.
+Hence, by identifying the point in the Pareto front that the algorithm should converge to,
+it is possible to execute the action that leads to that point.
+
+In order to select the point in the Pareto front, a weight vector $W (\forall w_i \in W, w_i \in [0,1])$ can be provided with the number of objectives as its dimension.
+Two different mechanisms are proposed, for reasons that will be explained in the experiments section (\ref{sec:exp}):
+
+\begin{itemize}
+\item Weighted Sum: the action chosen is the one that maximizes the weighted sum of the solution vector multiplied by $W$, for each point in the front.
+\item Euclidean distance: the euclidean distance from each point in the Pareto front (normalized in $[0,1]$) to the vector $W$ is calculated.
+The action to choose would be the one that leads to the point in the Pareto front with the smaller distance to $W$.
+\end{itemize}
+
+Note that, in the vanilla MCTS, there is no Pareto front obtained as a solution.
+Typically, in this case, rewards are calculated as a weighted sum of the objectives and a weight vector $W$, and then the action is chosen following any of the mechanisms usually employed in the literature: the action taken more often from the root, the one that leads to the best reward found, the move with the highest expected reward or the action that maximizes Equation~\ref{eq:ucb1} in the root.
+
 \section{Benchmarks} \label{sec:bench}
@@
 \begin{center}
-\includegraphics[width=0.9\columnwidth]{img/optDst.png}
+\includegraphics[width=1.0\columnwidth]{img/optDst.png}
 \end{center}
-\caption{Optimal Pareto Front of the Deep Sea Treasure, with both objectives to be maximized. From~\cite{Weijia2012}.}
+\caption{Optimal Pareto Front of the Deep Sea Treasure, with both objectives to be maximized.}
 \label{fig:optDst}
@@
 In order for these games to be real-time, the time budget allowed is close to $40$ milliseconds.
-With the objective of avoiding congestion peaks at the machine where the experiments are run, an average of the number of evaluations doable in $40$ ms.
+With the objective of avoiding congestion peaks at the machine where the experiments are run, the average of number of evaluations doable in $40$ ms.
 is calculated and employed in the tests.
-This leaded to $4500$ evaluations in the DST and $500$ evaluations for MO-PTSP, in the same server where the PTSP and MO-PTSP competitions were run\footnote{Intel Core i5, 230GHz, 4GB of RAM}.
+This leaded to $4500$ evaluations in the DST, and $500$ evaluations for MO-PTSP, using the same server where the PTSP and MO-PTSP competitions were run\footnote{Intel Core i5, 230GHz, 4GB of RAM}.
 
@@
 
-Figure~\ref{fig:extCase}
+As the optimal Pareto front of the DST is known, a measure of performance can be obtained by observing the percentage of times these solutions are found by the agents.
+As the solution that the algorithms converge to depends on the weights vector employed during the search (see end of Section~\ref{sec:momcts}), the approach taken here is to provide different weight vectors $W$ and analyze them separately.
+
+The weight vector for DST has two dimensions.
+This vector is here referred to as $W = (w_m, w_v)$, where $w_m$ weights moves and $w_v$ the treasure value, and $w_v = 1 - w_m$.
+$w_m$ takes values between $0$ and $1$, with a precision of $0.01$, and $100$ experiments have been run for each pair $(w_m, w_v)$.
+Hence, the game has been played a total of $10000$ times, for each algorithm.
 
@@
 	\includegraphics[width=2.1\columnwidth]{img/dstAll}
-	\caption{Results DST}
+	\caption{Results in DST: percentages of each optima found during $100$ games played with different weight vectors.
+Scalarization approaches converge to the edges of the optimal front, whereas Pareto approaches are able to find all optimal solutions.
+The proposed algorithm, MO-MCTS, finds these solutions significantly more often than NSGA-II.}
 	\label{fig:extCase}
@@
 
+Figure~\ref{fig:extCase} shows the results obtained after these experiments were performed.
+The first point to notice is that MCTS only converges (mostly) to the two optimal points located at the edges of the optimal Pareto front (\textit{OPT-1} and \textit{OPT-10}, see also Figure~\ref{fig:optDst}).
+This is an expected result, as K.
+Deb.
+suggested in~\cite{Deb2001} and was also discussed before in Section~\ref{sec:moo}: linear scalarization approaches only converge to the edges of the optimal Pareto front if its shape is non-convex.
+
+The results show clearly how approximating the optimal Pareto front allows for finding all possible solutions.
+Both NSGA-II and MO-MCTS approaches are able to converge to any solution in the front given the appropriate weight vector.
+It is important to highlight that these two algorithms employed the euclidean distance action selection (see Section~\ref{sec:momcts}).
+Experiments shown that weighted sum action selection provides similar results as in MCTS (this is, convergence to the edges of the front).
+The crucial distinction to make is that both algorithms NSGA-II and MO-MCTS allow for different action selection mechanisms, that are able to overcome this problem, by approximating a global Pareto front.
+
+Finally, in the comparison between NSGA-II and MO-MCTS, the latter algorithm obtains higher percentages for each one of the points in the front.
+This result suggests that, with a limited number of iterations/evaluations, the proposed algorithm is able to explore the search space more efficiently than a state of the art algorithm such NSGA-II.
+
 \subsection{Results in MO-PTSP} \label{ssec:resMOPTSP}
 
-Table~\ref{tab:resMOPTSP}
+The same three algorithms have been tested in the MO-PTSP domain.
+The framework~\footnote{www.ptsp-game.net} comes with $10$ maps, used in this research to run the algorithms in.
+In this case, the optimal Pareto front is not known in advance, and normally it will be different from one map to another.
+Hence, the mechanisms to compare the performance of the algorithms tested need to be different than the one used for DST.
+
+The idea is as follows.
+First of all, three different weight vectors are tested: $W_1 = (0.33,0.33,0.33)$, $W_2 = (0.1,0.3,0.6)$ and $W_3 = (0.1,0.6,0.3)$, where each $w_i$ corresponds to the weight for an objective ($w_t$ for time, $w_f$ for fuel and $w_d$ for damage).
+$W_1$ treats all objectives with the same weight, while $W_2$ and $W_3$ give more relevance to damage and fuel, respectively (note that Equations~\ref{eq:fuel} and~\ref{eq:dam} also give weight to the time objective, so having a vector with $w_t=0.6$ is not that relevant).
+All $W_1$, $W_2$ and $W_3$ provide then a wide spectrum for weights in this benchmark.
+In this case, MO-MCTS uses this weights to select an action based on a weighted sum, which has shown in preliminary experiments better performance than the Euclidean distance mechanism.
+
+The first point to check is if the different weight vectors affect the solutions obtained by MO-MCTS.
+Table~\ref{tab:weights} shows the results of executing the MO-MCTS controller during $30$ games in the $10$ maps of the MO-PTSP, for every weight vector.
+It can be seen that the highest weight in $W$ leads, in most of the cases, to the best solution in that objective in the map.
+
+\begin{table}[!t]
+\begin{center}
+%\begin{tabular}{|>{\centering\arraybackslash}m{0.65cm}|c|c|c|c|}
+\begin{tabular}{|p{0.8cm}|p{2.2cm}|m{1.13cm}|m{1.13cm}|m{1.13cm}|}
+\hline
+\textbf{Map}  & \textbf{$W:(w_t, w_f, w_d)$} & \textbf{Time} &  \textbf{Fuel} & \textbf{Damage} \\ 
+\hline
+ \multirow{3}{*}{\textbf{Map 1}} & $(0.33,0.33,0.33)$ & $1654 \pm 7$ & $131 \pm 2$& $846 \pm 13$\\
+ & $(0.1,0.3,0.6)$ & $1657 \pm 8$ & $130 \pm 2$ & \bm{$773 \pm 11$}\\
+ & $(0.1,0.6,0.3)$ & $1681 \pm 11$ & $131 \pm 2$ & $837 \pm 15$\\
+% & $(0.6,0.1,0.3)$ & $1627 \pm 7$ & $138 \pm 3$ & $812 \pm 14$\\
+\hline
+ \multirow{3}{*}{\textbf{Map 2}} & $(0.33,0.33,0.33)$ & $1409 \pm 7$ & $235 \pm 4$& $364 \pm 3$\\
+ & $(0.1,0.3,0.6)$ & $1402 \pm 6$ & $236 \pm 5$ & \bm{$354 \pm 2$}\\
+ & $(0.1,0.6,0.3)$ & $1416 \pm 8$ & \bm{$219 \pm 4$} & $360 \pm 3$\\
+% & $(0.6,0.1,0.3)$ & $1389 \pm 7$ & $241 \pm 5$ & $355 \pm 2$\\
+\hline
+ \multirow{3}{*}{\textbf{Map 3}} & $(0.33,0.33,0.33)$ & $1373 \pm 6$ & $221 \pm 3$& $301 \pm 7$\\
+ & $(0.1,0.3,0.6)$ & $1378 \pm 5$ & $211 \pm 3$ & \bm{$268 \pm 5$}\\
+ & $(0.1,0.6,0.3)$ & $1385 \pm 6$ & \bm{$203 \pm 4$} & $291 \pm 7$\\
+% & $(0.6,0.1,0.3)$ & $1383 \pm 6$ & $245 \pm 5$ & $300 \pm 7$\\
+\hline
+ \multirow{3}{*}{\textbf{Map 4}} & $(0.33,0.33,0.33)$ & $1383 \pm 6$ & $291 \pm 5$& $565 \pm 5$\\
+ & $(0.1,0.3,0.6)$ & $1385 \pm 7$ & $304 \pm 4$ & \bm{$542 \pm 4$}\\
+ & $(0.1,0.6,0.3)$ & $1423 \pm 6$ & \bm{$273 \pm 4$} & $583 \pm 5$\\
+% & $(0.6,0.1,0.3)$ & $1398 \pm 7$ & $325 \pm 5$ & $587 \pm 6$\\
+\hline
+ \multirow{3}{*}{\textbf{Map 5}} & $(0.33,0.33,0.33)$ & $1405 \pm 7$ & $467 \pm 4$& $559 \pm 4$\\
+ & $(0.1,0.3,0.6)$ & $1431 \pm 9$ & $447 \pm 4$ & \bm{$541 \pm 4$}\\
+ & $(0.1,0.6,0.3)$ & $1467 \pm 9$ & \bm{$411 \pm 5$} & $567 \pm 5$\\
+% & $(0.6,0.1,0.3)$ & $1410 \pm 6$ & $480 \pm 5$ & $546 \pm 4$\\
+\hline
+ \multirow{3}{*}{\textbf{Map 6}} & $(0.33,0.33,0.33)$ & $1575 \pm 7$ & $549 \pm 5$& $303 \pm 4$\\
+ & $(0.1,0.3,0.6)$ & $1626 \pm 9$ & $540 \pm 6$ & \bm{$286 \pm 5$}\\
+ & $(0.1,0.6,0.3)$ & $1703 \pm 11$ & \bm{$499 \pm 4$} & $316 \pm 7$\\
+% & $(0.6,0.1,0.3)$ & $1563 \pm 7$ & $560 \pm 5$ & $285 \pm 4$\\
+\hline
+ \multirow{3}{*}{\textbf{Map 7}} & $(0.33,0.33,0.33)$ & $1434 \pm 5$ & $599 \pm 6$& $284 \pm 6$\\
+ & $(0.1,0.3,0.6)$ & $1475 \pm 10$ & $602 \pm 5$ & \bm{$243 \pm 6$}\\
+ & $(0.1,0.6,0.3)$ & $1489 \pm 12$ & \bm{$549 \pm 3$} & $264 \pm 6$\\
+% & $(0.6,0.1,0.3)$ & $1409 \pm 7$ & $617 \pm 5$ & $259 \pm 5$\\
+\hline
+ \multirow{3}{*}{\textbf{Map 8}} & $(0.33,0.33,0.33)$ & $1761 \pm 9$ & $254 \pm 5$& $382 \pm 3$\\
+ & $(0.1,0.3,0.6)$ & $1804 \pm 10$ & $269 \pm 4$ & \bm{$357 \pm 4$}\\
+ & $(0.1,0.6,0.3)$ & $1826 \pm 10$ & \bm{$230 \pm 3$} & $392 \pm 7$\\
+% & $(0.6,0.1,0.3)$ & $1750 \pm 9$ & $304 \pm 7$ & $375 \pm 4$\\
+\hline
+ \multirow{3}{*}{\textbf{Map 9}} & $(0.33,0.33,0.33)$ & $2501 \pm 14$ & $926 \pm 6$& $574 \pm 9$\\
+ & $(0.1,0.3,0.6)$ & $2503 \pm 10$ & $921 \pm 10$ & \bm{$524 \pm 8$}\\
+ & $(0.1,0.6,0.3)$ & $2641 \pm 14$ & \bm{$833 \pm 5$} & $574 \pm 14$\\
+% & $(0.6,0.1,0.3)$ & $2442 \pm 12$ & $958 \pm 7$ & $559 \pm 8$\\
+\hline
+ \multirow{3}{*}{\textbf{Map 10}} & $(0.33,0.33,0.33)$ & $1430 \pm 8$ & $630 \pm 4$& $205 \pm 2$\\
+ & $(0.1,0.3,0.6)$ & $1493 \pm 13$ & $615 \pm 4$ & $209 \pm 2$\\
+ & $(0.1,0.6,0.3)$ & $1542 \pm 10$ & \bm{$554 \pm 4$} & $229 \pm 5$\\
+% & $(0.6,0.1,0.3)$ & $1394 \pm 10$ & $667 \pm 6$ & $202 \pm 2$\\
+\hline
+\end{tabular}
+\caption{MO-MCTS results in MO-PTSP with different weight vectors. Significantly better results per map are in bold.}
+\label{tab:weights}
+\end{center}
+\end{table}
+
+The few exceptions where this does not happen can actually been explained looking at the map: in map 1, a fuel canister is always collected near the end, restoring the fuel level to its maximum.
+This leaves too few cycles to make a difference in the controller (note also that this is the map with the fewest fuel consumption).
+The second exception happens in map 10 with the damage objective: but precisely map 10 is a map with no obstacles to damage the ship (thus only lava lakes deals damage).
+This also results in this map being the one with lowest damage overall.
+
+These results suggest that the weights effectively work on making the algorithm converge to different points in the Pareto front discovered by the algorithm.
+Now, it is time to compare these results with the other algorithms.
+In order to do this, the same number of experiments is run for NSGA-II and MCTS in the $10$ maps of the benchmark.
+
+In order to compare several algorithms between each other, the results are compared in pairs, in terms of dominance.
+The procedure is as follows: once all games on a single map have been run, the Mann–Whitney–Wilcoxon non-parametric test with $95\%$ confidence is calculated on the three objectives.
+If the measures on all objectives are assumed to be drawn from different distributions, their averages are compared for a dominance test.
+If one result dominates another, then the algorithm dominates the other in that particular map.
+
+Extending this comparison to all maps, each pair of algorithms ends with a triplet $(D, \O, d)$, where $D$ is the number of maps where the first algorithm dominates the second, $\O$ is the amount of maps where no dominance can be established, and $d$ states the number of maps where the first is dominated by the second.
+Table~\ref{tab:resMOPTSP} summarizes these results for all algorithms tested.
 
@@
 \hline
- \multirow{3}{*}{\textbf{MO-MCTS}} & $W_0: (0.33,0.33,0.33)$ & \multirow{3}{*}{$-$} & $(8,2,0)$ & $(8,2,0)$ & $(0,5,5)$\\
- & $W_1: (0.1,0.3,0.6)$ & & $(10,0,0)$ & $(4,6,0)$ & $(0,6,4)$\\
- & $W_2: (0.1,0.6,0.3)$ & & $(8,2,0)$ & $(7,3,0)$ & $(0,5,5)$\\
-\hline
- \multirow{3}{*}{\textbf{MCTS}} & $W_0: (0.33,0.33,0.33)$& $(0,8,2)$ & \multirow{3}{*}{$-$} & $(0,2,8)$ & $(0,2,8)$\\
- & $W_1: (0.1,0.3,0.6)$ & $(0,0,10)$ & & $(4,2,4)$ & $(0,3,7)$\\
- & $W_2: (0.1,0.6,0.3)$ & $(0,2,8)$ & & $(0,1,9)$ & $(0,6,4)$\\
-\hline
- \multirow{3}{*}{\textbf{NSGA-II}} & $W_0: (0.33,0.33,0.33)$& $(0,2,8)$ & $(8,2,0)$ & \multirow{3}{*}{$-$} & $(0,4,6)$\\
- & $W_1: (0.1,0.3,0.6)$ & $(0,6,4)$ & $(4,2,4)$ & & $(0,4,6)$\\
- & $W_2: (0.1,0.6,0.3)$ & $(0,3,7)$ & $(9,1,0)$ & & $(0,5,5)$\\
-\hline
- \multirow{3}{*}{\textbf{PurofMovio}} & $W_0: (0.33,0.33,0.33)$& $(5,5,0)$ & $(8,2,0)$ & $(6,4,0)$ & \multirow{3}{*}{$-$}\\
- & $W_1: (0.1,0.6,0.3)$ & $(4,6,0)$ & $(7,3,0)$ & $(6,4,0)$ & \\
- & $W_2: (0.1,0.3,0.6)$ & $(5,5,0)$ & $(6,4,0)$ & $(5,5,0)$ & \\
+ \multirow{3}{*}{\textbf{MO-MCTS}} & $W_1: (0.33,0.33,0.33)$ & \multirow{3}{*}{$-$} & $(8,2,0)$ & $(8,2,0)$ & $(0,5,5)$\\
+ & $W_2: (0.1,0.3,0.6)$ & & $(10,0,0)$ & $(4,6,0)$ & $(0,6,4)$\\
+ & $W_3: (0.1,0.6,0.3)$ & & $(8,2,0)$ & $(7,3,0)$ & $(0,5,5)$\\
+\hline
+ \multirow{3}{*}{\textbf{MCTS}} & $W_1: (0.33,0.33,0.33)$& $(0,8,2)$ & \multirow{3}{*}{$-$} & $(0,2,8)$ & $(0,2,8)$\\
+ & $W_2: (0.1,0.3,0.6)$ & $(0,0,10)$ & & $(4,2,4)$ & $(0,3,7)$\\
+ & $W_3: (0.1,0.6,0.3)$ & $(0,2,8)$ & & $(0,1,9)$ & $(0,6,4)$\\
+\hline
+ \multirow{3}{*}{\textbf{NSGA-II}} & $W_1: (0.33,0.33,0.33)$& $(0,2,8)$ & $(8,2,0)$ & \multirow{3}{*}{$-$} & $(0,4,6)$\\
+ & $W_2: (0.1,0.3,0.6)$ & $(0,6,4)$ & $(4,2,4)$ & & $(0,4,6)$\\
+ & $W_3: (0.1,0.6,0.3)$ & $(0,3,7)$ & $(9,1,0)$ & & $(0,5,5)$\\
+\hline
+ \multirow{3}{*}{\textbf{PurofMovio}} & $W_1: (0.33,0.33,0.33)$& $(5,5,0)$ & $(8,2,0)$ & $(6,4,0)$ & \multirow{3}{*}{$-$}\\
+ & $W_2: (0.1,0.6,0.3)$ & $(4,6,0)$ & $(7,3,0)$ & $(6,4,0)$ & \\
+ & $W_3: (0.1,0.3,0.6)$ & $(5,5,0)$ & $(6,4,0)$ & $(5,5,0)$ & \\
 \hline
 \end{tabular}
-\caption{Results MO-PTSP }
+\caption{Results in MO-PTSP: Each cell indicates the triplet $(D, \O, d)$, where $D$ is the number of maps where the row algorithm dominates the column one, $\O$ is the amount of maps where no dominance can be established, and $d$ states the number of maps where the row algorithm is dominated by the column one.
+All algorithms followed the same route (order of waypoints and fuel canisters) in every map tested.}
 \label{tab:resMOPTSP}
@@
 
-\subsection{A step further in MO-PTSP: dynamically varying weights} \label{ssec:resVarW}
+One of the first things to notice is that MO-MCTS dominates MCTS and NSGA-II in most of the maps, and it is never dominated by these.
+Particularly, the dominance of MO-MCTS over MCTS is outstanding, even dominating in all $10$ maps for one of the weight vectors.
+MO-MCTS also dominates NSGA-II in more maps than in those where there is no dominance, and it is as well never dominated by NSGA-II in any map.
+
+It is also interesting to see that NSGA-II dominates also the weighted sum version of MCTS, although for the vector $W_2$ there is a technical draw, as they dominate each other in $4$ different maps each and there is no dominance in the other $2$.
+
+Additionally, Table~\ref{tab:resMOPTSP} contains a fourth entry, \texttt{PurofMovio}, that the algorithms are compare against.
+\texttt{PurofMovio} is the winner entry of the 2013 MO-PTSP competition, a controller based in a weighted-sum MCTS approach (see~\cite{Powley2013} for details of its implementation).
+As can be seen, \texttt{PurofMovio} obtains better results than the algorithm proposed in this paper.
+
+However, it is very important to highlight that \texttt{PurofMovio} is not using the same heuristics as the ones presented in this research.
+Hence, nothing can be concluded from making pairwise comparisons directly with the winning entry of the competition.
+It is likely that \texttt{PurofMovio}'s heuristics are more efficient that the ones presented here, but the goal of this paper is not to develop the best heuristics for the MO-PTSP, but to provide an insight of how a multi-objective version of MCTS compares to other algorithms using the same heuristics.
+
+Anyway, the inclusion of \texttt{PurofMovio} in this comparison is not pointless.
+It is possible to asses the quality of the $3$ algorithms tested here by comparing their performance relatively, against a high quality entry.
+Attending to this criteria, it can be seen how MO-MCTS is the algorithm that is dominated less often by him, producing similar results in approximately $5$ out of the $10$ maps.
+MCTS and NSGA-II are dominated more often than MO-MCTS, being dominated in approximately $6$ and $7$ maps, respectively.
+This comparative result shows again that MO-MCTS is achieving the best results among the algorithms compared.
+
+\subsection{A step further in MO-PTSP: segments and weights} \label{ssec:resVarW}
+
+There is another aspect that can still be improved in the MO-PTSP benchmark, but that it is also applicable to other domains.
+It is naive to think that an unique weight vector will be the ideal one for the whole game.
+Specifically in the MO-PTSP, there are regions of the map where there are more obstacles on lava lakes, hence most likely to damage the ship.
+Also, the route followed during the game affects the relative ideal speed between waypoints, or sometimes a fuel canister is going to be picked up, which affects how the fuel objective can be managed.
+In general, many real-time games go through different phases, with different objectives and priorities.
+
+A way to provide different weights at different times in MO-PTSP is straightforward.
+Given the route of waypoints (or fuel canisters) that is being followed, one can divide it into \textit{segments}, where is segment starts and ends with a waypoint (or fuel canister).
+Then, each segment can be assigned a particular weight vector $W$.
+
+The question is then how to assign these weight vectors. Three different ways can be devised:
+
+\begin{itemize}
+\item Manually set the weight vectors. This was attempted and it proved to be a non trivial task.
+\item Setting the appropriate weight for each segment dynamically, based on the segment's characteristics.
+This involves the creation or discovery of features and some kind of function approximation to assign the values.
+\item Learn, for each specific map, the combination of weight vectors that produces better results.
+\end{itemize}
+
+This section details the efforts made in order to test the third variant, by using a stochastic hill climbing algorithm on each one of the maps.
+The goal is to check if by varying the weight vectors between segments, better solutions can be achieved.
+
+A solution is identified by a string of integers, where each integer refers to one of the weight vectors utilized in the previous sections.
+The solution is evaluated playing a particular map $10$ times, and its fitness is obtained by calculating the average of those runs.
+A population of $10$ individuals is kept, and the solutions of the initial population are created either randomly, or mutated from base individuals (where all segments have the same weight vector $W_1$, $W_2$, or $W_3$).
+
+The best solution (determined by dominance) is promoted to the next generation and it is mutated to generate other individuals.
+Also, a portion of individuals is generated randomly at every generation, until the end of the algorithm.
+Table~\ref{tab:resVarW} shows the results obtained on each run, one per map.
+
+\begin{table}[!t]
+\begin{center}
+%\begin{tabular}{|>{\centering\arraybackslash}m{0.65cm}|c|c|c|c|}
+\begin{tabular}{|p{0.8cm}|p{2.2cm}|m{1.13cm}|m{1.13cm}|m{1.13cm}|}
+\hline
+\textbf{Map}  & \textbf{Weight genome} & \textbf{Time} &  \textbf{Fuel} & \textbf{Damage} \\ 
+\hline
+ \multirow{3}{*}{\textbf{Map 1}} & $0000000000000$ & $1654 \pm 7$ & $131 \pm 2$& $846 \pm 13$\\
+ & $2010201001112$ & $1657 \pm 8$ & $130 \pm 2$ & $773 \pm 11$\\
+\hline
+ \multirow{3}{*}{\textbf{Map 1}} & $0000000000000$ & $1654 \pm 7$ & $131 \pm 2$& $846 \pm 13$\\
+ & $2010201001112$ & $1657 \pm 8$ & $130 \pm 2$ & $773 \pm 11$\\
+\hline
+ \multirow{3}{*}{\textbf{Map 1}} & $0000000000000$ & $1654 \pm 7$ & $131 \pm 2$& $846 \pm 13$\\
+ & $2010201001112$ & $1657 \pm 8$ & $130 \pm 2$ & $773 \pm 11$\\
+\hline
+ \multirow{3}{*}{\textbf{Map 1}} & $0000000000000$ & $1654 \pm 7$ & $131 \pm 2$& $846 \pm 13$\\
+ & $2010201001112$ & $1657 \pm 8$ & $130 \pm 2$ & $773 \pm 11$\\
+\hline
+ \multirow{3}{*}{\textbf{Map 1}} & $0000000000000$ & $1654 \pm 7$ & $131 \pm 2$& $846 \pm 13$\\
+ & $2010201001112$ & $1657 \pm 8$ & $130 \pm 2$ & $773 \pm 11$\\
+\hline
+ \multirow{3}{*}{\textbf{Map 1}} & $0000000000000$ & $1654 \pm 7$ & $131 \pm 2$& $846 \pm 13$\\
+ & $2010201001112$ & $1657 \pm 8$ & $130 \pm 2$ & $773 \pm 11$\\
+\hline
+ \multirow{3}{*}{\textbf{Map 1}} & $0000000000000$ & $1654 \pm 7$ & $131 \pm 2$& $846 \pm 13$\\
+ & $2010201001112$ & $1657 \pm 8$ & $130 \pm 2$ & $773 \pm 11$\\
+\hline
+ \multirow{3}{*}{\textbf{Map 1}} & $0000000000000$ & $1654 \pm 7$ & $131 \pm 2$& $846 \pm 13$\\
+ & $2010201001112$ & $1657 \pm 8$ & $130 \pm 2$ & $773 \pm 11$\\
+\hline
+ \multirow{3}{*}{\textbf{Map 1}} & $0000000000000$ & $1654 \pm 7$ & $131 \pm 2$& $846 \pm 13$\\
+ & $2010201001112$ & $1657 \pm 8$ & $130 \pm 2$ & $773 \pm 11$\\
+\hline
+ \multirow{3}{*}{\textbf{Map 1}} & $0000000000000$ & $1654 \pm 7$ & $131 \pm 2$& $846 \pm 13$\\
+ & $2010201001112$ & $1657 \pm 8$ & $130 \pm 2$ & $773 \pm 11$\\
+\hline
+\end{tabular}
+\caption{MO-PTSP Results with different weights: Each row corresponds to a run in the associated map.
+It gives results for two weights: the one below is the best after the run, while the one on the top is that from Table~\ref{tab:resMOPTSP} with the most similar genome.
+Each genome is a string of the form $xyz$, that represents $W_xW_yW_z$, where each value is a weight used in that particular segment.}
+\label{tab:resVarW}
+\end{center}
+\end{table}
 

### 2013-12-26 paper/paper.tex
introduction in

@@
 
-Here it goes, the introduction.
+Multi-objective optimization has been a field of study of manufacturing, engineering~\cite{Marler2004} and finance~\cite{Coello2006}, while having little impact in games research.
+This paper shows that multi-objective optimization has much to offer in developing game strategies that allow for a fine-grain control of alternative policies.
+The application of such approaches to this field can provide interesting results, especially in games that are long or complex enough where long-term planning is not trivial and achieving a good level of play requires balancing strategies.
+
+From a simple point of view, many competitive games can be seen as a scenario where two or more opponents have the single objective of winning.
+Achieving victory is usually complex in interesting games, and successful approaches normally provide a state value by combining features of the game state as a weighted sum.
+An example is a chess heuristic that assigns different weights to each piece according to their value.
+
+Multi-objective approaches can be applied in these scenarios, where several factors must be taken into account to achieve the victory.
+The algorithms can balance between different objectives, in order to provide a wide range of strategies well suited to the different stages of the game being played, or to face the existing opponents.
+Application of such approaches could be real-time strategy games, where long term planning must be carried out in order to balance aspects such as attack units, defensive structures and resource gathering.
+
+This paper proposes a multi-objective real-time algorithm version of Monte Carlo Tree Search (MCTS), a popular reinforcement learning approach within the last decade.
+The proposed algorithm is tested in two different games: a real-time version of the Deep Sea Treasure (DST; a classical multi-objective problem), and the Multi-Objective Physical Travelling Salesman Problem (MO-PTSP).
+The algorithm is also compared to two other approaches: a single-objective MCTS (that uses a weighted sum of features to value the state) and a rolling horizon Non-dominated Sorting Evolutionary Algorithm II (NSGA-II).
+This paper extends and formalizes our previous work described in~\cite{Perez2013a}.
+
+Two main goals can be identified in this paper: first, the proposed algorithm must be applicable to real-time domains (those where the next move to make must be decided within a small time budget) and it should obtain better or at least the same performance than the other state of the art algorithms.
+It is important to highlight that all three algorithms tested employ the same heuristic functions to evaluate the features of a given game state.
+Thus, the focus of this research is set on how the algorithm explores the search space, instead of providing the best possible solution to each given problem.
+
+Secondly, the algorithm must be able to provide solutions across the multi-objective spectrum: by parametrizing the algorithm, it must be possible to prioritize one objective over the others and therefore converge to solutions according to these preferences.
+
+The paper is structured as follows.
+First, Sections~\ref{sec:mcts} and~\ref{sec:moo} provide the necessary background for MCTS and multi-objective optimization, respectively.
+The algorithm proposed in this research is described in detail in Section~\ref{sec:momcts}.
+Then, Section~\ref{sec:bench} defines the games used to test the algorithms, with the results discussed in Section~\ref{sec:exp}.
+Finally, some conclusions and possible extensions of this work are drawn in Section~\ref{sec:conc}.
 
@@
 The idea is as follows.
-First of all, three different weight vectors are tested: $W_1 = (0.33,0.33,0.33)$, $W_2 = (0.1,0.3,0.6)$ and $W_3 = (0.1,0.6,0.3)$, where each $w_i$ corresponds to the weight for an objective ($w_t$ for time, $w_f$ for fuel and $w_d$ for damage).
-$W_1$ treats all objectives with the same weight, while $W_2$ and $W_3$ give more relevance to damage and fuel, respectively (note that Equations~\ref{eq:fuel} and~\ref{eq:dam} also give weight to the time objective, so having a vector with $w_t=0.6$ is not that relevant).
-All $W_1$, $W_2$ and $W_3$ provide then a wide spectrum for weights in this benchmark.
+First of all, $4$ different weight vectors are tested: $W_1 = (0.33,0.33,0.33)$, $W_2 = (0.1,0.3,0.6)$, $W_3 = (0.1,0.6,0.3)$ and $W_4 = (0.6,0.1,0.3)$, where each $w_i$ corresponds to the weight for an objective ($w_t$ for time, $w_f$ for fuel and $w_d$ for damage).
+$W_1$ treats all objectives with the same weight, while the other three give more relevance to different objectives.
+These vectors provide then a wide spectrum for weights in this benchmark.
 In this case, MO-MCTS uses this weights to select an action based on a weighted sum, which has shown in preliminary experiments better performance than the Euclidean distance mechanism.
@@
  & $(0.1,0.6,0.3)$ & $1681 \pm 11$ & $131 \pm 2$ & $837 \pm 15$\\
-% & $(0.6,0.1,0.3)$ & $1627 \pm 7$ & $138 \pm 3$ & $812 \pm 14$\\
+ & $(0.6,0.1,0.3)$ & \bm{$1649 \pm 8$} & $132 \pm 2$ & $833 \pm 13$\\
 \hline
@@
  & $(0.1,0.6,0.3)$ & $1416 \pm 8$ & \bm{$219 \pm 4$} & $360 \pm 3$\\
-% & $(0.6,0.1,0.3)$ & $1389 \pm 7$ & $241 \pm 5$ & $355 \pm 2$\\
+ & $(0.6,0.1,0.3)$ & $1396 \pm 8$ & $245 \pm 5$ & $361 \pm 2$\\
 \hline
@@
  & $(0.1,0.6,0.3)$ & $1385 \pm 6$ & \bm{$203 \pm 4$} & $291 \pm 7$\\
-% & $(0.6,0.1,0.3)$ & $1383 \pm 6$ & $245 \pm 5$ & $300 \pm 7$\\
+ & $(0.6,0.1,0.3)$ & \bm{$1363 \pm 4$} & $229 \pm 4$ & $285 \pm 7$\\
 \hline
@@
  & $(0.1,0.6,0.3)$ & $1423 \pm 6$ & \bm{$273 \pm 4$} & $583 \pm 5$\\
-% & $(0.6,0.1,0.3)$ & $1398 \pm 7$ & $325 \pm 5$ & $587 \pm 6$\\
+ & $(0.6,0.1,0.3)$ & $1388 \pm 6$ & $309 \pm 4$ & $559 \pm 5$\\
 \hline
@@
  & $(0.1,0.6,0.3)$ & $1467 \pm 9$ & \bm{$411 \pm 5$} & $567 \pm 5$\\
-% & $(0.6,0.1,0.3)$ & $1410 \pm 6$ & $480 \pm 5$ & $546 \pm 4$\\
+ & $(0.6,0.1,0.3)$ & \bm{$1399 \pm 9$} & $469 \pm 4$ & $547 \pm 3$\\
 \hline
@@
  & $(0.1,0.6,0.3)$ & $1703 \pm 11$ & \bm{$499 \pm 4$} & $316 \pm 7$\\
-% & $(0.6,0.1,0.3)$ & $1563 \pm 7$ & $560 \pm 5$ & $285 \pm 4$\\
+ & $(0.6,0.1,0.3)$ & \bm{$1571 \pm 7$} & $559 \pm 5$ & $294 \pm 4$\\
 \hline
@@
  & $(0.1,0.6,0.3)$ & $1489 \pm 12$ & \bm{$549 \pm 3$} & $264 \pm 6$\\
-% & $(0.6,0.1,0.3)$ & $1409 \pm 7$ & $617 \pm 5$ & $259 \pm 5$\\
+ & $(0.6,0.1,0.3)$ & \bm{$1407 \pm 8$} & $618 \pm 5$ & $270 \pm 6$\\
 \hline
@@
  & $(0.1,0.6,0.3)$ & $1826 \pm 10$ & \bm{$230 \pm 3$} & $392 \pm 7$\\
-% & $(0.6,0.1,0.3)$ & $1750 \pm 9$ & $304 \pm 7$ & $375 \pm 4$\\
+ & $(0.6,0.1,0.3)$ & \bm{$1732 \pm 9$} & $311 \pm 8$ & $379 \pm 6$\\
 \hline
@@
  & $(0.1,0.6,0.3)$ & $2641 \pm 14$ & \bm{$833 \pm 5$} & $574 \pm 14$\\
-% & $(0.6,0.1,0.3)$ & $2442 \pm 12$ & $958 \pm 7$ & $559 \pm 8$\\
+ & $(0.6,0.1,0.3)$ & \bm{$2470 \pm 9$} & $956 \pm 5$ & $573 \pm 8$\\
 \hline
@@
  & $(0.1,0.6,0.3)$ & $1542 \pm 10$ & \bm{$554 \pm 4$} & $229 \pm 5$\\
-% & $(0.6,0.1,0.3)$ & $1394 \pm 10$ & $667 \pm 6$ & $202 \pm 2$\\
+ & $(0.6,0.1,0.3)$ & \bm{$1378 \pm 5$} & $663 \pm 6$ & $202 \pm 4$\\
 \hline
@@
 
-The few exceptions where this does not happen can actually been explained looking at the map: in map 1, a fuel canister is always collected near the end, restoring the fuel level to its maximum.
+Some exceptions can actually be explained analyzing the map: in map 1, a fuel canister is always collected near the end, restoring the fuel level to its maximum.
 This leaves too few cycles to make a difference in the controller (note also that this is the map with the fewest fuel consumption).
-The second exception happens in map 10 with the damage objective: but precisely map 10 is a map with no obstacles to damage the ship (thus only lava lakes deals damage).
+Also, in map 10, there is no difference in the damage objective: but precisely map 10 is a map with no obstacles to damage the ship (thus only lava lakes deals damage).
 This also results in this map being the one with lowest damage overall.
+It can also be seen that, in those maps were time has priority, the results in this objective are not as dominant as the other two.
+This can be explained by the fact that the time heuristic is actually part of the fuel and damage heuristics (Equation~\ref{eq:dist} in Eqs.~\ref{eq:fuel} and~\ref{eq:dam}).
 
@@
  & $W_3: (0.1,0.6,0.3)$ & & $(8,2,0)$ & $(7,3,0)$ & $(0,5,5)$\\
+ & $W_4: (0.6,0.1,0.3)$ & & $(10,0,0)$ & $(3,7,0)$ & $(0,3,7)$\\
 \hline
@@
  & $W_3: (0.1,0.6,0.3)$ & $(0,2,8)$ & & $(0,1,9)$ & $(0,6,4)$\\
+ & $W_4: (0.6,0.1,0.3)$ & $(0,0,10)$ &  & $(3,3,4)$ & $(0,1,9)$\\
 \hline
@@
  & $W_3: (0.1,0.6,0.3)$ & $(0,3,7)$ & $(9,1,0)$ & & $(0,5,5)$\\
+ & $W_4: (0.6,0.1,0.3)$ & $(0,7,3)$ & $(4,3,3)$ &  & $(0,4,6)$\\
 \hline
@@
  & $W_3: (0.1,0.3,0.6)$ & $(5,5,0)$ & $(6,4,0)$ & $(5,5,0)$ & \\
+ & $W_4: (0.6,0.1,0.3)$ & $(7,3,0)$ & $(9,1,0)$ & $(6,4,0)$ & \\
 \hline
@@
 One of the first things to notice is that MO-MCTS dominates MCTS and NSGA-II in most of the maps, and it is never dominated by these.
-Particularly, the dominance of MO-MCTS over MCTS is outstanding, even dominating in all $10$ maps for one of the weight vectors.
+Particularly, the dominance of MO-MCTS over MCTS is outstanding, even dominating in all $10$ maps for two of the weight vectors.
 MO-MCTS also dominates NSGA-II in more maps than in those where there is no dominance, and it is as well never dominated by NSGA-II in any map.
@@
 
-Anyway, the inclusion of \texttt{PurofMovio} in this comparison is not pointless.
-It is possible to asses the quality of the $3$ algorithms tested here by comparing their performance relatively, against a high quality entry.
-Attending to this criteria, it can be seen how MO-MCTS is the algorithm that is dominated less often by him, producing similar results in approximately $5$ out of the $10$ maps.
-MCTS and NSGA-II are dominated more often than MO-MCTS, being dominated in approximately $6$ and $7$ maps, respectively.
+Anyway, the inclusion of \texttt{PurofMovio} in this comparison is not pointless: it is possible to asses the quality of the three algorithms tested here by comparing their performance relatively, against this high quality entry.
+Attending to this criteria, it can be seen how MO-MCTS is the algorithm that is dominated less often by him, producing similar results on an average of $4.75$ out of the $10$ maps, and being dominated in $5.25$ maps.
+MCTS and NSGA-II are dominated more often than MO-MCTS, being dominated on averages of $7.5$ and $5.75$ of the maps, respectively.
 This comparative result shows again that MO-MCTS is achieving the best results among the algorithms compared.
@@
 The solution is evaluated playing a particular map $10$ times, and its fitness is obtained by calculating the average of those runs.
-A population of $10$ individuals is kept, and the solutions of the initial population are created either randomly, or mutated from base individuals (where all segments have the same weight vector $W_1$, $W_2$, or $W_3$).
+A population of $10$ individuals is kept, and the solutions of the initial population are created either randomly, or mutated from base individuals.
+These base individuals have all segments with the same weight vector $W_1$, $W_2$, or $W_3$ ($W_4$ is out of this experiment, as it shown not to be that influential).
 
@@
 
+%\begin{table}[!t]
+%\begin{center}
+%%\begin{tabular}{|>{\centering\arraybackslash}m{0.65cm}|c|c|c|c|}
+%\begin{tabular}{|p{0.8cm}|p{2.2cm}|m{1.13cm}|m{1.13cm}|m{1.13cm}|}
+%\hline
+%\textbf{Map}  & \textbf{Weight genome} & \textbf{Time} &  \textbf{Fuel} & \textbf{Damage} \\ 
+%\hline
+% \multirow{3}{*}{\textbf{Map 1}} & $0000000000000$ & $1654 \pm 7$ & $131 \pm 2$& $846 \pm 13$\\
+% & $2010201001112$ & $1657 \pm 8$ & $130 \pm 2$ & $773 \pm 11$\\
+%\hline
+% \multirow{3}{*}{\textbf{Map 1}} & $0000000000000$ & $1654 \pm 7$ & $131 \pm 2$& $846 \pm 13$\\
+% & $2010201001112$ & $1657 \pm 8$ & $130 \pm 2$ & $773 \pm 11$\\
+%\hline
+% \multirow{3}{*}{\textbf{Map 1}} & $0000000000000$ & $1654 \pm 7$ & $131 \pm 2$& $846 \pm 13$\\
+% & $2010201001112$ & $1657 \pm 8$ & $130 \pm 2$ & $773 \pm 11$\\
+%\hline
+% \multirow{3}{*}{\textbf{Map 1}} & $0000000000000$ & $1654 \pm 7$ & $131 \pm 2$& $846 \pm 13$\\
+% & $2010201001112$ & $1657 \pm 8$ & $130 \pm 2$ & $773 \pm 11$\\
+%\hline
+% \multirow{3}{*}{\textbf{Map 1}} & $0000000000000$ & $1654 \pm 7$ & $131 \pm 2$& $846 \pm 13$\\
+% & $2010201001112$ & $1657 \pm 8$ & $130 \pm 2$ & $773 \pm 11$\\
+%\hline
+% \multirow{3}{*}{\textbf{Map 1}} & $0000000000000$ & $1654 \pm 7$ & $131 \pm 2$& $846 \pm 13$\\
+% & $2010201001112$ & $1657 \pm 8$ & $130 \pm 2$ & $773 \pm 11$\\
+%\hline
+% \multirow{3}{*}{\textbf{Map 1}} & $0000000000000$ & $1654 \pm 7$ & $131 \pm 2$& $846 \pm 13$\\
+% & $2010201001112$ & $1657 \pm 8$ & $130 \pm 2$ & $773 \pm 11$\\
+%\hline
+% \multirow{3}{*}{\textbf{Map 1}} & $0000000000000$ & $1654 \pm 7$ & $131 \pm 2$& $846 \pm 13$\\
+% & $2010201001112$ & $1657 \pm 8$ & $130 \pm 2$ & $773 \pm 11$\\
+%\hline
+% \multirow{3}{*}{\textbf{Map 1}} & $0000000000000$ & $1654 \pm 7$ & $131 \pm 2$& $846 \pm 13$\\
+% & $2010201001112$ & $1657 \pm 8$ & $130 \pm 2$ & $773 \pm 11$\\
+%\hline
+% \multirow{3}{*}{\textbf{Map 1}} & $0000000000000$ & $1654 \pm 7$ & $131 \pm 2$& $846 \pm 13$\\
+% & $2010201001112$ & $1657 \pm 8$ & $130 \pm 2$ & $773 \pm 11$\\
+%\hline
+%\end{tabular}
+%\caption{MO-PTSP Results with different weights: Each row corresponds to a run in the associated map.
+It gives results for two weights: the one below is the best after the run, while the one on the top is that from Table~\ref{tab:resMOPTSP} with the most similar genome.
+Each genome is a string of the form $xyz$, that represents $W_xW_yW_z$, where each value is a weight used in that particular segment.}
+%\label{tab:resVarW}
+%\end{center}
+%\end{table}
+
 \begin{table}[!t]
@@
 %\begin{tabular}{|>{\centering\arraybackslash}m{0.65cm}|c|c|c|c|}
-\begin{tabular}{|p{0.8cm}|p{2.2cm}|m{1.13cm}|m{1.13cm}|m{1.13cm}|}
-\hline
-\textbf{Map}  & \textbf{Weight genome} & \textbf{Time} &  \textbf{Fuel} & \textbf{Damage} \\ 
-\hline
- \multirow{3}{*}{\textbf{Map 1}} & $0000000000000$ & $1654 \pm 7$ & $131 \pm 2$& $846 \pm 13$\\
- & $2010201001112$ & $1657 \pm 8$ & $130 \pm 2$ & $773 \pm 11$\\
-\hline
- \multirow{3}{*}{\textbf{Map 1}} & $0000000000000$ & $1654 \pm 7$ & $131 \pm 2$& $846 \pm 13$\\
- & $2010201001112$ & $1657 \pm 8$ & $130 \pm 2$ & $773 \pm 11$\\
-\hline
- \multirow{3}{*}{\textbf{Map 1}} & $0000000000000$ & $1654 \pm 7$ & $131 \pm 2$& $846 \pm 13$\\
- & $2010201001112$ & $1657 \pm 8$ & $130 \pm 2$ & $773 \pm 11$\\
-\hline
- \multirow{3}{*}{\textbf{Map 1}} & $0000000000000$ & $1654 \pm 7$ & $131 \pm 2$& $846 \pm 13$\\
- & $2010201001112$ & $1657 \pm 8$ & $130 \pm 2$ & $773 \pm 11$\\
-\hline
- \multirow{3}{*}{\textbf{Map 1}} & $0000000000000$ & $1654 \pm 7$ & $131 \pm 2$& $846 \pm 13$\\
- & $2010201001112$ & $1657 \pm 8$ & $130 \pm 2$ & $773 \pm 11$\\
-\hline
- \multirow{3}{*}{\textbf{Map 1}} & $0000000000000$ & $1654 \pm 7$ & $131 \pm 2$& $846 \pm 13$\\
- & $2010201001112$ & $1657 \pm 8$ & $130 \pm 2$ & $773 \pm 11$\\
-\hline
- \multirow{3}{*}{\textbf{Map 1}} & $0000000000000$ & $1654 \pm 7$ & $131 \pm 2$& $846 \pm 13$\\
- & $2010201001112$ & $1657 \pm 8$ & $130 \pm 2$ & $773 \pm 11$\\
-\hline
- \multirow{3}{*}{\textbf{Map 1}} & $0000000000000$ & $1654 \pm 7$ & $131 \pm 2$& $846 \pm 13$\\
- & $2010201001112$ & $1657 \pm 8$ & $130 \pm 2$ & $773 \pm 11$\\
-\hline
- \multirow{3}{*}{\textbf{Map 1}} & $0000000000000$ & $1654 \pm 7$ & $131 \pm 2$& $846 \pm 13$\\
- & $2010201001112$ & $1657 \pm 8$ & $130 \pm 2$ & $773 \pm 11$\\
-\hline
- \multirow{3}{*}{\textbf{Map 1}} & $0000000000000$ & $1654 \pm 7$ & $131 \pm 2$& $846 \pm 13$\\
- & $2010201001112$ & $1657 \pm 8$ & $130 \pm 2$ & $773 \pm 11$\\
+\begin{tabular}{|p{0.8cm}|p{1.95cm}|m{1.13cm}|m{1.13cm}|m{1.13cm}|m{0.1cm}|}
+\hline
+\textbf{Map}  & \textbf{Weight genome} & \textbf{Time} &  \textbf{Fuel} & \textbf{Damage} & \textbf{D}\\ 
+\hline
+ \multirow{3}{*}{\textbf{Map 1}} & $11111111111111$ & $1654 \pm 7$ & $131 \pm 2$& $846 \pm 13$ & $\preceq$ \\
+ & $22222222222222$ & $1657 \pm 8$ & $130 \pm 2$ & $773 \pm 11$ & $\preceq$ \\
+ & $33333333333333$ & $1681 \pm 11$ & $131 \pm 2$ & $837 \pm 15$ & $\preceq$ \\
+ & $21201101220001$ & $1619 \pm $ & $130 \pm $ & $744 \pm $ & \\
+\hline
+ \multirow{3}{*}{\textbf{Map 2}} & $11111111111111$ & $1409 \pm 7$ & $235 \pm 4$& $364 \pm 3$ & $\preceq$ \\
+ & $22222222222222$ & $1402 \pm 6$ & $236 \pm 5$ & \bm{$354 \pm 2$} & $\preceq$ \\
+ & $33333333333333$ & $1416 \pm 8$ & \bm{$219 \pm 4$} & $360 \pm 3$ & $\preceq$ \\
+ & $12020201212102$ & $1390 \pm $ & $210 \pm $ & $353 \pm $ & \\
+\hline
+ \multirow{3}{*}{\textbf{Map 3}} & $11111111111111$ & $1373 \pm 6$ & $221 \pm 3$& $301 \pm 7$ & $\preceq$ \\
+ & $22222222222222$ & $1378 \pm 5$ & $211 \pm 3$ & \bm{$268 \pm 5$} & $\O$ \\
+ & $33333333333333$ & $1385 \pm 6$ & \bm{$203 \pm 4$} & $291 \pm 7$ & $\O$ \\
+ & $00001112001120$ & $1364 \pm $ & $223 \pm $ & $269 \pm $ & \\
+\hline
+ \multirow{3}{*}{\textbf{Map 4}} & $11111111111111$ & $1383 \pm 6$ & $291 \pm 5$& $565 \pm 5$ & $\O$ \\
+ & $22222222222222$ & $1385 \pm 7$ & $304 \pm 4$ & \bm{$542 \pm 4$} & $\O$ \\
+ & $33333333333333$ & $1423 \pm 6$ & \bm{$273 \pm 4$} & $583 \pm 5$ & $\preceq$ \\
+ & $22212221201211$ & $1408 \pm $ & $271 \pm $ & $584 \pm $ & \\
+\hline
+ \multirow{3}{*}{\textbf{Map 5}} & $11111111111111$ & $1405 \pm 7$ & $467 \pm 4$& $559 \pm 4$ & $\preceq$ \\
+ & $22222222222222$ & $1431 \pm 9$ & $447 \pm 4$ & \bm{$541 \pm 4$} & $\preceq$ \\
+ & $33333333333333$ & $1467 \pm 9$ & \bm{$411 \pm 5$} & $567 \pm 5$ & $\O$ \\
+ & $10200102000100$ & $1397 \pm $ & $448 \pm $ & $535 \pm $ & \\
+\hline
+ \multirow{3}{*}{\textbf{Map 6}} & $11111111111111$ & $1575 \pm 7$ & $549 \pm 5$& $303 \pm 4$ & $\preceq$ \\
+ & $22222222222222$ & $1626 \pm 9$ & $540 \pm 6$ & \bm{$286 \pm 5$} & $\preceq$ \\
+ & $33333333333333$ & $1703 \pm 11$ & \bm{$499 \pm 4$} & $316 \pm 7$ & $\O$ \\
+ & $20010201000000$ & $1570 \pm $ & $535 \pm $ & $266 \pm $ & \\
+\hline
+ \multirow{3}{*}{\textbf{Map 7}} & $11111111111111$ & $1434 \pm 5$ & $599 \pm 6$& $284 \pm 6$ & $\preceq$ \\
+ & $22222222222222$ & $1475 \pm 10$ & $602 \pm 5$ & \bm{$243 \pm 6$} & $\preceq$ \\
+ & $33333333333333$ & $1489 \pm 12$ & \bm{$549 \pm 3$} & $264 \pm 6$ & $\O$ \\
+ & $00221100210221$ & $1401 \pm $ & $563 \pm $ & $230 \pm $ & \\
+\hline
+ \multirow{3}{*}{\textbf{Map 8}} & $11111111111111$ & $1761 \pm 9$ & $254 \pm 5$& $382 \pm 3$ & $\preceq$ \\
+ & $22222222222222$ & $1804 \pm 10$ & $269 \pm 4$ & \bm{$357 \pm 4$} & $\O$ \\
+ & $33333333333333$ & $1826 \pm 10$ & \bm{$230 \pm 3$} & $392 \pm 7$ & $\O$ \\
+ & $12110020202212$ & $1747 \pm $ & $247 \pm $ & $363 \pm $ & \\
+\hline
+ \multirow{3}{*}{\textbf{Map 9}} & $11111111111111$ & $2501 \pm 14$ & $926 \pm 6$& $574 \pm 9$ & $\preceq$ \\
+ & $22222222222222$ & $2503 \pm 10$ & $921 \pm 10$ & \bm{$524 \pm 8$} & $\preceq$ \\
+ & $33333333333333$ & $2641 \pm 14$ & \bm{$833 \pm 5$} & $574 \pm 14$ & $\O$ \\
+ & $10021220222112$ & $2463 \pm $ & $891 \pm $ & $523 \pm $ & \\
+\hline
+ \multirow{3}{*}{\textbf{Map 10}} & $11111111111111$ & $1430 \pm 8$ & $630 \pm 4$& $205 \pm 2$ & $\O$ \\
+ & $22222222222222$ & $1493 \pm 13$ & $615 \pm 4$ & $209 \pm 2$ & $\O$ \\
+ & $33333333333333$ & $1542 \pm 10$ & \bm{$554 \pm 4$} & $229 \pm 5$ & $\O$ \\
+ & $00200000011100$ & $1418 \pm $ & $623 \pm $ & $197 \pm $ & \\
 \hline

### 2013-12-27 paper/paper.tex
all sections written. whole review pending

@@
 \begin{abstract}
-Abstract... 
+Multi-objective optimization has been traditionally a matter of study in domains like engineering or finance, with little impact on games research.
+However, action-decision based on a multi-objective evaluation may be beneficial to obtain a high quality level of play.
+This paper presents a Multi-objective Monte Carlo Tree Search algorithm for planning and control in real-time game domains, those where the time budget to decide the next move to make is significantly reduced.
+A comparison is made between the proposed algorithm, a single-objective version of Monte Carlo Tree Search and a rolling horizon implementation of Non-dominated Sorting Evolutionary Algorithm II.
+Two different benchmarks are employed, the Deep Sea Treasure and the Multi-Objective Physical Travelling Salesman Problem.
+Using the same heuristics on each game, the analysis is focused on how well the algorithms explore the search space.
+The results described in this paper show that the proposed approach provides better results than the other algorithms it is compared with.
+Additionally, it is also shown that the algorithm is able to converge to different solutions of the optimal, or the best found during the search, Pareto front.
 \end{abstract}
@@
 
-This section details the efforts made in order to test the third variant, by using a stochastic hill climbing algorithm on each one of the maps.
-The goal is to check if by varying the weight vectors between segments, better solutions can be achieved.
-
-A solution is identified by a string of integers, where each integer refers to one of the weight vectors utilized in the previous sections.
+This section shows some initial results in order to test the third variant, by using a stochastic hill climbing algorithm on each one of the maps.
+The goal is to check if, by varying the weight vectors between segments, better solutions can be achieved.
+
+A solution is identified by a string of integers, where each integer refers to one of the weight vectors utilized in the previous sections ( $1 = W_1 = (0.33,0.33,0.33)$, $2 = W_2 = (0.1,0.3,0.6)$ and $3 = W_3 = (0.1,0.6,0.3)$.
+$W_4$ has been left out of this experiment, as it shown not to be that influential).
 The solution is evaluated playing a particular map $10$ times, and its fitness is obtained by calculating the average of those runs.
 A population of $10$ individuals is kept, and the solutions of the initial population are created either randomly, or mutated from base individuals.
-These base individuals have all segments with the same weight vector $W_1$, $W_2$, or $W_3$ ($W_4$ is out of this experiment, as it shown not to be that influential).
-
-The best solution (determined by dominance) is promoted to the next generation and it is mutated to generate other individuals.
-Also, a portion of individuals is generated randomly at every generation, until the end of the algorithm.
+These base individuals have all segments with the same weight vector $W_1$, $W_2$, or $W_3$.
+
+The best solution, determined by dominance, is kept and promoted to the next generation, where it is mutated to generate other individuals of the population.
+Also, a portion of the individuals of the next population is created randomly at every generation, until the end of the algorithm, which is established at $50$ generations.
+
 Table~\ref{tab:resVarW} shows the results obtained on each run, one per map.
-
-%\begin{table}[!t]
-%\begin{center}
-%%\begin{tabular}{|>{\centering\arraybackslash}m{0.65cm}|c|c|c|c|}
-%\begin{tabular}{|p{0.8cm}|p{2.2cm}|m{1.13cm}|m{1.13cm}|m{1.13cm}|}
-%\hline
-%\textbf{Map}  & \textbf{Weight genome} & \textbf{Time} &  \textbf{Fuel} & \textbf{Damage} \\ 
-%\hline
-% \multirow{3}{*}{\textbf{Map 1}} & $0000000000000$ & $1654 \pm 7$ & $131 \pm 2$& $846 \pm 13$\\
-% & $2010201001112$ & $1657 \pm 8$ & $130 \pm 2$ & $773 \pm 11$\\
-%\hline
-% \multirow{3}{*}{\textbf{Map 1}} & $0000000000000$ & $1654 \pm 7$ & $131 \pm 2$& $846 \pm 13$\\
-% & $2010201001112$ & $1657 \pm 8$ & $130 \pm 2$ & $773 \pm 11$\\
-%\hline
-% \multirow{3}{*}{\textbf{Map 1}} & $0000000000000$ & $1654 \pm 7$ & $131 \pm 2$& $846 \pm 13$\\
-% & $2010201001112$ & $1657 \pm 8$ & $130 \pm 2$ & $773 \pm 11$\\
-%\hline
-% \multirow{3}{*}{\textbf{Map 1}} & $0000000000000$ & $1654 \pm 7$ & $131 \pm 2$& $846 \pm 13$\\
-% & $2010201001112$ & $1657 \pm 8$ & $130 \pm 2$ & $773 \pm 11$\\
-%\hline
-% \multirow{3}{*}{\textbf{Map 1}} & $0000000000000$ & $1654 \pm 7$ & $131 \pm 2$& $846 \pm 13$\\
-% & $2010201001112$ & $1657 \pm 8$ & $130 \pm 2$ & $773 \pm 11$\\
-%\hline
-% \multirow{3}{*}{\textbf{Map 1}} & $0000000000000$ & $1654 \pm 7$ & $131 \pm 2$& $846 \pm 13$\\
-% & $2010201001112$ & $1657 \pm 8$ & $130 \pm 2$ & $773 \pm 11$\\
-%\hline
-% \multirow{3}{*}{\textbf{Map 1}} & $0000000000000$ & $1654 \pm 7$ & $131 \pm 2$& $846 \pm 13$\\
-% & $2010201001112$ & $1657 \pm 8$ & $130 \pm 2$ & $773 \pm 11$\\
-%\hline
-% \multirow{3}{*}{\textbf{Map 1}} & $0000000000000$ & $1654 \pm 7$ & $131 \pm 2$& $846 \pm 13$\\
-% & $2010201001112$ & $1657 \pm 8$ & $130 \pm 2$ & $773 \pm 11$\\
-%\hline
-% \multirow{3}{*}{\textbf{Map 1}} & $0000000000000$ & $1654 \pm 7$ & $131 \pm 2$& $846 \pm 13$\\
-% & $2010201001112$ & $1657 \pm 8$ & $130 \pm 2$ & $773 \pm 11$\\
-%\hline
-% \multirow{3}{*}{\textbf{Map 1}} & $0000000000000$ & $1654 \pm 7$ & $131 \pm 2$& $846 \pm 13$\\
-% & $2010201001112$ & $1657 \pm 8$ & $130 \pm 2$ & $773 \pm 11$\\
-%\hline
-%\end{tabular}
-%\caption{MO-PTSP Results with different weights: Each row corresponds to a run in the associated map.
-It gives results for two weights: the one below is the best after the run, while the one on the top is that from Table~\ref{tab:resMOPTSP} with the most similar genome.
-Each genome is a string of the form $xyz$, that represents $W_xW_yW_z$, where each value is a weight used in that particular segment.}
-%\label{tab:resVarW}
-%\end{center}
-%\end{table}
+Each row corresponds to a run in the associated map, and it provides different results depending on the weights vector.
+The top three are the base individuals, taken from Table~\ref{tab:resMOPTSP} for comparison.
+The forth result on each row is the best one after the run.
+Each genome is a string of the form $xyz$, that represents $W_xW_yW_z$, where each value is a weight used in that particular segment.
+The last column indicates if the evolved genome dominates ($\preceq$) or not ($\O$) each one of the base genomes for that particular map.
 
@@
 \begin{center}
-%\begin{tabular}{|>{\centering\arraybackslash}m{0.65cm}|c|c|c|c|}
 \begin{tabular}{|p{0.8cm}|p{1.95cm}|m{1.13cm}|m{1.13cm}|m{1.13cm}|m{0.1cm}|}
@@
  & $33333333333333$ & $1681 \pm 11$ & $131 \pm 2$ & $837 \pm 15$ & $\preceq$ \\
- & $21201101220001$ & $1619 \pm $ & $130 \pm $ & $744 \pm $ & \\
+ & $32312212331112$ & $1619 \pm 12$ & $130 \pm 2$ & $744 \pm 15$ & \\
 \hline
@@
  & $33333333333333$ & $1416 \pm 8$ & \bm{$219 \pm 4$} & $360 \pm 3$ & $\preceq$ \\
- & $12020201212102$ & $1390 \pm $ & $210 \pm $ & $353 \pm $ & \\
+ & $23131312323213$ & $1390 \pm 10$ & $210 \pm 3$ & $353 \pm 3$ & \\
 \hline
  \multirow{3}{*}{\textbf{Map 3}} & $11111111111111$ & $1373 \pm 6$ & $221 \pm 3$& $301 \pm 7$ & $\preceq$ \\
- & $22222222222222$ & $1378 \pm 5$ & $211 \pm 3$ & \bm{$268 \pm 5$} & $\O$ \\
+ & $22222222222222$ & $1378 \pm 5$ & $211 \pm 3$ & \bm{$268 \pm 5$} & $\preceq$ \\
  & $33333333333333$ & $1385 \pm 6$ & \bm{$203 \pm 4$} & $291 \pm 7$ & $\O$ \\
- & $00001112001120$ & $1364 \pm $ & $223 \pm $ & $269 \pm $ & \\
-\hline
- \multirow{3}{*}{\textbf{Map 4}} & $11111111111111$ & $1383 \pm 6$ & $291 \pm 5$& $565 \pm 5$ & $\O$ \\
- & $22222222222222$ & $1385 \pm 7$ & $304 \pm 4$ & \bm{$542 \pm 4$} & $\O$ \\
- & $33333333333333$ & $1423 \pm 6$ & \bm{$273 \pm 4$} & $583 \pm 5$ & $\preceq$ \\
- & $22212221201211$ & $1408 \pm $ & $271 \pm $ & $584 \pm $ & \\
+ & $11122222112231$ & $1358 \pm 9$ & $219 \pm 7$ & $263 \pm 12$ & \\
+\hline
+ \multirow{3}{*}{\textbf{Map 4}} & $11111111111111$ & $1383 \pm 6$ & $291 \pm 5$& $565 \pm 5$ & $\preceq$ \\
+ & $22222222222222$ & $1385 \pm 7$ & $304 \pm 4$ & \bm{$542 \pm 4$} & $\preceq$ \\
+ & $33333333333333$ & $1423 \pm 6$ & \bm{$273 \pm 4$} & $583 \pm 5$ &  $\O$ \\
+ & $11121131212112$ & $1360 \pm 4$ & $282 \pm 5$ & $540 \pm 4$ & \\
 \hline
@@
  & $33333333333333$ & $1467 \pm 9$ & \bm{$411 \pm 5$} & $567 \pm 5$ & $\O$ \\
- & $10200102000100$ & $1397 \pm $ & $448 \pm $ & $535 \pm $ & \\
+ & $21311213111211$ & $1397 \pm 11$ & $448 \pm 10$ & $535 \pm 5$ & \\
 \hline
@@
  & $33333333333333$ & $1703 \pm 11$ & \bm{$499 \pm 4$} & $316 \pm 7$ & $\O$ \\
- & $20010201000000$ & $1570 \pm $ & $535 \pm $ & $266 \pm $ & \\
+ & $31121312111111$ & $1570 \pm 16$ & $535 \pm 10$ & $266 \pm 6$ & \\
 \hline
@@
  & $33333333333333$ & $1489 \pm 12$ & \bm{$549 \pm 3$} & $264 \pm 6$ & $\O$ \\
- & $00221100210221$ & $1401 \pm $ & $563 \pm $ & $230 \pm $ & \\
+ & $11332211321332$ & $1401 \pm 12$ & $563 \pm 4$ & $230 \pm 12$ & \\
 \hline
  \multirow{3}{*}{\textbf{Map 8}} & $11111111111111$ & $1761 \pm 9$ & $254 \pm 5$& $382 \pm 3$ & $\preceq$ \\
- & $22222222222222$ & $1804 \pm 10$ & $269 \pm 4$ & \bm{$357 \pm 4$} & $\O$ \\
+ & $22222222222222$ & $1804 \pm 10$ & $269 \pm 4$ & \bm{$357 \pm 4$} & $\preceq$ \\
  & $33333333333333$ & $1826 \pm 10$ & \bm{$230 \pm 3$} & $392 \pm 7$ & $\O$ \\
- & $12110020202212$ & $1747 \pm $ & $247 \pm $ & $363 \pm $ & \\
+ & $23221131313323$ & $1747 \pm 11$ & $247 \pm 9$ & $363 \pm 9$ & \\
 \hline
@@
  & $33333333333333$ & $2641 \pm 14$ & \bm{$833 \pm 5$} & $574 \pm 14$ & $\O$ \\
- & $10021220222112$ & $2463 \pm $ & $891 \pm $ & $523 \pm $ & \\
-\hline
- \multirow{3}{*}{\textbf{Map 10}} & $11111111111111$ & $1430 \pm 8$ & $630 \pm 4$& $205 \pm 2$ & $\O$ \\
- & $22222222222222$ & $1493 \pm 13$ & $615 \pm 4$ & $209 \pm 2$ & $\O$ \\
+ & $21132331333223$ & $2463 \pm 19$ & $891 \pm 8$ & $523 \pm 9$ & \\
+\hline
+ \multirow{3}{*}{\textbf{Map 10}} & $11111111111111$ & $1430 \pm 8$ & $630 \pm 4$& $205 \pm 2$ & $\preceq$ \\
+ & $22222222222222$ & $1493 \pm 13$ & $615 \pm 4$ & $209 \pm 2$ & $\preceq$ \\
  & $33333333333333$ & $1542 \pm 10$ & \bm{$554 \pm 4$} & $229 \pm 5$ & $\O$ \\
- & $00200000011100$ & $1418 \pm $ & $623 \pm $ & $197 \pm $ & \\
+ & $11311111322231$ & $1418 \pm 9$ & $623 \pm 9$ & $197 \pm 2$ & \\
 \hline
 \end{tabular}
-\caption{MO-PTSP Results with different weights: Each row corresponds to a run in the associated map.
-It gives results for two weights: the one below is the best after the run, while the one on the top is that from Table~\ref{tab:resMOPTSP} with the most similar genome.
-Each genome is a string of the form $xyz$, that represents $W_xW_yW_z$, where each value is a weight used in that particular segment.}
+\caption{MO-PTSP Results with different weights.}
 \label{tab:resVarW}
@@
 
+These results suggest some interesting concepts.
+First of all, it is indeed possible to obtain better results by varying the weights of the objectives along the route: in $22$ out of the $30$ comparisons made, the evolved solution dominates the base individuals and, in the other $8$, both solutions in the comparison would be in the same Pareto front of solutions.
+
+If the focus is set on what particular weights are more successful, it is interesting to see that the best configurations found are better, in all maps, than the base individuals with all weights equal to $W_1$ ($w_t = 0.33, w_f = 0.33, w_d = 0.33$) and $W_2$ ($w_t = 0.1, w_f = 0.3, w_d = 0.6$).
+In other words, it was always possible to find a combination of weight vectors that performed better than approaches that gave the same weight to all objectives, and also the ones prioritizing low damage.
+
+It is also worthwhile to mention that those cases where dominance over the base individual was not achieved, this base was the one that prioritized fuel.
+Actually, it can be seen that in these cases, the result in the fuel objective is the one that avoids the dominance to happen, obtaining a better value than the evolved solution in this particular objective.
+
 \section{Conclusions and Future Work} \label{sec:conc}
+
+This paper presents a multi-objective version of Monte Carlo Tree Search (MO-MCTS) for real-time games, scenarios where the time budget allowed for finding out the best possible action is significantly reduced.
+MO-MCTS is tested, in comparison with a single MCTS algorithm and a rolling horizon NSGA-II, in two different real-time games, the Deep Sea Treasure (DST) and the Multi-Objective Physical Travelling Salesman Problem (MO-PTSP).
+This comparison is made by using the same heuristics that determine the value of a given state, so the focus is set on how the algorithms explore the search space on each problem.
+
+The results obtained in this study show the proficiency of the algorithm proposed, as it provides better solutions than the ones obtained by the other algorithms in comparison.
+Also, the MO-MCTS approach samples successfully across the different solutions in the Pareto front (optimal front in the DST, the best found in MO-PTSP), depending on the weights provided to each objective.
+Finally, some initial results are obtained on the idea of using distinct weight vectors for the objective in different situations on the game, showing that it is possible to improve the results by performing these variations.
+
+The stochastic hill climbing algorithm used for creating strings of weight vectors for the MO-PTSP is relatively simple, although it was able to obtain very good results.
+However, it could be desirable to model behaviours that dynamically change the weight vectors according to measurements of the game state, instead of evolving a different solution for each particular map.
+A possible extension for this work is to analyze and discover features in the game state that allows to establish relationships between game situations and the weight vectors to use for each objective.
+This mechanism would be general enough to be applicable to many different maps without requiring specific learning in any particular level.
+
+Finally, the results this paper allow to presume that multi-objective approaches can provide high quality level of play in real-time games, and not only in the ones used as benchmarks in this research.
+Games (and other domains) where the goal can be decomposed in several simultaneous tasks, especially when these tasks are in conflict with each other, can benefit from approaches like the one presented here.
 
@@
 This work was supported by EPSRC grants EP/H048588/1, under the project entitled ``UCT for Games and Beyond''.
+The authors would also like to thank Edward Powley and Daniel Whitehouse for their work in the controller \texttt{PurofMovio} for the MO-PTSP.
 

### 2013-12-27 paper/paper.tex
first full paper review

@@
 Using the same heuristics on each game, the analysis is focused on how well the algorithms explore the search space.
-The results described in this paper show that the proposed approach provides better results than the other algorithms it is compared with.
+The results show that the proposed approach provides better results than the other algorithms it is compared with.
 Additionally, it is also shown that the algorithm is able to converge to different solutions of the optimal, or the best found during the search, Pareto front.
@@
 This paper shows that multi-objective optimization has much to offer in developing game strategies that allow for a fine-grain control of alternative policies.
-The application of such approaches to this field can provide interesting results, especially in games that are long or complex enough where long-term planning is not trivial and achieving a good level of play requires balancing strategies.
+The application of such approaches to this field can provide interesting results, especially in games that are long or complex enough where long-term planning is not trivial, and achieving a good level of play requires balancing strategies.
 
@@
 The proposed algorithm is tested in two different games: a real-time version of the Deep Sea Treasure (DST; a classical multi-objective problem), and the Multi-Objective Physical Travelling Salesman Problem (MO-PTSP).
-The algorithm is also compared to two other approaches: a single-objective MCTS (that uses a weighted sum of features to value the state) and a rolling horizon Non-dominated Sorting Evolutionary Algorithm II (NSGA-II).
+The algorithm is also compared with another two approaches: a single-objective MCTS (that uses a weighted sum of features to value the state) and a rolling horizon Non-dominated Sorting Evolutionary Algorithm II (NSGA-II).
 This paper extends and formalizes our previous work described in~\cite{Perez2013a}.
 
-Two main goals can be identified in this paper: first, the proposed algorithm must be applicable to real-time domains (those where the next move to make must be decided within a small time budget) and it should obtain better or at least the same performance than the other state of the art algorithms.
+Two main goals can be identified in this paper: first, the proposed algorithm must be applicable to real-time domains (those where the next move to make must be decided within a small time budget) and it should obtain better or at least the same performance than the other state-of-the-art algorithms.
 It is important to highlight that all three algorithms tested employ the same heuristic functions to evaluate the features of a given game state.
@@
 A commonly used value is $\sqrt{2}$, as it balances both facets of the search when the rewards are normalized between $0$ and $1$.
-It is worth noting that MCTS, when combined with UCB1 reaches asymptotically logarithmic regret~\cite{coquelin2007bandit}.
+It is worth noting that MCTS, when combined with UCB1, reaches asymptotically logarithmic regret~\cite{coquelin2007bandit}.
 
@@
 The general form of a MOOP is formally described as a maximization function $F_m(x)$, that transforms points in the decision space ($X$) to points in the solution space ($F$).
-The elements of the decision space are vectors of $n$ variables of the form $x = (x_1, x_2, \dots, x_n)$, while elements in the solution space are vectors with a dimension $m$: $F_m(x) = (f_1(x), f_2(x), \dots, f_m(x))$.
+The elements of the decision space are vectors of $n$ variables of the form $x = (x_1, x_2, \dots, x_n)$, while elements in the solution space are vectors of dimension $m$: $F_m(x) = (f_1(x), f_2(x), \dots, f_m(x))$.
 Therefore, each solution provides $m$ different scores (or rewards, or fitness) that are meant to be optimized.
@@
 
-It is said that a solution $F_m(x)$ \textit{dominates} another solution $F_m(y)$ if:
+It is said that a solution $F_m(x)$ \textit{dominates} another solution $F_m(y)$ if and only if:
 
@@
 However, there are some cases where it cannot be said that $F_m(x) \preceq F_m(y)$ or $F_m(y) \preceq F_m(x)$.
-This situation occurs when one of the objectives is better in $F_m(x)$ but a different objective is better in $F_m(y)$ (for instance, when $F_1(x) < F_1(y)$ but $F_2(x) > F_2(y)$).
+This situation occurs when one of the objectives is better in $F_m(x)$ but a different objective is better in $F_m(y)$ (for instance, when $f_1(x) < f_1(y)$ but $f_2(x) > f_2(y)$).
 In this case, it is said that these solutions are non-dominated with respect to each other.
@@
 	\includegraphics[scale=0.235]{img/moopDec.png}
-	\caption{Decision and Solution spaces in a MOOP with two variables ($x_1$ and $x_2$) and two objectives ($f_1$ and $f_2$).
+	\caption{Decision and solution spaces in a MOOP with two variables ($x_1$ and $x_2$) and two objectives ($f_1$ and $f_2$).
 In the objective space, yellow dots are non-optimal objective vectors, while blue dots form the Pareto-optimal front.}
@@
 
-As many Pareto front can be formed by several points in the solution space, it is important to devise a mechanism to assess the quality of a given front.
+As many Pareto fronts can be formed by several points in the solution space, it is important to devise a mechanism to assess the quality of a given front.
 A possibility is to use the Hypervolume Indicator (HV): given a Pareto front $P$, $HV(P)$ is defined as the volume of the objective space dominated by $P$.
@@
 Many different algorithms have been proposed to tackle multi-objective optimization problems in the literature.
-One of the most widely known and used methods is the weighted-sum approach.
-The procedure consists of giving a weight to each one of the objective and produce a single result as the linear combination of objectives and weights.
+One of the most popular methods used is the weighted-sum approach.
+The procedure consists of giving a weight to each one of the objectives and produce a single result as the linear combination of objectives and weights.
 By varying the weights provided, it is possible to converge to different solutions of the optimal Pareto front, if this is convex.
@@
 Multi-objective Reinforcement Learning (MORL)~\cite{Vamplew2010} changes this definition by using instead a vector $R = {r_0, r_1, \dots, r_n}$ as rewards of the problem.
-Thus, MORL problems differ from RL in having more than one objective objectives ($n$) that must be maximized.
+Thus, MORL problems differ from RL in having more than one objective ($n$) that must be maximized.
 If the objectives are independent or they do not oppose each other, scalarization technique approaches, as described above, could be suitable to tackle the problem.
-Essentially, this would mean to use a conventional RL algorithm on a single objective obtained by a weighted-sum of the multiple rewards.
+Essentially, this would mean to use a conventional RL algorithm on a single objective where the global reward is obtained from a weighted-sum of the multiple rewards.
 However, this is not always the case, as it is usual that the objectives are in conflict and the policy ($\pi$) must balance among them.
@@
 The second type of algorithms, \textit{multiple-policy}, target to approximate the optimal Pareto front of the problem.
+The algorithm proposed in this paper would fit in this category.
 An example of this type of algorithm is the one given by L.
@@
 The next step is to define the value for each pair of state and action, $W(s,a)$, as in Equation~\ref{eq:wang}.
-$\overline{r}^p_{sa}$ is the projection of $\overline{r}_{sa}$ into the piecewise linear surface defined by the Pareto archive $P$ (see Figure~\ref{fig:wangHV}.
+$\overline{r}^p_{sa}$ is the projection of $\overline{r}_{sa}$ into the piecewise linear surface defined by the Pareto archive $P$ (see Figure~\ref{fig:wangHV}).
 Then, $HV(P \cup \overline{r}_{sa})$ is declared as the HV of $P$ plus the point $\overline{r}_{sa}$.
@@
 
-The proposed algorithm was employed successfully in two domains: the DST and the Grid Scheduling problem, matching state of the art results in both domains, at the expense of a high computational cost.
+The proposed algorithm was employed successfully in two domains: the DST and the Grid Scheduling problem, matching state-of-the-art results in both domains, at the expense of a high computational cost.
 
@@
 In the vanilla algorithm, each node would use this vector $\overline{r}$ to update its own accumulated reward vector $\overline{R}$.
-Instead of doing this, each node in the MO-MCTS algorithm keeps a local Pareto front ($P$), updated at each iteration with the reward vector $\overline{r}$ obtained at the end of the simulation.
-Algorithm~\ref{alg:updPareto} describes how the node statistics are updated.
+In the algorithm proposed here, each node in the MO-MCTS algorithm also keeps a local Pareto front ($P$), updated at each iteration with the reward vector $\overline{r}$ obtained at the end of the simulation.
+Algorithm~\ref{alg:updPareto} describes how the node statistics are updated in MO-MCTS.
 
@@
 
-In this algorithm, if $\overline{r}$ is not dominated by the local Pareto front, it is added to the front and $\overline{r}$ is propagated to its parent.
-In case $\overline{r}$ is dominated by the node's Pareto front, the local Pareto front and there is no need to keep this propagation up the tree.
-
+Here, if $\overline{r}$ is not dominated by the local Pareto front, it is added to the front and $\overline{r}$ is propagated to its parent.
+In case $\overline{r}$ is dominated by the node's Pareto front, the local Pareto front does not change and there is no need to keep this propagation up the tree.
 Three observations can be made about the mechanism described here:
@@
 \item Each node in the tree has an estimate of the quality of the solutions reachable from there, both as an average (as in the baseline MCTS) and as the best case scenario (by keeping the non-dominated front $P$).
-\item By construction, if a reward $\overline{r}$ is dominated by the local front of a node, it is a given that it will be dominated by the nodes above in the tree, so there is no need to update the fronts of the upper nodes, producing little cost in the efficiency of the algorithm.
+\item By construction, if a reward $\overline{r}$ is dominated by the local front of a node, it is a given that it will be dominated by the nodes above in the tree, so there is no need to update the fronts of the upper nodes, producing little impact on the computational cost of the algorithm.
 \item It is easy to infer, from the last point, that the Pareto front of a node cannot be worse than the front of its children (in other words, the front of a child will never dominate that of its parent).
@@
 
-This algorithm, same as NSGA-II due to their multi-objective nature, provide a non-dominated front as a solution.
+This algorithm, same as in NSGA-II due to their multi-objective nature, provide a non-dominated front as a solution.
 However, in planning and control scenarios like the games analyzed in this research, an action must be provided to perform a move in the next step.
@@
 
-As shown before, it is straightforward to obtain which actions lead to what points in the front given as a solution: first gene in an NSGA-II individual, a root's child in MO-MCTS.
+As shown before, it is straightforward to obtain which actions lead to what points in the front given as a solution: first gene in the best NSGA-II individual, a root's child in MO-MCTS.
 Hence, by identifying the point in the Pareto front that the algorithm should converge to,
@@
 
-In order to select the point in the Pareto front, a weight vector $W (\forall w_i \in W, w_i \in [0,1])$ can be provided with the number of objectives as its dimension.
+In order to select a point in the Pareto front, a weight vector $W$ can be defined, with a dimension $n$ equal to the number of objectives ($W = (w_1, w_2, \dots, w_n); \sum_{i}^{n} w_i = 1$).
 Two different mechanisms are proposed, for reasons that will be explained in the experiments section (\ref{sec:exp}):
@@
 \begin{itemize}
-\item Weighted Sum: the action chosen is the one that maximizes the weighted sum of the solution vector multiplied by $W$, for each point in the front.
+\item Weighted Sum: the action chosen is the one that maximizes the weighted sum of the reward vector multiplied by $W$, for each point in the front.
 \item Euclidean distance: the euclidean distance from each point in the Pareto front (normalized in $[0,1]$) to the vector $W$ is calculated.
@@
 
-Two different are used in this research to analyze the performance of the algorithm proposed.
+Two different games are used in this research to analyze the performance of the algorithm proposed.
 
@@
 
-The Deep Sea Treasure (DST) is a well known multi-objective problem introduced by Vamplew et al.~\cite{Vamplew2010}.
+The Deep Sea Treasure (DST) is a popular multi-objective problem introduced by Vamplew et al.~\cite{Vamplew2010}.
 In this single-player puzzle, the agent moves a submarine with the objective of finding a treasure located at the bottom of the ship.
@@
 There are three types of cells: empty cells (or water), that the submarine can traverse; ground cells that, as the edges of the grid, cannot be traversed; and treasure cells, that provide different rewards and finish the game.
-Figure~\ref{fig:dstBase} (on the left) shows the environment of the DST.
+Figure~\ref{fig:dstBase} shows the DST environment.
 
@@
 There are two objectives in this game: the number of moves performed by the ship, that must be minimized, and the value of the treasure found, that should be maximized.
-As can be seen in Figure~\ref{fig:dstBase}, the most valuable treasures are at a greater distance from the initial position, making these objectives orthogonal.
+As can be seen in Figure~\ref{fig:dstBase}, the most valuable treasures are at a greater distance from the initial position, so the objectives are in conflict.
 
@@
 
-Section~\ref{sec:moo} introduced the problems of linear scalarization approaches when facing non-convex optimal Pareto front.
-The concave shape of the DST's optimal front makes those approximations to converge to the non dominated solutions located at the edges of the Pareto front ($(-19,1)$,$(-1,124)$).
+Section~\ref{sec:moo} introduced the problems of linear scalarization approaches when facing non-convex optimal Pareto fronts.
+The concave shape of the DST's optimal front makes those approximations to converge to the non dominated solutions located at the edges of the Pareto front: $(81,124)$ and $(99,1)$.
 Note that this happens independently from the weights chosen for the linear approximation: some solutions of the front just can't be reached with these approaches.
@@
 In this example, the submarine starts in the initial position ($P1$, root of the tree) and makes a sequence of moves that places it in $P2$.
-From this location, if the vessel wishes to move to $P3$, two optimal trajectories would be route \textit{a} and \textit{b}.
+From this location, two optimal trajectories to move to $P3$ would be route \textit{a} and \textit{b}.
 In a tree that does not use TT, there would be two different nodes to represent $P3$, although the location and number of moves performed up to this point are the same (thus, the states are equivalent).
@@
 The MO-PTSP is a real-time game where the agent navigates a ship and must visit $10$ waypoints scattered around the maze.
-All waypoints must be visited to consider a game as complete, and a game ticks timer is reset every time a waypoint is collected, finishing the game prematurely (and unsuccessfully) if it reaches $800$ game steps before visiting another waypoint.
+All waypoints must be visited to consider a game as complete, and a game-tick counter is reset every time a waypoint is collected, finishing the game prematurely (and unsuccessfully) if it reaches $800$ game steps before visiting another waypoint.
 
@@
 In the former case, the ship can collide with normal obstacles (that subtract $10$ units of damage) and specially damaging obstacles ($30$ units).
-In the latter, lava lakes are abundant in the MO-PTSP levels and, in contrast with normal surfaces, they deal one unit of damage for each time step the ship spends over this type of surface.
-All this subtractions are deduced from an initial counter of $5000$ points of damage.
+In the latter, lava lakes are abundant in MO-PTSP levels and, in contrast with normal surfaces, they deal one unit of damage for each time step the ship spends over this type of surface.
+All these subtractions are deduced from an initial counter of $5000$ points of damage.
 
@@
 The green ellipses represent fuel canisters, normal surfaces are drawn in brown and lava lakes are printed as red-dotted yellow surfaces.
-Normal obstacles black, damaging obstacles are drawn in red.
+Normal obstacles are black and damaging obstacles are drawn in red.
 Blue obstacles are elastic walls that produce no damage to the ship.
@@
 	\begin{center}
-	\includegraphics[width=0.7\columnwidth]{img/moptspmap}
+	\includegraphics[width=0.8\columnwidth]{img/moptspmap}
 	\caption{Sample MO-PTSP map.}
@@
 The main advantage of macro-actions is that the control algorithms can see further in the future, and then reduce the implication of open-endedness of real-time games (see Section~\ref{sec:mcts}).
-As this reduces the search space significantly, better algorithm performance is allowed.
+As this reduces the search space significantly, a better algorithm performance is allowed.
 A consequence of using macro-actions is also that the algorithm can employ $L$ consecutive steps to plan the next move (the next macro-action) to make.
@@
 
-In order to evaluate a game state in MO-PTSP, three different measures or rewards are taken, $r_t, r_f, r_d$, one for each one of the objectives time, fuel and damage, respectively.
+In order to evaluate a game state in MO-PTSP, three different measures or rewards are taken, $r_t, r_f, r_d$, one for each one of the objectives: time, fuel and damage, respectively.
 All these rewards are defined so they have to be maximized.
@@
 $d_s$ is the damage suffered so far in the game, and $d_m$ is the maximum damage the ship can take.
-In this case, three different variables are used to regulate the bahaviour of this objective: $\gamma$, $\beta_{1}$ and $\beta_{2}$.
-Both $\beta_{1}$ and $\beta_{2}$ have the same role as $\alpha$ in Equation~ref{eq:fuel}: they balance between the time objective and the damage measure.
+In this case, three different variables are used to regulate the behaviour of this objective: $\gamma$, $\beta_{1}$ and $\beta_{2}$.
+Both $\beta_{1}$ and $\beta_{2}$ have the same role as $\alpha$ in Equation~\ref{eq:fuel}: they balance between the time objective and the damage measure.
 The difference is that $\beta_{1}$ is used in high speeds, while $\beta_{2}$ is employed with low velocities.
@@
 
+It is important to highlight again that these three rewards, $r_t$, $r_f$ and $r_d$, are the same for the three algorithms tested in the experiments.
+
 \section{Experimentation} \label{sec:exp}
@@
 This NSGA-II version evolves a population where the individuals are sequences of actions (macro-actions in the MO-PTSP case), obtaining the fitness from the state of the game after applying such sequence.
+The populations sizes, determined empirically, were set to $20$ and $50$ individuals for DST and MO-PTSP respectively.
+The value of $C$ in Equation~\ref{eq:ucb1} and~\ref{eq:moucb} is set to $\sqrt{2}$.
 
@@
 is calculated and employed in the tests.
-This leaded to $4500$ evaluations in the DST, and $500$ evaluations for MO-PTSP, using the same server where the PTSP and MO-PTSP competitions were run\footnote{Intel Core i5, 230GHz, 4GB of RAM}.
+This leaded to $4500$ evaluations in the DST, and $500$ evaluations for MO-PTSP, using the same server where the PTSP and MO-PTSP competitions were run\footnote{Intel Core i5, 230GHz, 4GB of RAM.}.
 
@@
 
-As the optimal Pareto front of the DST is known, a measure of performance can be obtained by observing the percentage of times these solutions are found by the agents.
+As the optimal Pareto front of the DST is known, a measure of performance can be obtained by observing the percentage of times these solutions are found by the players.
 As the solution that the algorithms converge to depends on the weights vector employed during the search (see end of Section~\ref{sec:momcts}), the approach taken here is to provide different weight vectors $W$ and analyze them separately.
@@
 It is important to highlight that these two algorithms employed the euclidean distance action selection (see Section~\ref{sec:momcts}).
-Experiments shown that weighted sum action selection provides similar results as in MCTS (this is, convergence to the edges of the front).
-The crucial distinction to make is that both algorithms NSGA-II and MO-MCTS allow for different action selection mechanisms, that are able to overcome this problem, by approximating a global Pareto front.
+Other experiments, not included in this paper, showed that weighted sum action selection provides similar results as in MCTS (this is, convergence to the edges of the front).
+The crucial distinction to make here is that both algorithms NSGA-II and MO-MCTS allow for better action selection mechanisms, that are able to overcome this problem, by approximating a global Pareto front.
 
 Finally, in the comparison between NSGA-II and MO-MCTS, the latter algorithm obtains higher percentages for each one of the points in the front.
-This result suggests that, with a limited number of iterations/evaluations, the proposed algorithm is able to explore the search space more efficiently than a state of the art algorithm such NSGA-II.
+This result suggests that, with a limited number of iterations/evaluations, the proposed algorithm is able to explore the search space more efficiently than a rolling horizon version of a state-of-the-art NSGA-II.
 
@@
 The same three algorithms have been tested in the MO-PTSP domain.
-The framework~\footnote{www.ptsp-game.net} comes with $10$ maps, used in this research to run the algorithms in.
+The experiments are performed in the $10$ maps available within the framework~\footnote{Available at \url{www.ptsp-game.net}}.
 In this case, the optimal Pareto front is not known in advance, and normally it will be different from one map to another.
@@
 The idea is as follows.
-First of all, $4$ different weight vectors are tested: $W_1 = (0.33,0.33,0.33)$, $W_2 = (0.1,0.3,0.6)$, $W_3 = (0.1,0.6,0.3)$ and $W_4 = (0.6,0.1,0.3)$, where each $w_i$ corresponds to the weight for an objective ($w_t$ for time, $w_f$ for fuel and $w_d$ for damage).
+First of all, $4$ different weight vectors are tested: $W_1 = (0.33,0.33,0.33)$, $W_2 = (0.1,0.3,0.6)$, $W_3 = (0.1,0.6,0.3)$ and $W_4 = (0.6,0.1,0.3)$, where each $w_i$ corresponds to the weight for an objective ($w_t$ for time, $w_f$ for fuel and $w_d$ for damage, in this order).
 $W_1$ treats all objectives with the same weight, while the other three give more relevance to different objectives.
@@
 The procedure is as follows: once all games on a single map have been run, the Mann–Whitney–Wilcoxon non-parametric test with $95\%$ confidence is calculated on the three objectives.
-If the measures on all objectives are assumed to be drawn from different distributions, their averages are compared for a dominance test.
+If the measures on all objectives are assumed to be drawn from different distributions, their averages are compared for a dominance test (if not, then no dominance relationship can be derived from the results).
 If one result dominates another, then the algorithm dominates the other in that particular map.
@@
 MCTS and NSGA-II are dominated more often than MO-MCTS, being dominated on averages of $7.5$ and $5.75$ of the maps, respectively.
-This comparative result shows again that MO-MCTS is achieving the best results among the algorithms compared.
+This comparative result shows again that MO-MCTS is achieving the best results among the three algorithms compared here.
 
@@
 It is naive to think that an unique weight vector will be the ideal one for the whole game.
-Specifically in the MO-PTSP, there are regions of the map where there are more obstacles on lava lakes, hence most likely to damage the ship.
+Specifically in the MO-PTSP, there are regions of the map where there are more obstacles or lava lakes, hence the ship is most likely to suffer a higher damage.
 Also, the route followed during the game affects the relative ideal speed between waypoints, or sometimes a fuel canister is going to be picked up, which affects how the fuel objective can be managed.
@@
 A way to provide different weights at different times in MO-PTSP is straightforward.
-Given the route of waypoints (or fuel canisters) that is being followed, one can divide it into \textit{segments}, where is segment starts and ends with a waypoint (or fuel canister).
+Given the route of waypoints (and fuel canisters) that is being followed, one can divide it into \textit{segments}, where is segment starts and ends with a waypoint (or fuel canister).
 Then, each segment can be assigned a particular weight vector $W$.
@@
 
-This section shows some initial results in order to test the third variant, by using a stochastic hill climbing algorithm on each one of the maps.
+This section shows some initial results in order to test the third variant, by using a stochastic hill climbing algorithm on each map.
 The goal is to check if, by varying the weight vectors between segments, better solutions can be achieved.
 
-A solution is identified by a string of integers, where each integer refers to one of the weight vectors utilized in the previous sections ( $1 = W_1 = (0.33,0.33,0.33)$, $2 = W_2 = (0.1,0.3,0.6)$ and $3 = W_3 = (0.1,0.6,0.3)$.
-$W_4$ has been left out of this experiment, as it shown not to be that influential).
+An individual is identified by a string of integers, where each integer refers to one of the weight vectors utilized in the previous sections ( $1 = W_1 = (0.33,0.33,0.33)$, $2 = W_2 = (0.1,0.3,0.6)$ and $3 = W_3 = (0.1,0.6,0.3)$.
+$W_4$ has been left out of this experiment, as it has shown to be the less influential weight vector).
 The solution is evaluated playing a particular map $10$ times, and its fitness is obtained by calculating the average of those runs.
@@
 The best solution, determined by dominance, is kept and promoted to the next generation, where it is mutated to generate other individuals of the population.
-Also, a portion of the individuals of the next population is created randomly at every generation, until the end of the algorithm, which is established at $50$ generations.
+Also, a portion of the individuals of the next population is created uniformly at random at every generation, until the end of the algorithm, which is established at $50$ generations.
 
@@
 The forth result on each row is the best one after the run.
-Each genome is a string of the form $xyz$, that represents $W_xW_yW_z$, where each value is a weight used in that particular segment.
-The last column indicates if the evolved genome dominates ($\preceq$) or not ($\O$) each one of the base genomes for that particular map.
+Each genome is a string of the form $abc \dots z$, that represents $W_aW_bW_c \dots W_z$, where each element is a weight vector used in that particular segment.
+The last column indicates if the evolved individual dominates ($\preceq$) or not ($\O$) each one of the base genomes for that particular map.
 
@@
  \multirow{3}{*}{\textbf{Map 2}} & $11111111111111$ & $1409 \pm 7$ & $235 \pm 4$& $364 \pm 3$ & $\preceq$ \\
- & $22222222222222$ & $1402 \pm 6$ & $236 \pm 5$ & \bm{$354 \pm 2$} & $\preceq$ \\
- & $33333333333333$ & $1416 \pm 8$ & \bm{$219 \pm 4$} & $360 \pm 3$ & $\preceq$ \\
+ & $22222222222222$ & $1402 \pm 6$ & $236 \pm 5$ & $354 \pm 2$ & $\preceq$ \\
+ & $33333333333333$ & $1416 \pm 8$ & $219 \pm 4$ & $360 \pm 3$ & $\preceq$ \\
  & $23131312323213$ & $1390 \pm 10$ & $210 \pm 3$ & $353 \pm 3$ & \\
@@
  \multirow{3}{*}{\textbf{Map 3}} & $11111111111111$ & $1373 \pm 6$ & $221 \pm 3$& $301 \pm 7$ & $\preceq$ \\
- & $22222222222222$ & $1378 \pm 5$ & $211 \pm 3$ & \bm{$268 \pm 5$} & $\preceq$ \\
- & $33333333333333$ & $1385 \pm 6$ & \bm{$203 \pm 4$} & $291 \pm 7$ & $\O$ \\
+ & $22222222222222$ & $1378 \pm 5$ & $211 \pm 3$ & $268 \pm 5$ & $\preceq$ \\
+ & $33333333333333$ & $1385 \pm 6$ & $203 \pm 4$ & $291 \pm 7$ & $\O$ \\
  & $11122222112231$ & $1358 \pm 9$ & $219 \pm 7$ & $263 \pm 12$ & \\
@@
  \multirow{3}{*}{\textbf{Map 4}} & $11111111111111$ & $1383 \pm 6$ & $291 \pm 5$& $565 \pm 5$ & $\preceq$ \\
- & $22222222222222$ & $1385 \pm 7$ & $304 \pm 4$ & \bm{$542 \pm 4$} & $\preceq$ \\
- & $33333333333333$ & $1423 \pm 6$ & \bm{$273 \pm 4$} & $583 \pm 5$ &  $\O$ \\
+ & $22222222222222$ & $1385 \pm 7$ & $304 \pm 4$ & $542 \pm 4$ & $\preceq$ \\
+ & $33333333333333$ & $1423 \pm 6$ & $273 \pm 4$ & $583 \pm 5$ &  $\O$ \\
  & $11121131212112$ & $1360 \pm 4$ & $282 \pm 5$ & $540 \pm 4$ & \\
@@
  \multirow{3}{*}{\textbf{Map 5}} & $11111111111111$ & $1405 \pm 7$ & $467 \pm 4$& $559 \pm 4$ & $\preceq$ \\
- & $22222222222222$ & $1431 \pm 9$ & $447 \pm 4$ & \bm{$541 \pm 4$} & $\preceq$ \\
- & $33333333333333$ & $1467 \pm 9$ & \bm{$411 \pm 5$} & $567 \pm 5$ & $\O$ \\
+ & $22222222222222$ & $1431 \pm 9$ & $447 \pm 4$ &  $541 \pm 4$ & $\preceq$ \\
+ & $33333333333333$ & $1467 \pm 9$ & $411 \pm 5$ & $567 \pm 5$ & $\O$ \\
  & $21311213111211$ & $1397 \pm 11$ & $448 \pm 10$ & $535 \pm 5$ & \\
@@
  \multirow{3}{*}{\textbf{Map 6}} & $11111111111111$ & $1575 \pm 7$ & $549 \pm 5$& $303 \pm 4$ & $\preceq$ \\
- & $22222222222222$ & $1626 \pm 9$ & $540 \pm 6$ & \bm{$286 \pm 5$} & $\preceq$ \\
- & $33333333333333$ & $1703 \pm 11$ & \bm{$499 \pm 4$} & $316 \pm 7$ & $\O$ \\
+ & $22222222222222$ & $1626 \pm 9$ & $540 \pm 6$ & $286 \pm 5$ & $\preceq$ \\
+ & $33333333333333$ & $1703 \pm 11$ & $499 \pm 4$ & $316 \pm 7$ & $\O$ \\
  & $31121312111111$ & $1570 \pm 16$ & $535 \pm 10$ & $266 \pm 6$ & \\
@@
  \multirow{3}{*}{\textbf{Map 7}} & $11111111111111$ & $1434 \pm 5$ & $599 \pm 6$& $284 \pm 6$ & $\preceq$ \\
- & $22222222222222$ & $1475 \pm 10$ & $602 \pm 5$ & \bm{$243 \pm 6$} & $\preceq$ \\
- & $33333333333333$ & $1489 \pm 12$ & \bm{$549 \pm 3$} & $264 \pm 6$ & $\O$ \\
+ & $22222222222222$ & $1475 \pm 10$ & $602 \pm 5$ & $243 \pm 6$ & $\preceq$ \\
+ & $33333333333333$ & $1489 \pm 12$ & $549 \pm 3$ & $264 \pm 6$ & $\O$ \\
  & $11332211321332$ & $1401 \pm 12$ & $563 \pm 4$ & $230 \pm 12$ & \\
@@
  \multirow{3}{*}{\textbf{Map 8}} & $11111111111111$ & $1761 \pm 9$ & $254 \pm 5$& $382 \pm 3$ & $\preceq$ \\
- & $22222222222222$ & $1804 \pm 10$ & $269 \pm 4$ & \bm{$357 \pm 4$} & $\preceq$ \\
- & $33333333333333$ & $1826 \pm 10$ & \bm{$230 \pm 3$} & $392 \pm 7$ & $\O$ \\
+ & $22222222222222$ & $1804 \pm 10$ & $269 \pm 4$ & $357 \pm 4$ & $\preceq$ \\
+ & $33333333333333$ & $1826 \pm 10$ & $230 \pm 3$ & $392 \pm 7$ & $\O$ \\
  & $23221131313323$ & $1747 \pm 11$ & $247 \pm 9$ & $363 \pm 9$ & \\
@@
  \multirow{3}{*}{\textbf{Map 9}} & $11111111111111$ & $2501 \pm 14$ & $926 \pm 6$& $574 \pm 9$ & $\preceq$ \\
- & $22222222222222$ & $2503 \pm 10$ & $921 \pm 10$ & \bm{$524 \pm 8$} & $\preceq$ \\
- & $33333333333333$ & $2641 \pm 14$ & \bm{$833 \pm 5$} & $574 \pm 14$ & $\O$ \\
+ & $22222222222222$ & $2503 \pm 10$ & $921 \pm 10$ & $524 \pm 8$ & $\preceq$ \\
+ & $33333333333333$ & $2641 \pm 14$ & $833 \pm 5$ & $574 \pm 14$ & $\O$ \\
  & $21132331333223$ & $2463 \pm 19$ & $891 \pm 8$ & $523 \pm 9$ & \\
@@
  & $22222222222222$ & $1493 \pm 13$ & $615 \pm 4$ & $209 \pm 2$ & $\preceq$ \\
- & $33333333333333$ & $1542 \pm 10$ & \bm{$554 \pm 4$} & $229 \pm 5$ & $\O$ \\
+ & $33333333333333$ & $1542 \pm 10$ & $554 \pm 4$ & $229 \pm 5$ & $\O$ \\
  & $11311111322231$ & $1418 \pm 9$ & $623 \pm 9$ & $197 \pm 2$ & \\
@@
 
-These results suggest some interesting concepts.
+The results suggest some interesting ideas.
 First of all, it is indeed possible to obtain better results by varying the weights of the objectives along the route: in $22$ out of the $30$ comparisons made, the evolved solution dominates the base individuals and, in the other $8$, both solutions in the comparison would be in the same Pareto front of solutions.
@@
 If the focus is set on what particular weights are more successful, it is interesting to see that the best configurations found are better, in all maps, than the base individuals with all weights equal to $W_1$ ($w_t = 0.33, w_f = 0.33, w_d = 0.33$) and $W_2$ ($w_t = 0.1, w_f = 0.3, w_d = 0.6$).
-In other words, it was always possible to find a combination of weight vectors that performed better than approaches that gave the same weight to all objectives, and also the ones prioritizing low damage.
+In other words, it was always possible to find a combination of weight vectors that performed better than approaches that gave the same weight to all objectives, and also than the ones prioritizing low damage.
 
@@
 
-This paper presents a multi-objective version of Monte Carlo Tree Search (MO-MCTS) for real-time games, scenarios where the time budget allowed for finding out the best possible action is significantly reduced.
-MO-MCTS is tested, in comparison with a single MCTS algorithm and a rolling horizon NSGA-II, in two different real-time games, the Deep Sea Treasure (DST) and the Multi-Objective Physical Travelling Salesman Problem (MO-PTSP).
+This paper presents a multi-objective version of Monte Carlo Tree Search (MO-MCTS) for real-time games, scenarios where the time budget allowed for finding out the best possible action to apply next is significantly reduced.
+MO-MCTS is tested, in comparison with a single-objective MCTS algorithm and a rolling horizon NSGA-II, in two different real-time games, the Deep Sea Treasure (DST) and the Multi-Objective Physical Travelling Salesman Problem (MO-PTSP).
 This comparison is made by using the same heuristics that determine the value of a given state, so the focus is set on how the algorithms explore the search space on each problem.
@@
 The results obtained in this study show the proficiency of the algorithm proposed, as it provides better solutions than the ones obtained by the other algorithms in comparison.
-Also, the MO-MCTS approach samples successfully across the different solutions in the Pareto front (optimal front in the DST, the best found in MO-PTSP), depending on the weights provided to each objective.
-Finally, some initial results are obtained on the idea of using distinct weight vectors for the objective in different situations on the game, showing that it is possible to improve the results by performing these variations.
-
-The stochastic hill climbing algorithm used for creating strings of weight vectors for the MO-PTSP is relatively simple, although it was able to obtain very good results.
+Also, the MO-MCTS approach samples successfully across the different solutions in the Pareto front (optimal front in the DST, the best found in MO-PTSP), depending on the weights provided for each objective.
+Finally, some initial results are obtained applying the idea of using distinct weight vectors for the objectives in different situations within the game, showing that it is possible to improve the performance of the algorithms.
+
+The stochastic hill climbing algorithm used for creating strings of weight vectors for the MO-PTSP is relatively simple, although it was able to obtain very good results in few iterations.
 However, it could be desirable to model behaviours that dynamically change the weight vectors according to measurements of the game state, instead of evolving a different solution for each particular map.
@@
 
-Finally, the results this paper allow to presume that multi-objective approaches can provide high quality level of play in real-time games, and not only in the ones used as benchmarks in this research.
+Finally, the results described in this paper allow to presume that multi-objective approaches can provide high quality level of play in real-time games, and not only in the ones used as benchmarks in this research.
 Games (and other domains) where the goal can be decomposed in several simultaneous tasks, especially when these tasks are in conflict with each other, can benefit from approaches like the one presented here.

### 2014-01-08 paper/paper.tex
First Commit - not working

@@
 
-\usepackage{bm,subfigure,graphicx,url,times,multirow,amsmath,amssymb,algorithm,xspace,epsfig,todonotes,array,caption,color}
+\usepackage{bm,subfigure,graphicx,url,times,multirow,amsmath,amssymb,algorithm,xspace,epsfig,todonotes,array,caption,color,todonotes}
 \usepackage{algorithm, mathtools}
@@
 Multi-objective optimization has been traditionally a matter of study in domains like engineering or finance, with little impact on games research.
-However, action-decision based on a multi-objective evaluation may be beneficial to obtain a high quality level of play.
-This paper presents a Multi-objective Monte Carlo Tree Search algorithm for planning and control in real-time game domains, those where the time budget to decide the next move to make is significantly reduced.
-A comparison is made between the proposed algorithm, a single-objective version of Monte Carlo Tree Search and a rolling horizon implementation of Non-dominated Sorting Evolutionary Algorithm II.
-Two different benchmarks are employed, the Deep Sea Treasure and the Multi-Objective Physical Travelling Salesman Problem.
+However, action-decision based on a multi-objective evaluation may be beneficial in order to obtain a high quality level of play.
+This paper presents a Multi-objective Monte Carlo Tree Search algorithm for planning and control in real-time game domains, those where the time budget to decide the next move to make is close to 40ms.
+A comparison is made between the proposed algorithm, a single-objective version of Monte Carlo Tree Search and a rolling horizon implementation of Non-dominated Sorting Evolutionary Algorithm II (NDSAA-II).
+Two different benchmarks are employed, Deep Sea Treasure and Multi-Objective Physical Travelling Salesman Problem.
 Using the same heuristics on each game, the analysis is focused on how well the algorithms explore the search space.
-The results show that the proposed approach provides better results than the other algorithms it is compared with.
-Additionally, it is also shown that the algorithm is able to converge to different solutions of the optimal, or the best found during the search, Pareto front.
+Results show that the algorithm proposed outperforms (NDSAA-II).
+Additionally, it is also shown that the algorithm is able to converge to different optimal solutions or the optimal Pareto front (if achieved during search).
 \end{abstract}
@@
 Multi-objective optimization has been a field of study of manufacturing, engineering~\cite{Marler2004} and finance~\cite{Coello2006}, while having little impact in games research.
-This paper shows that multi-objective optimization has much to offer in developing game strategies that allow for a fine-grain control of alternative policies.
+This paper shows that multi-objective optimization has much to offer in developing game strategies that allow for a fine-grained control of alternative policies.
 The application of such approaches to this field can provide interesting results, especially in games that are long or complex enough where long-term planning is not trivial, and achieving a good level of play requires balancing strategies.
@@
 From a simple point of view, many competitive games can be seen as a scenario where two or more opponents have the single objective of winning.
-Achieving victory is usually complex in interesting games, and successful approaches normally provide a state value by combining features of the game state as a weighted sum.
-An example is a chess heuristic that assigns different weights to each piece according to their value.
-
-Multi-objective approaches can be applied in these scenarios, where several factors must be taken into account to achieve the victory.
-The algorithms can balance between different objectives, in order to provide a wide range of strategies well suited to the different stages of the game being played, or to face the existing opponents.
+Achieving victory is usually complex in interesting games, and successful approaches normally assign some form of value in a state, value being long term expected reward (i.e.
+expectation of victory).
+Due to a massive search space, heuristics are often used for this value function.
+An example is a chess heuristic that assigns different weights to each piece according to an estimated value.
+
+Multi-objective approaches can be applied to scenarios where several factors must be taken into account to achieve victory.
+The algorithms can balance between different objectives, in order to provide a wide range of strategies well suited to the different stages of the game being played, or to face existing opponents.
 Application of such approaches could be real-time strategy games, where long term planning must be carried out in order to balance aspects such as attack units, defensive structures and resource gathering.
-
-This paper proposes a multi-objective real-time algorithm version of Monte Carlo Tree Search (MCTS), a popular reinforcement learning approach within the last decade.
+This should allow allow smart balancing of heuristics.
+
+This paper proposes a multi-objective real-time algorithm version of Monte Carlo Tree Search (MCTS), a popular reinforcement learning approach that emerged the last decade.
 The proposed algorithm is tested in two different games: a real-time version of the Deep Sea Treasure (DST; a classical multi-objective problem), and the Multi-Objective Physical Travelling Salesman Problem (MO-PTSP).
@@
 
-Monte Carlo Tree Search (MCTS) is a tree search algorithm that was originally applied to board games, concretely to the two-players game of Go.
+Monte Carlo Tree Search (MCTS) is a tree search algorithm that was originally applied to board games, gaining momentum and popularity in the game of Go.
 This game is played in a square grid board, with a size of $19 \times 19$ in the original game, and $9 \times 9$ in its reduced version.
@@
 This is true independently from how many iterations the algorithm was able to make (although, in general, more iterations usually produce better results).
-This differs from other algorithms (such A*) that normally provide the next ply only after they have finished.
+This differs from other algorithms (such A*) that normally provide the next ply only after they have finished \todo[inline]{Spyros: A* Cannot play two-player games.
+Min-Max is used as an anytime algorithm with ``iterative deepening'' }.
 This makes MCTS a suitable candidate for real-time domains, where the decision time budget is limited, affecting the number of iterations that can be performed.

### 2014-01-08 paper/paper.tex
Spyros changes in

@@
 This paper presents a Multi-objective Monte Carlo Tree Search algorithm for planning and control in real-time game domains, those where the time budget to decide the next move to make is close to 40ms.
-A comparison is made between the proposed algorithm, a single-objective version of Monte Carlo Tree Search and a rolling horizon implementation of Non-dominated Sorting Evolutionary Algorithm II (NDSAA-II).
+A comparison is made between the proposed algorithm, a single-objective version of Monte Carlo Tree Search and a rolling horizon implementation of Non-dominated Sorting Evolutionary Algorithm II (NSGA-II).
 Two different benchmarks are employed, Deep Sea Treasure and Multi-Objective Physical Travelling Salesman Problem.
 Using the same heuristics on each game, the analysis is focused on how well the algorithms explore the search space.
-Results show that the algorithm proposed outperforms (NDSAA-II).
+Results show that the algorithm proposed outperforms (NSGA-II).
 Additionally, it is also shown that the algorithm is able to converge to different optimal solutions or the optimal Pareto front (if achieved during search).
@@
 In every one of this action decisions, MCTS balances between exploitation and exploration.
-In other words, this chooses between taking an action that leads to states with the best outcome found so far, and performing a move to go to less explored game states, respectively.
+In other words, it chooses between taking an action that leads to states with the best outcome found so far, and performing a move to go to less explored game states, respectively.
 In order to achieve this, MCTS uses Upper Confidence Bound (UCB1, see Equation~\ref{eq:ucb1}) as a \textit{Tree Policy}.
@@
 
-The balance between exploration and exploitation is achieved by setting the value of $C$.
+The balance between exploration and exploitation can be tampered by modifying $C$.
 Higher values of $C$ weight more the second term of the UCB1 Equation~\ref{eq:ucb1}, giving preference to those actions that have been explored less, at the expense of taking actions with the highest average reward $Q(s,a)$.
 A commonly used value is $\sqrt{2}$, as it balances both facets of the search when the rewards are normalized between $0$ and $1$.
-It is worth noting that MCTS, when combined with UCB1, reaches asymptotically logarithmic regret~\cite{coquelin2007bandit}.
+It is worth noting that MCTS, when combined with UCB1, reaches asymptotically logarithmic regret~\cite{coquelin2007bandit} \todo[inline]{I though only UCB1 had logarithmic regret, UCT is far worse, are you sure ?
+}.
 
@@
 
-MCTS has been employed extensively in real-time games in the literature.
+MCTS has been employed extensively in real-time games.
 A clear example of this is the popular real-time game \textit{Ms.
@@
 The consequence of this is that MCTS, in its vanilla form, it is not able to know if a given ply will lead to a win or a loss game end state.
-Robles et al~\cite{Robles2009} solved this problem by including hand-coded heuristics that guided MCTS simulations towards more promising portions of the search space.
-Other authors also included domain knowledge to bias the search in MCTS, such as in~\cite{Samothrakis, Ikehata2010}.
+Robles et al.~\cite{Robles2009} solved this problem by including hand-coded heuristics that guided search towards more promising portions of the search space.
+This approach has been followed in order to add heuristics knowledge to MCTS, as in~\cite{Samothrakis, Ikehata2010}.
 
@@
 
-Many different algorithms have been proposed to tackle multi-objective optimization problems in the literature.
+Many different algorithms have been proposed to tackle multi-objective optimization problems.
 One of the most popular methods used is the weighted-sum approach.
@@
 A popular choice for multi-objective optimization problems are evolutionary multi-objective optimization (EMOA) algorithms~\cite{Coello1999, Zhou2011}.
-One of the most well known algorithms in the literature is the Non-dominated Sorting Evolutionary Algorithm 2 (NSGA-2), which pseudocode is shown in Algorithm~\ref{alg:nsga2}.
+One of the most well known algorithms in the literature is the Non-dominated Sorting Evolutionary Algorithm II (NSGA-II), pseudocode of which is shown in Algorithm~\ref{alg:nsga2}.
 As in any evolutionary algorithm, NSGA-2 evolves a set of individuals or solutions to the problem, with the difference that here they are ranked according to dominance criteria and crowding distance (distances between members of the pareto fronts).
@@
 
-Reinforcement Learning (RL) algorithms have also tackled Multi-objective optimization in some scenarios.
+Reinforcement Learning (RL) algorithms have also tackled Multi-objective optimization in some settings.
 RL~\cite{Sutton1998} is a broad field in Machine Learning that studies real-time planning and control situations where  an agent has to find out the actions (or sequences of actions) that should be applied in order to maximize the reward from the environment.
 
-An RL problem can be defined as a tuple ($S$, $A$, $T$, $R$, $\pi$).
-$S$ is the set of possible states in the problem (or game), and $s_0$ is the initial state.
+The dynamics of an RL problem are usually captured by a Markov Decision Process (MDP).
+The (finite, undiscounted) is a tuple $(S, A, T, R)$, where $S$ is the set of possible states in the problem (or game), and $s_0$ is the initial state.
 $A$ is the set of available actions the agent can make at any given time, and the transition model $T(s_i, a_i, s_{i+1})$ determines the probability of reaching the state $s_{i+1}$ when action $a_i$ is applied in state $s_i$.
 The reward function $R(s_i)$ provides a single value (\textit{reward}) that the agent must optimize, representing the desirability of the state $s_i$ reached.
-Finally, a decision policy $\pi(s_i) = a_i$ determines which actions $a_i$ must be chosen from each state $s_i \in S$.
-One of the most important challenges in RL, as shown in Section~\ref{sec:mcts}, is the trade-off between exploration and exploitation.
-The decision policy can choose between following actions that provided good rewards in the past and exploring new parts of the search space by selecting new actions.
-
-Multi-objective Reinforcement Learning (MORL)~\cite{Vamplew2010} changes this definition by using instead a vector $R = {r_0, r_1, \dots, r_n}$ as rewards of the problem.
+Finally, a policy $\pi(s_i) = a_i$ determines which actions $a_i$ must be chosen from each state $s_i \in S$\footnote{This is a deterministic policy and only valid in during acting, not leaning}.
+One of the most important challenges in RL, as shown in Section~\ref{sec:mcts}, is the trade-off between exploration and exploitation while trying to act.
+While learning, a  policy must choose between following actions that provided good rewards in the past and exploring new parts of the search space by selecting new actions.
+
+Multi-objective Reinforcement Learning (MORL)~\cite{Vamplew2010} changes this definition by using a vector $R = {r_0, r_1, \dots, r_n}$ as rewards of the problem (instead of a scalar).
 Thus, MORL problems differ from RL in having more than one objective ($n$) that must be maximized.
@@
 Essentially, this would mean to use a conventional RL algorithm on a single objective where the global reward is obtained from a weighted-sum of the multiple rewards.
-However, this is not always the case, as it is usual that the objectives are in conflict and the policy ($\pi$) must balance among them.
+However, this is not always the case, as it is usual that the objectives are in conflict and the policy $\pi$ must balance among them.
 
@@
 
-The proposed algorithm was employed successfully in two domains: the DST and the Grid Scheduling problem, matching state-of-the-art results in both domains, at the expense of a high computational cost.
+The proposed algorithm was employed successfully in two domains: the DST and the Grid Scheduling problem, matching state-of-the-art results in both cases, at the expense of a high computational cost.
 
@@
 \begin{center}
-\includegraphics[width=1.0\columnwidth]{img/optDst.png}
+\includegraphics[width=1.0\columnwidth]{img/optDST.png}
 \end{center}
@@
 
-Some exceptions can actually be explained analyzing the map: in map 1, a fuel canister is always collected near the end, restoring the fuel level to its maximum.
+Some exceptions can actually be explained analysing the map: in map 1, a fuel canister is always collected near the end, restoring the fuel level to its maximum.
 This leaves too few cycles to make a difference in the controller (note also that this is the map with the fewest fuel consumption).
@@
 
-One of the first things to notice is that MO-MCTS dominates MCTS and NSGA-II in most of the maps, and it is never dominated by these.
+One of the first things to notice is that MO-MCTS dominates MCTS and NSGA-II in most of the maps, and it is never dominated in any.
 Particularly, the dominance of MO-MCTS over MCTS is outstanding, even dominating in all $10$ maps for two of the weight vectors.
@@
 
-It is also interesting to see that NSGA-II dominates also the weighted sum version of MCTS, although for the vector $W_2$ there is a technical draw, as they dominate each other in $4$ different maps each and there is no dominance in the other $2$.
+It is also interesting to see that NSGA-II dominates the weighted sum version of MCTS, although for the vector $W_2$ there is a technical draw, as they dominate each other in $4$ different maps each and there is no dominance in the other $2$.
 
@@
 
-This paper presents a multi-objective version of Monte Carlo Tree Search (MO-MCTS) for real-time games, scenarios where the time budget allowed for finding out the best possible action to apply next is significantly reduced.
+This paper presents a multi-objective version of Monte Carlo Tree Search (MO-MCTS) for real-time games, scenarios where the time budget allowed for finding out the best possible action to apply next is clost to ~$40$ms.
 MO-MCTS is tested, in comparison with a single-objective MCTS algorithm and a rolling horizon NSGA-II, in two different real-time games, the Deep Sea Treasure (DST) and the Multi-Objective Physical Travelling Salesman Problem (MO-PTSP).
@@
 
-The results obtained in this study show the proficiency of the algorithm proposed, as it provides better solutions than the ones obtained by the other algorithms in comparison.
+The results obtained in this study show the strength of the algorithm proposed, as it provides better solutions than the ones obtained by the other algorithms in comparison.
 Also, the MO-MCTS approach samples successfully across the different solutions in the Pareto front (optimal front in the DST, the best found in MO-PTSP), depending on the weights provided for each objective.
@@
 However, it could be desirable to model behaviours that dynamically change the weight vectors according to measurements of the game state, instead of evolving a different solution for each particular map.
-A possible extension for this work is to analyze and discover features in the game state that allows to establish relationships between game situations and the weight vectors to use for each objective.
+A possible extension for this work is to analyse and discover features in the game state that allows to establish relationships between game situations and the weight vectors to use for each objective.
 This mechanism would be general enough to be applicable to many different maps without requiring specific learning in any particular level.
 
-Finally, the results described in this paper allow to presume that multi-objective approaches can provide high quality level of play in real-time games, and not only in the ones used as benchmarks in this research.
-Games (and other domains) where the goal can be decomposed in several simultaneous tasks, especially when these tasks are in conflict with each other, can benefit from approaches like the one presented here.
+Finally, the results described in this paper allow us to presume that multi-objective approaches can provide a high quality level of play in real-time games.
+Multi-objective problems pop up in different settings and by using MCTS we can balance not only between exploration and exploitation but between multiple objectives as well.
 
@@
 
-\begin{biography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{photos/lucas.png}}]{Simon Lucas}
+\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{photos/lucas.png}}]{Simon Lucas}
 (SMIEEE) is a professor of Computer Science at the University of Essex (UK) where he leads the Game Intelligence Group.
@@
 He is the inventor of the scanning n-tuple classifier, and is the founding Editor-in-Chief of the IEEE Transactions on Computational Intelligence and AI in Games.
-\end{biography}
+\end{IEEEbiography}
 

### 2014-01-13 paper/paper.tex
ssamot changes and comments in

@@
 This is true independently from how many iterations the algorithm was able to make (although, in general, more iterations usually produce better results).
-This differs from other algorithms (such A*) that normally provide the next ply only after they have finished \todo[inline]{Spyros: A* Cannot play two-player games.
-Min-Max is used as an anytime algorithm with ``iterative deepening'' }.
+This differs from other algorithms (such A* in single player games, and standard Min-Max for two player games) that normally provide the next ply only after they have finished.
 This makes MCTS a suitable candidate for real-time domains, where the decision time budget is limited, affecting the number of iterations that can be performed.
@@
 A commonly used value is $\sqrt{2}$, as it balances both facets of the search when the rewards are normalized between $0$ and $1$.
-It is worth noting that MCTS, when combined with UCB1, reaches asymptotically logarithmic regret~\cite{coquelin2007bandit} \todo[inline]{I though only UCB1 had logarithmic regret, UCT is far worse, are you sure ?
-}.
+It is worth noting that MCTS, when combined with UCB1, reaches asymptotically logarithmic regret on each node of the tree~\cite{coquelin2007bandit}.
 

### 2014-01-28 paper/paper.tex
first Sanaz changes

@@
 In order to compare several algorithms between each other, the results are compared in pairs, in terms of dominance.
-The procedure is as follows: once all games on a single map have been run, the Mann–Whitney–Wilcoxon non-parametric test with $95\%$ confidence is calculated on the three objectives.
+The procedure is as follows: once all games on a single map have been run, the Mann-Whitney-Wilcoxon non-parametric test with $95\%$ confidence is calculated on the three objectives.
 If the measures on all objectives are assumed to be drawn from different distributions, their averages are compared for a dominance test (if not, then no dominance relationship can be derived from the results).

### 2014-02-09 paper/paper.tex
Sanaz changes in (first part) and variables changed for consistency

@@
 
-\author{Diego Perez, {\it Student Member, IEEE}, Sanaz Mostaghim, Spyridon Samothrakis, {\it Student Member, IEEE}, Simon M. Lucas, {\it Senior Member, IEEE}}
+\author{Diego Perez, {\it Student Member, IEEE}, Sanaz Mostaghim {Member, IEEE}, Spyridon Samothrakis, {\it Student Member, IEEE}, Simon M. Lucas, {\it Senior Member, IEEE}}
 
@@
 This is true independently from how many iterations the algorithm was able to make (although, in general, more iterations usually produce better results).
-This differs from other algorithms (such A* in single player games, and standard Min-Max for two player games) that normally provide the next ply only after they have finished.
+This differs from other algorithms (such A* in single player games, and standard Min-Max for two player games) that normally provide the next play only after they have finished.
 This makes MCTS a suitable candidate for real-time domains, where the decision time budget is limited, affecting the number of iterations that can be performed.
@@
 
-A multi-objective optimization problem (MOOP) represents a scenario where two or more objective functions are to be optimized (either maximized or minimized).
-The general form of a MOOP is formally described as a maximization function $F_m(x)$, that transforms points in the decision space ($X$) to points in the solution space ($F$).
-The elements of the decision space are vectors of $n$ variables of the form $x = (x_1, x_2, \dots, x_n)$, while elements in the solution space are vectors of dimension $m$: $F_m(x) = (f_1(x), f_2(x), \dots, f_m(x))$.
-Therefore, each solution provides $m$ different scores (or rewards, or fitness) that are meant to be optimized.
-Without loss of generality, it is assumed from now on that all objectives must be maximized.
-
-It is said that a solution $F_m(x)$ \textit{dominates} another solution $F_m(y)$ if and only if:
-
+A multi-objective optimization problem (MOP) represents a scenario where two or more conflicting objective functions are to be optimized at the same time and is defined as: 
+\begin{eqnarray}
+optimize ~~~\{ f_1(\vec x), f_2(\vec x),\cdots , f_m(\vec x) \} 
+\end{eqnarray}
+subject to $\vec x \in \Omega$, involving $m (\geq 2)$ conflicting objective functions $f_i : \Re ^n \rightarrow \Re$. The \textit{decision vectors} 
+$\vec x = (x_1, x_2, \cdots, x_n)^T$ belong to the feasible region $\Omega \subset \Re ^n$. We denote the image of the feasible region by $Z\subset \Re ^m$ and call it a feasible objective region. The
+elements of $Z$ are called objective vectors and they consist of $m$ objective (function) values $\vec f(\vec x)= (f_1(\vec x), f_2(\vec x), \cdots, f_m(\vec x))$.
+Therefore, each solution $\vec x$ provides $m$ different scores (or rewards, or fitness) that are meant to be optimized.
+Without loss of generality, it is assumed from now on that all objectives are to be maximized.
+
+It is said that a solution $\vec x$ \textit{dominates} another solution $\vec y$ if and only if:
 \begin{enumerate}
-\item $F_m(x)$ is not worse than $F_m(y)$ in all objectives for all $i = 1, 2, \dots, m$.
-\item At least one objective of $F_m(x)$ is better than its analogous counterpart in $F_m(y)$.
+\item $f_i(\vec x)$ is not worse than $f_i(\vec y)$, $\forall i = 1, 2, \dots, m$.
+\item For at least one objective $j$: $f_j(\vec x)$ is better than its analogous counterpart in $f_j(\vec y)$.
 \end{enumerate}
 
-When this two conditions apply, it is said that $F_m(x) \preceq F_m(y)$ ($F_m(x)$ dominates $F_m(y)$, and $F_m(x)$ is non-dominated by $F_m(y)$).
-The \textit{dominance} condition provides a partial ordering between points in the solution space: if $F_m(x) \preceq F_m(y)$, then $F_m(x)$ is considered to be better than $F_m(y)$.
-
-However, there are some cases where it cannot be said that $F_m(x) \preceq F_m(y)$ or $F_m(y) \preceq F_m(x)$.
-This situation occurs when one of the objectives is better in $F_m(x)$ but a different objective is better in $F_m(y)$ (for instance, when $f_1(x) < f_1(y)$ but $f_2(x) > f_2(y)$).
+When these two conditions apply, it is said that $\vec x \preceq \vec y$ ($\vec x$ dominates $\vec y$).
+The \textit{dominance} condition provides a partial ordering between solutions in the objective space.
+
+However, there are some cases where it cannot be said that $\vec x \preceq \vec y$ or $\vec y \preceq \vec x$. 
 In this case, it is said that these solutions are non-dominated with respect to each other.
-Solutions that are not dominated by each other are grouped in a \textit{non-dominated set}.
-Given a non-dominated set $P$, it is said that $P$ is the \textit{optimal Pareto front} if there is no other solution in the solution space that dominates any member of $P$.
-The relation between decision and objective space, dominance and Pareto fronts is depicted in Figure~\ref{fig:moop}.
+Solutions that are not dominated can be grouped in a \textit{non-dominated set}.
+Given a non-dominated set $P$, it is said that $P$ is the \textit{Pareto-set} if there is no other solution in the decision space that dominates any member of $P$.
+The corresponding objective vectors of the members in the Pareto-set build a so called \textit{Pareto-front}.
+The relations between decision and objective space, dominance and the non-dominated set are depicted in Figure~\ref{fig:moop}.
 
@@
 	\includegraphics[scale=0.235]{img/moopDec.png}
-	\caption{Decision and solution spaces in a MOOP with two variables ($x_1$ and $x_2$) and two objectives ($f_1$ and $f_2$).
-In the objective space, yellow dots are non-optimal objective vectors, while blue dots form the Pareto-optimal front.}
+	\caption{Decision and objective spaces in a MOOP with two variables ($x_1$ and $x_2$) and two objectives ($f_1$ and $f_2$).
+In the objective space, yellow dots are non-optimal objective vectors, while blue dots form a non-dominated set.}
 	\label{fig:moop}
@@
 
-As many Pareto fronts can be formed by several points in the solution space, it is important to devise a mechanism to assess the quality of a given front.
-A possibility is to use the Hypervolume Indicator (HV): given a Pareto front $P$, $HV(P)$ is defined as the volume of the objective space dominated by $P$.
+There are many different algorithms in the litrature which are proposed to tackle multi-objective optimization problems~\cite{Deb2001}.
+Weighted-sum approach is one of the traditional methods in which the objectives are weighted according to user preferences.
+The sum of the weighted objectives builds one objective function which is to be optimized.
+The solution to this single-objective problem is one certain solution ideally on a Paret-front.
+By varying the weights, it is possible to converge to different optimal solutions.
+Such linear scalarization approaches fail to find optimal solutions for the problems with non-convex shape of Pareto-fronts~\cite{Deb2001}.
+
+A popular choice for multi-objective optimization problems are evolutionary multi-objective optimization (EMO) algorithms~\cite{Coello1999, Zhou2011}. 
+The goal of EMO algorithms is to find a set of optimal solutions with a gooad convergence to the Pareto-front and additionally with a good spread and diversity along the Pareto-front.
+One of the most well-known algorithms in the literature is the Non-dominated Sorting Evolutionary Algorithm II (NSGA-{II})\textbf{~\cite{Deb2002}}, pseudocode of which is shown in Algorithm~\ref{alg:nsga2}.
+As in any evolutionary algorithm, NSGA-{II} evolves a set of $N$ individuals in a population denoted as $Pop$ with the difference that they are ranked according to dominance criterion and a crowding distance measure which is used to maintain a good diversity of solutions.
+
+\begin{algorithm}[!t]
+\begin{algorithmic}[1]
+%\Function{NSGA-{II}}{}
+\State \textbf{Input:} MOP, N
+\State \textbf{Output:} Non-dominated Set $F_0$
+	\State $t = 0$
+	\State $Pop(t) = NewRandomPopulation$	
+	\State $Q(t) = breed (Pop(t))$ \% Generate offspring
+	\While{Termination criterion not met}
+		\State $U(t) = Pop(t) \cup Q(t)$
+		\State $F = \Call{fastNonDominatedSort}{U(t)}$
+	    \State $Pop(t+1) = \emptyset$, $i = 0$		
+		\While{$|Pop(t+1)|+ |F_i| \leq N$}
+			\State $\Call{crowdingDistanceAssignment}{F_i}$	
+			\State $Pop(t+1) = Pop(t+1) \cup F_i$
+			\State $i = i+1$
+		\EndWhile
+		\State $\Call{sort}{Pop(t+1)}$
+		\State $Pop(t+1) = Pop(t+1) \cup F_i[1:(N-|Pop(t+1)|)]$
+		\State $Q(t+1) = breed(Pop(t+1))$	
+		\State $t = t+1$
+	\EndWhile
+	\Return $F_0$
+%\EndFunction
+\end{algorithmic}
+\caption{NSGA-{II} Algorithm \cite{Deb2002}}
+\label{alg:nsga2}
+\end{algorithm}
+
+The three main pillars of the NSGA-2 algorithm are:
+\begin{itemize}
+\item Fast non-dominated sorting: The function $F = \Call{fastNonDominatedSort}{U(t)}$ performs a non-dominated sorting and ranks the individuals stored in a set $U$ into several non-dominated fronts denoted as $F_i$, where the solutions in $F_0$ are the non-dominated solutions in the entire set.
+$F_i$ is the set of non-dominated solutions in $R \setminus (F_0 \cup \cdots \cup F_{i-1})$.
+\item Crowding distance: The function $\Call{crowdingDistanceAssignment}{F_i}$ assigns to each one of the individuals in $F_i$ a crowding distance value, which is the distance between the individual and its neighbours.
+The individuals with the smallest crowding distances are selected with a lower probability than the ones with the large values.
+\item Elitism: The solutions in the first front $F_0$ always survive to the next population. 
+\end{itemize}
+
+%%%%%%%%%%%%%
+% DO WE NEED THIS AT ALL??? 
+%A more recent approach called Multi-Objective Evolutionary Algorithm based on Decomposition (MOEA/D)~\cite{Zhang2007} is a combination of traditional linear scalirization approach and EMO.
+This approach is decomposes the MOP into several single-objective optimization sub-problems and an evolutionary algorithm optimizes them all simultaneously.
+Information is shared between neighbouring sub-problems in order to guide evolution.
+%The authors show that MOEA/D performs similarly than, and sometimes even outperforms, NSGA-II in the scenarios tested.
+%%%%%%%%%%%%%
+For more details about EMO approaches, please refer to \cite{Deb2001}.
+
+\subsection{Multi-objective Reinforcement Learning (MORL)}
+
+Reinforcement Learning (RL) algorithms have also used to deal with MOPs.
+RL~\cite{Sutton1998} is a broad field in Machine Learning that studies real-time planning and control situations where an agent has to find out the actions (or sequences of actions) that should be used in order to maximize the reward from the environment.
+
+The dynamics of an RL problem are usually captured by a Markov Decision Process (MDP) which is a tuple $(S, A, T, R)$, where $S$ is the set of possible states in the problem (or game), and $s_0$ is the initial state.
+$A$ is the set of available actions the agent can make at any given time, and the transition model $T(s_i, a_i, s_{i+1})$ determines the probability of reaching the state $s_{i+1}$ when action $a_i$ is applied in state $s_i$.
+The reward function $R(s_i)$ provides a single value (\textit{reward}) that the agent must optimize, representing the desirability of the state $s_i$ reached.
+Finally, a policy $\pi(s_i) = a_i$ determines which actions $a_i$ must be chosen from each state $s_i \in S$\footnote{This is a deterministic policy and only valid during acting, not learning}.
+One of the most important challenges in RL, as shown in Section~\ref{sec:mcts}, is the trade-off between exploration and exploitation while trying to act.
+While learning, a policy must choose between selecting the actions that provided good rewards in the past and exploring new parts of the search space by selecting new actions.
+
+Multi-objective Reinforcement Learning (MORL)~\cite{Vamplew2010} changes this definition by using a vector $R = {r_1, \dots, r_m}$ as rewards of the problem (instead of a scalar).
+Thus, MORL problems differ from RL in having more than one objective (here $m$) that must be maximized.
+If the objectives are independent or non-conflicting, scalarization approaches such as weighted sum approach, could be suitable to tackle the problem.
+Essentially, this would mean to use a conventional RL algorithm on a single objective where the global reward is obtained from a weighted-sum of the multiple rewards.
+However, this is not always the case, as usually the objectives are conflicting and the policy $\pi$ must balance among them.
+
+Vamplew et al.~\cite{Vamplew2010} propose a \textit{single-policy} algorithm which uses a preference order in the objectives (either given by the user or by the nature of the problem). 
+An example of this type of algorithm can be found at~\cite{Gabor1998}, where the authors introduce an order of preferences in the objectives and constrain the value of the desired rewards.
+Scalarization approaches would also fit in this category, as the work performed by S.
+Natarajan et al.~\cite{Natarajan2005}.
+
+The second type of algorithms, \textit{multiple-policy}, target to approximate the optimal Pareto-front of the problem.
+The algorithm proposed in this paper would fit in this category.
+An example of this type of algorithm is the one given by L.
+Barrett~\cite{Barrett2008}, who proposes the Convex Hull Iteration Algorithm.
+This algorithm provides the optimal policy for any linear preference function, by learning all policies that define the convex hull of the Pareto-front.
+
+\subsection{Metrics}
+The quality of an obtained non-dominated set can be measured using different diversity or/and convergence metrics~\cite{Deb2001}.
+The Hypervolume Indicator (HV) is a popular metric for measuring both the diversity and convergence of non-dominated solutions~\cite{Zitzler-Smetric}.
+This approach can be additionally used to compare different non-dominated sets.
+Given a Pareto front $P$, $HV(P)$ is defined as the volume of the objective space dominated by $P$.
 More formally, $HV(P) = \mu (x \in \mathbb{R}^{d} : \exists r \in P \hspace{0.25cm} s.t \hspace{0.25cm} r \preceq x)$, where $\mu$ is the de Lebesgue measure on $\mathbb{R}^{d}$.
 If the objectives are to be maximized, the higher the $HV(P)$, the better the front calculated.
-Figure~\ref{fig:hv} shows an example of $HV(P)$ where the objective dimension space is $2$.
+Figure~\ref{fig:hv} shows an example of $HV(P)$ where the objective dimension space is $m=2$.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%
+% I DON'T GET HOW DO WE DO THIS:  $HV(P) = \mu (x \in \mathbb{R}^{d} : \exists r \in P \hspace{0.25cm} s.t \hspace{0.25cm} r \preceq x)$ ????????????????
+%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 \end{center}
-\caption{$HV(P)$ of a given pareto front $P$.}
+\caption{$HV(P)$ of a given Pareto-front $P$}
 \label{fig:hv}
@@
 
-For more extensive descriptions, definitions, properties and multi-objective optimization in general, the reader can consult the work by K. Deb~\cite{Deb2001}.
-
-Many different algorithms have been proposed to tackle multi-objective optimization problems.
-One of the most popular methods used is the weighted-sum approach.
-The procedure consists of giving a weight to each one of the objectives and produce a single result as the linear combination of objectives and weights.
-By varying the weights provided, it is possible to converge to different solutions of the optimal Pareto front, if this is convex.
-However, K.
-Deb~\cite{Deb2001} explains how linear scalarization approaches fail in those scenarios where the optimal Pareto front is non-convex.
-
-A popular choice for multi-objective optimization problems are evolutionary multi-objective optimization (EMOA) algorithms~\cite{Coello1999, Zhou2011}.
-One of the most well known algorithms in the literature is the Non-dominated Sorting Evolutionary Algorithm II (NSGA-II), pseudocode of which is shown in Algorithm~\ref{alg:nsga2}.
-As in any evolutionary algorithm, NSGA-2 evolves a set of individuals or solutions to the problem, with the difference that here they are ranked according to dominance criteria and crowding distance (distances between members of the pareto fronts).
-A full description of the algorithm can be found in~\cite{Deb2000}.
-
-\begin{algorithm}[!t]
-\begin{algorithmic}[1]
-\Function{NSGA-2}{}
-	\State $P = NewRandomPopulation$
-	\While{Termination criteria not met}
-		\State $R = P \cup Q$ 
-		\State $F = \Call{fastNonDominatedSort}{R}$
-		\While{$|P| < N$}
-			\State $\Call{crowdingDistanceAssignment}{F_i}$	
-			\State $P = P \cup F_i$
-		\EndWhile
-		\State $\Call{sort}{P}$
-		\State $P = P[0:N]$
-		\State $Q = breed(P)$	
-	\EndWhile
-\EndFunction
-\end{algorithmic}
-\caption{NSGA-2 Algorithm.}
-\label{alg:nsga2}
-\end{algorithm}
-
-The three main pillars of the NSGA-2 algorithm are:
-
-\begin{itemize}
-\item A \textit{fast non-dominated sorting} algorithm, that ranks the individuals of the population and groups them in Pareto fronts.
-\item A \textit{crowding distance}, assigned to each one of the individuals, that measures how close it is to its neighbours.
-The selection genetic operator chooses individuals based on the ranks of the individuals and their crowding distance.
-\item \textit{Elitism}, implemented so the algorithm automatically promotes the best $N$ individuals to the next generation.
-\end{itemize}
-
-A more recent approach, developed by Q.
-Zhang, is the Multi-Objective Evolutionary Algorithm based on Decomposition (MOEA/D)~\cite{Zhang2007}, that decomposes the problem into several single optimization sub-problems and an evolutionary algorithm optimizes them all simultaneously.
-Information is shared between neighbouring sub-problems in order to guide evolution.
-The authors show that MOEA/D performs similarly than, and sometimes even outperforms, NSGA-II in the scenarios tested.
-
-Reinforcement Learning (RL) algorithms have also tackled Multi-objective optimization in some settings.
-RL~\cite{Sutton1998} is a broad field in Machine Learning that studies real-time planning and control situations where  an agent has to find out the actions (or sequences of actions) that should be applied in order to maximize the reward from the environment.
-
-The dynamics of an RL problem are usually captured by a Markov Decision Process (MDP).
-The (finite, undiscounted) is a tuple $(S, A, T, R)$, where $S$ is the set of possible states in the problem (or game), and $s_0$ is the initial state.
-$A$ is the set of available actions the agent can make at any given time, and the transition model $T(s_i, a_i, s_{i+1})$ determines the probability of reaching the state $s_{i+1}$ when action $a_i$ is applied in state $s_i$.
-The reward function $R(s_i)$ provides a single value (\textit{reward}) that the agent must optimize, representing the desirability of the state $s_i$ reached.
-Finally, a policy $\pi(s_i) = a_i$ determines which actions $a_i$ must be chosen from each state $s_i \in S$\footnote{This is a deterministic policy and only valid in during acting, not leaning}.
-One of the most important challenges in RL, as shown in Section~\ref{sec:mcts}, is the trade-off between exploration and exploitation while trying to act.
-While learning, a  policy must choose between following actions that provided good rewards in the past and exploring new parts of the search space by selecting new actions.
-
-Multi-objective Reinforcement Learning (MORL)~\cite{Vamplew2010} changes this definition by using a vector $R = {r_0, r_1, \dots, r_n}$ as rewards of the problem (instead of a scalar).
-Thus, MORL problems differ from RL in having more than one objective ($n$) that must be maximized.
-If the objectives are independent or they do not oppose each other, scalarization technique approaches, as described above, could be suitable to tackle the problem.
-Essentially, this would mean to use a conventional RL algorithm on a single objective where the global reward is obtained from a weighted-sum of the multiple rewards.
-However, this is not always the case, as it is usual that the objectives are in conflict and the policy $\pi$ must balance among them.
-
-According to how $\pi$ approaches this problem, Vamplew et al.~\cite{Vamplew2010} proposed the following distinction for MORL: \textit{single-policy} algorithms are those that provide a preference order in the objectives available (given by the user or by the nature of the problem).
-An example of this type of algorithm can be found at~\cite{Gabor1998}, where the authors introduce an order of preference in the objectives treated and constraint the value of the rewards desired.
-Scalarization approaches would also fit in this category, as the work performed by S.
-Natarajan et al.~\cite{Natarajan2005}.
-
-The second type of algorithms, \textit{multiple-policy}, target to approximate the optimal Pareto front of the problem.
-The algorithm proposed in this paper would fit in this category.
-An example of this type of algorithm is the one given by L.
-Barrett~\cite{Barrett2008}, who propose the Convex Hull Iteration Algorithm.
-This algorithm provides the optimal policy for any linear preference function, by learning all policies that define the convex hull of the Pareto front.
-
 \section{Multi-Objective Monte Carlo Tree Search} \label{sec:momcts}
@@
 Adapting MCTS into Multi-Objective Monte Carlo Tree Search (MO-MCTS) requires the obvious modification of dealing with multiple rewards instead of just one.
-As these are collected at the end of a Monte Carlo simulation, the reward value $r$ now becomes a vector $R = {r_0, r_1, \dots, r_n}$, where $n$ is the number of objectives to optimize.
+As these are collected at the end of a Monte Carlo simulation, the reward value $r$ now becomes a vector $R = {r_0, r_1, \dots, r_m}$, where $m$ is the number of objectives to optimize.
 Derived from this change, the average value $Q(s,a)$ becomes a vector that stores the average reward of each objective.
@@
 
-The next step is to define the value for each pair of state and action, $W(s,a)$, as in Equation~\ref{eq:wang}.
+The next step is to define the value for each pair of state and action, $V(s,a)$, as in Equation~\ref{eq:wang}.
 $\overline{r}^p_{sa}$ is the projection of $\overline{r}_{sa}$ into the piecewise linear surface defined by the Pareto archive $P$ (see Figure~\ref{fig:wangHV}).
@@
 If $\overline{r}_{sa}$ is dominated by P, the distance between $\overline{r}_{sa}$ and $\overline{r}^p_{sa}$ is subtracted from the HV calculation.
-The tree policy selects actions based on a maximization of the value of $W(s,a)$.
+The tree policy selects actions based on a maximization of the value of $V(s,a)$.
 
 \begin{equation} \label{eq:wang}
-	W(s,a) = \left\{
+	V(s,a) = \left\{
 \begin{array}{l l}      
@@
 
-In order to select a point in the Pareto front, a weight vector $W$ can be defined, with a dimension $n$ equal to the number of objectives ($W = (w_1, w_2, \dots, w_n); \sum_{i}^{n} w_i = 1$).
+In order to select a point in the Pareto front, a weight vector $W$ can be defined, with a dimension $m$ equal to the number of objectives ($W = (w_1, w_2, \dots, w_m); \sum_{i}^{m} w_i = 1$).
 Two different mechanisms are proposed, for reasons that will be explained in the experiments section (\ref{sec:exp}):
@@
 
-Additionally, the agent can only make up to $100$ moves.
-This allows the problem to be defined as the maximization of two rewards: $(M,T) = (100 - moves, treasureValue)$.
+Additionally, the agent can only make up to $100$ plies or moves.
+This allows the problem to be defined as the maximization of two rewards: $(\rho_p,\rho_v) = (100 - plies, treasureValue)$.
 Should the ship perform all moves without reaching a treasure, the result would be ($0$,$0$).
@@
 
-As DST has two different objectives, number of moves and value of the treasure, the quality of a state can be assessed by two rewards, $r_m$ and $r_v$, for each objective respectively.
-$r_m$ is adjusted to be maximized using the maximum number of moves in the game, while $r_v$ is just the value in the cell with the treasure.
+As DST has two different objectives, number of moves and value of the treasure, the quality of a state can be assessed by two rewards, $\rho_p$ and $\rho_v$, for each objective respectively.
+$\rho_p$ is adjusted to be maximized using the maximum number of moves in the game, while $\rho_v$ is just the value in the cell with the treasure.
 Equation~\ref{eq:dst} summarizes these rewards:
@@
 \begin{split}
-r_m = 100 - moves \\
-r_v = treasureValue
+\rho_p = 100 - moves \\
+\rho_v = treasureValue
 \end{split}
@@
 
-In order to evaluate a game state in MO-PTSP, three different measures or rewards are taken, $r_t, r_f, r_d$, one for each one of the objectives: time, fuel and damage, respectively.
+In order to evaluate a game state in MO-PTSP, three different measures or rewards are taken, $\rho_t, \rho_f, \rho_d$, one for each one of the objectives: time, fuel and damage, respectively.
 All these rewards are defined so they have to be maximized.
@@
 \begin{equation}	\label{eq:dist}
-r_t = 1 - d/D
+\rho_t = 1 - d_t/d_M
 \end{equation}
@@
 The MO-PTSP framework includes a path-finding library that allows controllers to query the shortest distance of a route of waypoints.
-$d$ indicates the distance from the current position until the last waypoint, following the desired route, and $D$ is the distance of the whole route from the starting position.
+$d_t$ indicates the distance from the current position until the last waypoint, following the desired route, and $d_M$ is the distance of the whole route from the starting position.
 Minimizing the distance to the last waypoint will lead to the end of the game.
 
-Equation~\ref{eq:fuel} shows how the value of the fuel objective, $r_f$, is obtained:
+Equation~\ref{eq:fuel} shows how the value of the fuel objective, $\rho_f$, is obtained:
 
@@
 \begin{equation}	\label{eq:fuel}
-r_f = (1 - (f_c/f_0)) \times \alpha + r_t \times (1 - \alpha)
+\rho_f = (1 - (\lambda_t/\lambda_0)) \times \alpha + \rho_t \times (1 - \alpha)
 \end{equation}
 
-$f_c$ is the fuel consumed so far, and $f_0$ is the initial fuel at the start of the game.
+$\lambda_t$ is the fuel consumed so far, and $\lambda_0$ is the initial fuel at the start of the game.
 $\alpha$ is a value that balances between the fuel component and the time objective from Equation~\ref{eq:dist}.
@@
 
-Finally, Equation~\ref{eq:dam} gives the method to calculate the damage objective, $r_d$:
+Finally, Equation~\ref{eq:dam} gives the method to calculate the damage objective, $\rho_d$:
 
@@
 \begin{equation}	\label{eq:dam}
-r_d = 
+\rho_d = 
 \begin{cases}
-	(1 - (d_s/d_m)) \times \beta_{1} + r_t \times (1 - \beta_{1}), & s > \gamma\\
-	(1 - (d_s/d_m)) \times \beta_{2} + r_t \times (1 - \beta_{2}), & s \leq \gamma
+	(1 - (g_t/g_M)) \times \beta_{1} + \rho_t \times (1 - \beta_{1}), & sp > \gamma\\
+	(1 - (g_t/g_M)) \times \beta_{2} + \rho_t \times (1 - \beta_{2}), & sp \leq \gamma
 \end{cases}
@@
 
-$d_s$ is the damage suffered so far in the game, and $d_m$ is the maximum damage the ship can take.
+$g_t$ is the damage suffered so far in the game, and $g_M$ is the maximum damage the ship can take.
 In this case, three different variables are used to regulate the behaviour of this objective: $\gamma$, $\beta_{1}$ and $\beta_{2}$.
@@
 The difference is that $\beta_{1}$ is used in high speeds, while $\beta_{2}$ is employed with low velocities.
-This is distinguished by the parameter $\gamma$, that can be seen as a threshold value for the ship's speed ($s$).
+This is distinguished by the parameter $\gamma$, that can be seen as a threshold value for the ship's speed ($sp$).
 This differentiation is made in order to avoid low speeds in lava lakes, as this increase the damage suffered importantly.
@@
 
-It is important to highlight again that these three rewards, $r_t$, $r_f$ and $r_d$, are the same for the three algorithms tested in the experiments.
+It is important to highlight again that these three rewards, $\rho_t$, $\rho_f$ and $\rho_d$, are the same for the three algorithms tested in the experiments.
 
@@
 The weight vector for DST has two dimensions.
-This vector is here referred to as $W = (w_m, w_v)$, where $w_m$ weights moves and $w_v$ the treasure value, and $w_v = 1 - w_m$.
-$w_m$ takes values between $0$ and $1$, with a precision of $0.01$, and $100$ experiments have been run for each pair $(w_m, w_v)$.
+This vector is here referred to as $W = (w_p, w_v)$, where $w_p$ weights moves and $w_v$ the treasure value, and $w_v = 1 - w_p$.
+$w_p$ takes values between $0$ and $1$, with a precision of $0.01$, and $100$ experiments have been run for each pair $(w_p, w_v)$.
 Hence, the game has been played a total of $10000$ times, for each algorithm.

### 2014-02-20 paper/paper.tex
last Sanaz changes

@@
 The goal of EMO algorithms is to find a set of optimal solutions with a gooad convergence to the Pareto-front and additionally with a good spread and diversity along the Pareto-front.
-One of the most well-known algorithms in the literature is the Non-dominated Sorting Evolutionary Algorithm II (NSGA-{II})\textbf{~\cite{Deb2002}}, pseudocode of which is shown in Algorithm~\ref{alg:nsga2}.
+One of the most well-known algorithms in the literature is the Non-dominated Sorting Evolutionary Algorithm II (NSGA-{II})~\cite{Deb2002}, pseudocode of which is shown in Algorithm~\ref{alg:nsga2}.
 As in any evolutionary algorithm, NSGA-{II} evolves a set of $N$ individuals in a population denoted as $Pop$ with the difference that they are ranked according to dominance criterion and a crowding distance measure which is used to maintain a good diversity of solutions.
@@
 The Deep Sea Treasure (DST) is a popular multi-objective problem introduced by Vamplew et al.~\cite{Vamplew2010}.
-In this single-player puzzle, the agent moves a submarine with the objective of finding a treasure located at the bottom of the ship.
+In this single-player puzzle, the agent moves a submarine with the objective of finding a treasure located at the bottom of the sea.
 The world is divided into a grid of $10$ rows and $11$ columns, and the vessel starts at the top left board position.
@@
 {Diego Perez}
-received a B.Sc.
-and a M.Sc.
-in Computer Science from University Carlos III, Madrid, in 2007.
-He is currently pursuing a Ph.D.
+is currently pursuing a Ph.D.
 in Artificial Intelligence applied to games at the University of Essex, Colchester.
-He has published in the domain of Game AI, participated in several Game AI competitions and organized the Physical Travelling Salesman Problem competition, held in IEEE conferences during 2012.
+He has published in the domain of Game AI, participated in several Game AI competitions and organized the Physical Travelling Salesman Problem competitions, held in IEEE conferences.
 He also has programming experience in the videogames industry with titles published for game consoles and PC.
@@
 {Sanaz Mostaghim}
-BIO HERE.
+is a professor of computer science at the Otto von Guericke University Magdeburg, Germany.
+Her current research interests are in the field of evolutionary computation, particularly in the areas of evolutionary multi-objective optimization, swarm intelligence and their applications in science and industry.
+She is an active IEEE member and is serving as an associate editor for IEEE Transactions on Evolutionary Computation and IEEE Transactions on Cybernetics.
 \end{IEEEbiography}
 
-\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{./photos/spyros.jpg}}]{Spyridon Samothrakis}
+\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{./photos/spyro.jpg}}]{Spyridon Samothrakis}
  is currently pursuing a PhD in Computational Intelligence and Games at the University of Essex.

### 2014-02-26 paper/paper.tex
edits after Michelle's proof read

@@
 Multi-objective optimization has been traditionally a matter of study in domains like engineering or finance, with little impact on games research.
-However, action-decision based on a multi-objective evaluation may be beneficial in order to obtain a high quality level of play.
+However, action-decision based on multi-objective evaluation may be beneficial in order to obtain a high quality level of play.
 This paper presents a Multi-objective Monte Carlo Tree Search algorithm for planning and control in real-time game domains, those where the time budget to decide the next move to make is close to 40ms.
 A comparison is made between the proposed algorithm, a single-objective version of Monte Carlo Tree Search and a rolling horizon implementation of Non-dominated Sorting Evolutionary Algorithm II (NSGA-II).
-Two different benchmarks are employed, Deep Sea Treasure and Multi-Objective Physical Travelling Salesman Problem.
+Two different benchmarks are employed, Deep Sea Treasure and the Multi-Objective Physical Travelling Salesman Problem.
 Using the same heuristics on each game, the analysis is focused on how well the algorithms explore the search space.
-Results show that the algorithm proposed outperforms (NSGA-II).
+Results show that the algorithm proposed outperforms NSGA-II.
 Additionally, it is also shown that the algorithm is able to converge to different optimal solutions or the optimal Pareto front (if achieved during search).
@@
 
-Multi-objective optimization has been a field of study of manufacturing, engineering~\cite{Marler2004} and finance~\cite{Coello2006}, while having little impact in games research.
+Multi-objective optimization has been a field of study in manufacturing, engineering~\cite{Marler2004} and finance~\cite{Coello2006}, while having little impact on games research.
 This paper shows that multi-objective optimization has much to offer in developing game strategies that allow for a fine-grained control of alternative policies.
-The application of such approaches to this field can provide interesting results, especially in games that are long or complex enough where long-term planning is not trivial, and achieving a good level of play requires balancing strategies.
-
-From a simple point of view, many competitive games can be seen as a scenario where two or more opponents have the single objective of winning.
-Achieving victory is usually complex in interesting games, and successful approaches normally assign some form of value in a state, value being long term expected reward (i.e.
+The application of such approaches to this field can provide interesting results, especially in games that are long or complex enough that long-term planning is not trivial, and achieving a good level of play requires balancing strategies.
+
+At their most basic, many competitive games can be viewed as simply two or more opponents having the single objective of winning.
+Achieving victory is usually complex in interesting games, and successful approaches normally assign some form of value to a state, value being long term expected reward (i.e.
 expectation of victory).
@@
 
-This paper proposes a multi-objective real-time algorithm version of Monte Carlo Tree Search (MCTS), a popular reinforcement learning approach that emerged the last decade.
+This paper proposes a multi-objective real-time algorithm version of Monte Carlo Tree Search (MCTS), a popular reinforcement learning approach that emerged in the last decade.
 The proposed algorithm is tested in two different games: a real-time version of the Deep Sea Treasure (DST; a classical multi-objective problem), and the Multi-Objective Physical Travelling Salesman Problem (MO-PTSP).
@@
 This game is played in a square grid board, with a size of $19 \times 19$ in the original game, and $9 \times 9$ in its reduced version.
-The game is played in turns, and the objective is to surround the opponent's stones by placing stones in any available position in the board.
+The game is played in turns, and the objective is to surround the opponent's stones by placing stones in any available position on the board.
 Due to the very large branching factor of Go, this game is considered the drosophila of Game AI, and MCTS players have reached professional level play in the reduced board size version~\cite{Lee2009}.
@@
 
-MCTS is considered to be an \textit{anytime} algorithm, as it is able to provide a a valid next move to choose at any moment in time.
-This is true independently from how many iterations the algorithm was able to make (although, in general, more iterations usually produce better results).
-This differs from other algorithms (such A* in single player games, and standard Min-Max for two player games) that normally provide the next play only after they have finished.
+MCTS is considered to be an \textit{anytime} algorithm, as it is able to provide a valid next move to choose at any moment in time.
+This is true independently from how many iterations the algorithm is able to make (although, in general, more iterations usually produce better results).
+This differs from other algorithms (such as A* in single player games, and standard Min-Max for two player games) that normally provide the next play only after they have finished.
 This makes MCTS a suitable candidate for real-time domains, where the decision time budget is limited, affecting the number of iterations that can be performed.
@@
 Each node in the tree maintains statistics that indicate how often a move is played from a given state ($N(s,a)$), how many times each move is played from there ($N(s)$) and the average reward ($Q(s,a)$) obtained after applying move $a$ in state $s$.
-The tree is built iteratively by simulating actions in the game, making move choices based on the statistics store in the nodes.
+The tree is built iteratively by simulating actions in the game, making move choices based on statistics stored in the nodes.
 
@@
 
-In every one of this action decisions, MCTS balances between exploitation and exploration.
+In every one of these action decisions, MCTS balances between exploitation and exploration.
 In other words, it chooses between taking an action that leads to states with the best outcome found so far, and performing a move to go to less explored game states, respectively.
@@
 
-The balance between exploration and exploitation can be tampered by modifying $C$.
-Higher values of $C$ weight more the second term of the UCB1 Equation~\ref{eq:ucb1}, giving preference to those actions that have been explored less, at the expense of taking actions with the highest average reward $Q(s,a)$.
+The balance between exploration and exploitation can be tempered by modifying $C$.
+Higher values of $C$ give added weight to the second term of the UCB1 Equation~\ref{eq:ucb1}, giving preference to those actions that have been explored less, at the expense of taking actions with the highest average reward $Q(s,a)$.
 A commonly used value is $\sqrt{2}$, as it balances both facets of the search when the rewards are normalized between $0$ and $1$.
@@
 An important feature of this game is that it is \textit{open-ended}, as an end game situation is, most of the time, far ahead in the future and can not be devised by the algorithm during its iterations.
-The consequence of this is that MCTS, in its vanilla form, it is not able to know if a given ply will lead to a win or a loss game end state.
+The consequence of this is that MCTS, in its vanilla form, it is not able to know if a given ply will lead to a win or a loss in the end game state.
 Robles et al.~\cite{Robles2009} solved this problem by including hand-coded heuristics that guided search towards more promising portions of the search space.
-This approach has been followed in order to add heuristics knowledge to MCTS, as in~\cite{Samothrakis, Ikehata2010}.
+This approach enabled the addition of heuristics knowledge to MCTS, as in~\cite{Samothrakis, Ikehata2010}.
 
@@
 Another use of MCTS is in the popular puzzle Morpion Solitaire~\cite{Edelkamp2010}, a connection game where the goal is to link nodes of a graph with straight lines that must contain at least five vertices.
-Finally, the PTSP has also been addressed with MCTS, both in the single-objective~\cite{Perez2013, Powley2012} and the multi-objective versions~\cite{Powley2013}.
+Finally, the PTSP has also been addressed by MCTS, both in the single-objective~\cite{Perez2013, Powley2012} and the multi-objective versions~\cite{Powley2013}.
 These papers describe the entries that won both editions of the PTSP Competition.
@@
 There are many different algorithms in the litrature which are proposed to tackle multi-objective optimization problems~\cite{Deb2001}.
-Weighted-sum approach is one of the traditional methods in which the objectives are weighted according to user preferences.
+Weighted-sum approach is one of the traditional methods in which the objectives are weighted according to user preference.
 The sum of the weighted objectives builds one objective function which is to be optimized.
-The solution to this single-objective problem is one certain solution ideally on a Paret-front.
+The solution to this single-objective problem is one certain solution, ideally on a Pareto-front.
 By varying the weights, it is possible to converge to different optimal solutions.
-Such linear scalarization approaches fail to find optimal solutions for the problems with non-convex shape of Pareto-fronts~\cite{Deb2001}.
+Such linear scalarization approaches fail to find optimal solutions for problems with non-convex-shape Pareto-fronts~\cite{Deb2001}.
 
 A popular choice for multi-objective optimization problems are evolutionary multi-objective optimization (EMO) algorithms~\cite{Coello1999, Zhou2011}. 
-The goal of EMO algorithms is to find a set of optimal solutions with a gooad convergence to the Pareto-front and additionally with a good spread and diversity along the Pareto-front.
+The goal of EMO algorithms is to find a set of optimal solutions with a good convergence to the Pareto-front as well as a good spread and diversity along the Pareto-front.
 One of the most well-known algorithms in the literature is the Non-dominated Sorting Evolutionary Algorithm II (NSGA-{II})~\cite{Deb2002}, pseudocode of which is shown in Algorithm~\ref{alg:nsga2}.
@@
 \item Crowding distance: The function $\Call{crowdingDistanceAssignment}{F_i}$ assigns to each one of the individuals in $F_i$ a crowding distance value, which is the distance between the individual and its neighbours.
-The individuals with the smallest crowding distances are selected with a lower probability than the ones with the large values.
+The individuals with the smallest crowding distances are selected with a lower probability than the ones with the larger values.
 \item Elitism: The solutions in the first front $F_0$ always survive to the next population. 
@@
 
-Reinforcement Learning (RL) algorithms have also used to deal with MOPs.
-RL~\cite{Sutton1998} is a broad field in Machine Learning that studies real-time planning and control situations where an agent has to find out the actions (or sequences of actions) that should be used in order to maximize the reward from the environment.
+Reinforcement Learning (RL) algorithms have also been used with MOPs.
+RL~\cite{Sutton1998} is a broad field in Machine Learning that studies real-time planning and control situations where an agent has to find the actions (or sequences of actions) that should be used in order to maximize the reward from the environment.
 
@@
 If the objectives are independent or non-conflicting, scalarization approaches such as weighted sum approach, could be suitable to tackle the problem.
-Essentially, this would mean to use a conventional RL algorithm on a single objective where the global reward is obtained from a weighted-sum of the multiple rewards.
+Essentially, this would mean using a conventional RL algorithm on a single objective where the global reward is obtained from a weighted-sum of the multiple rewards.
 However, this is not always the case, as usually the objectives are conflicting and the policy $\pi$ must balance among them.
@@
 Vamplew et al.~\cite{Vamplew2010} propose a \textit{single-policy} algorithm which uses a preference order in the objectives (either given by the user or by the nature of the problem). 
-An example of this type of algorithm can be found at~\cite{Gabor1998}, where the authors introduce an order of preferences in the objectives and constrain the value of the desired rewards.
-Scalarization approaches would also fit in this category, as the work performed by S.
+An example of this type of algorithm can be found at~\cite{Gabor1998}, where the authors introduce an order of preference in the objectives and constrain the value of the desired rewards.
+Scalarization approaches would also fit in this category, as shown in the work performed by S.
 Natarajan et al.~\cite{Natarajan2005}.
 
-The second type of algorithms, \textit{multiple-policy}, target to approximate the optimal Pareto-front of the problem.
-The algorithm proposed in this paper would fit in this category.
+The second type of algorithms, \textit{multiple-policy}, aims to approximate the optimal Pareto-front of the problem.
+This is the aim of the algorithm proposed in this paper.
 An example of this type of algorithm is the one given by L.
@@
 Derived from this change, the average value $Q(s,a)$ becomes a vector that stores the average reward of each objective.
-Note that the other statistics ($N(s,a)$ and $N(s)$) do not need to change, as this are just node and action counters.
+Note that the other statistics ($N(s,a)$ and $N(s)$) do not need to change, as these are just node and action counters.
 The important question to answer next is how to adapt the vector $Q(s,a)$ to use it in the UCB1 formula (Equation~\ref{eq:ucb1}).
 
-An initial attempt on Multi-Objective MCTS was addressed by W.
+An initial attempt at Multi-Objective MCTS was addressed by W.
 Wang and Michele Sebag~\cite{Weijia2012, Weijia2013}.
 In their work, the authors employ a mechanism, based on the HV calculation, to replace the UCB1 equation.
-The algorithm keeps a Pareto archive (P) with the best solutions found in game end states.
+The algorithm keeps a Pareto archive (P) with the best solutions found in end game states.
 Every node in the tree defines $\overline{r}_{sa}$ as a vector of UCB1 values, in which each $\overline{r}_{sa, i}$ is the result of calculating UCB1 for each particular objective $i$.
 
-The next step is to define the value for each pair of state and action, $V(s,a)$, as in Equation~\ref{eq:wang}.
+The next step is to define the value for each state and action pair, $V(s,a)$, as in Equation~\ref{eq:wang}.
 $\overline{r}^p_{sa}$ is the projection of $\overline{r}_{sa}$ into the piecewise linear surface defined by the Pareto archive $P$ (see Figure~\ref{fig:wangHV}).
@@
 As mentioned before, the objective of this paper is to propose an MO-MCTS algorithm that is suitable for real-time domains.
-Although the work discussed here is influenced by Wang's approach, some modifications need to be done in order to overcome the high computational cost involved by their approach.
-
-In the algorithm proposed in this paper, the vector $\overline{r}$ of rewards that was obtained at the end of an Monte Carlo simulation is back-propagated through the nodes visited in the last iteration until the root is reached.
+Although the work discussed here is influenced by Wang's approach, some modifications need to be done in order to overcome the high computational cost involved in their approach.
+
+In the algorithm proposed in this paper, the reward vector $\overline{r}$ that was obtained at the end of a Monte Carlo simulation is back-propagated through the nodes visited in the last iteration until the root is reached.
 In the vanilla algorithm, each node would use this vector $\overline{r}$ to update its own accumulated reward vector $\overline{R}$.
@@
 This last detail is important for two main reasons.
-First of all, the algorithm allows the root to store information indicating which is the action to take in order to converge to any specific solution in the front discovered.
+First of all, the algorithm allows the root to store information as to which action to take in order to converge to any specific solution in the front discovered.
 This information can be used, when all iterations have been performed, to select the move to perform next.
-If weights for the different objectives are provided, this weights can be used to select the desired solution in the Pareto front of the root node, and hence select the action that leads to that point.
+If weights for the different objectives are provided, these weights can be used to select the desired solution in the Pareto front of the root node, and hence select the action that leads to that point.
 Secondly, the root's Pareto front can be used to measure the global quality of the search using the hypervolume calculation.
@@
 
-This algorithm, same as in NSGA-II due to their multi-objective nature, provide a non-dominated front as a solution.
+This algorithm, similar to NSGA-II due to their multi-objective nature, provides a non-dominated front as a solution.
 However, in planning and control scenarios like the games analyzed in this research, an action must be provided to perform a move in the next step.
-The question that arises is how to choose the move to make with the information available.
-
-As shown before, it is straightforward to obtain which actions lead to what points in the front given as a solution: first gene in the best NSGA-II individual, a root's child in MO-MCTS.
+The question that arises is how to choose the move to make based on the information available.
+
+As shown before, it is straightforward to discover which actions lead to what points in the front given as a solution: first gene in the best NSGA-II individual, a root's child in MO-MCTS.
 Hence, by identifying the point in the Pareto front that the algorithm should converge to,
@@
 \item Euclidean distance: the euclidean distance from each point in the Pareto front (normalized in $[0,1]$) to the vector $W$ is calculated.
-The action to choose would be the one that leads to the point in the Pareto front with the smaller distance to $W$.
+The action to choose would be the one that leads to the point in the Pareto front with the shortest distance to $W$.
 \end{itemize}
@@
 Note that, in the vanilla MCTS, there is no Pareto front obtained as a solution.
-Typically, in this case, rewards are calculated as a weighted sum of the objectives and a weight vector $W$, and then the action is chosen following any of the mechanisms usually employed in the literature: the action taken more often from the root, the one that leads to the best reward found, the move with the highest expected reward or the action that maximizes Equation~\ref{eq:ucb1} in the root.
+Typically, in this case, rewards are calculated as a weighted sum of the objectives and a weight vector $W$, and then the action is chosen following any of the mechanisms usually employed in the literature: the action taken more often from the root; the one that leads to the best reward found; the move with the highest expected reward; or the action that maximizes Equation~\ref{eq:ucb1} in the root.
 
@@
 The Deep Sea Treasure (DST) is a popular multi-objective problem introduced by Vamplew et al.~\cite{Vamplew2010}.
-In this single-player puzzle, the agent moves a submarine with the objective of finding a treasure located at the bottom of the sea.
+In this single-player puzzle, the agent controls a submarine with the objective of finding a treasure located at the bottom of the sea.
 The world is divided into a grid of $10$ rows and $11$ columns, and the vessel starts at the top left board position.
@@
 Section~\ref{sec:moo} introduced the problems of linear scalarization approaches when facing non-convex optimal Pareto fronts.
-The concave shape of the DST's optimal front makes those approximations to converge to the non dominated solutions located at the edges of the Pareto front: $(81,124)$ and $(99,1)$.
+The concave shape of the DST's optimal front means that those approximations converge to the non dominated solutions located at the edges of the Pareto front: $(81,124)$ and $(99,1)$.
 Note that this happens independently from the weights chosen for the linear approximation: some solutions of the front just can't be reached with these approaches.
@@
 \end{center}
-\caption{Example of two different sequences of actions (R: Right, D: Down) that lie in the same position in the map, but different node in the tree.}
+\caption{Example of two different sequences of actions (R: Right, D: Down) that lie in the same position in the map, but at a different node in the tree.}
 \label{fig:dstTable}
@@
 In a tree that does not use TT, there would be two different nodes to represent $P3$, although the location and number of moves performed up to this point are the same (thus, the states are equivalent).
-It is worthwhile to highlight that the coordinates of the vessel are not enough to identify two equivalent states, but also the number of moves is needed: imagine a third route from $P2$ to $P3$ with the moves: \textit{Up}, \textit{Right}, \textit{Right}, \textit{Down}, \textit{Down}.
+It is worthwhile highlighting that the coordinates of the vessel are not enough to identify two equivalent states, the number of moves is also needed: imagine a third route from $P2$ to $P3$ with the moves: \textit{Up}, \textit{Right}, \textit{Right}, \textit{Down}, \textit{Down}.
 As the submarine performs $5$ moves, the states are not the same, and the node where the ship is in $P3$ now would be two levels deeper in the tree.
 
-TT tables are implemented in the MCTS algorithms tested in this benchmark by using hash tables that stores a \textit{representative} node for each pair $(position, moves)$ found.
+TT tables are implemented in the MCTS algorithms tested in this benchmark by using hash tables that store a \textit{representative} node for each pair $(position, moves)$ found.
 The key of the hash map needs then to be obtained from three values: position coordinates $x$ and $y$ of the ship in the board, and number of moves, indicated by the depth of the node in the tree.
@@
 
-The Multi-Objective Physical Travelling Salesman Problem (MO-PTSP) is a game employed in a competition held in the IEEE Conference on Computational Intelligence in Games (CIG) in 2013, a modification of the Physical Travelling Salesman Problem (PTSP), previously introduced by Perez et al.~\cite{PerezCEC2012}.
+The Multi-Objective Physical Travelling Salesman Problem (MO-PTSP) is a game that was employed in a competition held at the IEEE Conference on Computational Intelligence in Games (CIG) in 2013.
+This was a modification of the Physical Travelling Salesman Problem (PTSP), previously introduced by Perez et al.~\cite{PerezCEC2012}.
 The MO-PTSP is a real-time game where the agent navigates a ship and must visit $10$ waypoints scattered around the maze.
-All waypoints must be visited to consider a game as complete, and a game-tick counter is reset every time a waypoint is collected, finishing the game prematurely (and unsuccessfully) if it reaches $800$ game steps before visiting another waypoint.
+All waypoints must be visited to consider a game as complete, and a game-tick counter is reset every time a waypoint is collected, finishing the game prematurely (and unsuccessfully) if $800$ game steps are reached before visiting another waypoint.
 
@@
 The ship starts with $5000$ units of fuel, and one unit is spent every time an action supplied has the throttle input \textit{on}.
-There are, however, two ways of collecting fuel: each waypoint visited grants the ship with $50$ more units of fuel.
-Also, fuel canisters are scattered around the maze and their collection provides with $250$ more units.
+There are, however, two ways of collecting fuel: each waypoint visited grants the ship $50$ more units of fuel; and there are also fuel canisters scattered around the maze that provide $250$ more units.
 
@@
 In the latter, lava lakes are abundant in MO-PTSP levels and, in contrast with normal surfaces, they deal one unit of damage for each time step the ship spends over this type of surface.
-All these subtractions are deduced from an initial counter of $5000$ points of damage.
+All these subtractions are deducted from an initial counter of $5000$ points of damage.
 
 Figure~\ref{fig:sampleMap} shows an example of an MO-PTSP map, as drawn by the framework.
-Waypoints yet not visited are painted as blue circles, while those already collected are depicted as empty circles.
+Waypoints not yet visited are painted as blue circles, while those already collected are depicted as empty circles.
 The green ellipses represent fuel canisters, normal surfaces are drawn in brown and lava lakes are printed as red-dotted yellow surfaces.
@@
 
-All algorithms tested in the MO-PTSP in this research employ macro-actions.
-Macro-actions is a concept that can be used for coarsening the action space by applying the action chosen by the control algorithm in several consecutive cycles, instead of just in the next one.
+All algorithms tested in the MO-PTSP in this research employ macro-actions, a concept that can be used for coarsening the action space by applying the action chosen by the control algorithm in several consecutive cycles, instead of just in the next one.
 
@@
 A macro-action of length $L$ is defined as a repetition of a given action during $L$ consecutive time steps.
-The main advantage of macro-actions is that the control algorithms can see further in the future, and then reduce the implication of open-endedness of real-time games (see Section~\ref{sec:mcts}).
+The main advantage of macro-actions is that the control algorithms can see further in the future, and then reduce the implication of open-endedness in real-time games (see Section~\ref{sec:mcts}).
 As this reduces the search space significantly, a better algorithm performance is allowed.
@@
 $\alpha$ is a value that balances between the fuel component and the time objective from Equation~\ref{eq:dist}.
-Note that, as waypoints needs to keep being visited, it is necessary to include a distance measure for this reward.
+Note that, as waypoints need to keep being visited, it is necessary to include a distance measure for this reward.
 Otherwise, an approach that prioritizes this objective would not minimize distance to waypoints at all (the ship could just stand still: no fuel consumed is optimal), and therefore could not complete the game.
@@
 This is distinguished by the parameter $\gamma$, that can be seen as a threshold value for the ship's speed ($sp$).
-This differentiation is made in order to avoid low speeds in lava lakes, as this increase the damage suffered importantly.
-The values for these variables have been determined also empirically, and they are set to $\gamma = 0.8$, $\beta_{1} = 0.75$ and $\beta_{2} = 0.25$.
+This differentiation is made in order to avoid low speeds in lava lakes, as this increase the damage suffered significantly.
+The values for these variables have also been determined empirically, and they are set to $\gamma = 0.8$, $\beta_{1} = 0.75$ and $\beta_{2} = 0.25$.
 
@@
 The experiments performed in this research compare three different algorithms in the two benchmarks presented in Section~\ref{sec:bench}: a single objective MCTS (referred to here simply as \textit{MCTS}), the Multi-Objective MCTS (\textit{MO-MCTS}) and a rolling horizon version of the NSGA-II algorithm described in Section~\ref{sec:moo} (\textit{NSGA-II}).
-This NSGA-II version evolves a population where the individuals are sequences of actions (macro-actions in the MO-PTSP case), obtaining the fitness from the state of the game after applying such sequence.
+This NSGA-II version evolves a population where the individuals are sequences of actions (macro-actions in the MO-PTSP case), obtaining the fitness from the state of the game after applying the sequence.
 The populations sizes, determined empirically, were set to $20$ and $50$ individuals for DST and MO-PTSP respectively.
@@
 In order for these games to be real-time, the time budget allowed is close to $40$ milliseconds.
-With the objective of avoiding congestion peaks at the machine where the experiments are run, the average of number of evaluations doable in $40$ ms.
-is calculated and employed in the tests.
-This leaded to $4500$ evaluations in the DST, and $500$ evaluations for MO-PTSP, using the same server where the PTSP and MO-PTSP competitions were run\footnote{Intel Core i5, 230GHz, 4GB of RAM.}.
+With the objective of avoiding congestion peaks at the machine where the experiments are run, the average number of evaluations possible in $40$ ms is calculated and employed in the tests.
+This leads to $4500$ evaluations in the DST, and $500$ evaluations for MO-PTSP, using the same server where the PTSP and MO-PTSP competitions were run\footnote{Intel Core i5, 230GHz, 4GB of RAM.}.
 
@@
 The weight vector for DST has two dimensions.
-This vector is here referred to as $W = (w_p, w_v)$, where $w_p$ weights moves and $w_v$ the treasure value, and $w_v = 1 - w_p$.
+This vector is here referred to as $W = (w_p, w_v)$, where $w_p$ weights moves; and $w_v$ weights the treasure value ($w_v = 1 - w_p$).
 $w_p$ takes values between $0$ and $1$, with a precision of $0.01$, and $100$ experiments have been run for each pair $(w_p, w_v)$.
@@
 Figure~\ref{fig:extCase} shows the results obtained after these experiments were performed.
-The first point to notice is that MCTS only converges (mostly) to the two optimal points located at the edges of the optimal Pareto front (\textit{OPT-1} and \textit{OPT-10}, see also Figure~\ref{fig:optDst}).
+The first point to note is that MCTS only converges (mostly) to the two optimal points located at the edges of the optimal Pareto front (\textit{OPT-1} and \textit{OPT-10}, see also Figure~\ref{fig:optDst}).
 This is an expected result, as K.
@@
 It is important to highlight that these two algorithms employed the euclidean distance action selection (see Section~\ref{sec:momcts}).
-Other experiments, not included in this paper, showed that weighted sum action selection provides similar results as in MCTS (this is, convergence to the edges of the front).
-The crucial distinction to make here is that both algorithms NSGA-II and MO-MCTS allow for better action selection mechanisms, that are able to overcome this problem, by approximating a global Pareto front.
+Other experiments, not included in this paper, showed that weighted sum action selection provides similar results to MCTS (that is, convergence to the edges of the front).
+The crucial distinction to make here is that both algorithms, NSGA-II and MO-MCTS, allow for better action selection mechanisms, that are able to overcome this problem by approximating a global Pareto front.
 
@@
 The experiments are performed in the $10$ maps available within the framework~\footnote{Available at \url{www.ptsp-game.net}}.
-In this case, the optimal Pareto front is not known in advance, and normally it will be different from one map to another.
+In this case, the optimal Pareto front is not known in advance, and normally it will differ from one map to another.
 Hence, the mechanisms to compare the performance of the algorithms tested need to be different than the one used for DST.
@@
 $W_1$ treats all objectives with the same weight, while the other three give more relevance to different objectives.
-These vectors provide then a wide spectrum for weights in this benchmark.
-In this case, MO-MCTS uses this weights to select an action based on a weighted sum, which has shown in preliminary experiments better performance than the Euclidean distance mechanism.
+These vectors then provide a wide spectrum for weights in this benchmark.
+In this case, MO-MCTS uses these weights to select an action based on a weighted sum, which in preliminary experiments has shown better performance than the Euclidean distance mechanism.
 
 The first point to check is if the different weight vectors affect the solutions obtained by MO-MCTS.
-Table~\ref{tab:weights} shows the results of executing the MO-MCTS controller during $30$ games in the $10$ maps of the MO-PTSP, for every weight vector.
+Table~\ref{tab:weights} shows the results of executing the MO-MCTS controller during $30$ games in each one of the $10$ maps of the MO-PTSP, for every weight vector.
 It can be seen that the highest weight in $W$ leads, in most of the cases, to the best solution in that objective in the map.
@@
 
-Some exceptions can actually be explained analysing the map: in map 1, a fuel canister is always collected near the end, restoring the fuel level to its maximum.
-This leaves too few cycles to make a difference in the controller (note also that this is the map with the fewest fuel consumption).
-Also, in map 10, there is no difference in the damage objective: but precisely map 10 is a map with no obstacles to damage the ship (thus only lava lakes deals damage).
-This also results in this map being the one with lowest damage overall.
-It can also be seen that, in those maps were time has priority, the results in this objective are not as dominant as the other two.
+Some exceptions can be explained by analysing the map: in map 1, a fuel canister is always collected near the end, restoring the fuel level to its maximum.
+This leaves too few cycles to make a difference in the controller (note also that this is the map with the smallest fuel consumption).
+Also, in map 10, there is no difference in the damage objective: but map 10 is a map with no obstacles to damage the ship (thus only lava lakes deal damage).
+This also results in this map being the one with the lowest overall damage.
+It can also be seen that, in those maps where time has priority, the results for this objective are not as dominant as the other two.
 This can be explained by the fact that the time heuristic is actually part of the fuel and damage heuristics (Equation~\ref{eq:dist} in Eqs.~\ref{eq:fuel} and~\ref{eq:dam}).
@@
 
-In order to compare several algorithms between each other, the results are compared in pairs, in terms of dominance.
+In order to compare the several algorithms, the results are examined in pairs, in terms of dominance.
 The procedure is as follows: once all games on a single map have been run, the Mann-Whitney-Wilcoxon non-parametric test with $95\%$ confidence is calculated on the three objectives.
@@
 Particularly, the dominance of MO-MCTS over MCTS is outstanding, even dominating in all $10$ maps for two of the weight vectors.
-MO-MCTS also dominates NSGA-II in more maps than in those where there is no dominance, and it is as well never dominated by NSGA-II in any map.
+MO-MCTS also dominates NSGA-II in more maps than in those where there is no dominance, and neither is it ever dominated by NSGA-II in any map.
 
@@
 
-Additionally, Table~\ref{tab:resMOPTSP} contains a fourth entry, \texttt{PurofMovio}, that the algorithms are compare against.
-\texttt{PurofMovio} is the winner entry of the 2013 MO-PTSP competition, a controller based in a weighted-sum MCTS approach (see~\cite{Powley2013} for details of its implementation).
+Additionally, Table~\ref{tab:resMOPTSP} contains a fourth entry, \texttt{PurofMovio}, that the algorithms are compared against.
+\texttt{PurofMovio} was the winning entry of the 2013 MO-PTSP competition, a controller based on a weighted-sum MCTS approach (see~\cite{Powley2013} for details of its implementation).
 As can be seen, \texttt{PurofMovio} obtains better results than the algorithm proposed in this paper.
@@
 Hence, nothing can be concluded from making pairwise comparisons directly with the winning entry of the competition.
-It is likely that \texttt{PurofMovio}'s heuristics are more efficient that the ones presented here, but the goal of this paper is not to develop the best heuristics for the MO-PTSP, but to provide an insight of how a multi-objective version of MCTS compares to other algorithms using the same heuristics.
+It is likely that \texttt{PurofMovio}'s heuristics are more efficient than the ones presented here, but the goal of this paper is not to develop the best heuristics for the MO-PTSP, but to provide an insight into how a multi-objective version of MCTS compares to other algorithms using the same heuristics.
 
 Anyway, the inclusion of \texttt{PurofMovio} in this comparison is not pointless: it is possible to asses the quality of the three algorithms tested here by comparing their performance relatively, against this high quality entry.
-Attending to this criteria, it can be seen how MO-MCTS is the algorithm that is dominated less often by him, producing similar results on an average of $4.75$ out of the $10$ maps, and being dominated in $5.25$ maps.
-MCTS and NSGA-II are dominated more often than MO-MCTS, being dominated on averages of $7.5$ and $5.75$ of the maps, respectively.
+Attending to this criteria, it can be seen how MO-MCTS is the algorithm that is dominated less often by \texttt{PurofMovio}, producing similar results on an average of $4.75$ out of the $10$ maps, and being dominated in $5.25$ maps.
+MCTS and NSGA-II are dominated more often than MO-MCTS, being dominated in an average of $7.5$ and $5.75$ of the maps, respectively.
 This comparative result shows again that MO-MCTS is achieving the best results among the three algorithms compared here.
@@
 
-There is another aspect that can still be improved in the MO-PTSP benchmark, but that it is also applicable to other domains.
-It is naive to think that an unique weight vector will be the ideal one for the whole game.
-Specifically in the MO-PTSP, there are regions of the map where there are more obstacles or lava lakes, hence the ship is most likely to suffer a higher damage.
-Also, the route followed during the game affects the relative ideal speed between waypoints, or sometimes a fuel canister is going to be picked up, which affects how the fuel objective can be managed.
+There is another aspect that can be further improved in the MO-PTSP benchmark, and is also applicable to other domains.
+It is naive to think that a unique weight vector will be the ideal one for the whole game.
+Specifically in the MO-PTSP, there are regions of the map where there are more obstacles or lava lakes, hence the ship is most likely to suffer higher damage there.
+Also, the route followed during the game affects the relative ideal speed between waypoints, or perhaps a fuel canister is going to be picked up, which will affect how the fuel objective can be managed.
 In general, many real-time games go through different phases, with different objectives and priorities.
@@
 A way to provide different weights at different times in MO-PTSP is straightforward.
-Given the route of waypoints (and fuel canisters) that is being followed, one can divide it into \textit{segments}, where is segment starts and ends with a waypoint (or fuel canister).
+Given the route of waypoints (and fuel canisters) that is being followed, one can divide it into \textit{segments}, where each segment starts and ends with a waypoint (or fuel canister).
 Then, each segment can be assigned a particular weight vector $W$.
@@
 An individual is identified by a string of integers, where each integer refers to one of the weight vectors utilized in the previous sections ( $1 = W_1 = (0.33,0.33,0.33)$, $2 = W_2 = (0.1,0.3,0.6)$ and $3 = W_3 = (0.1,0.6,0.3)$.
-$W_4$ has been left out of this experiment, as it has shown to be the less influential weight vector).
+$W_4$ has been left out of this experiment, as it has shown to be the least influential weight vector).
 The solution is evaluated playing a particular map $10$ times, and its fitness is obtained by calculating the average of those runs.
 A population of $10$ individuals is kept, and the solutions of the initial population are created either randomly, or mutated from base individuals.
-These base individuals have all segments with the same weight vector $W_1$, $W_2$, or $W_3$.
+These base individuals all have segments with the same weight vector $W_1$, $W_2$, or $W_3$.
 
@@
 \end{tabular}
-\caption{MO-PTSP Results with different weights.}
+\caption{MO-PTSP Results with different weights. The last column indicates if the evolved individual dominates ($\preceq$) or not ($\O$) each one of the base genomes for that particular map.}
 \label{tab:resVarW}
@@
 If the focus is set on what particular weights are more successful, it is interesting to see that the best configurations found are better, in all maps, than the base individuals with all weights equal to $W_1$ ($w_t = 0.33, w_f = 0.33, w_d = 0.33$) and $W_2$ ($w_t = 0.1, w_f = 0.3, w_d = 0.6$).
-In other words, it was always possible to find a combination of weight vectors that performed better than approaches that gave the same weight to all objectives, and also than the ones prioritizing low damage.
-
-It is also worthwhile to mention that those cases where dominance over the base individual was not achieved, this base was the one that prioritized fuel.
-Actually, it can be seen that in these cases, the result in the fuel objective is the one that avoids the dominance to happen, obtaining a better value than the evolved solution in this particular objective.
+In other words, it was always possible to find a combination of weight vectors that performed better than approaches that gave the same weight to all objectives, and also better than the ones that prioritize low damage.
+
+It is also worthwhile mentioning that in those cases where dominance over the base individual was not achieved, the base was the one that prioritized fuel.
+Actually, it can be seen that in these cases, the result in the fuel objective is the one that prevents the dominance from happening, obtaining a better value than the evolved solution in this particular objective.
 
@@
 
-This paper presents a multi-objective version of Monte Carlo Tree Search (MO-MCTS) for real-time games, scenarios where the time budget allowed for finding out the best possible action to apply next is clost to ~$40$ms.
+This paper presents a multi-objective version of Monte Carlo Tree Search (MO-MCTS) for real-time games: scenarios where the time budget allowed for finding out the best possible action to apply next is close to $40$ms.
 MO-MCTS is tested, in comparison with a single-objective MCTS algorithm and a rolling horizon NSGA-II, in two different real-time games, the Deep Sea Treasure (DST) and the Multi-Objective Physical Travelling Salesman Problem (MO-PTSP).
@@
 However, it could be desirable to model behaviours that dynamically change the weight vectors according to measurements of the game state, instead of evolving a different solution for each particular map.
-A possible extension for this work is to analyse and discover features in the game state that allows to establish relationships between game situations and the weight vectors to use for each objective.
+A possible extension for this work is to analyse and discover features in the game state that allows the establishment of relationships between game situations and the weight vectors to use for each objective.
 This mechanism would be general enough to be applicable to many different maps without requiring specific learning in any particular level.
@@
 This work was supported by EPSRC grants EP/H048588/1, under the project entitled ``UCT for Games and Beyond''.
-The authors would also like to thank Edward Powley and Daniel Whitehouse for their work in the controller \texttt{PurofMovio} for the MO-PTSP.
+The authors would also like to thank Edward Powley and Daniel Whitehouse for their work in the controller \texttt{PurofMovio} for the MO-PTSP competition.
 

### 2014-02-27 paper/paper.tex
last changes before submission

@@
 
-If, during the \textit{tree selection} phase, a node has less children than the available number of actions from a given position, a new node is added as a child of the current one (\textit{expansion} phase) and the \textit{simulation} step starts.
+If, during the \textit{tree selection} phase, a node has fewer children than the available number of actions from a given position, a new node is added as a child of the current one (\textit{expansion} phase) and the \textit{simulation} step starts.
 At this point, MCTS executes a Monte Carlo simulation (or roll-out; \textit{default policy}) from the expanded node.
@@
 
-Anyway, the inclusion of \texttt{PurofMovio} in this comparison is not pointless: it is possible to asses the quality of the three algorithms tested here by comparing their performance relatively, against this high quality entry.
+Nonetheless, the inclusion of \texttt{PurofMovio} in this comparison is not pointless: it is possible to asses the quality of the three algorithms tested here by comparing their performance relatively, against this high quality entry.
 Attending to this criteria, it can be seen how MO-MCTS is the algorithm that is dominated less often by \texttt{PurofMovio}, producing similar results on an average of $4.75$ out of the $10$ maps, and being dominated in $5.25$ maps.

### 2014-03-22 code/src/lunarcode/HOOClass.java
Almost working

@@ -1,0 +1,43 @@
+package lunarcode;
+
+import framework.core.Game;
+import ssamot.mcts.ucb.optimisation.ContinuousProblem;
+
+/**
+ * Created with IntelliJ IDEA.
+ * User: ssamot
+ */
+public class HOOClass extends ContinuousProblem {
+
+
+    private final  LunarGame game;
+    double max = Integer.MAX_VALUE;
+    double min = Integer.MIN_VALUE;
+
+
+    public HOOClass(LunarGame cGame) {
+
+        game = cGame;
+    }
+
+
+
+    @Override
+    public double getFtarget() {
+        // TODO Auto-generated method stub
+        return 0;
+    }
+
+    @Override
+    public double evaluate(double[] x) {
+       LunarLanderObjective obj = new LunarLanderObjective(game);
+
+       double score = obj.valueOf(x);
+       min = Math.min(min,score);
+       max = Math.max(max, score); 
+       
+       //return -(score-min)/(max-min);
+        return -score;
+    }
+
+}

### 2014-03-22 code/src/lunarcode/LunarGame.java

@@ -7,6 +7,8 @@ import framework.utils.Vector2d;
 import java.awt.*;
 import java.util.HashMap;
 import java.util.List;
+import java.util.Random;
+;
 
 /**
  * Created by Samuel Roberts, 2013
@@ -17,6 +19,8 @@ public class LunarGame extends Game {
     public LunarTerrain terrain;
     public LunarShip ship;
     public int ticks;
+    public Random random = new Random();
+
 
     // did the ship contact the landscape?
     public boolean landed;
@@ -35,6 +39,8 @@ public class LunarGame extends Game {
 
         // add gravity to the ship
         ship.v.add(0, LunarParams.lunarGravity * LunarParams.dt);
+        thrust+=random.nextGaussian()*0.01;
+        spin += random.nextGaussian()*0.0001;
         ship.updateCont(thrust,spin);
 
         // wrap ship around screen based on centre x (basic but works)
@@ -44,7 +50,13 @@ public class LunarGame extends Game {
         // check for collision
         if(terrain.isShipColliding(ship)) {
             landed = true;
-            if(terrain.onLandingPad(ship) && ship.v.mag() <= LunarParams.survivableVelocity) landedSuccessfully = true;
+            Vector2d nv = ship.d.copy();
+            nv.normalise();
+
+            double anglePoints = nv.dot(LunarParams.landingFacing) / (nv.mag() * LunarParams.landingFacing.mag());
+            anglePoints = Math.acos(anglePoints);
+            //System.out.println(anglePoints);
+            if(terrain.onLandingPad(ship) && ship.v.mag() <= LunarParams.survivableVelocity && (anglePoints < 0.1)) landedSuccessfully = true;
         }
 
         ticks++;

### 2014-03-22 code/src/lunarcode/LunarLanderEvoCont.java

@@ -15,14 +15,31 @@ import java.util.Arrays;
  */
 public class LunarLanderEvoCont {
 
+    private double[] solution;
+    private int s_i = 0;
+    private int macro_counter = 0;
+
     public double[] getAction(Game copy, ElapsedCpuTimer ect) {
 
-        CMAEvolutionStrategy cma = new CMAEvolutionStrategy();
-        cma.setDimension(10); // overwrite some loaded properties
-        cma.parameters.setPopulationSize(20);
-        cma.setInitialX(0.5); // in each dimension, also setTypicalX can be used
-        cma.setInitialStandardDeviation(0.3);
         LunarLanderObjective fitfun = new LunarLanderObjective((LunarGame) copy.getCopy());
+
+        if(solution!=null) {
+            double[] tbr = fitfun.convertInputs(solution[s_i],solution[s_i+1]);
+            macro_counter+=1;
+            if(macro_counter >= fitfun.MACRO_LENGTH) {
+
+                s_i+=2;
+                macro_counter = 0;
+            }
+            return tbr;
+        }
+
+        CMAEvolutionStrategy cma = new CMAEvolutionStrategy();
+        cma.setDimension(8); // overwrite some loaded properties
+        //cma.parameters.setPopulationSize(10);
+        cma.setInitialX(0.5); // in each dimension, also setTypicalX can be used
+        cma.setInitialStandardDeviation(0.2);
+
         double[] fitness = cma.init();
 
         while (!ect.exceededMaxTime()) {
@@ -49,21 +66,21 @@ public class LunarLanderEvoCont {
         double[] best = cma.getBestX().clone();
 
 
-//        for (int i = 0; i < best.length ; i++) {
-//
-//            if(best[i] < 0 ) {
-//                best[i] = 0.0;
-//            }
-//
-//            if(best[i] > 1 ) {
-//                best[i] = 1.0;
-//            }
-//        }
-
         System.out.println("best" + Arrays.toString(best));
         System.out.println(cma.getBestFunctionValue() + " " + cma.getCountEval()) ;
         //System.exit(0);
         double[] tbr = fitfun.convertInputs(best[0],best[1]);
+        
+        if(cma.getBestFunctionValue() < -LunarLanderObjective.WIN_BASE) {
+            //LunarLanderObjective.MACRO_LENGTH -=1;
+
+            solution = best;
+            s_i = 0;
+            macro_counter = 1;
+
[…]

### 2014-03-22 code/src/lunarcode/LunarLanderObjective.java

@@ -3,20 +3,19 @@ package lunarcode;
 import fr.inria.optimization.cmaes.fitness.IObjectiveFunction;
 import framework.utils.Vector2d;
 
-import java.util.Arrays;
-
 /**
  * Created with IntelliJ IDEA.
  * User: ssamot
  * Date: 17/03/14
  * Time: 13:08
- * This is a Java port from Tom Schaul's VGDL - https://github.com/schaul/py-vgdl
  */
 public class LunarLanderObjective implements IObjectiveFunction {
 
     private final LunarGame game;
-    private static final int MACRO_LENGHT = 4;
-    Vector2d desiredDirection = new Vector2d(1, 0);
+    public static int MACRO_LENGTH = 8;
+    Vector2d vertical = new Vector2d(1, 0);
+    Vector2d horizontal = new Vector2d(0, -1);
+    public static final int WIN_BASE = 999999999;
 
 
     public LunarLanderObjective(LunarGame game) {
@@ -28,93 +27,128 @@ public class LunarLanderObjective implements IObjectiveFunction {
         x = x.clone();
         LunarGame cGame = (LunarGame) game.getCopy();
 
-//        for (int i = 0; i < x.length; i++) {
-//
-//            if (x[i] < 0) {
-//                x[i] = 0.0;
-//            }
-//
-//            if (x[i] > 1) {
-//                x[i] = 1.0;
-//            }
-//        }
+        Vector2d nearestPad = cGame.terrain.getNearestSafeLandingPoint(game.getShip().s);
+        double distFromPad = cGame.getShip().s.dist(nearestPad);
+        double xdist_oring =   Math.abs(game.getShip().s.x - nearestPad.x);
+        double y_maxspeed = 2;
+        double x_ignorspeed = 1;
+        if (distFromPad > 40) {
+            y_maxspeed = 30;
+            x_ignorspeed = 10;
+        }
+
+
+        int phase = 0;
+        if ( xdist_oring < 50) {
+            phase = 1;
+        }
+
         int i = 0;
-        forward: {
-        for (i = 0; i < x.length - 1; i += 2) {
+        double score = 0;
 
-            //System.out.println(thrust + " " + spin);
-            double[] realVulues = convertInputs(x[i], x[i + 1]);
-            for (int j = 0; j < MACRO_LENGHT; j++) {
-                cGame.tickCont(realVulues[0], realVulues[1]);
-                if (cGame.isEnded()) {
+        forward:
+        {
[…]

### 2014-03-22 code/src/lunarcode/LunarLanderΗΟΟCont.java

@@ -1,0 +1,62 @@
+package lunarcode;
+
+import fr.inria.optimization.cmaes.CMAEvolutionStrategy;
+import framework.core.Game;
+import framework.utils.ElapsedCpuTimer;
+import ssamot.mcts.MCTS;
+import ssamot.mcts.ucb.optimisation.EDECBOptimiser;
+import ssamot.mcts.ucb.optimisation.HOOOptimiser;
+
+import java.util.Arrays;
+
+/**
+ * Created with IntelliJ IDEA.
+ * User: ssamot
+ * Date: 17/03/14
+ * Time: 13:57
+ */
+public class LunarLanderΗΟΟCont {
+
+    public double[] getAction(Game copy, ElapsedCpuTimer ect) {
+
+
+        int min = 0;
+        int max = 1;
+        int gamma = 1;
+        MCTS.DEBUG = false;
+
+        HOOClass func = new HOOClass((LunarGame) copy);
+        int dimension = 2;
+        int iterations = 10000;
+        HOOOptimiser hoo = new HOOOptimiser(func, dimension, iterations, min,
+                max, gamma);
+
+//        EDECBOptimiser hoo = new EDECBOptimiser(func, dimension, iterations, min,
+//                max, gamma);
+        LunarLanderObjective fitfun = new LunarLanderObjective((LunarGame) copy.getCopy());
+
+
+        int i = 0;
+        while (!ect.exceededMaxTime()) {
+            hoo.runForSim(1);
+            i++;
+        }
+
+
+        double[] best = hoo.getBestNode().sampleAction();
+        System.out.println("best" + Arrays.toString(best));
+        //System.out.println(hoo + " " + i) ;
+        //System.exit(0);
+        double[] tbr = fitfun.convertInputs(best[0],best[1]);
+        //LunarGame g = (LunarGame) copy.getCopy();
+        //g.tickCont(tbr[0],tbr[1]);
+        //System.out.println(g.getShip().v.mag() + " sdfdf")      ;
+
+
+        System.out.println("tbtr" + Arrays.toString(tbr));
+
+        return tbr;
+
+    }
+
+}

### 2014-03-22 code/src/lunarcode/LunarParams.java

@@ -49,6 +49,6 @@ public class LunarParams {
     // make the landscape totally flat?
     public static boolean flatLandscape = false;
 
-    public static Vector2d startingPoint = new Vector2d(400, 200);
-    public static Vector2d landingFacing = new Vector2d(0, 1);
+    public static Vector2d startingPoint = new Vector2d(700, 100);
+    public static Vector2d landingFacing = new Vector2d(1,0);
 }

### 2014-03-22 code/src/lunarcode/LunarRun.java

@@ -6,6 +6,7 @@ import framework.core.Game;
 import framework.core.PTSPConstants;
 import framework.utils.ElapsedCpuTimer;
 import framework.utils.JEasyFrame;
+import framework.utils.Vector2d;
 
 /**
  * Created by Samuel Roberts, 2013
@@ -17,6 +18,7 @@ public class LunarRun {
 
     public static Game game;
     public static LunarLanderEvoCont cont;
+    //public static LunarLanderΗΟΟCont cont;
 
     public static void main(String[] args) {
         runGame();
@@ -25,6 +27,7 @@ public class LunarRun {
     public static void prepareGame() {
         game = new LunarGame();
         cont = new LunarLanderEvoCont();
+        //cont = new LunarLanderΗΟΟCont();
     }
 
     public static void runGame() {
@@ -44,6 +47,13 @@ public class LunarRun {
             ((LunarGame) game).tickCont(actionToExecute[0], actionToExecute[1]);
             System.out.println("speed = " + ((LunarGame) game).getShip().v.mag() + " " + ((LunarGame) game).landedSuccessfully);
             view.repaint();
+
+            Vector2d nv = ((LunarGame) game).ship.d.copy();
+            nv.normalise();
+
+            double anglePoints = nv.dot(LunarParams.landingFacing) / (nv.mag() * LunarParams.landingFacing.mag());
+            anglePoints = Math.acos(anglePoints);
+            System.out.println("anglePoints = " + anglePoints);
             try {
                 Thread.sleep(LunarParams.delay);
             } catch (InterruptedException e) {
@@ -52,7 +62,7 @@ public class LunarRun {
 
         }
         System.out.println("Ship landed with velocity " + ((LunarShip) game.getShip()).v.mag() + "Victory?" + ((LunarGame) game).landedSuccessfully);
-        System.exit(0);
+        //System.exit(0);
     }
 
 

### 2014-03-22 code/src/lunarcode/LunarLanderHOOCont.java
Minor Fix

@@ -15,7 +15,7 @@ import java.util.Arrays;
  * Date: 17/03/14
  * Time: 13:57
  */
-public class LunarLanderΗΟΟCont {
+public class LunarLanderHOOCont {
 
     public double[] getAction(Game copy, ElapsedCpuTimer ect) {
 

### 2014-03-22 code/src/lunarcode/LunarRun.java

@@ -18,7 +18,7 @@ public class LunarRun {
 
     public static Game game;
     public static LunarLanderEvoCont cont;
-    //public static LunarLanderΗΟΟCont cont;
+    //public static LunarLanderHOOCont cont;
 
     public static void main(String[] args) {
         runGame();
@@ -27,7 +27,7 @@ public class LunarRun {
     public static void prepareGame() {
         game = new LunarGame();
         cont = new LunarLanderEvoCont();
-        //cont = new LunarLanderΗΟΟCont();
+        //cont = new LunarLanderHOOCont();
     }
 
     public static void runGame() {

### 2014-03-24 lunarcontpaper/lunarcont.tex
paper now actually hopefully in repo

\documentclass[conference]{IEEEtran}
% If the IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it.  e.g.
% \documentclass[conference]{./IEEEtran}

% Add and required packages here
\usepackage{graphicx,times,amsmath,fontenc,multirow,url,caption}
% I don't even remember which of these packages are even being used 
% but I dare not remove any in case the entire thing breaks

% To create the author's affliation portion using \thanks
\IEEEoverridecommandlockouts

\begin{document}

% Paper title: keep the \ \\ \LARGE\bf in it to leave enough margin.
\title{\ \\ \LARGE\bf Continuous Action Solutions in the {\itshape Lunar Lander} Problem } 
% TODO - better title

\author{Samuel A. Roberts, Spyridon Samothrakis, Simon M. Lucas}

% Uncomment out the following line for invited papers
%\specialpapernotice{(Invited Paper)}

% Make the title area
\maketitle

\begin{abstract}

% come back to this after the rest of the paper's done

\end{abstract}

% No keywords

\section{Introduction}

This paper describes an investigation into solving a continuous problem based on a familiar arcade game, {\itshape Lunar Lander}.

% TODO - actually determine what the core essence of what it is we're doing here before writing the introduction

\section{Motivation}

% TODO - THIS SECTION COULD DO WITH SOME FLESHING OUT

% building upon previous work in MCTS but using a continuous action space?

\subsection{Solving Continuous Problems with Continuous Actions} % this title is awful and it needs changing

Previous research into using MCTS to provide solutions for continuous physics-based problems have examined cases such as the Physical Travelling Salesman Problem \cite{perez14}. Within these continuous contexts, discrete actions have proven to be quite effective, especially with use of macro-actions and some degree of higher level planning \cite{powley12}.  % this section seems clunky

{\itshape Lunar Lander} is an excellent test case to study for many reasons, due to its inherently multi-objective nature. With the objectives to both land quickly while also minimising velocity changes, which cost fuel which must be conserved, it has been an interesting scenario to examine the properties and constraints imposed just by the environment alone. While there have been attempts to solve this deceptively complex problem before using discrete macro-actions \cite{roberts13}, in this paper the problem is being approached as one that can be better solved with continuous actions.

% TODO: this feels unfinished and only barely scratching the surface

% must find more examples of relevant previous research

\section{Methodology}

\subsection{Software Implementation}

% TODO

% anything notable to mention? section worth keeping?

\subsection{Environment and Physics}

% this is a heavily cut down rewritten version of the physics described in sam's 2013 cig paper
% more detail will be added depending on if it is appropriate or not

\subsubsection{Environmental Properties}

The properties of the base environment of {\itshape Lunar Lander} are that it is frictionless, and that it is a two-dimensional plane with horizontal wrapping. This can also be conceptualised as a cylinder. Anything that passes from the left edge of the playing field moves to the right instantaneously, and vice versa.

The other feature of note is the jagged landscape, a simplistic representation of the noisy, crater-filled lunar landscape the game represents. This landscape is constructed as a series of line segments, with each vertex of the jagged landscape being distributed equally horizontally, and randomly vertically. % unclear?

% TODO: DIAGRAM - jagged lunar landscape might be useful (reuse of old one probably not sufficient)

% not sure how much detail should be gone into here? should just cite generation equations from old paper or copy over??

Each vertex of this landscape is generated sequentially. % INCOMPLETE due to uncertainty listed above

\subsubsection{Spaceship Physics}

The spaceship within this game world is modelled as an a circular mass with some basic physical properties, such as a position within the game world $\boldsymbol {s}$, a velocity $\boldsymbol {v}$, an orientation $\theta$, an angular velocity $\omega$ and a radius of a bounding sphere $r$ used for collision detection with the landscape.

This collision detection method is based on the lowest point of the spaceship's bounding circle coming into intersection with the highest point of the landscape for the line perpendicular between the ship and the bottom of the playing field, as can be seen in Figure \ref{fig:diagram_landscapepoint}.

% this needs a diagram for clarity: reuse the old one? reusing it temporarily
\begin{figure}[hbtp]
\centering
\includegraphics[scale=0.4]{img/landscapepoint}
\caption{The point closest to the spaceship on the landscape lies in between two of the defined vertices of the landscape, and requires interpolation to calculate the y co-ordinate from the x co-ordinate of the ship.}
\label{fig:diagram_landscapepoint}
\end{figure}

As the landscape is stored as a series of line segments, interpolation must be used in the event the ship is not perfectly aligned with one of the landscape axes. The point that lies on the line segment that the ship is being checked against is calculated through simple linear interpolation based on which two vertices the ship is horizontally closest to. For the ship's centre $\boldsymbol {s}$, the left nearest landscape vertex $\boldsymbol {p}^{l}$ and the right nearest landscape vertex $\boldsymbol {p}^{r}$, the point of collision $\boldsymbol {p}^{c}$ against the landscape is calculated as

% does this constitute self-plaigarism???

\begin{equation}
\boldsymbol {p}_{x}^{c} = \boldsymbol {s}_{x}
\boldsymbol {p}_{y}^{c} = \boldsymbol {p}_{y}^{l} + v(\boldsymbol {p}_{y}^{r} - \boldsymbol {p}_{y}^{l})
\end{equation}

where $v$ is a value between $0$ to $1$ used for interpolation, and can be calculated as follows.

\begin{equation}
v = \frac{ {\boldsymbol {s}_{x} - \boldsymbol {p}_{x}^{l}}}{ {\boldsymbol {p}_{x}^{r} - \boldsymbol {p}_{x}^{l}} }
\end{equation}

Collision is then true if the following statement is true:

\begin{equation}
\boldsymbol {s}_{y} + r \geq \boldsymbol {p}^{c}_{y}
\end{equation}

The ship colliding with the landscape constitutes the end of the {\itshape Lunar Lander} game. The conditions surrounding this collision, including speed, orientation of the ship, and fuel used, constitute whether the nature of the collision is a success or a failure.

\subsection{Continuous Action MCTS} % this can be renamed something much better presumably

\subsubsection{Heuristic} % this too

% TODO - meeting or discussion might be good to be entirely certain on what's going on in the code
% DISCUSS IN MEETING: use of a two-stage approach: might be good to mention this?? temporary holdover?????

\section{Results}

% pending
% DISCUSS IN MEETING: we know that an effective two-stage solution exists, but how can this be written up?
% DISCUSS IN MEETING: the solution works even with noise as well, can the amount of time taken to reach the landing pad be improved?
% DISCUSS IN MEETING: is there any useful information to report, tables, graphs, etc? my methodology is too empirical, can't think of a better way of discussing results without graphs and tables
% DISCUSS IN MEETING: any alternate approaches feasible to compare against?

\section{Conclusion}

% pending
% DISCUSS IN MEETING: this entire section

% Use \section* for the acknowledgments
\section*{Acknowledgements}
Samuel Roberts is supported by an EPSRC PhD Studentship.
% any other acknowledgements??

\bibliographystyle{IEEEtran}
\bibliography{lunarcontbib}
\end{document}

### 2014-03-25 code/src/lunarcode/LunarGame.java
Latest code

@@ -39,8 +39,8 @@ public class LunarGame extends Game {
 
         // add gravity to the ship
         ship.v.add(0, LunarParams.lunarGravity * LunarParams.dt);
-        thrust+=random.nextGaussian()*0.01;
-        spin += random.nextGaussian()*0.0001;
+        thrust+=random.nextGaussian()*0.1;
+        spin += random.nextGaussian()*0.01;
         ship.updateCont(thrust,spin);
 
         // wrap ship around screen based on centre x (basic but works)

### 2014-03-25 code/src/lunarcode/LunarLanderObjective.java

@@ -32,7 +32,7 @@ public class LunarLanderObjective implements IObjectiveFunction {
         double xdist_oring =   Math.abs(game.getShip().s.x - nearestPad.x);
         double y_maxspeed = 2;
         double x_ignorspeed = 1;
-        if (distFromPad > 40) {
+        if (distFromPad > 60) {
             y_maxspeed = 30;
             x_ignorspeed = 10;
         }
@@ -68,7 +68,7 @@ public class LunarLanderObjective implements IObjectiveFunction {
                             double anglePoints = nv.dot(vertical) / (nv.mag() * vertical.mag());
                             anglePoints = Math.acos(anglePoints);
 
-                            if (anglePoints < 0.5) {
+                            if (anglePoints < 0.8) {
                                 anglePoints = 0.0;
                             }
                             score -= anglePoints;

### 2014-03-30 lpaper/Template_Latex_CIG2012/SamplePaper.tex
1/3 Done

\documentclass[conference]{IEEEtran}
% If the IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it.  e.g.
% \documentclass[conference]{./IEEEtran}

% Add and required packages here
\usepackage{graphicx,times,amsmath, fontenc}

% Correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor IEEEtran}

% To create the author's affliation portion using \thanks
\IEEEoverridecommandlockouts

\textwidth 178mm
\textheight 239mm
\oddsidemargin -7mm
\evensidemargin -7mm
\topmargin -6mm
\columnsep 5mm

\begin{document}

% Paper title: keep the \ \\ \LARGE\bf in it to leave enough margin.
\title{\ \\ \LARGE\bf Sample Paper for CIG'12: The 2012 IEEE
Conference on Computational Intelligence and Games \thanks{P. L{\aa}ngstrump
is with The Villa Villekula Institute, South Seas
(phone: +34-1-2345-6789; fax: +34-9-8765-4321; email: {\tt
pippi@vvkula.ss}); H. Granger is with the Hogwarts School of Witchcraft and Wizardry; F. Pi is with the T.I.A.}}

\author{Pippilotta L{\aa}ngstrump, {\it Senior Member, IEEE}, Hermione Granger, {\it Member, IEEE}, and Filem\'{o}n Pi}

% Uncomment out the following line for invited papers
%\specialpapernotice{(Invited Paper)}

% Make the title area
\maketitle

\begin{abstract}
The abstract goes here.  Please try to make it less than 150 words.
We suggest that you read this document carefully before you begin
preparing your manuscript.  As the IEEE does not want conference papers
to contain keywords, ensure you do not include them in your paper.
Also, at this time, the IEEE only has some general guidelines about the
format for conference papers.  It is up to each individual conference
to decide which format to use.  In order to have a uniform look for all
papers published in the conference proceedings, we require that every
author follow the format of this sample paper.

This template is for LaTeX users and is based on the CEC'07 template
written by K.C. Tan. Many thanks go to him for his hard work.
Authors should use this sample paper as a guide in the production of
their paper. Word users should download and use the template files
posted on the conference website.
\end{abstract}

% No keywords

\section{Introduction}
% No \PARstart
If you have an introduction for your paper, put it here.

This sample file is intended to serve as a ``starter file.''  You
need to replace the text in this file with the text that makes up
your paper.

\subsection{Subsection Heading Here}
If applicable, subsection text goes here.  Note that you need to use
$\backslash$subsection.  You may or may not have any subsections.
That is okay.

\subsubsection{Subsubsection Heading}
Insert any subsubsection text here.  Same thing as before --- you may
or may not have any subsubsections.

\subsubsection{About This Template}
This template is for LaTeX users and is based on the CEC'07 template
written by K.C. Tan.  Many thanks go to him for his hard work.  Authors
should use this sample paper as a guide in the production of their paper.
Word users should download and use the template files posted on the
conference website.

\subsection{Page Layout}
\begin{itemize}
\item The IEEE now only accepts 100$\%$ Xplore compliant papers prepared
in PDF format.  Please make sure that you follow these guidelines in
preparing your PDF files.  Violations of any of these specifications
may result in the rejection of your paper.
\item Paper size: US letter format ($8.5\times 11$ in or $216\times
278$ mm).
\item Paper length: maximum 8 pages, including figures, tables, and
references.  In exceptional circumstances, up to two additional pages
will be permitted for a charge of USD\$100 per additional page.
\item Paper format: double column, single spaced, 10pt font.
\item Text width: 7.0 in (178 mm).
\item Text height: 9.375 in (240 mm).
All text and figures must be contained in the $178 \times 240$ mm
image area.
\item The left/right/bottom margin must be 0.75 in (19 mm).
\item The top margin must be 0.75 in (19 mm), except for the title page,
where it must be 1 in (25 mm).
\item Text should appear in two columns, each 3.4 in (86.5 mm) wide, with
0.2 in (5 mm) space between columns.
\item Do {\bf NOT} number the pages in the manuscript.
\item Unix LaTeX users should use the following commands to ``compile''
their paper:
    \begin{itemize}
    \item latex mypaper
    \item dvips -Ppdf -G0 -tletter mypaper.dvi
    \item ps2pdf -dEmbedAllFonts=true mypaper.ps mypaper.pdf
    \end{itemize}
\end{itemize}

The page size and margin settings in IEEEtran.cls are set for IEEE
Transactions papers.  Some adjustments have been made to produce this
sample paper.

{\it Also, please note, IEEE PDF eXpress will be made available to assist
in creating IEEE Xplore compliant PDF files for camera-ready submission.}

\section{Results}
The main results and findings go here.

Do not number an equation if it will not be directly cited
in the paper.  In order to avoid numbered equations, use
$\backslash$begin\{equation*\}--$\backslash$end\{equation*\},
$\backslash$[ --$\backslash$], or \$\$--\$\$.  For example:
\begin{equation*}
a = b + c,
\end{equation*}
$$\dot x = f(x,u) + g(x,u),$$
or
\[\ddot s=G(s,t)\]
where $f$, $g$, and $G$ are functions.

Note that Equation (\ref{equation-eqn1}) below is numbered!  It is
produced using $\backslash$begin\{equation\}--$\backslash$end\{equation\}:
\begin{equation}
F_i(P_i)=a_{i}+b_{i}P_i+c_{i}P_{i}^2
\label{equation-eqn1}
\end{equation}
where $a_{i},\ b_{i}$, and $c_{i}$ are coefficients of unit $i$, and $P_i$
represents some value for unit $i$.

Aligning equations can be done with
either the align or eqnarray commands.  Recently,
$\backslash$begin\{align\}--$\backslash$end\{align\} has gained popularity
over $\backslash$begin\{eqnarray\}--$\backslash$end\{eqnarray\}.

Equation (\ref{equation-eqn2}) is produced using
$\backslash$begin\{align\}--$\backslash$end\{align\}:
\begin{align}
\dot {x}_l=& \sum_{i = 1}^m {\frac{c_{P_{x_i} } e^{k_{x_i}\bar{x}_i} + c_{N_{x_i} }
e^{ -  k_{x_i} \bar{x}_i}}{e^{k_{x_i} \bar{x}_i} + e^{ - k_{x_i} \bar{x}_i}}} \nonumber\\
& + \frac{1}{2}\sum\limits_j^q (c_{P{u_j }} + c_{N _{u_j }} ) \nonumber\\
y=& \ A_0 + A_1 \tanh (K_x \bar {x}) + B\tanh (K_u \bar {u}) \nonumber\\
 =& \ F(x),
\label{equation-eqn2}
\end{align}
where $F(x)$ is a function.

Equation (\ref{equation-eqn3}) represents the same equation produced
using $\backslash$begin\{eqnarray\}--$\backslash$end\{eqnarray\}:
\begin{eqnarray}
\dot {x}_l&=& \sum_{i = 1}^m {\frac{c_{P_{x_i} } e^{k_{x_i}\bar{x}_i} + c_{N_{x_i} }
e^{ - k_{x_i} \bar{x}_i}}{e^{k_{x_i} \bar{x}_i} + e^{ - k_{x_i} \bar{x}_i}}} \nonumber\\
&&+ \frac{1}{2}\sum\limits_j^q (c_{P{u_j }} + c_{N _{u_j }} ) \nonumber\\
y&=& \ A_0 + A_1 \tanh (K_x \bar {x}) + B\tanh (K_u \bar {u})\nonumber\\
&=& \ F(x),
\label{equation-eqn3}
\end{eqnarray}
where $F(x)$ is a function.  You get the idea!

\subsection{Example of a Figure}
Below is an example of a floating figure using the graphicx
package.  Note that $\backslash$label must occur AFTER (or within)
$\backslash$caption.  For figures, $\backslash$caption should occur
after the $\backslash$includegraphics.  To reference a figure, use
the word Figure followed by the figure number.  Here is an example:
Figure~\ref{figure-fig1}.

\begin{figure}[htp]
\centerline{\includegraphics[width=0.6\columnwidth]{pippi_langstrumpf.eps}}
\caption{The first author and a member of her team}
\label{figure-fig1}
\end{figure}

\subsection{Figures and Tables}
Please follow the style in this sample paper when generating your figures
and tables.

\subsection{What Sections to Include}
Usually, your paper should contain an Introduction, Background,
Methodology, Results, and Conclusions section.  You may also add
Acknowledgments if you like.  After that, you should list your references
in a References section.

\subsection{Page Limit and Overlength Page Charges}
A paper submitted to this conference should be prepared in a
single-spaced, two-column format.  Its length must be kept to 8
pages or less.  In exceptional circumstances, up to two additional
pages will be permitted for a charge of USD\$100 per additional page.
Table~\ref{table-tab1} shows the page limit and page charge schedule.

% An example of a floating table.  Note that, for IEEE style tables, the
% \caption command should come BEFORE the table.  Table text will default
% to \footnotesize as IEEE normally uses this smaller font for tables.
% The \label must come after \caption as always.
\begin{table}
\begin{center}
%% Increase table row spacing; adjust to taste
\renewcommand{\arraystretch}{1.3}
\caption{Page Limit}
\label{table-tab1}
% The array package and the MDW tools package offers better commands
% for making tables than plain LaTeX2e's tabular which is used here.
\begin{tabular}{|c|c|}
\hline
Page limit: & 8\\
\hline
Excess page charge: & USD\$100/page\\
\hline
\end{tabular}
\end{center}
\end{table}

Another example of a table is shown in Table~\ref{table-tab2}.

\begin{table}[h]
\caption{A second table}
\begin{center}
\begin{tabular}{|c|c|c|c|c|c|}
\hline
\multicolumn{1}{|c|}{\raisebox{-1.50ex}[0cm][0cm]{\!Method\!}}
& \multicolumn{1}{|c|}{Mean}
& \multicolumn{1}{|c|}{Best}
& \multicolumn{1}{|c|}{Mean}
& \multicolumn{1}{|c|}{Maximum}
& \multicolumn{1}{|c|}{Minimum} \\
& time & time & cost & cost & cost\\ \hline
A      &  $928.36$  &  $926.20$  &  $124793.5$ & $126902.9$ & $123488.3$ \\ \hline
B      &  $646.16$  &  $644.28$  &  $124119.4$ & $127245.9$ & $122679.7$ \\ \hline
C      &  $1056.8$  &  $1054.2$  &  $123489.7$ & $124356.5$ & $122647.6$ \\ \hline
D      &  $632.67$  &  $630.36$  &  $123382.0$ & $125740.6$ & $122624.4$ \\ \hline
\end{tabular}
\label{table-tab2}
\end{center}
\end{table}

Citations are included like so~\cite{book}.
Multiple citations appear like this~\cite{conf,article}.

\section{Conclusions}
The conclusion goes here.

This template is for LaTeX users. This CIG'12 LaTeX template is
based on the CIG'08 template written by Luigi Barone and the CEC'07
template written by K.C. Tan. Many thanks go to them for their hard
work. Authors should use this sample paper as a guide in the
production of their paper. Word users should download and use the
template files posted on the conference website.

% Conference papers do not normally have an appendix, but if you must,
% use \section*
\section*{Appendix}
Put your appendix here if you have any.

% Use \section* for the acknowledgments
\section*{Acknowledgments}
The authors would like to thank Mr. XYZ for his/her help.
This work was supported in part by the National Science Foundation
under grant no. XXXXX, etc.

% Trigger a \newpage just before a given reference number in order to
% balance the columns on the last page.  Adjust the value as needed;
% it may need to be readjusted if the document is modified later.
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% The references section can either be generated by hand or by an
% automatic tool like BibTeX.  If using BibTex, use the standard IEEEtran
% bibliography style.
%\bibliographystyle{IEEEtran.bst}
%
% The argument to \bibliography is/are the name(s) of your BibTeX file(s)
% that contains string definitions and bibliography database(s).
%\bibliography{IEEEabrv,SamplePaper}
%
% If you generate the bibliography by hand, or if you copy in the
% resultant .bbl file, set the second argument of \begin to the number of
% references in the bibliography (used to reserve space for the reference
% number labels box).

\begin{thebibliography}{3}
\bibitem{book}
A.~Great, \emph{This is the book title}.\hskip 1em plus 0.5em minus 0.4em\relax
  This is the name of the publisher, 2006.

\bibitem{conf}
F.~Author, S.~Author, and T.~NonRelatedAuthor, ``This is the paper title,'' in
  \emph{This is the proceedings title}, 2008, pp. 1--8.

\bibitem{article}
B.~Myself, ``This is the title of the journal article,'' \emph{This is the name
  of the journal}, pp. 1--30, 2007.
\end{thebibliography}

% That's all folks...
\end{document}

### 2014-03-30 lpaper/lunarcont.tex
1/3 Done

\documentclass[conference]{IEEEtran}
\usepackage{graphicx,times,amsmath,algorithm,caption, subcaption,nicefrac,bm}
\usepackage{multicol}
\usepackage[noend]{algpseudocode}
\usepackage{mathtools}
\usepackage{bbm}
\usepackage{amssymb}

\pdfminorversion=3

\newcommand{\defeq}{\vcentcolon=}
\newcommand{\eqdef}{=\vcentcolon}
\newcommand{\mathsc}[1]{{\normalfont\textsc{#1}}}
\newcommand{\argmax}{\operatornamewithlimits{argmax}}
\DeclareMathOperator*{\argmin}{arg\,min}
% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}
% % \usepackage[table]{xcolor}

\DeclareMathOperator{\dis}{d}
\DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}

\newcommand{\twopartdef}[4]
{
	\left\{
		\begin{array}{ll}
			#1 & \mbox{if } #2 \\
			#3 & \mbox{if } #4
		\end{array}
	\right.
}

\IEEEoverridecommandlockouts

\begin{document}

% Paper title: keep the \ \\ \LARGE\bf in it to leave enough margin.
\title{\ \\ \LARGE\bf Rolling Horizon methods for Games with Continuous Actions } 
% TODO - better title

\author{Samuel A. Roberts, Spyridon Samothrakis, Diego Perez, Simon M. Lucas}

% Uncomment out the following line for invited papers
%\specialpapernotice{(Invited Paper)}

% Make the title area
\maketitle

\begin{abstract}
It is often the case that games have continuous dynamics and allow for continuous actions. For larger games with complicated dynamics, having agents learn offline behaviours is a daunting task. On the other hand, provided a forward model is available, one might try to spread the cost of search in a rolling horizon fashion (e.g. as in example MCTS). In this paper we compare HOLOP, an open loop planning algorithm and a version evolutionary planning using CMA-ES. He show that rolling horizon CMA-ES outperforms HOLOP in two classic benchmark problems (Cart Balancing and <problem name> ) and Lunar Lander, a classic arcade game. We conclude that off-the-shelf evolutionary algorithms can be used successfully in a rolling horizon setting.

\end{abstract}

% No keywords

\section{Introduction} \label{Introduction}
One of the most common uses of function optimisation is for solving control problems. If an agent has access to a generative model of the world (i.e. a simulator) in which it is going to act, a control problem becomes a search/optimisation problem, usually referred to as ``simulation based planning''. Recently, a family of algorithms, mostly known as Monte Carlo Tree Search (MCTS)~\cite{browne2012survey}, has been applied to planning problems of discrete states and actions with considerable success. The focus of these algorithms is to attack extremely large state spaces of perfect information games, where one can sample rewards from the state space easily (e.g. Computer Go~\cite{gelly2006modification}). With access to a generative model (from which one can easily sample) and perfect sensor information, the following procedure works well: the agent receives sensor information, formulates a plan of action, performs the first action of that plan, receives a new state, re-formulates a plan, acts again (using the first action of its plan),  ad infinitum.

The most important reason for continuous re-planning is the fact that most planning algorithms' computational complexity is linear (or worse) in the number of states, whereas the number of states increases exponentially at each time step. One thus hopes to perform an action that looks good now, act, and replan, effectively creating a smaller problem or a ``closer to action'' horizon. This kind of behaviour is known as rolling horizon, sample based or model predictive control/planning~\cite{chang2007simulation}. Most of these algorithms are not exact; at each time step, due to the large state space, planning takes place in a Monte Carlo fashion, with random ``rollouts'' (i.e. simulations) guiding the algorithm in a best first search manner. All this planning and replanning is computationally intensive, thus very efficient use of samples should be made.  In the case of discrete state and actions, bandit algorithms~\cite{auer2002finite} provided a formidable solution to this problem.
An example of exploiting this efficient sampling can be found in Samothrakis et al.~\cite{samothrakis2011fast} where a high-performance MCTS Pac-Man agent
was developed using only $50 - 300$ simulations per action (the game simulator requested an action
from the controller every 40ms).

Recently, algorithms stemming from, or at least partially inspired by, MCTS aiming at solving planning problems involving continuous states and actions have come to light. A typical such algorithm is Hierarchical Optimistic Open Loop Planing (HOLOP)~\cite{weinstein2012bandit}. HOLOP is based partially on a strong~\cite{hoo2011} real-valued optimisation algorithm called HOO (for ``Hierarchical Optimistic Optimisation''). In this paper two major contributions are made plus a smaller ``add-on''; the first contribution is the comparison of two versions of a very popular and strong evolutionary strategy: Covariance Matrix Adaptation Evolutionary Strategy (CMA-ES)~\cite{hansen2003reducing} and its Uncertainty Handling version UH-CMA-ES~\cite{hansen2009tec} with Truncated Hierarchical Optimistic Optimisation (T-HOO)~\cite{hoo2011} on the
Black Box Optimisation Problem (BBOP) benchmark suit~\cite{bbopnoisy}.  T-HOO is a faster version of HOO~\cite{hoo2011}, an algorithm reminiscent of Monte Carlo Tree Search. The experiments are conducted using similar conditions to those an agent would face if it was to solve planning problems. Our second contribution is to incorporate CMA-ES and UH-CMA-ES into three versions of an algorithm called EOLOP, for Evolutionary Open Loop Optimistic Planning.  This algorithm uses evolution instead of tree search for implementing sample-based control algorithms. We then proceed to test the algorithms proposed on noisy versions of standard reinforcement learning (RL)
benchmarks, namely the inverted pendulum and the double integrator problems. Our ``extra'' contribution is an add-on for evolutionary algorithms (transfered from HOLOP) that can help when the number of function evaluations is very small.

While evolution is commonly used to evolve a reactive neural network controller (in a process commonly called neuro-evolution) for these and other RL problems,
it should be noted that evolution is being applied in a very different way here. In this paper, evolution is applied to perform each action given the current state.  This approach can only be applied when a generative
model is available, but has the advantage of offering immediate good performance without any
prior learning.  The disadvantage, compared to neuro-evolution, is that every action performed requires
CPU time for the simulations.  This is described in more detail in section \ref{Planning}.

The rest of this paper is structured as follows.  Section \ref{Optimisation} discusses the relevant base algorithms that are going to be used in this paper, ((Truncated) Hierarchical Online Optimisation~\cite{hoo2011} and CMA-ES~\cite{hansen2003reducing}).
Section~\ref{Planning} explains why and how the above algorithms can be used in the context of planning. Section \ref{evo} describes how one can incorporate evolution into this, making this our main contribution section from an algorithmic standpoint. In Section \ref{Methodology} the experimental setup is described. In Section~\ref{sec:Experiments} a number of experiments are portrayed and analysed.  These form the bulk of this paper's contribution. We conclude with a short discussion in Section \ref{Conclusion}.

% TODO - actually determine what the core essence of what it is we're doing here before writing the introduction

\section{Motivation}

\subsection{Solving Continuous Problems with Continuous Actions} % this title is awful and it needs changing

Previous research into using MCTS to provide solutions for continuous physics-based problems have examined cases such as the Physical Travelling Salesman Problem \cite{perez14}. Within these continuous contexts, discrete actions have proven to be quite effective, especially with use of macro-actions and some degree of higher level planning \cite{powley12}.  % this section seems clunky

{\itshape Lunar Lander} is an excellent test case to study for many reasons, due to its inherently multi-objective nature. With the objectives to both land quickly while also minimising velocity changes, which cost fuel which must be conserved, it has been an interesting scenario to examine the properties and constraints imposed just by the environment alone. While there have been attempts to solve this deceptively complex problem before using discrete macro-actions \cite{roberts13}, in this paper the problem is being approached as one that can be better solved with continuous actions.

\section{Planning}  \label{Planning}
In this section a formalised version of planing will be presented and an explanation provided on how the algorithms described in the previous section can be used to attack the problem.
\subsection{Markov Decision Processes}

The main decision theoretic abstraction for planning/control is the Markov Decision Process (MDP)~\cite{howard1960dynamic}. Formally,  an MDP is a tuple $\langle S,A,T,R,\gamma\rangle$, where:

\begin{itemize}
\item $S$ is a set of states, $s \in S$, where $s_j^t$ is state $j$ at time $t$ and $s'_j$ is the state j at time $t+1$.
\item $A$ is a set of actions, each action named $a_j$.
\item  $T:S \times \S \times  A  \rightarrow [0,1]$ is the probability of moving from state $s$ to state $s'$  after action $a \in A$ has taken place. $T(s'|s,a)$ denotes this probability.
\item  $R:S \rightarrow \Re$, $R(s)$ is a reward function at each state.
\item $\gamma$, a discount factor.
\end{itemize}

The MDP defines a single agent environment, fully observable to the agent. In an MDP, the Markov property holds (hence the name), which means that all the information an agent needs in order to act is embedded in the current state.
A possible route of action an agent might take is known as the \textit{policy} $\pi$. A policy is a probabilistic mapping between state and actions, $\pi:S \times A \rightarrow [0,1]$, thus $\pi(s,a) \in [0,1]$,  $\sum_{a\in A}{\pi(s,a)} = 1$. The set of all policies is denoted as $\Pi$. The goal of an agent in an MDP environment is to maximise its long term value $V(\pi,s_0) = R(s) + \sum\limits_{s\in S}\sum\limits_{a\in A} \gamma \pi(s,a)T(s^\prime|s,a)  V(\pi,s^\prime) $, by assigning the probability of actions to take at each state.

\begin{figure}[ht]
  \centering
  \includegraphics[width=0.45\textwidth]{graphics/gamma-crop.pdf}
  \caption{Examples of how rewards are affected by $\gamma$.}
  \label{fig:gamma}
\end{figure}

\subsection{Taxonomy of Planning Algorithms}
The terminology of Chang~\cite{chang2007simulation} is followed when trying to classify planning algorithms. If the policy $\pi:S \times A \rightarrow [0,1]$ is followed, what is known as ``closed loop'' planning takes place. The agent takes an action, senses its environment, takes another action etc. An alternative to closed-loop planning, ``open-loop'' planning, requires defining time. An agent sees a ``time'' when it's about to act instead of a state, i.e it doesn't have access to $s_t$, but just $t$. An open loop policy (or plan or control) is when an agent learns a policy with the form $\pi_o:S \times \mathcal{T} \rightarrow [0,1]$, where  $\mathcal{T}$ is an ordered set of time steps. Thus, an agent takes actions irrespective of the current state it is in. Intuitively this means that a sequence of actions is formed by the agent and the agent will take these actions in sequence (e.g. $(a_1,a_2,a_3,a_4,a_5,a_6,\dots,a_n)$). Obviously this is not optimal\footnote{Not optimal in the general case. It is optimal under the condition that the MDP is deterministic, i.e. each action leads to one specific state with probability one, and everything else has a probability zero}, although sometimes is much easier to do.

Another point of interest is planning, and when it happens. In open loop planning (and what is usually termed ``planning'' without any further qualifications), an agent forms a plan once and follows it until the end.

If the agent replans at every step, discarding or augmenting the plan received from previous steps, it is known as ``rolling horizon'' planning. The idea here is that, if the agent can plan as well as possible up to a  certain point, it can perform an action, move the planning horizon one step forward and replan. For example, Monte Carlo Tree Search for infinite MDPs can be seen as an approximate rolling horizon planning version of $TD(1)$\cite{silver2009reinforcement}.  The term ``Simulation based'' is used when the planning happens for an MDP that has a tree like structure and rewards are only visible at the end of the tree, a situation common in many games.

When learning how to act using a generative model, one can make a third distinction as to the type of sets represented by $A$ and $S$. In this paper the focus is on an $A$ and $S$ that come from a metric space. This means  that both sets' elements define a notion of distance and, for all practical purposes, have an infinite amount of elements. More specifically, both sets are drawn from a bounded set of real numbers, $\Re^n$.

Since an open loop policy is now a real-valued vector of actions, algorithms like T-HOO and CMA-ES can be used to find such a policy. Hence the title of this paper:  Rolling Horizon Open-loop Policy Evolutionary Learning. To the best of our knowledge this is the first time evolution has been used in such a setting. There are examples of closed loop rolling horizon papers, but these are beyond the scope of this article~\cite{xiang2012differential,samothrakis2010planning}.

\subsection{Planning using a Generative Model}

The Markov Decision Process simply describes an environment alongside the rewards that said environment provides to an agent. A problem central to Artificial Intelligence is what actions an agent must take to maximise its long term reward. The most basic assumption one can make is that the agent knows nothing about the environment. In this case the agent gets thrown into the environment and interacts (``trains'') for some time. After the training is complete, the agent should be able to act intelligently (presumably with some success).

Another approach one can take is to assume that the agent has a copy of the MDP in its ``mind'' and thus is able to plan accordingly. This is usually termed dynamic programming~\cite{sutton1998reinforcement}, however it cannot be used in the case of continuous states or actions without the use of a function approximator (FA) or some kind of discretization. Another approach, and one which has made a significant impact recently, arguably presented for the first time by Kearns~\cite{citeulike:1584916} and popularised mostly as Upper Confidence Bounds for Trees (UCT)~\cite{kocsis2006bandit}, is to use a generative model.

A generative model is nothing more than a simulator of future events. This is a less of a requirement than knowing the full model (which entails knowing the full function $T$), but still requires the agent to have some knowledge of the world. This category of algorithms, which are broadly termed ``Rolling Horizon'', has the following steps. From the current state, a number of simulations of possible futures is run by executing a set of action sequences (all of this in the ``mind'' of the agent). The simulations last for a certain number of steps (hence the ``rolling horizon'') or continue until an end state is encountered (i.e. when the MDP has ``absorbing'' states that cannot be acted from).

Once a satisfactory plan/policy has been estimated (or the agent runs out of time, iterations etc. ) it proceeds by executing the first action of the plan, thus moving to the next state. Once there, the agent performs the same sequence of ``mental'' actions, (i.e. plans and executes the first action), until some predefined stopping condition.  In this paper we will concern ourselves \textit{only} with this method of planning. What changes from method to method is how the plan is formed.

In standard UCT, a procedure similar to temporal difference learning is followed, where one learns the value of each state. An agent then forms a policy by choosing a value greedily. In our case, because our state space has continuous states and actions, this is impossible without the use of a function approximator (or an infinite discretization of space), which is not attempted here. What is attempted though is to learn the policy function directly without reference to state, while planning continuously from the current state. So while the agent forms open-loop plans at each clock tick, in reality it is executing a closed-loop policy. While the strategy is not provably optimal, it can work well in practice~\cite{weinstein2012bandit}. This will be seen by our experimental results (see Section \ref{sec:Experiments}).

\begin{algorithm}[!ht]
\caption{Rolling Horizon Open Loop Planning}\label{algo:ooop}
\begin{algorithmic}

\State 	\textbf{define external function} $\mathsc{FindOpenLoopPlan} (depth) $ \Comment{Try to find the optimal policy. You can use either evolution or HOO here}

	\Procedure{$\mathsc{OOOP}$} {}
		\While{$\mathit{True}$}
		    \State $a \gets \mathsc{FindOpenLoopPlan}(depth)[0]$ \Comment{Execute first action of the policy}
		    \State $s \sim T(s'|s,a)$ \Comment{Stop only at an absorbing state}
		\EndWhile
		\State
	\EndProcedure

\end{algorithmic}
\end{algorithm}

\section{Methodology}

\subsection{Simple Planning Benchmarking Problems}

% % \todo[inline]{Simon to do this}.

For direct comparison with recent work on sample based planning using trees, in particular
Weinstein and Littman~\cite{weinstein2012bandit} and Pazis and Lagoudakis \cite{PazisBAS},
we used the same problems: double integrator, and inverted pendulum.  The basic
experimental setup followed Weinstein and Littman~\cite{weinstein2012bandit}
and Pazis and Lagoudakis \cite{PazisBAS}
closely, except that they used a fixed noise level for each experiment whereas we varied the
noise level to explore how performance was affected.

Both problems are modelled
using continuous state discrete time simulations.  In each case the desired
acceleration is corrupted by additive uniform random noise
before being limited within the specified range.  The noise levels are described in the next section.

\subsubsection{Double Integrator}

The double integrator problem is to control a mass along a single dimension.  The state space
is 2-dimensional, consisting of $(p,v)$ where $p$ is position and $v$ is velocity.
The goal is to change the state from $(1,0)$ to $(0,0)$.  Both position and velocity
are clamped to be within the range $-2$ to $+2$.  At each time step the controller selects
the desired acceleration $a'$, which then is noise corrupted and range limited between $-1.5N$ to $+1.5$N to yield the
applied acceleration $a$.

Euler integration is then used to update the position and velocity, given the time step $\delta_t$:

\begin{eqnarray}
v_{t+\delta_t} & = & v_{t} + a \delta_t \\
p_{t+\delta_t} & = & p_{t} + v_{t+\delta_t} \delta_t
\end{eqnarray}

At each time step the reward $r$ is:
\begin{equation}
 r = -(p^2 + a^2)
\end{equation}

The problem becomes harder as the time interval $\delta_t$ at %word missing?
each action is applied increases,
and following Pazis and Lagoudakis we set $(\delta_t = 0.5s)$.

% \begin{equation}
% R = -(p^2 + a^2)
% \end{equation}

\subsubsection{Inverted Pendulum}

The inverted pendulum problem is also known as the pole balancing
or cart-pole problem.  The goal is to keep the pendulum / pole as upright and as still as possible.
This is achieved by accelerating the cart which affects the freely pivoted pendulum.

In this instance of the problem the controller only sees a two dimensional
state space, consisting of the
angle $\theta$ that the inverted pendulum deviates from the vertical, and the angular velocity
of the pendulum $\dot{\theta}$.
The controller selects the desired angular force $F$,
which is then corrupted by additive uniform random noise and limited in the range $-50N$ to $+50$N
so as to yield the applied angular force $u$.  The angular acceleration $\ddot{\theta}$ is then calculated
as follows:

\begin{equation}
\ddot{\theta} = \frac{g sin(\theta) - \alpha m l (\dot{\theta})^2 sin(2 \theta) / 2 - \alpha cos(\theta) u}
                     {4l/3 - \alpha m l  cos^2(\theta)}
\end{equation}

where $g$ is the acceleration due to gravity
$(g = 9.8ms^{-2})$,
$m$ is the mass of the pendulum
$(m = 2kg)$, $(\alpha = 1 / (m + M))$, $M$ is the mass of the cart
$(M = 8kg)$ and $l$ is the length of the pendulum $(l = 0.5m)$.
The time interval $\delta_t$ was set to $0.1s$: i.e.\ the simulation is
updated 10 times per second:

\begin{eqnarray}
\dot{\theta}_{t+\delta_t} & = &  \dot{\theta}_{t} + \ddot{\theta} \delta_t \\
\theta_{t+\delta_t}  & = & \theta_{t} + \dot{\theta}_{t+\delta_t} \delta_t
\end{eqnarray}

At each timestep the reward $r$ punishes deviation from the vertical, high speed, and high force:
\begin{equation}
r = -((2 \theta / \pi)^2 + \dot{\theta}^2 + (F / 50)^2 )
\label{eq:rew:t}
\end{equation}

\subsection{Environment and Physics}

% this is a heavily cut down rewritten version of the physics described in sam's 2013 cig paper
% more detail will be added depending on if it is appropriate or not

\subsubsection{Environmental Properties}

The properties of the base environment of {\itshape Lunar Lander} are that it is frictionless, and that it is a two-dimensional plane with horizontal wrapping. This can also be conceptualised as a cylinder. Anything that passes from the left edge of the playing field moves to the right instantaneously, and vice versa.

The other feature of note is the jagged landscape, a simplistic representation of the noisy, crater-filled lunar landscape the game represents. This landscape is constructed as a series of line segments, with each vertex of the jagged landscape being distributed equally horizontally, and randomly vertically. % unclear?

% TODO: DIAGRAM - jagged lunar landscape might be useful (reuse of old one probably not sufficient)

% not sure how much detail should be gone into here? should just cite generation equations from old paper or copy over??

Each vertex of this landscape is generated sequentially. % INCOMPLETE due to uncertainty listed above

\subsubsection{Spaceship Physics}

The spaceship within this game world is modelled as an a circular mass with some basic physical properties, such as a position within the game world $\boldsymbol {s}$, a velocity $\boldsymbol {v}$, an orientation $\theta$, an angular velocity $\omega$ and a radius of a bounding sphere $r$ used for collision detection with the landscape.

This collision detection method is based on the lowest point of the spaceship's bounding circle coming into intersection with the highest point of the landscape for the line perpendicular between the ship and the bottom of the playing field, as can be seen in Figure \ref{fig:diagram_landscapepoint}.

% this needs a diagram for clarity: reuse the old one? reusing it temporarily
\begin{figure}[hbtp]
\centering
\includegraphics[scale=0.4]{graphics/landscapepoint}
\caption{The point closest to the spaceship on the landscape lies in between two of the defined vertices of the landscape, and requires interpolation to calculate the y co-ordinate from the x co-ordinate of the ship.}
\label{fig:diagram_landscapepoint}
\end{figure}

As the landscape is stored as a series of line segments, interpolation must be used in the event the ship is not perfectly aligned with one of the landscape axes. The point that lies on the line segment that the ship is being checked against is calculated through simple linear interpolation based on which two vertices the ship is horizontally closest to. For the ship's centre $\boldsymbol {s}$, the left nearest landscape vertex $\boldsymbol {p}^{l}$ and the right nearest landscape vertex $\boldsymbol {p}^{r}$, the point of collision $\boldsymbol {p}^{c}$ against the landscape is calculated as

% does this constitute self-plaigarism???

\begin{equation}
\boldsymbol {p}_{x}^{c} = \boldsymbol {s}_{x}
\boldsymbol {p}_{y}^{c} = \boldsymbol {p}_{y}^{l} + v(\boldsymbol {p}_{y}^{r} - \boldsymbol {p}_{y}^{l})
\end{equation}

where $v$ is a value between $0$ to $1$ used for interpolation, and can be calculated as follows.

\begin{equation}
v = \frac{ {\boldsymbol {s}_{x} - \boldsymbol {p}_{x}^{l}}}{ {\boldsymbol {p}_{x}^{r} - \boldsymbol {p}_{x}^{l}} }
\end{equation}

Collision is then true if the following statement is true:

\begin{equation}
\boldsymbol {s}_{y} + r \geq \boldsymbol {p}^{c}_{y}
\end{equation}

The ship colliding with the landscape constitutes the end of the {\itshape Lunar Lander} game. The conditions surrounding this collision, including speed, orientation of the ship, and fuel used, constitute whether the nature of the collision is a success or a failure.

\subsection{Continuous Action MCTS} % this can be renamed something much better presumably

\subsubsection{Heuristic} % this too

% TODO - meeting or discussion might be good to be entirely certain on what's going on in the code
% DISCUSS IN MEETING: use of a two-stage approach: might be good to mention this?? temporary holdover?????

\section{Results}

\begin{figure*}[ht]
  %\centering

  \centering
        \begin{subfigure}[b]{0.32\textwidth}
                \centering
                \includegraphics[width=1.0\textwidth]{graphics/online-IP-crop.pdf}
                \caption{Scores for a noise level of $l=1$ under variable function evaluations.\vspace{0.14in} }
                \label{fig:IP}
        \end{subfigure}   \begin{subfigure}[b]{0.32\textwidth}
                \centering
                \includegraphics[width=1.0\textwidth]{graphics/online-noise-IP_50-crop.pdf}
                \caption{Scores for 50 iterations under variable noise level $l$. Notice how UH-EOLOP fails to balance the pole.}
                \label{fig:IP-N-50}
        \end{subfigure}  \begin{subfigure}[b]{0.32\textwidth}
                \centering
                \includegraphics[width=1.0\textwidth]{graphics/online-noise-IP_100-crop.pdf}
                \caption{Score for 100 iterations under variable noise level $l$.\vspace{0.14in}}
                \label{fig:IP-N-100}
        \end{subfigure}%

         \caption{Performance of all algorithms in the Inverted Pendulum benchmark (error bars for the 95th percentile). Higher scores are better}
         \label{fig:IP-all}

\end{figure*}

        \begin{figure*}[ht]

        \begin{subfigure}[b]{0.32\textwidth}
                \centering
                \includegraphics[width=1.0\textwidth]{graphics/online-DI-crop.pdf}
                \caption{Scores for a noise level of 1 under variable function evaluations.}
                \label{fig:DI}
        \end{subfigure}   \begin{subfigure}[b]{0.32\textwidth}
                \centering
                \includegraphics[width=1.0\textwidth]{graphics/online-noise-DI_50-crop.pdf}
                  \caption{Scores for 50 iterations under variable noise level $l$. }
                \label{fig:DI-N-50}
        \end{subfigure}   \begin{subfigure}[b]{0.32\textwidth}
                \centering
                \includegraphics[width=1.0\textwidth]{graphics/online-noise-DI_100-crop.pdf}
                \caption{Score for 100 iterations an variable noise level $l$.}
                \label{fig:DI-N-100}
        \end{subfigure}

  \label{fig:DI-all}

       \caption{Performance of all algorithms in the Double Integrator benchmark (error bars for the 95th percentile). Higher scores are better}

\end{figure*}

\section{Conclusion}

% pending
% DISCUSS IN MEETING: this entire section

% Use \section* for the acknowledgments
\section*{Acknowledgements}
Samuel Roberts is supported by an EPSRC PhD Studentship.
% any other acknowledgements??

\bibliographystyle{IEEEtran}
\bibliography{lunarcontbib}
\end{document}

### 2014-03-30 lpaper/graphics/crop.sh

@@ -1,0 +1,3 @@
+find ./ -name "*.pdf" | while read FILE; do
+pdfcrop "$FILE"
+done;

### 2014-04-01 lpaper/lunarcont.tex
Closer then ever

@@
 			#1 & \mbox{if } #2 \\
-			#3 & \mbox{if } #4
+			#3 & \mbox{otherwise } 
 		\end{array}
@@
 % Paper title: keep the \ \\ \LARGE\bf in it to leave enough margin.
-\title{\ \\ \LARGE\bf Rolling Horizon methods for Games with Continuous Actions } 
+\title{\ \\ \LARGE\bf Rolling Horizon methods for Games with Continuous States and Actions } 
 % TODO - better title
@@
 \begin{abstract}
-It is often the case that games have continuous dynamics and allow for continuous actions.
-For larger games with complicated dynamics, having agents learn offline behaviours is a daunting task.
-On the other hand, provided a forward model is available, one might try to spread the cost of search in a rolling horizon fashion (e.g.
-as in example MCTS).
-In this paper we compare HOLOP, an open loop planning algorithm and a version evolutionary planning using CMA-ES.
-He show that rolling horizon CMA-ES outperforms HOLOP in two classic benchmark problems (Cart Balancing and <problem name> ) and Lunar Lander, a classic arcade game.
-We conclude that off-the-shelf evolutionary algorithms can be used successfully in a rolling horizon setting.
+It is often the case that games have continuous dynamics and allow for continuous actions, possibly with with some added noise.
+For larger games with complicated dynamics, having agents learn offline behaviours in such a setting is a daunting task.
+On the other hand, provided a forward model is available, one might try to spread the cost of search/learning in a rolling horizon fashion (e.g.
+as in Monte Carlo Tree Search).
+In this paper we compare T-HOLOP (Truncated Hierarchical Open Loop Planning), an open loop planning algorithm at least partially inspired by MCTS with a version evolutionary planning using CMA-ES (which we call EVO-P) in two planning benchmark problems (Cart Balancing and the Double Integrator) and Lunar Lander, a classic arcade game.
+He show that EVO-P outperforms T-HOLOP in almost all settings.
+We conclude that off-the-shelf evolutionary algorithms can be used successfully in a rolling horizon setting, and might offer considerable advantages over more formal methods.
 
@@
 HOLOP is based partially on a strong~\cite{hoo2011} real-valued optimisation algorithm called HOO (for ``Hierarchical Optimistic Optimisation'').
-In this paper two major contributions are made plus a smaller ``add-on''; the first contribution is the comparison of two versions of a very popular and strong evolutionary strategy: Covariance Matrix Adaptation Evolutionary Strategy (CMA-ES)~\cite{hansen2003reducing} and its Uncertainty Handling version UH-CMA-ES~\cite{hansen2009tec} with Truncated Hierarchical Optimistic Optimisation (T-HOO)~\cite{hoo2011} on the
-Black Box Optimisation Problem (BBOP) benchmark suit~\cite{bbopnoisy}.
-T-HOO is a faster version of HOO~\cite{hoo2011}, an algorithm reminiscent of Monte Carlo Tree Search.
-The experiments are conducted using similar conditions to those an agent would face if it was to solve planning problems.
-Our second contribution is to incorporate CMA-ES and UH-CMA-ES into three versions of an algorithm called EOLOP, for Evolutionary Open Loop Optimistic Planning.
-This algorithm uses evolution instead of tree search for implementing sample-based control algorithms.
-We then proceed to test the algorithms proposed on noisy versions of standard reinforcement learning (RL)
-benchmarks, namely the inverted pendulum and the double integrator problems.
-Our ``extra'' contribution is an add-on for evolutionary algorithms (transfered from HOLOP) that can help when the number of function evaluations is very small.
+In this paper two contributions are made; the first contribution is the comparison of two versions of a very popular and strong evolutionary strategy: EVO-P, the open loop planning algorithm using Covariance Matrix Adaptation Evolutionary Strategy (CMA-ES)~\cite{hansen2003reducing} with a version of HOLOP, which stands for Hierarchical Open Loop Optimistic Planning~\cite{hoo2011} on two standard benchmark problems, the double integrator and the reverse pendulum.
+For speed purposed, we use T-HOO (Truncated Hierarchical Optimistic Optimisation), a faster version of HOO~\cite{hoo2011}, as the engine of HOLOP, hence we name the algorithm used T-HOLOP.
+Our second contribution is the use of the these algorithms in Lunar Lander, a popular arcade game, coupled with a strong two-stage heuristic.
+We show that while EVO-P performs really well, T-HOLOP struggles.
 
@@
 prior learning.  The disadvantage, compared to neuro-evolution, is that every action performed requires
-CPU time for the simulations.  This is described in more detail in section \ref{Planning}.
+CPU time for the simulations.
+This is described in more detail in section \ref{sec:Planning} and in line with what is being done in Perez et al.
+\cite{perez2013rolling}, although this time in a continuous setting.
 
 The rest of this paper is structured as follows.
-Section \ref{Optimisation} discusses the relevant base algorithms that are going to be used in this paper, ((Truncated) Hierarchical Online Optimisation~\cite{hoo2011} and CMA-ES~\cite{hansen2003reducing}).
-Section~\ref{Planning} explains why and how the above algorithms can be used in the context of planning.
-Section \ref{evo} describes how one can incorporate evolution into this, making this our main contribution section from an algorithmic standpoint.
-In Section \ref{Methodology} the experimental setup is described.
+Section \ref{sec:Optimisation} discusses (briefly and epigrammatically) the relevant base algorithms that are going to be used in this paper, ((Truncated) Hierarchical Online Optimisation~\cite{hoo2011} and CMA-ES~\cite{hansen2003reducing}).
+Section~\ref{sec:Planning} explains why and how the above algorithms can be used in the context of planning (and thus transformed into HOLOP and EVO-P respectively).
+In Section \ref{sec:methodology} the experimental setup is described.
 In Section~\ref{sec:Experiments} a number of experiments are portrayed and analysed.
 These form the bulk of this paper's contribution.
-We conclude with a short discussion in Section \ref{Conclusion}.
+We conclude with a short discussion in Section \ref{sec:conclusion}.
 
-% TODO - actually determine what the core essence of what it is we're doing here before writing the introduction
+\section{Algorithms for Continuous Function Optimisation} \label{sec:Optimisation}
 
-\section{Motivation}
+This is the first background section of this paper, which aims to introduce the two core optimisation algorithms on which our contributions are made.
+We will present both algorithms from an implementer's viewpoint, without explicitly justifying the design and mathematical choices behind them, nor going into too much detail.
+Interested readers should look at the cited bibliography for extensive discussions.
 
-\subsection{Solving Continuous Problems with Continuous Actions} % this title is awful and it needs changing
+In general, the optimisation problem can be defined as: Given a function $f:\bm{X}\rightarrow Y $, where $X \in \mathcal{R}^n$, with $n$ being the length of vector $\bm{X}$.
+The domain $\bm{X}$ is usually called the search space.
+Our goal is to find a vector $\bm{x}_0$, for which $f(\bm{x}) > f(\bm{x_0}) $ for all $x$ (also known as ``minimisation'') or $f(\bm{x}) < f(\bm{x_0})$ for all $x$ (known as ``maximisation'').
+There are an enormous number of algorithms devoted to this problem, but only two will be presented and compared here, which are both considered state of the art in their respective fields and do not require the calculation of a gradient.
+If one has an analytic expression of the gradient or if the functions proposed are smooth enough, it might be better to look elsewhere for solutions (e.g.
+Conjugate Gradient Methods~\cite{shewchuk1994introduction}).
 
-Previous research into using MCTS to provide solutions for continuous physics-based problems have examined cases such as the Physical Travelling Salesman Problem \cite{perez14}.
-Within these continuous contexts, discrete actions have proven to be quite effective, especially with use of macro-actions and some degree of higher level planning \cite{powley12}.
-% this section seems clunky
+\subsection{CMA-ES} \label{sec:algCma}
 
-{\itshape Lunar Lander} is an excellent test case to study for many reasons, due to its inherently multi-objective nature.
-With the objectives to both land quickly while also minimising velocity changes, which cost fuel which must be conserved, it has been an interesting scenario to examine the properties and constraints imposed just by the environment alone.
-While there have been attempts to solve this deceptively complex problem before using discrete macro-actions \cite{roberts13}, in this paper the problem is being approached as one that can be better solved with continuous actions.
+The Covariance Matrix Adaptation - Evolutionary Strategy (CMA-ES) is an evolutionary algorithm designed for continuous domains, specially suited for non-linear and non-convex optimization problems~\cite{Hansen2006}.
+In general, this algorithm is applied to those problems that are not constrained and are made of up to $100$ dimensions.%~\footnote{\url{http://www.lri.fr/~hansen/cmatutorial-crop-crop.pdf}}
 
-\section{Planning}  \label{Planning}
+A multivariate normal distribution (MND) is a generalization of the univariate normal distribution, which is a probability distribution that has a bell-shaped probability density function (Gaussian function).
+A multivariate vector $X = (x_0, x_1, \dots x_N), x_i \in \mathcal{R}$ is said to have an MND if it satisfies that any linear combination from its components, $w_0x_0 + w_1x_1 + \dots + w_Nx_N$, is normally distributed.
+
+CMA-ES creates a population of individuals by sampling from an MND: $\mathcal{N}(m,C)$, which is uniquely defined by the distribution mean $m \in \mathcal{R}^{n}$ and its covariance matrix $C \in \mathcal{R}^{n \times n}$.
+The top of the density function, which corresponds to $m$, also determines the translation of the distribution.
+The covariance matrix $C$, positive definite and symmetric, determines the shape of the distribution and its graphical interpretation: it defines an iso-density ellipsoid $\{ x_i \in \mathcal{R} | (x - m)^T C^{-1} (x - m) = 1\}$.
+Through the classic evolutionary setup and by making clever use of population statistics, the algorithm proceeds from generation to generation of sampling until some convergence criteria has been reached.
+
+\subsection{T-HOO}
+%$\bm{x^{C_l}_{\mathit{min}}} = \bm{x^{C}_{\mathit{min}}} $,
+Recently, a family of algorithms based on discretizing the search domain has been proposed.
+Only one of these algorithms is going to be discussed here, Truncated Hierarchical Optimistic Optimisation (T-HOO)~\cite{hoo2011}.
+The T-HOO algorithm was built from the ground up to perform optimisation of noisy functions.
+It is an iterative, anytime algorithm, which is run for $n_0$ iterations.
+The algorithm is based on a number of components.
+The first one is a data structure called a node and designated $\nu$.
+The node includes the maximum and minimum values of our search space, vectors $\bm{x_{\mathit{max}}}$ and $\bm{x_{\mathit{min}}}$ respectively.
+The node also has a property called $\mathit{count}$, which initially is set to zero and a property called $\mathit{sum}$, which, again, is initially set to zero.
+An unvisited node (i.e.
+$\mathit{count}=0$) is called a leaf node.
+``Splitting'' a node is the procedure of creating two children $C(\nu) = \{C_l(\nu),C_r(\nu)\}$.
+Both children are initialised $\bm{x_{\mathit{max}}}$ and $\bm{x_{\mathit{min}}}$ to their parents' nodes values.
+When there are multiple dimensions to choose from, at  at $k$ according to some random distribution.
+Thus, for child $C_l(\nu)$ ,   $\bm{x^{C_l}_{\mathit{max,k}}} =\bm{x^{C}_{\mathit{min,k}}} +  \left(\bm{x^{C}_{\mathit{min,k}}} - \bm{x^{C}_{\mathit{max,k}}}\right)/2 $.
+For the other child,  $\bm{x^{C_r}_{\mathit{min,k}}} = \bm{x^{C}_{\mathit{min,k}}} +  \left(\bm{x^{C}_{\mathit{min,k}}} - \bm{x^{C}_{\mathit{max,k}}}\right)/2 $.
+In other words, the father node is split into two children nodes, choosing which dimension to split uniform randomly (or according to some other criterion).
+Each node also has an associated value called $B(\nu)$, which is defined recursively as:  $B(\nu) = min \{U(\nu),max \{B(C_l(\nu)),B(C_r(\nu))\}\}$.
+Part of this equation is the term $U(\nu)$, which is defined as in Equation \ref{eq:U}:
+
+\begin{equation}
+U(\nu) = S(\nu)/N(\nu) + \sqrt{2ln n_0/N(\nu)} + \upsilon_1 \rho^h
+\label{eq:U}
+\end{equation}
+
+If a dissimilarity metric is defined between different values of $\bm{X}$ as ${\lVert \bm{x}_1 - \bm{x}_2) \rVert}^\alpha$, one can set $\upsilon_1 =\sqrt{|A|}/2)^\alpha$ and $\rho = 2^{-\alpha/|A|}$.
+In our experiments $a = 2$.
+If $N(\nu) = 0$, $B(\nu) = U(\nu) = \infty$, i.e.
+unvisited nodes have priority.
+If a node has not been visited then it is called a leaf node.
+If a node is a leaf node, one can ``draw'' a sample from it, by uniformly randomly sampling between $\bm{x_{\mathit{max}}}$ and $\bm{x_{\mathit{min}}}$.
+
+The algorithm works by starting at the root node (the initial node).
+Since the leaf node and the root node coincide, HOOT samples from it and splits it.
+It can now proceed to the second iteration, where the algorithm follows a route using the maximum B-value of its children nodes, which it then samples and splits.
+The process repeats until the maximum number of iterations $n_0$ is over.
+If a certain depth is exceeded there is no point splitting nodes any more (given our iteration budget $n_0$).
+
+\section{Planning}  \label{sec:Planning}
 In this section a formalised version of planing will be presented and an explanation provided on how the algorithms described in the previous section can be used to attack the problem.
@@
 \item $A$ is a set of actions, each action named $a_j$.
-\item  $T:S \times \S \times  A  \rightarrow [0,1]$ is the probability of moving from state $s$ to state $s'$  after action $a \in A$ has taken place. $T(s'|s,a)$ denotes this probability.
+\item  $T:S \times S \times  A  \rightarrow [0,1]$ is the probability of moving from state $s$ to state $s'$  after action $a \in A$ has taken place. $T(s'|s,a)$ denotes this probability.
 \item  $R:S \rightarrow \Re$, $R(s)$ is a reward function at each state.
@@
 A possible route of action an agent might take is known as the \textit{policy} $\pi$.
-A policy is a probabilistic mapping between state and actions, $\pi:S \times A \rightarrow [0,1]$, thus $\pi(s,a) \in [0,1]$,  $\sum_{a\in A}{\pi(s,a)} = 1$.
+A policy is a probabilistic mapping between state and actions, $\pi:S \times A \rightarrow [0,1]$, thus $\pi(s,a) \in [0,1]$,  $\sum\limits_{a\in A}{\pi(s,a)} = 1$.
 The set of all policies is denoted as $\Pi$.
-The goal of an agent in an MDP environment is to maximise its long term value $V(\pi,s_0) = R(s) + \sum\limits_{s\in S}\sum\limits_{a\in A} \gamma \pi(s,a)T(s^\prime|s,a)  V(\pi,s^\prime) $, by assigning the probability of actions to take at each state.
+The goal of an agent in an MDP environment is to maximise its long term value.
+Note that since here we are dealing with continuous actions and states, sets $A$ and $S$ are metric (and infinitive).
+Also note that we are NOT dealing with continuous time MDPs in this paper, although there is a notion of time.
 
@@
 
-Since an open loop policy is now a real-valued vector of actions, algorithms like T-HOO and CMA-ES can be used to find such a policy.
-Hence the title of this paper:  Rolling Horizon Open-loop Policy Evolutionary Learning.
+Since an open loop policy is now a real-valued vector of actions, algorithms like T-HOO and CMA-ES can be used to find such a policy, hence the title of this paper.
 To the best of our knowledge this is the first time evolution has been used in such a setting.
@@
 
-\subsection{Planning using a Generative Model}
+\subsection{Notes on planning using a Generative Model}
 
@@
 What changes from method to method is how the plan is formed.
+The main version of the algorithm used can be seen in Algorithm \ref{algo:ooop}.
 
-In standard UCT, a procedure similar to temporal difference learning is followed, where one learns the value of each state.
+In standard UCT, a procedure similar to temporal difference learning is followed (or, more closely, monte carlo methods), where one learns the value of each state.
 An agent then forms a policy by choosing a value greedily.
@@
 
-\State 	\textbf{define external function} $\mathsc{FindOpenLoopPlan} (depth) $ \Comment{Try to find the optimal policy. You can use either evolution or HOO here}
+\State 	\textbf{define external function} $\mathsc{FindOpenLoopPlan} (depth, T',S') $ \Comment{Try to find the optimal policy. You can use either evolution or T-HOO here}
 
-	\Procedure{$\mathsc{OOOP}$} {}
-		\While{$\mathit{True}$}
-		    \State $a \gets \mathsc{FindOpenLoopPlan}(depth)[0]$ \Comment{Execute first action of the policy}
+	\Procedure{$\mathsc{OOP}$} {}
+		\While{$!StoppingCondition$}
+		    \State $a \gets \mathsc{FindOpenLoopPlan}(depth, T',S')[0]$ \Comment{Execute first action of the policy}
 		    \State $s \sim T(s'|s,a)$ \Comment{Stop only at an absorbing state}
@@
 
-\section{Methodology}
+\section{Methodology} \label{sec:methodology}
+Our methodology is simple, yet efficient. We test T-HOLOP and EVO-P in two simple benchmark problems (see below for more) and Lunar Lander, a classic arcade game.
 
@@
 
-\subsection{Environment and Physics}
-
-% this is a heavily cut down rewritten version of the physics described in sam's 2013 cig paper
-% more detail will be added depending on if it is appropriate or not
+\subsection{Lunar Lander}
+We describe the properties of Lunar Lander.
 
@@
 
-\begin{equation}
-\boldsymbol {p}_{x}^{c} = \boldsymbol {s}_{x}
-\boldsymbol {p}_{y}^{c} = \boldsymbol {p}_{y}^{l} + v(\boldsymbol {p}_{y}^{r} - \boldsymbol {p}_{y}^{l})
-\end{equation}
-
-where $v$ is a value between $0$ to $1$ used for interpolation, and can be calculated as follows.
-
-\begin{equation}
-v = \frac{ {\boldsymbol {s}_{x} - \boldsymbol {p}_{x}^{l}}}{ {\boldsymbol {p}_{x}^{r} - \boldsymbol {p}_{x}^{l}} }
-\end{equation}
-
-Collision is then true if the following statement is true:
-
-\begin{equation}
-\boldsymbol {s}_{y} + r \geq \boldsymbol {p}^{c}_{y}
-\end{equation}
-
-The ship colliding with the landscape constitutes the end of the {\itshape Lunar Lander} game.
-The conditions surrounding this collision, including speed, orientation of the ship, and fuel used, constitute whether the nature of the collision is a success or a failure.
-
-\subsection{Continuous Action MCTS} % this can be renamed something much better presumably
-
-\subsubsection{Heuristic} % this too
-
-% TODO - meeting or discussion might be good to be entirely certain on what's going on in the code
-% DISCUSS IN MEETING: use of a two-stage approach: might be good to mention this?? temporary holdover?????
-
-\section{Results}
-
 \begin{figure*}[ht]
@@
                 \includegraphics[width=1.0\textwidth]{graphics/online-noise-IP_50-crop.pdf}
-                \caption{Scores for 50 iterations under variable noise level $l$. Notice how UH-EOLOP fails to balance the pole.}
+                \caption{Scores for 50 iterations under variable noise level $l$.}
                 \label{fig:IP-N-50}
@@
 
-\section{Conclusion}
+\begin{equation}
+\boldsymbol {p}_{x}^{c} = \boldsymbol {s}_{x}
+\boldsymbol {p}_{y}^{c} = \boldsymbol {p}_{y}^{l} + v(\boldsymbol {p}_{y}^{r} - \boldsymbol {p}_{y}^{l})
+\end{equation}
 
-% pending
-% DISCUSS IN MEETING: this entire section
+where $v$ is a value between $0$ to $1$ used for interpolation, and can be calculated as follows.
+
+\begin{equation}
+v = \frac{ {\boldsymbol {s}_{x} - \boldsymbol {p}_{x}^{l}}}{ {\boldsymbol {p}_{x}^{r} - \boldsymbol {p}_{x}^{l}} }
+\end{equation}
+
+Collision is then true if the following statement is true:
+
+\begin{equation}
+\boldsymbol {s}_{y} + r \geq \boldsymbol {p}^{c}_{y}
+\end{equation}
+
+The ship colliding with the landscape constitutes the end of the {\itshape Lunar Lander} game.
+The conditions surrounding this collision, including speed, orientation of the ship, and fuel used, constitute whether the nature of the collision is a success or a failure.
+
+An agent can modify the ship directions by applying linear thrust $\bm{a_T}$ and the angular impulse $\bm{a_I}$.
+The controllers have to identify a vector of real numbers $\bm{A}$, which each element of the set $a_i \in [0,1]$.
+Each even element of the vector represents the linear thrust, whereas each odd element represents angular velocity.
+Each linear thrust vector element is transformed by applying as follows:
+
+\begin{equation}
+\bm{a_I} = 5(\bm{a_I} - 0.5)
+\bm{a_T} = 300\bm{a_T} 
+\end{equation}
+
+Finally each linear thrust and angular impulse is repeated for $X$ actions, the effects of which we are going to explore below.
+
+\subsubsection{Heuristic} % this too
+
+Lunar Lander has a setup where a player gets rewards if and only if it actually lands on the pad.
+This creates the problem of rollouts (whether used by EVO-P or T-HOLOP) rarely will see a reward.
+In order to circumvent this, as it is common in other games, we use a heuristic for each state.
+In order to counter this tendency, as is common in games, we used a two-stage heuristic.
+The first stage involves the ship trying to reach the landing pad.
+
+\begin{multline}
+rng = -\sum{arccos(vl\cdot\theta)/||vl||*||\theta||} - |\boldsymbol {v_y} - V_{y}^{max}| \\
+h = \twopartdef{ -|(|\boldsymbol {v}| - C) - |\boldsymbol {s}_{x} - \boldsymbol {np}_{x}||}{||\boldsymbol {s} - \boldsymbol {np}|| < D}
+{rng - |\boldsymbol {v_x}|   }{Otherwise}
+\label{eq:h1}
+\end{multline}
+
+In the above Equation \ref{eq:h1} the constant C denotes the maximum magnitute of speed, which we set to $C=100$.
+$D = 60$ is the distance from the landing point $np$ where this heuristic takes precedence.
+Intuitively, the upper part of the heuristic $h$ tries to get the ship above the landing point as fast as possible while maintaining a certain speed.
+The second part of the heuristic tries to minimise velocity on the $x$ axis (as the ship tries to land), maintain a minimum landing speed on the $y$ axis and get a certain angle (the summation term $rng$) as close to vector $vl = (0,-1)$, i.e.
+help the ship stay upright through the duration of the rollout.
+
+\section{Results} \label{sec:Experiments}
+We present the results of our experiments below.
+
+\subsection{Benchmark Problems}
+
+For the second set of experiments we have defined the following variables.
+The first two involve running the experiments with a fixed rollout depth of 50, i.e.
+each algorithm at each step plans for 50 steps ahead.
+We set the noise levels to $0.1 (\mathcal{U}(0,1)-0.5)*2 l$ uniform noise for the double integrator experiment and $10 (\mathcal{U}(0,1)-0.5)*2 l$  for the Inverted Pendulum.
+The variable $l$ is the noise level which defines how noisy the environment is.
+Initially, we performed two experiments, with a variable number of iterations and a fixed noise level of one.
+
+An example of how the \textit{predicted} reward looks like for a $2$-step look-ahead (instead of $50$ we use in the experiments) can be seen in Figure \ref{fig:vis}.
+We plotted the full state-space by pushing the internal generative model two steps forward for all possible actions by discretizing each action.
+Notice how the predicted reward landscape changes over time as the agent moves.
+The continuous replanning should aim to perform the action with the highest future reward.
+
+The results of these experiments can be seen in graphs  \ref{fig:IP} and \ref{fig:DI}.
+These are close to the original experiments of Weinstein and Littman~\cite{weinstein2012bandit}, only this time we varied the number
+of iterations allowed for each algorithm.
+The original experiment had given the algorithms 200 value iterations.
+With this number of iterations, for the Inverted Pendulum experiment the mean reported (using HOLOP) was $-47.45$.
+We got a better mean using T-HOLOP
+% % \footnote{We believe there is a minor bug in the HOLOP implementation Weinstein et al use.}
+\footnote{This was due to a minor difference between our implementation of HOLOP and Weinstein's.}
+ of $-42.94$.
+
+ EVO-P got as score of $-41.69$.
+It also  seems to have a drop in performance as we increase function evaluations, presumably because it converges to a false minima without any uncertainty handling mechanism to stop this.
+This kind of behaviour has been observed previously in Evolutionary Algorithms~\cite{hansen2009tec} when having noisy function.
+
+For the double integrator experiment, we had slightly different results.
+The Weinstein paper reports a mean of $-2.72$ for 200 iterations.
+Using T-HOLOP we got a better mean of $-2.62$.
+A far better result of $-1.74$ is obtained by EVO-P.
+
+Results for the variable noise levels $l$ (Figures \ref{fig:IP-N-50}, \ref{fig:IP-N-100}, \ref{fig:DI-N-50}, \ref{fig:DI-N-100}) closely mimic the performance of their constant noise counterparts.
+For the inverted pendulum, using just 50 iterations (Figure \ref{fig:IP-N-50}) EVO-P is able to stabilise the pendulum with .
+This still applies if we increase the number of function evaluations to 100 (Figure \ref{fig:DI-N-100}).
+On the other hand, the results start to get closer to the fixed noise level experiments if we increase the number of iterations in the Inverted Pendulum experiment to 100 (Figure \ref{fig:IP-N-50}).
+
+\section{Conclusion} \label{sec:conclusion}
 
@@
 \section*{Acknowledgements}
-Samuel Roberts is supported by an EPSRC PhD Studentship.
-% any other acknowledgements??
+This work was supported by EPSRC grant EP/H048588/1 entitled:``UCT for Games and Beyond'' and an EPSRC Ph.D studentship.
 

### 2014-04-01 code/src/lunarcode/LunarLanderEvoCont.java

@@ -23,19 +23,19 @@ public class LunarLanderEvoCont {
 
         LunarLanderObjective fitfun = new LunarLanderObjective((LunarGame) copy.getCopy());
 
-        if(solution!=null) {
-            double[] tbr = fitfun.convertInputs(solution[s_i],solution[s_i+1]);
-            macro_counter+=1;
-            if(macro_counter >= fitfun.MACRO_LENGTH) {
-
-                s_i+=2;
-                macro_counter = 0;
-            }
-            return tbr;
-        }
+//        if(solution!=null) {
+//            double[] tbr = fitfun.convertInputs(solution[s_i],solution[s_i+1]);
+//            macro_counter+=1;
+//            if(macro_counter >= fitfun.MACRO_LENGTH) {
+//
+//                s_i+=2;
+//                macro_counter = 0;
+//            }
+//            return tbr;
+//        }
 
         CMAEvolutionStrategy cma = new CMAEvolutionStrategy();
-        cma.setDimension(8); // overwrite some loaded properties
+        cma.setDimension(9); // overwrite some loaded properties
         //cma.parameters.setPopulationSize(10);
         cma.setInitialX(0.5); // in each dimension, also setTypicalX can be used
         cma.setInitialStandardDeviation(0.2);
@@ -79,7 +79,7 @@ public class LunarLanderEvoCont {
             macro_counter = 1;
 
         }
-        System.out.println("LunarLanderObjective.MACRO_LENGTH = " + LunarLanderObjective.MACRO_LENGTH);
+        //System.out.println("LunarLanderObjective.MACRO_LENGTH = " + LunarLanderObjective.MACRO_LENGTH);
         System.out.println("cma.getBestFunctionValue() = " + cma.getBestFunctionValue());
         //LunarGame g = (LunarGame) copy.getCopy();
         //g.tickCont(tbr[0],tbr[1]);

### 2014-04-02 lpaper/lunarcont.tex
one more step

@@
                 \includegraphics[width=1.0\textwidth]{graphics/online-IP-crop.pdf}
-                \caption{Scores for a noise level of $l=1$ under variable function evaluations.\vspace{0.14in} }
+                \caption{Scores for a noise level of $l=1$ under variable function evaluations. }
                 \label{fig:IP}
@@
                 \includegraphics[width=1.0\textwidth]{graphics/online-noise-IP_100-crop.pdf}
-                \caption{Score for 100 iterations under variable noise level $l$.\vspace{0.14in}}
+                \caption{Score for 100 iterations under variable noise level $l$.}
                 \label{fig:IP-N-100}
@@
 
-         \caption{Performance of all algorithms in the Inverted Pendulum benchmark (error bars for the 95th percentile). Higher scores are better}
+         \caption{Performance of both T-HOLOP and EVO-P in the Inverted Pendulum benchmark (error bars for the 95th percentile). Higher scores are better}
          \label{fig:IP-all}
@@
 
+\begin{figure*}[ht]
+  %\centering
+
+  \centering
+        \begin{subfigure}{0.49\textwidth}
+                \centering
+                \includegraphics[width=1.0\textwidth]{./lunarexps/sample-crop.pdf}
+                \caption{Best value for each tick EVO-P could find. }
+                \label{fig:sample}
+        \end{subfigure}  
+        \begin{subfigure}{0.49\textwidth}
+                \centering
+                \includegraphics[width=1.0\textwidth]{./lunarexps/sample-detail-crop.pdf}
+                \caption{Detail of Figure \ref{fig:sample}, when the ship is closer to landing.}
+                \label{fig:sample-detail}
+
+        \end{subfigure}
+
+         \caption{Sample run for EVO-P. Best fitness discovered at each tick is shown. Notice the big ``bump'' in fitness when the ship gets close to the landing pad}
+         \label{fig:samples}
+
+\end{figure*}
+
 \subsection{Benchmark Problems}

### 2014-04-02 code/src/lunarcode/LunarLanderEvoCont.java

@@ -39,6 +39,7 @@ public class LunarLanderEvoCont {
         //cma.parameters.setPopulationSize(10);
         cma.setInitialX(0.5); // in each dimension, also setTypicalX can be used
         cma.setInitialStandardDeviation(0.2);
+        cma.options.verbosity = -1;
 
         double[] fitness = cma.init();
 
@@ -66,8 +67,8 @@ public class LunarLanderEvoCont {
         double[] best = cma.getBestX().clone();
 
 
-        System.out.println("best" + Arrays.toString(best));
-        System.out.println(cma.getBestFunctionValue() + " " + cma.getCountEval()) ;
+        //System.out.println("best" + Arrays.toString(best));
+        //System.out.println(cma.getBestFunctionValue() + " " + cma.getCountEval()) ;
         //System.exit(0);
         double[] tbr = fitfun.convertInputs(best[0],best[1]);
         
@@ -80,13 +81,13 @@ public class LunarLanderEvoCont {
 
         }
         //System.out.println("LunarLanderObjective.MACRO_LENGTH = " + LunarLanderObjective.MACRO_LENGTH);
-        System.out.println("cma.getBestFunctionValue() = " + cma.getBestFunctionValue());
+        //System.out.println(cma.getBestFunctionValue());
         //LunarGame g = (LunarGame) copy.getCopy();
         //g.tickCont(tbr[0],tbr[1]);
         //System.out.println(g.getShip().v.mag() + " sdfdf")      ;
 
 
-        System.out.println("tbtr" + Arrays.toString(tbr));
+        //System.out.println("tbtr" + Arrays.toString(tbr));
 
         return tbr;
 

### 2014-04-02 code/src/lunarcode/LunarLanderObjective.java

@@ -93,7 +93,7 @@ public class LunarLanderObjective implements IObjectiveFunction {
 
         if(phase == 0) {
             double xdist = Math.abs(cGame.getShip().s.x - nearestPad.x);
-            double ydist = Math.abs(cGame.getShip().s.y - nearestPad.y);
+            //double ydist = Math.abs(cGame.getShip().s.y - nearestPad.y);
             Vector2d speed = cGame.getShip().v;
             score = -Math.abs(speed.mag() - 100) - xdist;
             //score += getScorePhaseA(game, cGame, nearestPad, 2, 50);

### 2014-04-02 code/src/lunarcode/LunarRun.java

@@ -8,6 +8,11 @@ import framework.utils.ElapsedCpuTimer;
 import framework.utils.JEasyFrame;
 import framework.utils.Vector2d;
 
+import java.io.BufferedWriter;
+import java.io.File;
+import java.io.FileWriter;
+import java.io.IOException;
+
 /**
  * Created by Samuel Roberts, 2013
  */
@@ -21,7 +26,9 @@ public class LunarRun {
     //public static LunarLanderHOOCont cont;
 
     public static void main(String[] args) {
-        runGame();
+        //runGame();
+        runGames(1);
+
     }
 
     public static void prepareGame() {
@@ -32,11 +39,10 @@ public class LunarRun {
 
     public static void runGame() {
         prepareGame();
-        LunarView view = new LunarView((LunarGame)game, (LunarShip)game.getShip());
+        LunarView view = new LunarView((LunarGame) game, (LunarShip) game.getShip());
         JEasyFrame frame = new JEasyFrame(view, "Lunar Lander Cont");
 
-        while(!game.isEnded())
-        {
+        while (!game.isEnded()) {
 
             ElapsedCpuTimer ect = new ElapsedCpuTimer(ElapsedCpuTimer.TimerType.CPU_TIME);
             ect.setMaxTimeMillis(PTSPConstants.ACTION_TIME_MS);
@@ -64,63 +70,71 @@ public class LunarRun {
         System.out.println("Ship landed with velocity " + ((LunarShip) game.getShip()).v.mag() + "Victory?" + ((LunarGame) game).landedSuccessfully);
         //System.exit(0);
     }
+//
+//
+
+
+    public static void runGames(int trials) {
+        //Prepare the average results.
+
+
+        BufferedWriter writer = null;
+        try {
+
+            File logFile = new File("/home/ssamot/projects/MultiObjectiveMCTS/lpaper/lunarexps/noise_statistics.csv");
+            logFile.createNewFile();
 
 
 
 
+            //Close writer
+
+        } catch (Exception e) {
+            e.printStackTrace();
+        }
 
 
-//    public static void runGames(int trials) {
-//        //Prepare the average results.
-//
-//
-//        StatSummary ssWp = new StatSummary();
-//        StatSummary ssTime = new StatSummary();
-//        StatSummary ssFuel = new StatSummary();
-//
-//
-//        int numGamesPlayed = 0;
-//
-//
-//        //For each trial...
-//        for(int i=0;i<trials;i++)
-//        {
[…]

### 2014-04-02 lpaper/lunarexps/sample.py

@@ -1,0 +1,34 @@
+import numpy as np
+from matplotlib import cm
+import matplotlib.pyplot as plt
+from matplotlib.mlab import griddata
+import numpy as np
+
+ar = np.loadtxt(open("samplerun.csv","rb"),delimiter=",")
+print len(ar)
+fig = plt.figure()
+
+sample = plt.plot(ar, color = "b" ,label = "Sample Run")
+
+plt.xlabel('Game Tick')
+plt.ylabel('Best CMA score')
+#fig.set_size_inches(18.5,10.5)
+
+plt.savefig("sample.pdf")
+
+plt.hold(False)
+
+i = 650
+
+x = np.arange(i,len(ar),1)
+
+print len(ar[i:]), len(x)
+
+sample = plt.plot(x, ar[i:], color = "b" ,label = "Sample Run - Detail")
+#fig.set_size_inches(18.5,10.5)
+
+plt.xlabel('Game Tick')
+plt.ylabel('Best CMA score')
+
+plt.savefig("sample-detail.pdf")
+

### 2014-04-07 lpaper/lunarcont.tex
Experiments done, a bit more effort needed to cleanup

@@
 
+\begin{table}
+	\begin{tabular}{lrrrr}
+\hline
+ Map Name   &   Success Mean &   Success CI &   Ticks Mean &   Tick CI \\
+\hline
+ map2       &            1   &     0        &       2925.9 &   184.437 \\
+ map1-tall  &            1   &     0        &       2733   &   159.34  \\
+ map1-easy  &            1   &     0        &       2293   &   123.917 \\
+ map1-wall  &            0.9 &     0.226216 &       2421.6 &   682.893 \\
+\hline
+\end{tabular}
+\end{table}
+
+% \begin{figure*}[ht]
+%   %\centering
+
+%   \centering
+%         \begin{subfigure}{0.49\textwidth}
+%                 \centering
+%                 \includegraphics[width=1.0\textwidth]{./lunarexps/sample-crop.pdf}
+%                 \caption{Best value for each tick EVO-P could find. }
+%                 \label{fig:sample}
+%         \end{subfigure}  
+%         \begin{subfigure}{0.49\textwidth}
+%                 \centering
+%                 \includegraphics[width=1.0\textwidth]{./lunarexps/sample-detail-crop.pdf}
+%                 \caption{Detail of Figure \ref{fig:sample}, when the ship is closer to landing.}
+%                 \label{fig:sample-detail}
+
+%         \end{subfigure}
+
+%          \caption{Sample run for EVO-P. Best fitness discovered at each tick is shown. Notice the big ``bump'' in fitness when the ship gets close to the landing pad}
+%          \label{fig:samples}
+
+% \end{figure*}
+
 \begin{figure*}[ht]
-  %\centering
+  \centering
 
@@
                 \centering
-                \includegraphics[width=1.0\textwidth]{./lunarexps/sample-crop.pdf}
-                \caption{Best value for each tick EVO-P could find. }
-                \label{fig:sample}
+                \includegraphics[width=1.0\textwidth]{./graphics/map2.pdf}
+                \caption{Map. }
+                \label{fig:map2}
         \end{subfigure}  
@@
                 \centering
-                \includegraphics[width=1.0\textwidth]{./lunarexps/sample-detail-crop.pdf}
+                \includegraphics[width=1.0\textwidth]{./graphics/map1-easy.pdf}
+                \caption{sdfdf}
+                \label{fig:map1-easy}
+
+        \end{subfigure}
+\\
+        \begin{subfigure}{0.49\textwidth}
+                \centering
+                \includegraphics[width=1.0\textwidth]{./graphics/map1-tall.pdf}
                 \caption{Detail of Figure \ref{fig:sample}, when the ship is closer to landing.}
-                \label{fig:sample-detail}
+                \label{fig:map1-tall}\end{subfigure}\begin{subfigure}{0.49\textwidth}
+                \centering
+                \includegraphics[width=1.0\textwidth]{./graphics/map1-wall.pdf}
+                \caption{Detail of Figure \ref{fig:sample}, when the ship is closer to landing.}
+                \label{fig:map-wall}
 
@@
          \caption{Sample run for EVO-P. Best fitness discovered at each tick is shown. Notice the big ``bump'' in fitness when the ship gets close to the landing pad}
-         \label{fig:samples}
+         \label{fig:maps}
 

### 2014-04-07 code/exp.sh

@@ -1,0 +1,7 @@
+#!/bin/bash
+cd out/production/code
+for i in {1..10}
+do
+	java lunarcode.LunarRun
+   #echo "Welcome $i times"
+done

### 2014-04-07 code/src/lunarcode/HOOClass.java

@@ -33,10 +33,10 @@ public class HOOClass extends ContinuousProblem {
        LunarLanderObjective obj = new LunarLanderObjective(game);
 
        double score = obj.valueOf(x);
-       min = Math.min(min,score);
-       max = Math.max(max, score); 
-       
-       //return -(score-min)/(max-min);
+//       min = Math.min(min,score);
+//       max = Math.max(max, score);
+//
+//       //return -(score-min)/(max-min);
         return -score;
     }
 

### 2014-04-07 code/src/lunarcode/LunarGame.java

@@ -39,8 +39,8 @@ public class LunarGame extends Game {
 
         // add gravity to the ship
         ship.v.add(0, LunarParams.lunarGravity * LunarParams.dt);
-        thrust+=random.nextGaussian()*0.1;
-        spin += random.nextGaussian()*0.01;
+       // thrust+=random.nextGaussian();
+       // spin += random.nextGaussian();
         ship.updateCont(thrust,spin);
 
         // wrap ship around screen based on centre x (basic but works)

### 2014-04-07 code/src/lunarcode/LunarLanderEvoCont.java

@@ -35,10 +35,10 @@ public class LunarLanderEvoCont {
 //        }
 
         CMAEvolutionStrategy cma = new CMAEvolutionStrategy();
-        cma.setDimension(9); // overwrite some loaded properties
-        //cma.parameters.setPopulationSize(10);
+        cma.setDimension(8); // overwrite some loaded properties
+        //cma.parameters.setPopulationSize(20);
         cma.setInitialX(0.5); // in each dimension, also setTypicalX can be used
-        cma.setInitialStandardDeviation(0.2);
+        cma.setInitialStandardDeviation(0.5/3.0);
         cma.options.verbosity = -1;
 
         double[] fitness = cma.init();

### 2014-04-07 code/src/lunarcode/LunarLanderHOOCont.java

@@ -26,13 +26,13 @@ public class LunarLanderHOOCont {
         MCTS.DEBUG = false;
 
         HOOClass func = new HOOClass((LunarGame) copy);
-        int dimension = 2;
+        int dimension = 8;
         int iterations = 10000;
-        HOOOptimiser hoo = new HOOOptimiser(func, dimension, iterations, min,
-                max, gamma);
-
-//        EDECBOptimiser hoo = new EDECBOptimiser(func, dimension, iterations, min,
+//        HOOOptimiser hoo = new HOOOptimiser(func, dimension, iterations, min,
 //                max, gamma);
+
+        EDECBOptimiser hoo = new EDECBOptimiser(func, dimension, iterations, min,
+                max, gamma);
         LunarLanderObjective fitfun = new LunarLanderObjective((LunarGame) copy.getCopy());
 
 
@@ -42,9 +42,10 @@ public class LunarLanderHOOCont {
             i++;
         }
 
+        System.out.println("i = " + i);
 
         double[] best = hoo.getBestNode().sampleAction();
-        System.out.println("best" + Arrays.toString(best));
+        //System.out.println("best" + Arrays.toString(best));
         //System.out.println(hoo + " " + i) ;
         //System.exit(0);
         double[] tbr = fitfun.convertInputs(best[0],best[1]);
@@ -53,7 +54,7 @@ public class LunarLanderHOOCont {
         //System.out.println(g.getShip().v.mag() + " sdfdf")      ;
 
 
-        System.out.println("tbtr" + Arrays.toString(tbr));
+        //System.out.println("tbtr" + Arrays.toString(tbr));
 
         return tbr;
 

### 2014-04-07 code/src/lunarcode/LunarLanderObjective.java

@@ -3,6 +3,8 @@ package lunarcode;
 import fr.inria.optimization.cmaes.fitness.IObjectiveFunction;
 import framework.utils.Vector2d;
 
+import java.util.Vector;
+
 /**
  * Created with IntelliJ IDEA.
  * User: ssamot
@@ -12,9 +14,10 @@ import framework.utils.Vector2d;
 public class LunarLanderObjective implements IObjectiveFunction {
 
     private final LunarGame game;
-    public static int MACRO_LENGTH = 8;
+    public static int MACRO_LENGTH = 7;
     Vector2d vertical = new Vector2d(1, 0);
-    Vector2d horizontal = new Vector2d(0, -1);
+    Vector2d left = new Vector2d(1, 2);
+    Vector2d right = new Vector2d(1, -2);
     public static final int WIN_BASE = 999999999;
 
 
@@ -29,17 +32,20 @@ public class LunarLanderObjective implements IObjectiveFunction {
 
         Vector2d nearestPad = cGame.terrain.getNearestSafeLandingPoint(game.getShip().s);
         double distFromPad = cGame.getShip().s.dist(nearestPad);
-        double xdist_oring =   Math.abs(game.getShip().s.x - nearestPad.x);
+        double xdist_oring = Math.abs(game.getShip().s.x - nearestPad.x);
         double y_maxspeed = 2;
-        double x_ignorspeed = 1;
+        double x_ignorspeed = 5;
+
+
         if (distFromPad > 60) {
-            y_maxspeed = 30;
+            y_maxspeed = 10;
             x_ignorspeed = 10;
         }
 
+        double distanceToTerrain = getDistance(game);
 
         int phase = 0;
-        if ( xdist_oring < 50) {
+        if (xdist_oring < 50) {
             phase = 1;
         }
 
@@ -62,17 +68,57 @@ public class LunarLanderObjective implements IObjectiveFunction {
 
 //
                     else {
+                        // double rotvel = Math.abs(((LunarShip) cGame.getShip()).rotVel);
+
+                        Vector2d correctOrientation = vertical;
                         if (phase == 1) {
+
                             Vector2d nv = cGame.getShip().d.copy();
                             //nv.normalise();
-                            double anglePoints = nv.dot(vertical) / (nv.mag() * vertical.mag());
+                            double anglePoints = nv.dot(correctOrientation) / (nv.mag() * correctOrientation.mag());
                             anglePoints = Math.acos(anglePoints);
 
-                            if (anglePoints < 0.8) {
[…]

### 2014-04-07 code/src/lunarcode/LunarParams.java

@@ -37,7 +37,8 @@ public class LunarParams {
     public static double friction = 1.0;
 
     // terrain generation values
-    public static int randomSeed = 104353;
+    //public static int randomSeed = 20394; //104353;
+    public static int randomSeed = 104353; //104353;
     public static int numPoints = 100;
     public static int numLandingPads = 1;
     public static int landingPadSize = 5;
@@ -49,6 +50,15 @@ public class LunarParams {
     // make the landscape totally flat?
     public static boolean flatLandscape = false;
 
-    public static Vector2d startingPoint = new Vector2d(700, 100);
+    public static Vector2d startingPoint = new Vector2d(130, 200);
+
+    //public static Vector2d startingPoint = new Vector2d(600, 30);
+
+    //public static Vector2d startingPoint = new Vector2d(750, 300);
+
+    // map 2
+    //public static Vector2d startingPoint = new Vector2d(100, 400);
+
+
     public static Vector2d landingFacing = new Vector2d(1,0);
 }

### 2014-04-07 code/src/lunarcode/LunarRun.java

@@ -22,27 +22,30 @@ public class LunarRun {
     public static boolean m_verbose = false;
 
     public static Game game;
-    public static LunarLanderEvoCont cont;
-    //public static LunarLanderHOOCont cont;
+    //public static LunarLanderEvoCont cont;
+    public static LunarLanderHOOCont cont;
 
     public static void main(String[] args) {
-        //runGame();
-        runGames(1);
+        runGame();
+        //runGames(10);
 
     }
 
     public static void prepareGame() {
         game = new LunarGame();
-        cont = new LunarLanderEvoCont();
-        //cont = new LunarLanderHOOCont();
+        //cont = new LunarLanderEvoCont();
+        cont = new LunarLanderHOOCont();
     }
 
     public static void runGame() {
+
+
         prepareGame();
         LunarView view = new LunarView((LunarGame) game, (LunarShip) game.getShip());
         JEasyFrame frame = new JEasyFrame(view, "Lunar Lander Cont");
 
         while (!game.isEnded()) {
+
 
             ElapsedCpuTimer ect = new ElapsedCpuTimer(ElapsedCpuTimer.TimerType.CPU_TIME);
             ect.setMaxTimeMillis(PTSPConstants.ACTION_TIME_MS);
@@ -51,15 +54,23 @@ public class LunarRun {
             double[] actionToExecute = cont.getAction(game, ect);
 
             ((LunarGame) game).tickCont(actionToExecute[0], actionToExecute[1]);
-            System.out.println("speed = " + ((LunarGame) game).getShip().v.mag() + " " + ((LunarGame) game).landedSuccessfully);
+            //System.out.println("speed = " + ((LunarGame) game).getShip().v.mag() + " " + ((LunarGame) game).landedSuccessfully);
             view.repaint();
 
-            Vector2d nv = ((LunarGame) game).ship.d.copy();
-            nv.normalise();
+//            Vector2d nv = ((LunarGame) game).ship.d.copy();
+//            nv.normalise();
+//
+//            double anglePoints = nv.dot(LunarParams.landingFacing) / (nv.mag() * LunarParams.landingFacing.mag());
+//            anglePoints = Math.acos(anglePoints);
+//            //System.out.println("anglePoints = " + anglePoints);
+//
+//            Vector2d nearestPad = ((LunarGame)game).terrain.getNearestSafeLandingPoint(game.getShip().s);
+//            Vector2d s = game.getShip().s;
[…]

### 2014-04-07 lpaper/lunarexps/sample_means.py

@@ -1,0 +1,38 @@
+import numpy as np
+from tabulate import tabulate
+import numpy as np
+import scipy.stats
+import scipy as sp
+
+
+def mean_confidence_interval(data, confidence=0.95):
+    a = 1.0*np.array(data)
+    n = len(a)
+    m, se = np.mean(a), scipy.stats.sem(a)
+    h = se * sp.stats.t._ppf((1+confidence)/2., n-1)
+    return [m,h]
+
+map2 = np.loadtxt(open("map2.csv","rb"),delimiter=",")
+map1_tall = np.loadtxt(open("map1-tall.csv","rb"),delimiter=",")
+map1_easy = np.loadtxt(open("map1-easy.csv","rb"),delimiter=",")
+map1_wall = np.loadtxt(open("map1-wall.csv","rb"),delimiter=",")
+
+
+
+maps = {}
+maps["map2"] = map2
+maps["map1-tall"] = map1_tall
+maps["map1-easy"] = map1_easy
+maps["map1-wall"] = map1_wall
+
+print mean_confidence_interval(map2[:,0])
+
+table = [(["map2"] + mean_confidence_interval(map2[:,0]) + mean_confidence_interval(map2[:,1])) ,
+		 (["map1-tall"] + mean_confidence_interval(map1_tall[:,0]) + mean_confidence_interval(map1_tall[:,1])) ,
+		 (["map1-easy"] + mean_confidence_interval(map1_easy[:,0]) + mean_confidence_interval(map1_easy[:,1])) ,
+		 (["map1-wall"] + mean_confidence_interval(map1_wall[:,0]) + mean_confidence_interval(map1_wall[:,1])) ,
+]
+print tabulate(table, headers = ["Map Name", "Success Mean", "Success CI", "Ticks Mean", "Tick CI"], tablefmt="latex")
+
+
+

### 2014-04-07 lpaper/lunarcont.tex
8 full pages, need to re-read

@@
 \DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}
+
+\newcommand{\abs}[1]{\lvert#1\rvert}
+\newcommand{\norm}[1]{\lVert#1\rVert}
 
@@
 
-The spaceship within this game world is modelled as an a circular mass with some basic physical properties, such as a position within the game world $\boldsymbol {s}$, a velocity $\boldsymbol {v}$, an orientation $\theta$, an angular velocity $\omega$ and a radius of a bounding sphere $r$ used for collision detection with the landscape.
+The spaceship within this game world is modelled as an a circular mass with some basic physical properties, such as a position within the game world $\bm {s}$, a velocity $\bm {v}$, an orientation $\bm{d}$, an angular velocity $\omega$ and a radius of a bounding sphere $r$ used for collision detection with the landscape.
 
@@
 The point that lies on the line segment that the ship is being checked against is calculated through simple linear interpolation based on which two vertices the ship is horizontally closest to.
-For the ship's centre $\boldsymbol {s}$, the left nearest landscape vertex $\boldsymbol {p}^{l}$ and the right nearest landscape vertex $\boldsymbol {p}^{r}$, the point of collision $\boldsymbol {p}^{c}$ against the landscape is calculated as
+For the ship's centre $\bm {s}$, the left nearest landscape vertex $\bm {p}^{l}$ and the right nearest landscape vertex $\bm {p}^{r}$, the point of collision $\bm {p}^{c}$ against the landscape is calculated as
 
@@
 \begin{equation}
-\boldsymbol {p}_{x}^{c} = \boldsymbol {s}_{x}
-\boldsymbol {p}_{y}^{c} = \boldsymbol {p}_{y}^{l} + v(\boldsymbol {p}_{y}^{r} - \boldsymbol {p}_{y}^{l})
+\bm {p}_{x}^{c} = \bm {s}_{x}
+\bm {p}_{y}^{c} = \bm {p}_{y}^{l} + v(\bm {p}_{y}^{r} - \bm {p}_{y}^{l})
 \end{equation}
@@
 \begin{equation}
-v = \frac{ {\boldsymbol {s}_{x} - \boldsymbol {p}_{x}^{l}}}{ {\boldsymbol {p}_{x}^{r} - \boldsymbol {p}_{x}^{l}} }
+v = \frac{ {\bm {s}_{x} - \bm {p}_{x}^{l}}}{ {\bm {p}_{x}^{r} - \bm {p}_{x}^{l}} }
 \end{equation}
@@
 \begin{equation}
-\boldsymbol {s}_{y} + r \geq \boldsymbol {p}^{c}_{y}
+\bm {s}_{y} + r \geq \bm {p}^{c}_{y}
 \end{equation}
@@
 In order to circumvent this, as it is common in other games, we use a heuristic for each state.
-In order to counter this tendency, as is common in games, we used a two-stage heuristic.
-The first stage involves the ship trying to reach the landing pad.
+Initially, we can define a number of vectors that portray the
+
+\begin{align}
+& \bm{du} = [1,0]\notag\\ 
+& \bm{dl} = [1,2]\notag\\
+& \bm{dr} = [1,-2]
+\label{eq:vectors}
+\end{align}
+
+The vectors defined in Equation \ref{eq:vectors} define three directions for the ship.
+The first one, $\bm{du}$, define a ship standing upright, the second one, $\bm{dl}$, define tilting to the left, while the third one $\bm{dr}$ a ship tilting to the right.
+We make use of these vectors in order to help align the ship in the right direction, deepening on which way the landing point is.
 
 \begin{multline}
-rng = -\sum{arccos(vl\cdot\theta)/||vl||*||\theta||} - |\boldsymbol {v_y} - V_{y}^{max}| \\
-h = \twopartdef{ -|(|\boldsymbol {v}| - C) - |\boldsymbol {s}_{x} - \boldsymbol {np}_{x}||}{||\boldsymbol {s} - \boldsymbol {np}|| < D}
-{rng - |\boldsymbol {v_x}|   }{Otherwise}
+f = \twopartdef{-\sum{arccos(\bm{dl}\cdot\bm{d}/\norm{\bm{dl}}\norm{\bm{d}}})}{\bm{np_x} < \bm{s_x}}{-\sum{arccos(\bm{dr}\cdot\bm{d}/\norm{\bm{dr}}\norm{\bm{d}}})} {otherwise}\\
+l = -\sum{arccos(\bm{du}\cdot\bm{d}/\norm{\bm{du}}\norm{\bm{d}}}) - \abs{\bm {v_y} - V_{y}^{max}} \\
+h = \twopartdef{ \abs{-(\norm{\bm {v}} - C) - \abs{\bm {s}_{x} - \bm {np}_{x}}} - f}{\norm{\bm {s} - \bm {np}} < D}
+{l - |\bm {v_x}|   }{Otherwise}
 \label{eq:h1}
-\end{multline}
+\end{multline}rng
 
-In the above Equation \ref{eq:h1} the constant C denotes the maximum magnitute of speed, which we set to $C=100$.
+In the above Equation \ref{eq:h1} the constant C denotes the maximum magnitude of speed, which we set to $C=30$.
 $D = 60$ is the distance from the landing point $np$ where this heuristic takes precedence.
 Intuitively, the upper part of the heuristic $h$ tries to get the ship above the landing point as fast as possible while maintaining a certain speed.
-The second part of the heuristic tries to minimise velocity on the $x$ axis (as the ship tries to land), maintain a minimum landing speed on the $y$ axis and get a certain angle (the summation term $rng$) as close to vector $vl = (0,-1)$, i.e.
+The second part of the heuristic tries to minimise velocity on the $x$ axis (as the ship tries to land), maintain a minimum landing speed on the $y$ axis and get a certain angle (the summation term $f$) as close to vector $du$, i.e.
 help the ship stay upright through the duration of the rollout.
+It's worth noting the main reason why we use a heuristic of this kind, one that includes not only the final state, but makes use of the path of the state as well, is that there is not enough information to discern states if one takes into account direction only, without incorporating the angular velocity.
+The ship might be spinning madly, but the end of the rollout might find it in the correct direction.
 
@@
 We present the results of our experiments below.
-
-\begin{table}
-	\begin{tabular}{lrrrr}
-\hline
- Map Name   &   Success Mean &   Success CI &   Ticks Mean &   Tick CI \\
-\hline
- map2       &            1   &     0        &       2925.9 &   184.437 \\
- map1-tall  &            1   &     0        &       2733   &   159.34  \\
- map1-easy  &            1   &     0        &       2293   &   123.917 \\
- map1-wall  &            0.9 &     0.226216 &       2421.6 &   682.893 \\
-\hline
-\end{tabular}
-\end{table}
 
@@
                 \includegraphics[width=1.0\textwidth]{./graphics/map2.pdf}
-                \caption{Map. }
+                \caption{Map A,  where the ship is lower then the pad}
                 \label{fig:map2}
@@
                 \includegraphics[width=1.0\textwidth]{./graphics/map1-easy.pdf}
-                \caption{sdfdf}
+                \caption{Map B, an easy map where only the first heuristic is used}
                 \label{fig:map1-easy}
@@
                 \includegraphics[width=1.0\textwidth]{./graphics/map1-tall.pdf}
-                \caption{Detail of Figure \ref{fig:sample}, when the ship is closer to landing.}
+                \caption{Map C, where the ship is close to the top of the screen}
                 \label{fig:map1-tall}\end{subfigure}\begin{subfigure}{0.49\textwidth}
@@
                 \includegraphics[width=1.0\textwidth]{./graphics/map1-wall.pdf}
-                \caption{Detail of Figure \ref{fig:sample}, when the ship is closer to landing.}
+                \caption{Map D, where the ship is behind a wall }
                 \label{fig:map-wall}
@@
 
-         \caption{Sample run for EVO-P. Best fitness discovered at each tick is shown. Notice the big ``bump'' in fitness when the ship gets close to the landing pad}
+         \caption{The four different maps used in our experiments. The maps capture different position/properties}
          \label{fig:maps}
@@
 
+\subsection{Lunar Lander}
+
+\begin{table}[h]
+	\begin{tabular}{lrrrr}
+\hline
+ Map Name   &   Success Mean &   Success CI &   Ticks Mean &   Tick CI \\
+\hline
+ Map A       &            1   &     0        &       2925.9 &   $\pm 184.437$ \\
+ Map B  &            1   &     0        &       2733   &   $\pm 159.34$  \\
+ Map C  &            1   &     0        &       2293   &   $\pm 123.917$ \\
+ Map D  &            0.9 &     $\pm 0.226216$ &       2421.6 &   $\pm 682.893$ \\
+\hline
+\end{tabular}
+\label{tb:results}
+\end{table}
+
+For lunar lander, we run $10$ runs for each map in Figure \ref{fig:maps}, and the results for EVO-P can be seen in Table \ref{tb:results}.
+The results include the mean number of successful landings and the mean time it took to land the ship, alongside their confidence for the $95$th interval.
+For EVO-P we used the default CMA-ES population size, each individual was bound between, each gene had a length of $8$ and each action pair ($x[i], x[i+1]$) was repeated $7$ times.
+We derived these number experimentally, however we did not optimise for them (not even implicitly) to any serious degree.
+
+There are a number of things to notice here.
+The first one is that the only in Map D, where the ship starts the problem facing a wall do we get a failure to land, and this is due to a crash.
+This has to do with the fact that even a minor thrust can push you a bit forward and drive the ship to a position where there is no longer the possibility of recovery from a crush (or the combination of that recovery is extremely hard to get).
+The Second thing to observe is that we get there slowly, but this is by design, as we he have made the flying speed of the ship controllable.
+One can pottentially tweak all the default parameters we have used, plus the ones incorporated in CMA-ES as a kind of meta-heuristic search and achieve much better results.
+Another thing to note is that EVO-P lands the ship (on average) almost at the same time on MAPS B, C, D, although map has the same landscape as the other two but the ship is much closer.
+This is due to the fact that the minimum landing speed, set on one of our heuristics, is really low and The final thing we would like to note is that we have only included the results for EVO-P because T-HOLOP failed to land the ship using the same heuristic, and behaved badly.
+It's worth noting here that T-HOLOP requires rewards between certain bounds and this is not the case with our heuristic.
+For completeness purposes, we did change the UCB-like node selection policy of T-HOO to e-greedy, in which case we DID get the ship to behave reasonably, albeit the failure to land was evident here as well.
+This is important we think for a number of reasons.
+The first one is that it shows that different Heuristics might be needed for different algorithms.
+Although this might be self-evident, it shows that one must design heuristics with a specific algorithm in mind.
+
 \section{Conclusion} \label{sec:conclusion}
+We showed that it's possible to use evolution in a rolling horizon setting, making it possible to create agents using off-the-shelf evolutionary algorithms and a generative model.
+We used an experimental setup that included both simple benchmark problems and a popular arcade game, Lunar Lander.
+Unfortunately, it seems that different algorithms require different heuristics to work properly in the case that absorbing states are not met during rollouts.
+In future research we aim to improve MCTS-like algorithms for continuous settings by including a learning module, possibly in the form of actor-critic algorithms.
+This should automate learning the heuristics (after some ad-hoc played games) and help guide the rollouts towards more sensible solutions.
 

### 2014-04-07 code/src/lunarcode/HOOClass.java

@@ -11,8 +11,9 @@ public class HOOClass extends ContinuousProblem {
 
 
     private final  LunarGame game;
-    double max = Integer.MAX_VALUE;
-    double min = Integer.MIN_VALUE;
+    double max = Integer.MIN_VALUE;
+    double min = Integer.MAX_VALUE;
+    double[] best = null;
 
 
     public HOOClass(LunarGame cGame) {
@@ -33,11 +34,22 @@ public class HOOClass extends ContinuousProblem {
        LunarLanderObjective obj = new LunarLanderObjective(game);
 
        double score = obj.valueOf(x);
+       score = -score;
 //       min = Math.min(min,score);
+
 //       max = Math.max(max, score);
 //
 //       //return -(score-min)/(max-min);
-        return -score;
+
+        if(score > max) {
+            max = score;
+            best = x;
+        }
+
+        min = Math.min(min,score);
+
+
+        return score;
     }
 
 }

### 2014-04-07 code/src/lunarcode/LunarLanderHOOCont.java

@@ -36,15 +36,18 @@ public class LunarLanderHOOCont {
         LunarLanderObjective fitfun = new LunarLanderObjective((LunarGame) copy.getCopy());
 
 
-        int i = 0;
+        //int i = 0;
         while (!ect.exceededMaxTime()) {
-            hoo.runForSim(1);
-            i++;
+            hoo.runForSim(50);
+            //i++;
         }
 
-        System.out.println("i = " + i);
+        //System.out.println("i = " + i);
 
-        double[] best = hoo.getBestNode().sampleAction();
+        //double[] best = hoo.getBestNode().sampleAction();
+        double[] best = func.best;
+
+        //System.out.println(Arrays.toString(hoo.getBestNode().getMax()) + " " +  Arrays.toString(hoo.getBestNode().getMin()) ) ;
         //System.out.println("best" + Arrays.toString(best));
         //System.out.println(hoo + " " + i) ;
         //System.exit(0);

### 2014-04-08 lpaper/lunarcont.tex
One more reading and we are done

@@
 In this paper we compare T-HOLOP (Truncated Hierarchical Open Loop Planning), an open loop planning algorithm at least partially inspired by MCTS with a version evolutionary planning using CMA-ES (which we call EVO-P) in two planning benchmark problems (Cart Balancing and the Double Integrator) and Lunar Lander, a classic arcade game.
-He show that EVO-P outperforms T-HOLOP in almost all settings.
-We conclude that off-the-shelf evolutionary algorithms can be used successfully in a rolling horizon setting, and might offer considerable advantages over more formal methods.
+He show that EVO-P outperforms T-HOLOP in the classic benchmarks, while T-HOLOP is unable to find a solution using the same heuristic.
+We conclude that off-the-shelf evolutionary algorithms can be used successfully in a rolling horizon setting, and that a different type of heuristics might be needed under different optimisation algorithms.
 
@@
 Our second contribution is the use of the these algorithms in Lunar Lander, a popular arcade game, coupled with a strong two-stage heuristic.
-We show that while EVO-P performs really well, T-HOLOP struggles.
+We show that while EVO-P performs reasonably well, T-HOLOP struggles.
 
@@
 
-This is the first background section of this paper, which aims to introduce the two core optimisation algorithms on which our contributions are made.
+This is the first background section of this paper, which aims to introduce the two core optimisation algorithms on top of which our contributions are made.
 We will present both algorithms from an implementer's viewpoint, without explicitly justifying the design and mathematical choices behind them, nor going into too much detail.
@@
 Also note that we are NOT dealing with continuous time MDPs in this paper, although there is a notion of time.
+An example of the effects of the discount factor on rewards can be found in Figure \ref{fig:gamma}.
 
@@
 		\While{$!StoppingCondition$}
-		    \State $a \gets \mathsc{FindOpenLoopPlan}(depth, T',S')[0]$ \Comment{Execute first action of the policy}
+		    \State $a \gets \mathsc{FindOpenLoopPlan}(depth, T',S')[0]$ \Comment{Execute first action(s) of the policy}
 		    \State $s \sim T(s'|s,a)$ \Comment{Stop only at an absorbing state}
@@
 \subsection{Lunar Lander}
-We describe the properties of Lunar Lander.
+In this section we give a brief overview of the Game of Lunar Lander, a popular arcade game.
 
@@
 % unclear?
-
-% TODO: DIAGRAM - jagged lunar landscape might be useful (reuse of old one probably not sufficient)
-
-% not sure how much detail should be gone into here? should just cite generation equations from old paper or copy over??
-
-Each vertex of this landscape is generated sequentially. % INCOMPLETE due to uncertainty listed above
 
@@
                 \includegraphics[width=1.0\textwidth]{graphics/online-DI-crop.pdf}
-                \caption{Scores for a noise level of 1 under variable function evaluations.}
+                \caption{Scores for a noise level of $l=1$ under variable function evaluations.}
                 \label{fig:DI}
@@
 The ship colliding with the landscape constitutes the end of the {\itshape Lunar Lander} game.
-The conditions surrounding this collision, including speed, orientation of the ship, and fuel used, constitute whether the nature of the collision is a success or a failure.
+The conditions surrounding this collision, including speed, orientation of the ship constitute whether the nature of the collision is a success or a failure.
+In this experimental setup with accept a velocity of $5$ and an angle of $< 0.1$ radians as successful landing criteria.
 
@@
 
-Finally each linear thrust and angular impulse is repeated for $X$ actions, the effects of which we are going to explore below.
+Finally each linear thrust and angular impulse is repeated for $X = 7$ actions, the effects of which we are going to explore below.
 
@@
 The first one, $\bm{du}$, define a ship standing upright, the second one, $\bm{dl}$, define tilting to the left, while the third one $\bm{dr}$ a ship tilting to the right.
-We make use of these vectors in order to help align the ship in the right direction, deepening on which way the landing point is.
+We make use of these vectors in order to help align the ship in the right direction, depending on which way the landing point is.
 
@@
 h = \twopartdef{ \abs{-(\norm{\bm {v}} - C) - \abs{\bm {s}_{x} - \bm {np}_{x}}} - f}{\norm{\bm {s} - \bm {np}} < D}
-{l - |\bm {v_x}|   }{Otherwise}
+{l - \abs{\bm {v_x} -C_l }  }{Otherwise}
 \label{eq:h1}
-\end{multline}rng
+\end{multline}
 
@@
 Intuitively, the upper part of the heuristic $h$ tries to get the ship above the landing point as fast as possible while maintaining a certain speed.
-The second part of the heuristic tries to minimise velocity on the $x$ axis (as the ship tries to land), maintain a minimum landing speed on the $y$ axis and get a certain angle (the summation term $f$) as close to vector $du$, i.e.
+The second part of the heuristic tries to minimise velocity on the $x$ axis (as the ship tries to land with maximum speed $C_l = 2$), maintain a minimum landing speed on the $y$ axis and get a certain angle (the summation term $f$) as close to vector $du$, i.e.
 help the ship stay upright through the duration of the rollout.
@@
 Results for the variable noise levels $l$ (Figures \ref{fig:IP-N-50}, \ref{fig:IP-N-100}, \ref{fig:DI-N-50}, \ref{fig:DI-N-100}) closely mimic the performance of their constant noise counterparts.
-For the inverted pendulum, using just 50 iterations (Figure \ref{fig:IP-N-50}) EVO-P is able to stabilise the pendulum with .
+For the inverted pendulum, using just 50 iterations (Figure \ref{fig:IP-N-50}) EVO-P is able to stabilise the pendulum with $50$ iterations.
 This still applies if we increase the number of function evaluations to 100 (Figure \ref{fig:DI-N-100}).
@@
 
-For lunar lander, we run $10$ runs for each map in Figure \ref{fig:maps}, and the results for EVO-P can be seen in Table \ref{tb:results}.
+For lunar lander, we do $10$ runs for each map in Figure \ref{fig:maps}, and the results for EVO-P can be seen in Table \ref{tb:results}.
 The results include the mean number of successful landings and the mean time it took to land the ship, alongside their confidence for the $95$th interval.
-For EVO-P we used the default CMA-ES population size, each individual was bound between, each gene had a length of $8$ and each action pair ($x[i], x[i+1]$) was repeated $7$ times.
-We derived these number experimentally, however we did not optimise for them (not even implicitly) to any serious degree.
+For EVO-P we used the default CMA-ES population size, each individual was bound between, each gene had a length of $8$ and each action pair ($x[i], x[i+1]$) was repeated $7$ times (where $x[i]$ represents thrust and $x[i+1]$ angular thrust).
+We derived these numbers experimentally and we did not optimise for them (not even implicitly) to any serious degree.
 
@@
 \section{Conclusion} \label{sec:conclusion}
-We showed that it's possible to use evolution in a rolling horizon setting, making it possible to create agents using off-the-shelf evolutionary algorithms and a generative model.
+We showed that it is doable to use evolution in a rolling horizon setting, making it possible to create agents using off-the-shelf evolutionary algorithms and a generative model.
 We used an experimental setup that included both simple benchmark problems and a popular arcade game, Lunar Lander.
@@
 In future research we aim to improve MCTS-like algorithms for continuous settings by including a learning module, possibly in the form of actor-critic algorithms.
-This should automate learning the heuristics (after some ad-hoc played games) and help guide the rollouts towards more sensible solutions.
+This should automate learning the heuristics (after some ad-hoc played games) and help guide the rollouts towards more sensible solutions, even in the case where heuristics do not favour T-HOLOP.
 

### 2014-04-08 lpaper/lunarcont.tex
final

@@
 
-\author{Samuel A. Roberts, Spyridon Samothrakis, Diego Perez, Simon M. Lucas}
+\author{Spyridon Samothrakis, Samuel A. Roberts, Diego Perez, Simon M. Lucas}
 
@@
 Section \ref{sec:Optimisation} discusses (briefly and epigrammatically) the relevant base algorithms that are going to be used in this paper, ((Truncated) Hierarchical Online Optimisation~\cite{hoo2011} and CMA-ES~\cite{hansen2003reducing}).
-Section~\ref{sec:Planning} explains why and how the above algorithms can be used in the context of planning (and thus transformed into HOLOP and EVO-P respectively).
+Section~\ref{sec:Planning} explains why and how the above algorithms can be used in the context of planning (and thus transformed into T-HOLOP and EVO-P respectively).
 In Section \ref{sec:methodology} the experimental setup is described.
@@
 \section{Methodology} \label{sec:methodology}
-Our methodology is simple, yet efficient. We test T-HOLOP and EVO-P in two simple benchmark problems (see below for more) and Lunar Lander, a classic arcade game.
+Our methodology is simple, yet efficient.
+We test T-HOLOP and EVO-P in two simple benchmark problems (see below for more) and Lunar Lander, a classic arcade game.
+EVO-P is simply CMA-ES used in a rolling horizon setting, while T-HOLOP is T-HOO used in the same manner.
 
@@
 
-An example of how the \textit{predicted} reward looks like for a $2$-step look-ahead (instead of $50$ we use in the experiments) can be seen in Figure \ref{fig:vis}.
-We plotted the full state-space by pushing the internal generative model two steps forward for all possible actions by discretizing each action.
-Notice how the predicted reward landscape changes over time as the agent moves.
-The continuous replanning should aim to perform the action with the highest future reward.
-
 The results of these experiments can be seen in graphs  \ref{fig:IP} and \ref{fig:DI}.
@@
 \label{tb:results}
+\caption{EVO-P results for different lunar lander maps}
 \end{table}
 
-For lunar lander, we do $10$ runs for each map in Figure \ref{fig:maps}, and the results for EVO-P can be seen in Table \ref{tb:results}.
+For lunar lander, we do $10$ runs for each map in Figure \ref{fig:maps}, and the results for EVO-P can be seen in Table I.
 The results include the mean number of successful landings and the mean time it took to land the ship, alongside their confidence for the $95$th interval.
@@
 It's worth noting here that T-HOLOP requires rewards between certain bounds and this is not the case with our heuristic.
-For completeness purposes, we did change the UCB-like node selection policy of T-HOO to e-greedy, in which case we DID get the ship to behave reasonably, albeit the failure to land was evident here as well.
+For completeness purposes, we did change the UCB-like node selection policy of T-HOO to e-greedy, in which case we \emph{did} get the ship to behave reasonably, albeit the failure to land was evident here as well.
 This is important we think for a number of reasons.

### 2014-04-08 lpaper/lunarcont.tex
More minor changes

@@
 
-Lunar Lander has a setup where a player gets rewards if and only if it actually lands on the pad.
+Lunar Lander has a setup where an agent gets rewards if and only if she manages to land to actually land the ship on the pad.
 This creates the problem of rollouts (whether used by EVO-P or T-HOLOP) rarely will see a reward.
 In order to circumvent this, as it is common in other games, we use a heuristic for each state.
-Initially, we can define a number of vectors that portray the
+Initially, we need define a number of vectors that portray ship directions:
 
@@
 help the ship stay upright through the duration of the rollout.
-It's worth noting the main reason why we use a heuristic of this kind, one that includes not only the final state, but makes use of the path of the state as well, is that there is not enough information to discern states if one takes into account direction only, without incorporating the angular velocity.
+It's worth noting the main reason why we use a heuristic of this kind, one that includes not only the final state, but makes use of the path of the state as well, is that there is not enough information to discern states if one takes into account ship direction only; without incorporating the angular velocity.
 The ship might be spinning madly, but the end of the rollout might find it in the correct direction.
@@
 For lunar lander, we do $10$ runs for each map in Figure \ref{fig:maps}, and the results for EVO-P can be seen in Table I.
-The results include the mean number of successful landings and the mean time it took to land the ship, alongside their confidence for the $95$th interval.
+The results include the mean number of successful landings (\emph{Success Mean}) and the mean time it took to land the ship (\emph{Ticks Mean}), alongside their confidence for the $95$th interval.
 For EVO-P we used the default CMA-ES population size, each individual was bound between, each gene had a length of $8$ and each action pair ($x[i], x[i+1]$) was repeated $7$ times (where $x[i]$ represents thrust and $x[i+1]$ angular thrust).
@@
 The first one is that the only in Map D, where the ship starts the problem facing a wall do we get a failure to land, and this is due to a crash.
-This has to do with the fact that even a minor thrust can push you a bit forward and drive the ship to a position where there is no longer the possibility of recovery from a crush (or the combination of that recovery is extremely hard to get).
-The Second thing to observe is that we get there slowly, but this is by design, as we he have made the flying speed of the ship controllable.
-One can pottentially tweak all the default parameters we have used, plus the ones incorporated in CMA-ES as a kind of meta-heuristic search and achieve much better results.
+This has to do with the fact that even a minor thrust can push you a bit forward and drive the ship to a position where there is no longer the possibility of recovery from a crush (or the combination of that recovery is extremely hard to find).
+The second thing to observe is that we get there slowly, but this is by design, as we he have made the flying speed of the ship controllable.
+One can potentially tweak all the default parameters we have used, plus the ones incorporated in CMA-ES as a kind of meta-heuristic search and achieve much better results.
 Another thing to note is that EVO-P lands the ship (on average) almost at the same time on MAPS B, C, D, although map has the same landscape as the other two but the ship is much closer.
-This is due to the fact that the minimum landing speed, set on one of our heuristics, is really low and The final thing we would like to note is that we have only included the results for EVO-P because T-HOLOP failed to land the ship using the same heuristic, and behaved badly.
-It's worth noting here that T-HOLOP requires rewards between certain bounds and this is not the case with our heuristic.
+This is due to the fact that the minimum landing speed, set on one of our heuristics, is really low.
+The final thing we would like to note is that we have only included the results for EVO-P because T-HOLOP failed to land the ship using the same heuristic, and behaved badly.
+It's worth noting here that T-HOLOP requires rewards between certain bounds, so they can be normalised around $[0,1]$ and this is not the case with our heuristic.
 For completeness purposes, we did change the UCB-like node selection policy of T-HOO to e-greedy, in which case we \emph{did} get the ship to behave reasonably, albeit the failure to land was evident here as well.
@@
 Although this might be self-evident, it shows that one must design heuristics with a specific algorithm in mind.
+The second is, but we can speculate at the time, that T-HOLOP spends too much trying to solve a hard problem than the one at hand.
+It will converge eventually, but the the Guassian approximation of CMA-ES seems to outperform it in the short term.
 

### 2014-04-09 lpaper/lunarcont.tex
Simon's changes + some minor additions

@@
 \documentclass[conference]{IEEEtran}
-\usepackage{graphicx,times,amsmath,algorithm,caption, subcaption,nicefrac,bm}
+\usepackage{graphicx,times,amsmath,algorithm,caption, subcaption,nicefrac,bm, url}
 \usepackage{multicol}
@@
 An agent sees a ``time'' when it's about to act instead of a state, i.e it doesn't have access to $s_t$, but just $t$.
-An open loop policy (or plan or control) is when an agent learns a policy with the form $\pi_o:S \times \mathcal{T} \rightarrow [0,1]$, where  $\mathcal{T}$ is an ordered set of time steps.
+An open loop policy (or plan or control) is when an agent learns a policy with the form $\pi_o:A \times \mathcal{T} \rightarrow [0,1]$, where  $\mathcal{T}$ is an ordered set of time steps.
 Thus, an agent takes actions irrespective of the current state it is in.
@@
 
-Lunar Lander has a setup where an agent gets rewards if and only if she manages to land to actually land the ship on the pad.
+Lunar Lander has a setup where an agent gets rewards if and only if it manages to actually land the ship on the pad.
 This creates the problem of rollouts (whether used by EVO-P or T-HOLOP) rarely will see a reward.
 In order to circumvent this, as it is common in other games, we use a heuristic for each state.
+The heuristic was partially inspired by the two-stage approach found here : \url{http://www.cs.berkeley.edu/~jrs/4/lunar.html}\cite{lunar}.
 Initially, we need define a number of vectors that portray ship directions:
@@
 
-\begin{multline}
-f = \twopartdef{-\sum{arccos(\bm{dl}\cdot\bm{d}/\norm{\bm{dl}}\norm{\bm{d}}})}{\bm{np_x} < \bm{s_x}}{-\sum{arccos(\bm{dr}\cdot\bm{d}/\norm{\bm{dr}}\norm{\bm{d}}})} {otherwise}\\
-l = -\sum{arccos(\bm{du}\cdot\bm{d}/\norm{\bm{du}}\norm{\bm{d}}}) - \abs{\bm {v_y} - V_{y}^{max}} \\
-h = \twopartdef{ \abs{-(\norm{\bm {v}} - C) - \abs{\bm {s}_{x} - \bm {np}_{x}}} - f}{\norm{\bm {s} - \bm {np}} < D}
-{l - \abs{\bm {v_x} -C_l }  }{Otherwise}
+\begin{align}
+f &= \twopartdef{-\sum{arccos(\bm{dl}\cdot\bm{d}/\norm{\bm{dl}}\norm{\bm{d}}})}{\bm{np_x} < \bm{s_x}}{-\sum{arccos(\bm{dr}\cdot\bm{d}/\norm{\bm{dr}}\norm{\bm{d}}})} {otherwise}\notag\\
+l &= -\sum{arccos(\bm{du}\cdot\bm{d}/\norm{\bm{du}}\norm{\bm{d}}}) - \abs{\bm {v_y} - V_{y}^{max}}\notag\\
+h &= \twopartdef{ \abs{-(\norm{\bm {v}} - C) - \abs{\bm {s}_{x} - \bm {np}_{x}}} - f}{\norm{\bm {s} - \bm {np}} < D}
+{l - \abs{\bm {v_x} -C_l }  }{Otherwise}\notag\\
+h&\gets h + p 
 \label{eq:h1}
-\end{multline}
+\end{align}
 
@@
 help the ship stay upright through the duration of the rollout.
-It's worth noting the main reason why we use a heuristic of this kind, one that includes not only the final state, but makes use of the path of the state as well, is that there is not enough information to discern states if one takes into account ship direction only; without incorporating the angular velocity.
-The ship might be spinning madly, but the end of the rollout might find it in the correct direction.
+It's worth noting the main reason why we use a heuristic of this kind, one that includes not only the final state, but makes use of the path of the state as well, is that there is not enough information to discern states if one takes into account ship direction only; without incorporating the angular velocity,  the ship might be spinning madly, but the end of the rollout might find it in the correct direction.
+Finally notice that the heuristic $h$ gets the value $p=10000$ if the ship has crashed at any point during the rollout.
 

### 2014-04-11 lpaper/lunarcont.tex
added Michelle's fixes

@@
 as in Monte Carlo Tree Search).
-In this paper we compare T-HOLOP (Truncated Hierarchical Open Loop Planning), an open loop planning algorithm at least partially inspired by MCTS with a version evolutionary planning using CMA-ES (which we call EVO-P) in two planning benchmark problems (Cart Balancing and the Double Integrator) and Lunar Lander, a classic arcade game.
-He show that EVO-P outperforms T-HOLOP in the classic benchmarks, while T-HOLOP is unable to find a solution using the same heuristic.
+In this paper we compare T-HOLOP (Truncated Hierarchical Open Loop Planning), an open loop planning algorithm at least partially inspired by MCTS, with a version of evolutionary planning that uses CMA-ES (which we call EVO-P) in two planning benchmark problems (Cart Balancing and the Double Integrator) and in Lunar Lander, a classic arcade game.
+We show that EVO-P outperforms T-HOLOP in the classic benchmarks, while T-HOLOP is unable to find a solution using the same heuristics.
 We conclude that off-the-shelf evolutionary algorithms can be used successfully in a rolling horizon setting, and that a different type of heuristics might be needed under different optimisation algorithms.
@@
 \section{Introduction} \label{Introduction}
-One of the most common uses of function optimisation is for solving control problems.
+One of the most common uses of function optimisation is solving control problems.
 If an agent has access to a generative model of the world (i.e.
@@
 HOLOP is based partially on a strong~\cite{hoo2011} real-valued optimisation algorithm called HOO (for ``Hierarchical Optimistic Optimisation'').
-In this paper two contributions are made; the first contribution is the comparison of two versions of a very popular and strong evolutionary strategy: EVO-P, the open loop planning algorithm using Covariance Matrix Adaptation Evolutionary Strategy (CMA-ES)~\cite{hansen2003reducing} with a version of HOLOP, which stands for Hierarchical Open Loop Optimistic Planning~\cite{hoo2011} on two standard benchmark problems, the double integrator and the reverse pendulum.
-For speed purposed, we use T-HOO (Truncated Hierarchical Optimistic Optimisation), a faster version of HOO~\cite{hoo2011}, as the engine of HOLOP, hence we name the algorithm used T-HOLOP.
-Our second contribution is the use of the these algorithms in Lunar Lander, a popular arcade game, coupled with a strong two-stage heuristic.
+In this paper two contributions are made.
+The first contribution is the comparison of two versions of a very popular and strong evolutionary strategy: EVO-P, the open loop planning algorithm using Covariance Matrix Adaptation Evolutionary Strategy (CMA-ES)~\cite{hansen2003reducing} and a version of HOLOP, which stands for Hierarchical Open Loop Optimistic Planning~\cite{hoo2011} on two standard benchmark problems, the double integrator and the reverse pendulum.
+For speed purposes, we use T-HOO (Truncated Hierarchical Optimistic Optimisation), a faster version of HOO~\cite{hoo2011}, as the engine of HOLOP, hence we name the algorithm T-HOLOP.
+Our second contribution is the use of these algorithms in Lunar Lander, a popular arcade game, coupled with a strong two-stage heuristic.
 We show that while EVO-P performs reasonably well, T-HOLOP struggles.
@@
 CPU time for the simulations.
-This is described in more detail in section \ref{sec:Planning} and in line with what is being done in Perez et al.
+This is described in more detail in Section \ref{sec:Planning} and in line with what was done in Perez et al.
 \cite{perez2013rolling}, although this time in a continuous setting.
@@
 
-This is the first background section of this paper, which aims to introduce the two core optimisation algorithms on top of which our contributions are made.
+This is the first background section of this paper and aims to introduce the two core optimisation algorithms on which our contributions are based.
 We will present both algorithms from an implementer's viewpoint, without explicitly justifying the design and mathematical choices behind them, nor going into too much detail.
@@
 Our goal is to find a vector $\bm{x}_0$, for which $f(\bm{x}) > f(\bm{x_0}) $ for all $x$ (also known as ``minimisation'') or $f(\bm{x}) < f(\bm{x_0})$ for all $x$ (known as ``maximisation'').
-There are an enormous number of algorithms devoted to this problem, but only two will be presented and compared here, which are both considered state of the art in their respective fields and do not require the calculation of a gradient.
+There are an enormous number of algorithms devoted to this problem, but only two will be presented and compared here.
+Both of these are considered state of the art in their respective fields and do not require the calculation of a gradient.
 If one has an analytic expression of the gradient or if the functions proposed are smooth enough, it might be better to look elsewhere for solutions (e.g.
@@
 The covariance matrix $C$, positive definite and symmetric, determines the shape of the distribution and its graphical interpretation: it defines an iso-density ellipsoid $\{ x_i \in \mathcal{R} | (x - m)^T C^{-1} (x - m) = 1\}$.
-Through the classic evolutionary setup and by making clever use of population statistics, the algorithm proceeds from generation to generation of sampling until some convergence criteria has been reached.
+Through the classic evolutionary setup, and by making clever use of population statistics, the algorithm proceeds from generation to generation of sampling until some convergence criteria has been reached.
 
@@
 ``Splitting'' a node is the procedure of creating two children $C(\nu) = \{C_l(\nu),C_r(\nu)\}$.
-Both children are initialised $\bm{x_{\mathit{max}}}$ and $\bm{x_{\mathit{min}}}$ to their parents' nodes values.
-When there are multiple dimensions to choose from, at  at $k$ according to some random distribution.
+Both children are initialised $\bm{x_{\mathit{max}}}$ and $\bm{x_{\mathit{min}}}$ to their parents' node values when there are multiple dimensions to choose from, at $k$ according to some random distribution.
 Thus, for child $C_l(\nu)$ ,   $\bm{x^{C_l}_{\mathit{max,k}}} =\bm{x^{C}_{\mathit{min,k}}} +  \left(\bm{x^{C}_{\mathit{min,k}}} - \bm{x^{C}_{\mathit{max,k}}}\right)/2 $.
 For the other child,  $\bm{x^{C_r}_{\mathit{min,k}}} = \bm{x^{C}_{\mathit{min,k}}} +  \left(\bm{x^{C}_{\mathit{min,k}}} - \bm{x^{C}_{\mathit{max,k}}}\right)/2 $.
-In other words, the father node is split into two children nodes, choosing which dimension to split uniform randomly (or according to some other criterion).
+In other words, the father node is split into two children nodes, choosing which dimension to split uniform random (or according to some other criterion).
 Each node also has an associated value called $B(\nu)$, which is defined recursively as:  $B(\nu) = min \{U(\nu),max \{B(C_l(\nu)),B(C_r(\nu))\}\}$.
@@
 The goal of an agent in an MDP environment is to maximise its long term value.
-Note that since here we are dealing with continuous actions and states, sets $A$ and $S$ are metric (and infinitive).
+Note that, since we are dealing with continuous actions and states here, sets $A$ and $S$ are metric (and infinitive).
 Also note that we are NOT dealing with continuous time MDPs in this paper, although there is a notion of time.
@@
 
-Another point of interest is planning, and when it happens.
+Another point of interest is \textit{when} planning happens.
 In open loop planning (and what is usually termed ``planning'' without any further qualifications), an agent forms a plan once and follows it until the end.
@@
 A generative model is nothing more than a simulator of future events.
-This is a less of a requirement than knowing the full model (which entails knowing the full function $T$), but still requires the agent to have some knowledge of the world.
+This is less of a requirement than knowing the full model (which entails knowing the full function $T$), but still requires the agent to have some knowledge of the world.
 This category of algorithms, which are broadly termed ``Rolling Horizon'', has the following steps.
@@
 
-In standard UCT, a procedure similar to temporal difference learning is followed (or, more closely, monte carlo methods), where one learns the value of each state.
+In standard UCT, a procedure similar to temporal difference learning is followed (or, more closely, Monte Carlo methods), where one learns the value of each state.
 An agent then forms a policy by choosing a value greedily.
@@
 Our methodology is simple, yet efficient.
-We test T-HOLOP and EVO-P in two simple benchmark problems (see below for more) and Lunar Lander, a classic arcade game.
+We test T-HOLOP and EVO-P in two simple benchmark problems (see below for more detail) and n Lunar Lander, a classic arcade game.
 EVO-P is simply CMA-ES used in a rolling horizon setting, while T-HOLOP is T-HOO used in the same manner.
@@
 Weinstein and Littman~\cite{weinstein2012bandit} and Pazis and Lagoudakis \cite{PazisBAS},
-we used the same problems: double integrator, and inverted pendulum.  The basic
+we used the same problems: double integrator and inverted pendulum.  The basic
 experimental setup followed Weinstein and Littman~\cite{weinstein2012bandit}
@@
 
-The problem becomes harder as the time interval $\delta_t$ at %word missing?
-each action is applied increases,
+The problem becomes harder as the time interval $\delta_t$ that is applied at each action, increases,
 and following Pazis and Lagoudakis we set $(\delta_t = 0.5s)$.
@@
 \subsection{Lunar Lander}
-In this section we give a brief overview of the Game of Lunar Lander, a popular arcade game.
+In this section we give a brief overview of the game of Lunar Lander, a popular arcade game.
 
@@
 
-The spaceship within this game world is modelled as an a circular mass with some basic physical properties, such as a position within the game world $\bm {s}$, a velocity $\bm {v}$, an orientation $\bm{d}$, an angular velocity $\omega$ and a radius of a bounding sphere $r$ used for collision detection with the landscape.
+The spaceship within this game world is modelled as a circular mass with some basic physical properties, such as a position within the game world $\bm {s}$, a velocity $\bm {v}$, an orientation $\bm{d}$, an angular velocity $\omega$ and a radius of a bounding sphere $r$ used for collision detection with the landscape.
 
@@
 The ship colliding with the landscape constitutes the end of the {\itshape Lunar Lander} game.
-The conditions surrounding this collision, including speed, orientation of the ship constitute whether the nature of the collision is a success or a failure.
-In this experimental setup with accept a velocity of $5$ and an angle of $< 0.1$ radians as successful landing criteria.
+The conditions surrounding this collision, including speedand orientation of the ship, constitute whether the nature of the collision is a success or a failure.
+In this experimental setup we accept a velocity of $5$ and an angle of $< 0.1$ radians as successful landing criteria.
 
-An agent can modify the ship directions by applying linear thrust $\bm{a_T}$ and the angular impulse $\bm{a_I}$.
-The controllers have to identify a vector of real numbers $\bm{A}$, which each element of the set $a_i \in [0,1]$.
+An agent can modify the ship's direction by applying linear thrust $\bm{a_T}$ and angular impulse $\bm{a_I}$.
+The controllers have to identify a vector of real numbers $\bm{A}$, with each element of the set $a_i \in [0,1]$.
 Each even element of the vector represents the linear thrust, whereas each odd element represents angular velocity.
@@
 Lunar Lander has a setup where an agent gets rewards if and only if it manages to actually land the ship on the pad.
-This creates the problem of rollouts (whether used by EVO-P or T-HOLOP) rarely will see a reward.
-In order to circumvent this, as it is common in other games, we use a heuristic for each state.
+This creates the problem that rollouts (whether used by EVO-P or T-HOLOP) will rarely see a reward.
+In order to circumvent this, as it is a problem common to other games as well, we use a heuristic for each state.
 The heuristic was partially inspired by the two-stage approach found here : \url{http://www.cs.berkeley.edu/~jrs/4/lunar.html}\cite{lunar}.
-Initially, we need define a number of vectors that portray ship directions:
+Initially, we need to define a number of vectors that portray ship directions:
 
@@
 The vectors defined in Equation \ref{eq:vectors} define three directions for the ship.
-The first one, $\bm{du}$, define a ship standing upright, the second one, $\bm{dl}$, define tilting to the left, while the third one $\bm{dr}$ a ship tilting to the right.
-We make use of these vectors in order to help align the ship in the right direction, depending on which way the landing point is.
+The first one, $\bm{du}$, defines a ship standing upright, the second one, $\bm{dl}$, defines it tilting to the left, while the third one $\bm{dr}$ tilting to the right.
+We make use of these vectors in order to help align the ship in the right direction, depending on where the landing point is.
 
@@
 In the above Equation \ref{eq:h1} the constant C denotes the maximum magnitude of speed, which we set to $C=30$.
-$D = 60$ is the distance from the landing point $np$ where this heuristic takes precedence.
+$D = 60$ is the distance from the landing point $np$, where this heuristic takes precedence.
 Intuitively, the upper part of the heuristic $h$ tries to get the ship above the landing point as fast as possible while maintaining a certain speed.
-The second part of the heuristic tries to minimise velocity on the $x$ axis (as the ship tries to land with maximum speed $C_l = 2$), maintain a minimum landing speed on the $y$ axis and get a certain angle (the summation term $f$) as close to vector $du$, i.e.
+The second part of the heuristic tries to minimise velocity on the $x$ axis (as the ship tries to land with maximum speed $C_l = 2$), maintain a minimum landing speed on the $y$ axis and get a certain angle (the summation term $f$) as close to vector $du$ as possible, i.e.
 help the ship stay upright through the duration of the rollout.
-It's worth noting the main reason why we use a heuristic of this kind, one that includes not only the final state, but makes use of the path of the state as well, is that there is not enough information to discern states if one takes into account ship direction only; without incorporating the angular velocity,  the ship might be spinning madly, but the end of the rollout might find it in the correct direction.
-Finally notice that the heuristic $h$ gets the value $p=10000$ if the ship has crashed at any point during the rollout.
+It's worth noting that the main reason why we use a heuristic of this kind (one that includes not only the final state, but makes use of the path of the state as well) is that there is not enough information to discern states if one only takes into account ship direction.
+Without incorporating the angular velocity,  the ship might be spinning madly, but the end of the rollout might find it in the correct direction.
+Finally, notice that the heuristic $h$ gets the value $p=10000$ if the ship has crashed at any point during the rollout.
 
@@
 
- EVO-P got as score of $-41.69$.
-It also  seems to have a drop in performance as we increase function evaluations, presumably because it converges to a false minima without any uncertainty handling mechanism to stop this.
+ EVO-P got a score of $-41.69$.
+It also seems to experience a drop in performance as we increase function evaluations, presumably because it converges to a false minima without any uncertainty handling mechanism to stop this.
 This kind of behaviour has been observed previously in Evolutionary Algorithms~\cite{hansen2009tec} when having noisy function.
@@
 Using T-HOLOP we got a better mean of $-2.62$.
-A far better result of $-1.74$ is obtained by EVO-P.
+A far better result of $-1.74$ was obtained by EVO-P.
 
@@
 \label{tb:results}
-\caption{EVO-P results for different lunar lander maps}
+\caption{EVO-P results for different Lunar Lander maps}
 \end{table}
 
-For lunar lander, we do $10$ runs for each map in Figure \ref{fig:maps}, and the results for EVO-P can be seen in Table I.
-The results include the mean number of successful landings (\emph{Success Mean}) and the mean time it took to land the ship (\emph{Ticks Mean}), alongside their confidence for the $95$th interval.
-For EVO-P we used the default CMA-ES population size, each individual was bound between, each gene had a length of $8$ and each action pair ($x[i], x[i+1]$) was repeated $7$ times (where $x[i]$ represents thrust and $x[i+1]$ angular thrust).
-We derived these numbers experimentally and we did not optimise for them (not even implicitly) to any serious degree.
+For Lunar Lander, we perform $10$ runs for each map in Figure \ref{fig:maps}, and the results for EVO-P can be seen in Table I.
+The results include the mean number of successful landings (\emph{Success Mean}) and the mean time it took to land the ship (\emph{Ticks Mean}), alongside the confidence for the $95$th interval.
+For EVO-P we used the default CMA-ES population size, each individual gene bound between $[0,1]$, each gene had a length of $8$ and each action pair ($x[i], x[i+1]$) was repeated $7$ times (where $x[i]$ represents thrust and $x[i+1]$ angular thrust).
+We derived these numbers experimentally and did not optimise for them (not even implicitly) to any serious degree.
 
 There are a number of things to notice here.
-The first one is that the only in Map D, where the ship starts the problem facing a wall do we get a failure to land, and this is due to a crash.
+The first is that only in Map D, where the ship starts the problem facing a wall, do we get a failure to land, and this is due to a crash.
 This has to do with the fact that even a minor thrust can push you a bit forward and drive the ship to a position where there is no longer the possibility of recovery from a crush (or the combination of that recovery is extremely hard to find).
 The second thing to observe is that we get there slowly, but this is by design, as we he have made the flying speed of the ship controllable.
-One can potentially tweak all the default parameters we have used, plus the ones incorporated in CMA-ES as a kind of meta-heuristic search and achieve much better results.
-Another thing to note is that EVO-P lands the ship (on average) almost at the same time on MAPS B, C, D, although map has the same landscape as the other two but the ship is much closer.
+One can potentially tweak all the default parameters we have used, plus the ones incorporated in CMA-ES, as a kind of meta-heuristic search and achieve much better results.
+Another thing to note is that EVO-P lands the ship (on average) almost at the same time on MAPS B, C, D, although map B has the same landscape as the other two but the ship is much closer.
 This is due to the fact that the minimum landing speed, set on one of our heuristics, is really low.
-The final thing we would like to note is that we have only included the results for EVO-P because T-HOLOP failed to land the ship using the same heuristic, and behaved badly.
+The final thing we would like to point out is that we have only included the results for EVO-P because T-HOLOP failed to land the ship using the same heuristic, and behaved badly.
 It's worth noting here that T-HOLOP requires rewards between certain bounds, so they can be normalised around $[0,1]$ and this is not the case with our heuristic.
 For completeness purposes, we did change the UCB-like node selection policy of T-HOO to e-greedy, in which case we \emph{did} get the ship to behave reasonably, albeit the failure to land was evident here as well.
-This is important we think for a number of reasons.
+This, we think, is important for a number of reasons.
 The first one is that it shows that different Heuristics might be needed for different algorithms.
 Although this might be self-evident, it shows that one must design heuristics with a specific algorithm in mind.
-The second is, but we can speculate at the time, that T-HOLOP spends too much trying to solve a hard problem than the one at hand.
-It will converge eventually, but the the Guassian approximation of CMA-ES seems to outperform it in the short term.
+The second is, but we can speculate only at this time, that T-HOLOP spends too much effort trying to solve a hard problem than the one at hand.
+It will converge eventually, but the the Gaussian approximation of CMA-ES seems to outperform it in the short term.
 

### 2014-04-11 lpaper/lunarcont.tex
Added Diego Fixes

@@
 For larger games with complicated dynamics, having agents learn offline behaviours in such a setting is a daunting task.
-On the other hand, provided a forward model is available, one might try to spread the cost of search/learning in a rolling horizon fashion (e.g.
+On the other hand, provided a generative model is available, one might try to spread the cost of search/learning in a rolling horizon fashion (e.g.
 as in Monte Carlo Tree Search).
-In this paper we compare T-HOLOP (Truncated Hierarchical Open Loop Planning), an open loop planning algorithm at least partially inspired by MCTS, with a version of evolutionary planning that uses CMA-ES (which we call EVO-P) in two planning benchmark problems (Cart Balancing and the Double Integrator) and in Lunar Lander, a classic arcade game.
+In this paper we compare T-HOLOP (Truncated Hierarchical Open Loop Planning), an open loop planning algorithm at least partially inspired by MCTS, with a version of evolutionary planning that uses CMA-ES (which we call EVO-P) in two planning benchmark problems (Inverted Pendulum and the Double Integrator) and in Lunar Lander, a classic arcade game.
 We show that EVO-P outperforms T-HOLOP in the classic benchmarks, while T-HOLOP is unable to find a solution using the same heuristics.
@@
 
-Once a satisfactory plan/policy has been estimated (or the agent runs out of time, iterations etc.
-) it proceeds by executing the first action of the plan, thus moving to the next state.
+Once a satisfactory plan/policy has been estimated (or the agent runs out of time, iterations etc.) it proceeds by executing the first action of the plan, thus moving to the next state.
 Once there, the agent performs the same sequence of ``mental'' actions, (i.e.
@@
 Our methodology is simple, yet efficient.
-We test T-HOLOP and EVO-P in two simple benchmark problems (see below for more detail) and n Lunar Lander, a classic arcade game.
+We test T-HOLOP and EVO-P in two simple benchmark problems (see below for more detail) and in Lunar Lander, a classic arcade game.
 EVO-P is simply CMA-ES used in a rolling horizon setting, while T-HOLOP is T-HOO used in the same manner.
@@
 
-         \caption{Performance of both T-HOLOP and EVO-P in the Inverted Pendulum benchmark (error bars for the 95th percentile). Higher scores are better}
+         \caption{Performance of both T-HOLOP and EVO-P in the Inverted Pendulum benchmark (error bars for the 95th percentile). Higher scores are better.}
          \label{fig:IP-all}
@@
 
-       \caption{Performance of all algorithms in the Double Integrator benchmark (error bars for the 95th percentile). Higher scores are better}
+       \caption{Performance of all algorithms in the Double Integrator benchmark (error bars for the 95th percentile). Higher scores are better.}
 
@@
                 \includegraphics[width=1.0\textwidth]{./graphics/map2.pdf}
-                \caption{Map A,  where the ship is lower then the pad}
+                \caption{Map A,  where the ship is lower then the pad.}
                 \label{fig:map2}
@@
                 \includegraphics[width=1.0\textwidth]{./graphics/map1-easy.pdf}
-                \caption{Map B, an easy map where only the first heuristic is used}
+                \caption{Map B, an easy map where only the first heuristic is used.}
                 \label{fig:map1-easy}
@@
                 \includegraphics[width=1.0\textwidth]{./graphics/map1-tall.pdf}
-                \caption{Map C, where the ship is close to the top of the screen}
+                \caption{Map C, where the ship is close to the top of the screen.}
                 \label{fig:map1-tall}\end{subfigure}\begin{subfigure}{0.49\textwidth}
@@
                 \includegraphics[width=1.0\textwidth]{./graphics/map1-wall.pdf}
-                \caption{Map D, where the ship is behind a wall }
+                \caption{Map D, where the ship is behind a wall.}
                 \label{fig:map-wall}
@@
 
-         \caption{The four different maps used in our experiments. The maps capture different position/properties}
+         \caption{The four different maps used in our experiments. The maps capture different position/properties.}
          \label{fig:maps}
@@
 \label{tb:results}
-\caption{EVO-P results for different Lunar Lander maps}
+\caption{EVO-P results for different Lunar Lander maps.}
 \end{table}
@@
 \section*{Acknowledgements}
-This work was supported by EPSRC grant EP/H048588/1 entitled:``UCT for Games and Beyond'' and an EPSRC Ph.D studentship.
+This work was supported by EPSRC grant EP/H048588/1 entitled: ``UCT for Games and Beyond'' and an EPSRC Ph.D studentship.
 

### 2014-05-28 paper/paper.tex
last version
changes after proof read

@@
 
-\author{Diego Perez, {\it Student Member, IEEE}, Sanaz Mostaghim {Member, IEEE}, Spyridon Samothrakis, {\it Student Member, IEEE}, Simon M. Lucas, {\it Senior Member, IEEE}}
+\author{Diego Perez, {\it Student Member, IEEE}, Sanaz Mostaghim, {\it Member, IEEE}, Spyridon Samothrakis, {\it Student Member, IEEE}, Simon M. Lucas, {\it Senior Member, IEEE}}
 
@@
 	\includegraphics[scale=0.235]{img/moopDec.png}
-	\caption{Decision and objective spaces in a MOOP with two variables ($x_1$ and $x_2$) and two objectives ($f_1$ and $f_2$).
+	\caption{Decision and objective spaces in a MOP with two variables ($x_1$ and $x_2$) and two objectives ($f_1$ and $f_2$).
 In the objective space, yellow dots are non-optimal objective vectors, while blue dots form a non-dominated set.}
@@
 
-There are many different algorithms in the litrature which are proposed to tackle multi-objective optimization problems~\cite{Deb2001}.
-Weighted-sum approach is one of the traditional methods in which the objectives are weighted according to user preference.
-The sum of the weighted objectives builds one objective function which is to be optimized.
+There are many different algorithms in the literature that are proposed to tackle multi-objective optimization problems~\cite{Deb2001}.
+A weighted-sum approach is one of the traditional methods in which the objectives are weighted according to user preference.
+The sum of the weighted objectives builds one objective function to be optimized.
 The solution to this single-objective problem is one certain solution, ideally on a Pareto-front.
@@
 The goal of EMO algorithms is to find a set of optimal solutions with a good convergence to the Pareto-front as well as a good spread and diversity along the Pareto-front.
-One of the most well-known algorithms in the literature is the Non-dominated Sorting Evolutionary Algorithm II (NSGA-{II})~\cite{Deb2002}, pseudocode of which is shown in Algorithm~\ref{alg:nsga2}.
-As in any evolutionary algorithm, NSGA-{II} evolves a set of $N$ individuals in a population denoted as $Pop$ with the difference that they are ranked according to dominance criterion and a crowding distance measure which is used to maintain a good diversity of solutions.
+One of the most well-known algorithms in the literature is the Non-dominated Sorting Evolutionary Algorithm II (NSGA-{II})~\cite{Deb2002}, the pseudocode of which is shown in Algorithm~\ref{alg:nsga2}.
+As in any evolutionary algorithm, NSGA-{II} evolves a set of $N$ individuals in a population denoted as $Pop$ with the difference that they are ranked according to a dominance criterion and a crowding distance measure, which are used to maintain a good diversity of solutions.
 
@@
 \item Crowding distance: The function $\Call{crowdingDistanceAssignment}{F_i}$ assigns to each one of the individuals in $F_i$ a crowding distance value, which is the distance between the individual and its neighbours.
-The individuals with the smallest crowding distances are selected with a lower probability than the ones with the larger values.
+The individuals with the smallest crowding distances are selected with a lower probability than the ones with larger values.
 \item Elitism: The solutions in the first front $F_0$ always survive to the next population. 
@@
 
-The dynamics of an RL problem are usually captured by a Markov Decision Process (MDP) which is a tuple $(S, A, T, R)$, where $S$ is the set of possible states in the problem (or game), and $s_0$ is the initial state.
+The dynamics of an RL problem are usually captured by a Markov Decision Process (MDP) which is a tuple $(S, A, T, R)$, in which $S$ is the set of possible states in the problem (or game), and $s_0$ is the initial state.
 $A$ is the set of available actions the agent can make at any given time, and the transition model $T(s_i, a_i, s_{i+1})$ determines the probability of reaching the state $s_{i+1}$ when action $a_i$ is applied in state $s_i$.
@@
 Thus, MORL problems differ from RL in having more than one objective (here $m$) that must be maximized.
-If the objectives are independent or non-conflicting, scalarization approaches such as weighted sum approach, could be suitable to tackle the problem.
+If the objectives are independent or non-conflicting, scalarization approaches such as the weighted sum approach, could be suitable to tackle the problem.
 Essentially, this would mean using a conventional RL algorithm on a single objective where the global reward is obtained from a weighted-sum of the multiple rewards.
@@
 
-The proposed algorithm was employed successfully in two domains: the DST and the Grid Scheduling problem, matching state-of-the-art results in both cases, at the expense of a high computational cost.
+The proposed algorithm was employed successfully in two domains: the DST and the Grid Scheduling problem.
+In both cases, the results matched the state-of-the-art, albeit at the expense of a high computational cost.
 
 As mentioned before, the objective of this paper is to propose an MO-MCTS algorithm that is suitable for real-time domains.
-Although the work discussed here is influenced by Wang's approach, some modifications need to be done in order to overcome the high computational cost involved in their approach.
+Although the work discussed here is influenced by Wang's approach, some modifications need to be made in order to overcome the high computational cost involved in their approach.
 
@@
 Here, if $\overline{r}$ is not dominated by the local Pareto front, it is added to the front and $\overline{r}$ is propagated to its parent.
-In case $\overline{r}$ is dominated by the node's Pareto front, the local Pareto front does not change and there is no need to keep this propagation up the tree.
+If $\overline{r}$ is dominated by the node's Pareto front, the local Pareto front does not change and there is no need to maintain this propagation up the tree.
 Three observations can be made about the mechanism described here:
@@
 
-As shown before, it is straightforward to discover which actions lead to what points in the front given as a solution: first gene in the best NSGA-II individual, a root's child in MO-MCTS.
-Hence, by identifying the point in the Pareto front that the algorithm should converge to,
-it is possible to execute the action that leads to that point.
+As shown before, it is straightforward to discover which actions lead to what points in the front given as a solution: the first gene in the best NSGA-II individual, a root's child in MO-MCTS.
+Hence, by identifying the point in the Pareto front that the algorithm should converge to, it is possible to execute the action that leads to that point.
 
@@
 Note that, in the vanilla MCTS, there is no Pareto front obtained as a solution.
-Typically, in this case, rewards are calculated as a weighted sum of the objectives and a weight vector $W$, and then the action is chosen following any of the mechanisms usually employed in the literature: the action taken more often from the root; the one that leads to the best reward found; the move with the highest expected reward; or the action that maximizes Equation~\ref{eq:ucb1} in the root.
+Typically, in this case, rewards are calculated as a weighted sum of the objectives and a weight vector $W$.
+The action is then chosen following any of the mechanisms usually employed in the literature: the action taken more often from the root; the one that leads to the best reward found; the move with the highest expected reward; or the action that maximizes Equation~\ref{eq:ucb1} in the root.
 
@@
 The ship can perform four different moves: \textit{up}, \textit{down}, \textit{right}  and \textit{left}.
-If the action applied takes the ship off the grid or into the sea floor, the vessel's position does not change.
-There are two objectives in this game: the number of moves performed by the ship, that must be minimized, and the value of the treasure found, that should be maximized.
+If the action applied takes the ship off the grid or into the sea floor, the vessel's position will not change.
+There are two objectives in this game: the number of moves performed by the ship, which must be minimized, and the value of the treasure found, which should be maximized.
 As can be seen in Figure~\ref{fig:dstBase}, the most valuable treasures are at a greater distance from the initial position, so the objectives are in conflict.
@@
 
-The DST is a problem specially suited for the use of Transposition Tables (TT)~\cite{Childs2008} within MCTS.
-TT is a technique used to optimize three search based algorithms when two or more states can be found at different locations in the tree.
+The DST is a problem especially suited for the use of Transposition Tables (TT)~\cite{Childs2008} within MCTS.
+TT is a technique used to optimize tree search-based algorithms when two or more states can be found at different locations in the tree.
 It consists of sharing information between these equivalent states in a centralized manner, in order to avoid managing these positions as completely different states.
@@
 TT tables are implemented in the MCTS algorithms tested in this benchmark by using hash tables that store a \textit{representative} node for each pair $(position, moves)$ found.
-The key of the hash map needs then to be obtained from three values: position coordinates $x$ and $y$ of the ship in the board, and number of moves, indicated by the depth of the node in the tree.
+The key of the hash map then needs to be obtained from three values: position coordinates $x$ and $y$ of the ship in the board, and number of moves, indicated by the depth of the node in the tree.
 Hence, transpositions can only happen at the same depth within the tree, a feature that has been successfully tried before in the literature~\cite{Kozelek2009}.
@@
 As DST has two different objectives, number of moves and value of the treasure, the quality of a state can be assessed by two rewards, $\rho_p$ and $\rho_v$, for each objective respectively.
-$\rho_p$ is adjusted to be maximized using the maximum number of moves in the game, while $\rho_v$ is just the value in the cell with the treasure.
+$\rho_p$ is adjusted to be maximized using the maximum number of moves in the game, while $\rho_v$ is simply the value of the cell that holds the treasure.
+Therefore, the reward vector to maximize is defined as $\overline{r} = \{ \rho_p, \rho_v\}$.
 Equation~\ref{eq:dst} summarizes these rewards:
@@
 
-This needs to be done while optimizing three different objectives:  1) \textbf{Time}: the player must collect all waypoints scattered around the maze in as few time steps as possible; 2) \textbf{Fuel}: the fuel consumption by the end of the game must be minimized; and 3) \textbf{Damage}: the ship should end the game with as little damage as possible.
+This needs to be accomplished while optimizing three different objectives:  1) \textbf{Time}: the player must collect all waypoints scattered around the maze in as few time steps as possible; 2) \textbf{Fuel}: the fuel consumption by the end of the game must be minimized; and 3) \textbf{Damage}: the ship should end the game with as little damage as possible.
 
@@
 Regarding the third objective, the ship can suffer damage in two different ways: by colliding with obstacles and driving through lava.
-In the former case, the ship can collide with normal obstacles (that subtract $10$ units of damage) and specially damaging obstacles ($30$ units).
+In the former case, the ship can collide with normal obstacles (that subtract $10$ units of damage) and especially damaging obstacles ($30$ units).
 In the latter, lava lakes are abundant in MO-PTSP levels and, in contrast with normal surfaces, they deal one unit of damage for each time step the ship spends over this type of surface.
@@
 
-All algorithms tested in the MO-PTSP in this research employ macro-actions, a concept that can be used for coarsening the action space by applying the action chosen by the control algorithm in several consecutive cycles, instead of just in the next one.
+All the algorithms tested in the MO-PTSP in this research employ macro-actions, a concept that can be used for coarsening the action space by applying the action chosen by the control algorithm in several consecutive cycles, instead of just in the next one.
 
@@
 Note that, as waypoints need to keep being visited, it is necessary to include a distance measure for this reward.
-Otherwise, an approach that prioritizes this objective would not minimize distance to waypoints at all (the ship could just stand still: no fuel consumed is optimal), and therefore could not complete the game.
-The value of $\alpha$ has been determined empirically, in order to provide a good balance between these two components, and it is set to $0.66$.
-
-Finally, Equation~\ref{eq:dam} gives the method to calculate the damage objective, $\rho_d$:
+Otherwise, an approach that prioritizes this objective would not minimize distance to waypoints at all (the ship could just stand still: no fuel consumed is optimal), and therefore would not complete the game.
+The value of $\alpha$ has been determined empirically, in order to provide a good balance between these two components, and is set to $0.66$.
+
+Finally, Equation~\ref{eq:dam} gives the method used to calculate the damage objective, $\rho_d$:
 
@@
 This is distinguished by the parameter $\gamma$, that can be seen as a threshold value for the ship's speed ($sp$).
-This differentiation is made in order to avoid low speeds in lava lakes, as this increase the damage suffered significantly.
+This differentiation is made in order to avoid low speeds in lava lakes, as this significantly increases the damage suffered.
 The values for these variables have also been determined empirically, and they are set to $\gamma = 0.8$, $\beta_{1} = 0.75$ and $\beta_{2} = 0.25$.
 
-It is important to highlight again that these three rewards, $\rho_t$, $\rho_f$ and $\rho_d$, are the same for the three algorithms tested in the experiments.
+The final reward vector to be maximized is therefore $\overline{r} = \{ \rho_t, \rho_f, \rho_d \}$.
+It is important to highlight again that these three rewards are the same for all the algorithms tested in the experiments.
 
@@
 This NSGA-II version evolves a population where the individuals are sequences of actions (macro-actions in the MO-PTSP case), obtaining the fitness from the state of the game after applying the sequence.
-The populations sizes, determined empirically, were set to $20$ and $50$ individuals for DST and MO-PTSP respectively.
+The population sizes, determined empirically, were set to $20$ and $50$ individuals for DST and MO-PTSP respectively.
 The value of $C$ in Equation~\ref{eq:ucb1} and~\ref{eq:moucb} is set to $\sqrt{2}$.
 
-All algorithms have a limited number of evaluations before providing an action to perform.
+All the algorithms have a limited number of evaluations before providing an action to perform.
 In order for these games to be real-time, the time budget allowed is close to $40$ milliseconds.
@@
 The results show clearly how approximating the optimal Pareto front allows for finding all possible solutions.
-Both NSGA-II and MO-MCTS approaches are able to converge to any solution in the front given the appropriate weight vector.
-It is important to highlight that these two algorithms employed the euclidean distance action selection (see Section~\ref{sec:momcts}).
+Both the NSGA-II and MO-MCTS approaches are able to converge to any solution in the front given the appropriate weight vector.
+It is important to highlight that these two algorithms employed the Euclidean distance action selection (see Section~\ref{sec:momcts}).
 Other experiments, not included in this paper, showed that weighted sum action selection provides similar results to MCTS (that is, convergence to the edges of the front).
-The crucial distinction to make here is that both algorithms, NSGA-II and MO-MCTS, allow for better action selection mechanisms, that are able to overcome this problem by approximating a global Pareto front.
+The crucial distinction to make here is that both algorithms, NSGA-II and MO-MCTS, allow for better action selection mechanisms, which are able to overcome this problem by approximating a global Pareto front.
 
@@
 The experiments are performed in the $10$ maps available within the framework~\footnote{Available at \url{www.ptsp-game.net}}.
-In this case, the optimal Pareto front is not known in advance, and normally it will differ from one map to another.
+In this case, the optimal Pareto front is not known in advance, and normally differs from one map to another.
 Hence, the mechanisms to compare the performance of the algorithms tested need to be different than the one used for DST.
@@
 First of all, $4$ different weight vectors are tested: $W_1 = (0.33,0.33,0.33)$, $W_2 = (0.1,0.3,0.6)$, $W_3 = (0.1,0.6,0.3)$ and $W_4 = (0.6,0.1,0.3)$, where each $w_i$ corresponds to the weight for an objective ($w_t$ for time, $w_f$ for fuel and $w_d$ for damage, in this order).
-$W_1$ treats all objectives with the same weight, while the other three give more relevance to different objectives.
+$W_1$ treats all objectives as having the same weight, while the other three give more relevance to different objectives.
 These vectors then provide a wide spectrum for weights in this benchmark.
@@
 
-Some exceptions can be explained by analysing the map: in map 1, a fuel canister is always collected near the end, restoring the fuel level to its maximum.
+Some exceptions can be explained by analysing the specific maps: in map 1, a fuel canister is always collected near the end, restoring the fuel level to its maximum.
 This leaves too few cycles to make a difference in the controller (note also that this is the map with the smallest fuel consumption).
-Also, in map 10, there is no difference in the damage objective: but map 10 is a map with no obstacles to damage the ship (thus only lava lakes deal damage).
+Also, in map 10, there is no difference in the damage objective, however map 10 is a map with no obstacles to damage the ship (thus only lava lakes deal damage).
 This also results in this map being the one with the lowest overall damage.
@@
 Extending this comparison to all maps, each pair of algorithms ends with a triplet $(D, \O, d)$, where $D$ is the number of maps where the first algorithm dominates the second, $\O$ is the amount of maps where no dominance can be established, and $d$ states the number of maps where the first is dominated by the second.
-Table~\ref{tab:resMOPTSP} summarizes these results for all algorithms tested.
+Table~\ref{tab:resMOPTSP} summarizes these results for all the algorithms tested.
 
@@
 \caption{Results in MO-PTSP: Each cell indicates the triplet $(D, \O, d)$, where $D$ is the number of maps where the row algorithm dominates the column one, $\O$ is the amount of maps where no dominance can be established, and $d$ states the number of maps where the row algorithm is dominated by the column one.
-All algorithms followed the same route (order of waypoints and fuel canisters) in every map tested.}
+All the algorithms followed the same route (order of waypoints and fuel canisters) in every map tested.}
 \label{tab:resMOPTSP}
@@
 One of the first things to notice is that MO-MCTS dominates MCTS and NSGA-II in most of the maps, and it is never dominated in any.
-Particularly, the dominance of MO-MCTS over MCTS is outstanding, even dominating in all $10$ maps for two of the weight vectors.
-MO-MCTS also dominates NSGA-II in more maps than in those where there is no dominance, and neither is it ever dominated by NSGA-II in any map.
+In particular, the dominance of MO-MCTS over MCTS is outstanding, even dominating in all $10$ maps for two of the weight vectors.
+MO-MCTS also dominates NSGA-II in more maps than in those where there is no dominance, and  is never dominated by NSGA-II in any map.
 
@@
 
-Nonetheless, the inclusion of \texttt{PurofMovio} in this comparison is not pointless: it is possible to asses the quality of the three algorithms tested here by comparing their performance relatively, against this high quality entry.
+Nonetheless, the inclusion of \texttt{PurofMovio} in this comparison is not pointless: it is possible to assess the quality of the three algorithms tested here by comparing their performance relatively, against this high quality entry.
 Attending to this criteria, it can be seen how MO-MCTS is the algorithm that is dominated less often by \texttt{PurofMovio}, producing similar results on an average of $4.75$ out of the $10$ maps, and being dominated in $5.25$ maps.
@@
 Specifically in the MO-PTSP, there are regions of the map where there are more obstacles or lava lakes, hence the ship is most likely to suffer higher damage there.
-Also, the route followed during the game affects the relative ideal speed between waypoints, or perhaps a fuel canister is going to be picked up, which will affect how the fuel objective can be managed.
+Also, the route followed during the game affects the relative ideal speed between waypoints, or perhaps a fuel canister may be picked up, which will affect how the fuel objective will be managed.
 In general, many real-time games go through different phases, with different objectives and priorities.
@@
 A way to provide different weights at different times in MO-PTSP is straightforward.
-Given the route of waypoints (and fuel canisters) that is being followed, one can divide it into \textit{segments}, where each segment starts and ends with a waypoint (or fuel canister).
+Given the route of waypoints (and fuel canisters) being followed, one can divide it into \textit{segments}, where each segment starts and ends with a waypoint (or fuel canister).
 Then, each segment can be assigned a particular weight vector $W$.
@@
 
-This section shows some initial results in order to test the third variant, by using a stochastic hill climbing algorithm on each map.
+This section shows some initial results obtained when testing the third variant, using a stochastic hill climbing algorithm on each map.
 The goal is to check if, by varying the weight vectors between segments, better solutions can be achieved.
@@
 
-Finally, the results described in this paper allow us to presume that multi-objective approaches can provide a high quality level of play in real-time games.
+Finally, the results described in this paper allow us to assume that multi-objective approaches can provide a high quality level of play in real-time games.
 Multi-objective problems pop up in different settings and by using MCTS we can balance not only between exploration and exploitation but between multiple objectives as well.

### 2014-05-28 paper/latexDiff.sh
diff paper for reviewers

@@ -1,0 +1,2 @@
+#!/bin/bash
+latexdiff paperVersion1.tex paper.tex > diff/paperDiff.tex

### 2014-05-28 paper/paper.tex
starting revision, response to reviewers

@@
 Application of such approaches could be real-time strategy games, where long term planning must be carried out in order to balance aspects such as attack units, defensive structures and resource gathering.
-This should allow allow smart balancing of heuristics.
+This should allow smart balancing of heuristics.
 
@@
 Secondly, the algorithm must be able to provide solutions across the multi-objective spectrum: by parametrizing the algorithm, it must be possible to prioritize one objective over the others and therefore converge to solutions according to these preferences.
+Therefore, the main contribution of this paper is the proposal of a multi-objective version of MCTS that is applicable to real-time domains and is able to provide different solutions across the multi-objective spectrum.
 
@@
 
-Each iteration of MCTS can be divided into several steps~\cite{Gelly2006}: \textit{Tree selection}, \textit{Expansion}, \textit{Monte Carlo simulation} and \textit{Back-propagation} (all summarized in Figure~\ref{fig:mcts}).
+Each iteration of MCTS can be divided into several steps, as introduced by G.
+Chaslot et al.~\cite{Chaslot2007}: \textit{Tree selection}, \textit{Expansion}, \textit{Monte Carlo simulation} and \textit{Back-propagation} (all summarized in Figure~\ref{fig:mcts}).
 When the algorithm starts, the tree is formed only by the root node, which holds the current state of the game.
@@
 In other words, it chooses between taking an action that leads to states with the best outcome found so far, and performing a move to go to less explored game states, respectively.
-In order to achieve this, MCTS uses Upper Confidence Bound (UCB1, see Equation~\ref{eq:ucb1}) as a \textit{Tree Policy}.
+In order to achieve this, Kocsis and Szepesv\'{a}ri~\cite{kocsis2006bandit} applied Upper Confidence Bound (UCB1, see Equation~\ref{eq:ucb1}) as a \textit{Tree Policy}.
 
@@
 Higher values of $C$ give added weight to the second term of the UCB1 Equation~\ref{eq:ucb1}, giving preference to those actions that have been explored less, at the expense of taking actions with the highest average reward $Q(s,a)$.
-A commonly used value is $\sqrt{2}$, as it balances both facets of the search when the rewards are normalized between $0$ and $1$.
+A commonly used value for single-player games is $\sqrt{2}$, as it balances both facets of the search when the rewards are normalized between $0$ and $1$.
+The value of $C$ is application dependant, and it may vary from game to game.
 It is worth noting that MCTS, when combined with UCB1, reaches asymptotically logarithmic regret on each node of the tree~\cite{coquelin2007bandit}.
@@
 MCTS has been employed extensively in real-time games.
-A clear example of this is the popular real-time game \textit{Ms.
+An example of this is the popular real-time game \textit{Ms.
 PacMan}.
@@
 
-MCTS has also been applied to single-player games, like SameGame~\cite{Matsumoto2010}, where the player's goal is to destroy contiguous tiles of the same colour, distributed in a rectangular grid.
-Another use of MCTS is in the popular puzzle Morpion Solitaire~\cite{Edelkamp2010}, a connection game where the goal is to link nodes of a graph with straight lines that must contain at least five vertices.
+MCTS has also been applied to single-player games, like SameGame~\cite{Schadd2008}, where the player's goal is to destroy contiguous tiles of the same colour, distributed in a rectangular grid.
+Another use of MCTS is in the popular puzzle Morpion Solitaire~\cite{Cazenave2007}, a connection game where the goal is to link nodes of a graph with straight lines that must contain at least five vertices.
 Finally, the PTSP has also been addressed by MCTS, both in the single-objective~\cite{Perez2013, Powley2012} and the multi-objective versions~\cite{Powley2013}.
@@
 
-As mentioned before, the objective of this paper is to propose an MO-MCTS algorithm that is suitable for real-time domains.
-Although the work discussed here is influenced by Wang's approach, some modifications need to be made in order to overcome the high computational cost involved in their approach.
+As mentioned before, one of the objectives of this paper is to propose an MO-MCTS algorithm that is suitable for real-time domains.
+Therefore, an approach different from Wang's is needed, in order to overcome the high computational cost involved in their approach.
 
@@
 Here, if $\overline{r}$ is not dominated by the local Pareto front, it is added to the front and $\overline{r}$ is propagated to its parent.
+It may also happen that $\overline{r}$ dominates all or some solutions of the local front, in which case the new solution is included and the dominated solutions are removed from the Pareto front.
 If $\overline{r}$ is dominated by the node's Pareto front, the local Pareto front does not change and there is no need to maintain this propagation up the tree.
@@
 
-\subsubsection{Transposition Tables in DST} \label{ssec:transTables}
+\subsection{Heuristics for DST} \label{ssec:heurDST}
 
@@
 
-\subsubsection{Heuristics for DST} \label{sssec:heurDST}
-
 As DST has two different objectives, number of moves and value of the treasure, the quality of a state can be assessed by two rewards, $\rho_p$ and $\rho_v$, for each objective respectively.
@@
 
-\subsubsection{Macro-actions for MO-PTSP} \label{sssec:macro}
+\subsection{Heuristics for MO-PTSP} \label{sssec:heurMOPTSP}
 
@@
 For more information about macro-actions and their application to real-time navigation problems such as the PTSP, the reader is referred to~\cite{Perez2013}.
-
-\subsubsection{Heuristics for MO-PTSP} \label{sssec:heurMOPTSP}
 
@@
 With the objective of avoiding congestion peaks at the machine where the experiments are run, the average number of evaluations possible in $40$ ms is calculated and employed in the tests.
-This leads to $4500$ evaluations in the DST, and $500$ evaluations for MO-PTSP, using the same server where the PTSP and MO-PTSP competitions were run\footnote{Intel Core i5, 230GHz, 4GB of RAM.}.
+This leads to $4500$ evaluations in the DST, and $500$ evaluations for MO-PTSP, using the same server where the PTSP and MO-PTSP competitions were run\footnote{Intel Core i5, 2.90GHz, 4GB of RAM.}.
 
 \subsection{Results in DST} \label{ssec:resDST}
-
-As the optimal Pareto front of the DST is known, a measure of performance can be obtained by observing the percentage of times these solutions are found by the players.
-As the solution that the algorithms converge to depends on the weights vector employed during the search (see end of Section~\ref{sec:momcts}), the approach taken here is to provide different weight vectors $W$ and analyze them separately.
-
-The weight vector for DST has two dimensions.
-This vector is here referred to as $W = (w_p, w_v)$, where $w_p$ weights moves; and $w_v$ weights the treasure value ($w_v = 1 - w_p$).
-$w_p$ takes values between $0$ and $1$, with a precision of $0.01$, and $100$ experiments have been run for each pair $(w_p, w_v)$.
-Hence, the game has been played a total of $10000$ times, for each algorithm.
 
@@
 
+As the optimal Pareto front of the DST is known, a measure of performance can be obtained by observing the percentage of times these solutions are found by the players.
+As the solution that the algorithms converge to depends on the weights vector employed during the search (see end of Section~\ref{sec:momcts}), the approach taken here is to provide different weight vectors $W$ and analyze them separately.
+
+The weight vector for DST has two dimensions.
+This vector is here referred to as $W = (w_p, w_v)$, where $w_p$ weights moves; and $w_v$ weights the treasure value ($w_v = 1 - w_p$).
+$w_p$ takes values between $0$ and $1$, with a precision of $0.01$, and $100$ runs have been performed for each pair $(w_p, w_v)$.
+Hence, the game has been played a total of $10000$ times, for each algorithm.
+
 Figure~\ref{fig:extCase} shows the results obtained after these experiments were performed.
@@
  & $(0.1,0.6,0.3)$ & $1681 \pm 11$ & $131 \pm 2$ & $837 \pm 15$\\
- & $(0.6,0.1,0.3)$ & \bm{$1649 \pm 8$} & $132 \pm 2$ & $833 \pm 13$\\
+ & $(0.6,0.1,0.3)$ & $1649 \pm 8$ & $132 \pm 2$ & $833 \pm 13$\\
 \hline
@@
  & $(0.1,0.6,0.3)$ & $1467 \pm 9$ & \bm{$411 \pm 5$} & $567 \pm 5$\\
- & $(0.6,0.1,0.3)$ & \bm{$1399 \pm 9$} & $469 \pm 4$ & $547 \pm 3$\\
+ & $(0.6,0.1,0.3)$ & $1399 \pm 9$ & $469 \pm 4$ & $547 \pm 3$\\
 \hline
@@
  & $(0.1,0.6,0.3)$ & $1703 \pm 11$ & \bm{$499 \pm 4$} & $316 \pm 7$\\
- & $(0.6,0.1,0.3)$ & \bm{$1571 \pm 7$} & $559 \pm 5$ & $294 \pm 4$\\
+ & $(0.6,0.1,0.3)$ & $1571 \pm 7$ & $559 \pm 5$ & $294 \pm 4$\\
 \hline
@@
 Now, it is time to compare these results with the other algorithms.
-In order to do this, the same number of experiments is run for NSGA-II and MCTS in the $10$ maps of the benchmark.
+In order to do this, the same number of runs is performed for NSGA-II and MCTS in the $10$ maps of the benchmark.
 

### 2014-05-29 paper/paper.tex
More modifications after first review.

@@
 	\begin{center}
-	\includegraphics[scale=0.235]{img/moopDec.png}
+	\includegraphics[scale=0.235]{img/moopDec2.png}
 	\caption{Decision and objective spaces in a MOP with two variables ($x_1$ and $x_2$) and two objectives ($f_1$ and $f_2$).
-In the objective space, yellow dots are non-optimal objective vectors, while blue dots form a non-dominated set.}
+In the objective space, yellow dots are non-optimal objective vectors, while blue dots form a non-dominated set.
+From~\cite{Brockhoff2013}.}
 	\label{fig:moop}
@@
 
-Extending this comparison to all maps, each pair of algorithms ends with a triplet $(D, \O, d)$, where $D$ is the number of maps where the first algorithm dominates the second, $\O$ is the amount of maps where no dominance can be established, and $d$ states the number of maps where the first is dominated by the second.
+Extending this comparison to all maps, each pair of algorithms ends with a triplet $(D, \O, d)$.
+$D$ is the number of maps where the first algorithm dominates the second.
+$\O$ is the amount of maps where no dominance can be established, either because the Mann-Whitney-Wilcoxon non-parametric test failed, or because there is no dominance according to the dominance rules described in Section~\ref{sec:moo}.
+Finally, $d$ states the number of maps where the first is dominated by the second.
+For example, a triplet $(D, \O, d) = (8,2,0)$ comparing algorithms $A$ and $B$ would mean that the results obtained by $A$ dominate those from $B$ in $8$ of the $10$ maps, and that it is not possible to derive any dominance in the other $2$.
 Table~\ref{tab:resMOPTSP} summarizes these results for all the algorithms tested.

### 2014-05-31 paper/paper.tex
More changes. Before Sanaz's edits for revision 1.

@@
 The game is played in turns, and the objective is to surround the opponent's stones by placing stones in any available position on the board.
-Due to the very large branching factor of Go, this game is considered the drosophila of Game AI, and MCTS players have reached professional level play in the reduced board size version~\cite{Lee2009}.
+Due to the very large branching factor of the game and the absence of clear heuristics to tackle it, Go became object of study for many researchers.
+MCTS was the first algorithm able to reach professional level play in the reduced board size version~\cite{Lee2009}.
 After its success in Go, MCTS has been used extensively by many researchers in this and different domains.
@@
 As in any evolutionary algorithm, NSGA-{II} evolves a set of $N$ individuals in a population denoted as $Pop$ with the difference that they are ranked according to a dominance criterion and a crowding distance measure, which are used to maintain a good diversity of solutions.
+After each iteration of the algorithm, only the best $N$ individuals according to this ranking are maintained to the next generation.
 
@@
 The individuals with the smallest crowding distances are selected with a lower probability than the ones with larger values.
-\item Elitism: The solutions in the first front $F_0$ always survive to the next population. 
+\item Elitism: The individuals from the first front $F_0$ are always ranked first, according to the dominance criterion and their crowding distance.
+This ensures that the best solutions always survive to the next generation.
 \end{itemize}
@@
 
+As the optimal Pareto front of the DST is known, a measure of performance can be obtained by observing the percentage of times these solutions are found by the players.
+As the solution that the algorithms converge to depends on the weights vector employed during the search (see end of Section~\ref{sec:momcts}), the approach taken here is to provide different weight vectors $W$ and analyze them separately.
+
+The weight vector for DST has two dimensions.
+This vector is here referred to as $W = (w_p, w_v)$, where $w_p$ weights moves; and $w_v$ weights the treasure value ($w_v = 1 - w_p$).
+$w_p$ takes values between $0$ and $1$, with a precision of $0.01$, and $100$ runs have been performed for each pair $(w_p, w_v)$.
+Hence, the game has been played a total of $10000$ times, for each algorithm.
+
 \begin{figure*}[!t]
@@
 
-As the optimal Pareto front of the DST is known, a measure of performance can be obtained by observing the percentage of times these solutions are found by the players.
-As the solution that the algorithms converge to depends on the weights vector employed during the search (see end of Section~\ref{sec:momcts}), the approach taken here is to provide different weight vectors $W$ and analyze them separately.
-
-The weight vector for DST has two dimensions.
-This vector is here referred to as $W = (w_p, w_v)$, where $w_p$ weights moves; and $w_v$ weights the treasure value ($w_v = 1 - w_p$).
-$w_p$ takes values between $0$ and $1$, with a precision of $0.01$, and $100$ runs have been performed for each pair $(w_p, w_v)$.
-Hence, the game has been played a total of $10000$ times, for each algorithm.
-
 Figure~\ref{fig:extCase} shows the results obtained after these experiments were performed.
@@
 
+\begin{table}[!t]
+\begin{center}
+%\begin{tabular}{|>{\centering\arraybackslash}m{0.65cm}|c|c|c|c|}
+\begin{tabular}{|p{0.8cm}|p{2.2cm}|m{1.13cm}|m{1.13cm}|m{1.13cm}|}
+\hline
+\textbf{Map}  & \textbf{$W:(w_t, w_f, w_d)$} & \textbf{Time} &  \textbf{Fuel} & \textbf{Damage} \\ 
+\hline
+ \multirow{3}{*}{\textbf{Map 1}} & $(0.33,0.33,0.33)$ & $1654 \pm 7$ & $131 \pm 2$& $846 \pm 13$\\
+ & $(0.1,0.3,0.6)$ & $1657 \pm 8$ & $130 \pm 2$ & \bm{$773 \pm 11$}\\
+ & $(0.1,0.6,0.3)$ & $1681 \pm 11$ & $131 \pm 2$ & $837 \pm 15$\\
+ & $(0.6,0.1,0.3)$ & $1649 \pm 8$ & $132 \pm 2$ & $833 \pm 13$\\
+\hline
+ \multirow{3}{*}{\textbf{Map 2}} & $(0.33,0.33,0.33)$ & $1409 \pm 7$ & $235 \pm 4$& $364 \pm 3$\\
+ & $(0.1,0.3,0.6)$ & $1402 \pm 6$ & $236 \pm 5$ & \bm{$354 \pm 2$}\\
+ & $(0.1,0.6,0.3)$ & $1416 \pm 8$ & \bm{$219 \pm 4$} & $360 \pm 3$\\
+ & $(0.6,0.1,0.3)$ & $1396 \pm 8$ & $245 \pm 5$ & $361 \pm 2$\\
+\hline
+ \multirow{3}{*}{\textbf{Map 3}} & $(0.33,0.33,0.33)$ & $1373 \pm 6$ & $221 \pm 3$& $301 \pm 7$\\
+ & $(0.1,0.3,0.6)$ & $1378 \pm 5$ & $211 \pm 3$ & \bm{$268 \pm 5$}\\
+ & $(0.1,0.6,0.3)$ & $1385 \pm 6$ & \bm{$203 \pm 4$} & $291 \pm 7$\\
+ & $(0.6,0.1,0.3)$ & \bm{$1363 \pm 4$} & $229 \pm 4$ & $285 \pm 7$\\
+\hline
+ \multirow{3}{*}{\textbf{Map 4}} & $(0.33,0.33,0.33)$ & $1383 \pm 6$ & $291 \pm 5$& $565 \pm 5$\\
+ & $(0.1,0.3,0.6)$ & $1385 \pm 7$ & $304 \pm 4$ & \bm{$542 \pm 4$}\\
+ & $(0.1,0.6,0.3)$ & $1423 \pm 6$ & \bm{$273 \pm 4$} & $583 \pm 5$\\
+ & $(0.6,0.1,0.3)$ & $1388 \pm 6$ & $309 \pm 4$ & $559 \pm 5$\\
+\hline
+ \multirow{3}{*}{\textbf{Map 5}} & $(0.33,0.33,0.33)$ & $1405 \pm 7$ & $467 \pm 4$& $559 \pm 4$\\
+ & $(0.1,0.3,0.6)$ & $1431 \pm 9$ & $447 \pm 4$ & $541 \pm 4$\\
+ & $(0.1,0.6,0.3)$ & $1467 \pm 9$ & \bm{$411 \pm 5$} & $567 \pm 5$\\
+ & $(0.6,0.1,0.3)$ & $1399 \pm 9$ & $469 \pm 4$ & $547 \pm 3$\\
+\hline
+ \multirow{3}{*}{\textbf{Map 6}} & $(0.33,0.33,0.33)$ & $1575 \pm 7$ & $549 \pm 5$& $303 \pm 4$\\
+ & $(0.1,0.3,0.6)$ & $1626 \pm 9$ & $540 \pm 6$ & $286 \pm 5$\\
+ & $(0.1,0.6,0.3)$ & $1703 \pm 11$ & \bm{$499 \pm 4$} & $316 \pm 7$\\
+ & $(0.6,0.1,0.3)$ & $1571 \pm 7$ & $559 \pm 5$ & $294 \pm 4$\\
+\hline
+ \multirow{3}{*}{\textbf{Map 7}} & $(0.33,0.33,0.33)$ & $1434 \pm 5$ & $599 \pm 6$& $284 \pm 6$\\
+ & $(0.1,0.3,0.6)$ & $1475 \pm 10$ & $602 \pm 5$ & \bm{$243 \pm 6$}\\
+ & $(0.1,0.6,0.3)$ & $1489 \pm 12$ & \bm{$549 \pm 3$} & $264 \pm 6$\\
+ & $(0.6,0.1,0.3)$ & \bm{$1407 \pm 8$} & $618 \pm 5$ & $270 \pm 6$\\
+\hline
+ \multirow{3}{*}{\textbf{Map 8}} & $(0.33,0.33,0.33)$ & $1761 \pm 9$ & $254 \pm 5$& $382 \pm 3$\\
+ & $(0.1,0.3,0.6)$ & $1804 \pm 10$ & $269 \pm 4$ & \bm{$357 \pm 4$}\\
+ & $(0.1,0.6,0.3)$ & $1826 \pm 10$ & \bm{$230 \pm 3$} & $392 \pm 7$\\
+ & $(0.6,0.1,0.3)$ & \bm{$1732 \pm 9$} & $311 \pm 8$ & $379 \pm 6$\\
+\hline
+ \multirow{3}{*}{\textbf{Map 9}} & $(0.33,0.33,0.33)$ & $2501 \pm 14$ & $926 \pm 6$& $574 \pm 9$\\
+ & $(0.1,0.3,0.6)$ & $2503 \pm 10$ & $921 \pm 10$ & \bm{$524 \pm 8$}\\
+ & $(0.1,0.6,0.3)$ & $2641 \pm 14$ & \bm{$833 \pm 5$} & $574 \pm 14$\\
+ & $(0.6,0.1,0.3)$ & \bm{$2470 \pm 9$} & $956 \pm 5$ & $573 \pm 8$\\
+\hline
+ \multirow{3}{*}{\textbf{Map 10}} & $(0.33,0.33,0.33)$ & $1430 \pm 8$ & $630 \pm 4$& $205 \pm 2$\\
+ & $(0.1,0.3,0.6)$ & $1493 \pm 13$ & $615 \pm 4$ & $209 \pm 2$\\
+ & $(0.1,0.6,0.3)$ & $1542 \pm 10$ & \bm{$554 \pm 4$} & $229 \pm 5$\\
+ & $(0.6,0.1,0.3)$ & \bm{$1378 \pm 5$} & $663 \pm 6$ & $202 \pm 4$\\
+\hline
+\end{tabular}
+\caption{MO-MCTS results in MO-PTSP with different weight vectors. Values in bold indicate the best results on each map, when they are separated at least by $2$ standard errors from the others.}
+\label{tab:weights}
+\end{center}
+\end{table}
+
 The same three algorithms have been tested in the MO-PTSP domain.
@@
 It can be seen that the highest weight in $W$ leads, in most of the cases, to the best solution in that objective in the map.
-
-\begin{table}[!t]
-\begin{center}
-%\begin{tabular}{|>{\centering\arraybackslash}m{0.65cm}|c|c|c|c|}
-\begin{tabular}{|p{0.8cm}|p{2.2cm}|m{1.13cm}|m{1.13cm}|m{1.13cm}|}
-\hline
-\textbf{Map}  & \textbf{$W:(w_t, w_f, w_d)$} & \textbf{Time} &  \textbf{Fuel} & \textbf{Damage} \\ 
-\hline
- \multirow{3}{*}{\textbf{Map 1}} & $(0.33,0.33,0.33)$ & $1654 \pm 7$ & $131 \pm 2$& $846 \pm 13$\\
- & $(0.1,0.3,0.6)$ & $1657 \pm 8$ & $130 \pm 2$ & \bm{$773 \pm 11$}\\
- & $(0.1,0.6,0.3)$ & $1681 \pm 11$ & $131 \pm 2$ & $837 \pm 15$\\
- & $(0.6,0.1,0.3)$ & $1649 \pm 8$ & $132 \pm 2$ & $833 \pm 13$\\
-\hline
- \multirow{3}{*}{\textbf{Map 2}} & $(0.33,0.33,0.33)$ & $1409 \pm 7$ & $235 \pm 4$& $364 \pm 3$\\
- & $(0.1,0.3,0.6)$ & $1402 \pm 6$ & $236 \pm 5$ & \bm{$354 \pm 2$}\\
- & $(0.1,0.6,0.3)$ & $1416 \pm 8$ & \bm{$219 \pm 4$} & $360 \pm 3$\\
- & $(0.6,0.1,0.3)$ & $1396 \pm 8$ & $245 \pm 5$ & $361 \pm 2$\\
-\hline
- \multirow{3}{*}{\textbf{Map 3}} & $(0.33,0.33,0.33)$ & $1373 \pm 6$ & $221 \pm 3$& $301 \pm 7$\\
- & $(0.1,0.3,0.6)$ & $1378 \pm 5$ & $211 \pm 3$ & \bm{$268 \pm 5$}\\
- & $(0.1,0.6,0.3)$ & $1385 \pm 6$ & \bm{$203 \pm 4$} & $291 \pm 7$\\
- & $(0.6,0.1,0.3)$ & \bm{$1363 \pm 4$} & $229 \pm 4$ & $285 \pm 7$\\
-\hline
- \multirow{3}{*}{\textbf{Map 4}} & $(0.33,0.33,0.33)$ & $1383 \pm 6$ & $291 \pm 5$& $565 \pm 5$\\
- & $(0.1,0.3,0.6)$ & $1385 \pm 7$ & $304 \pm 4$ & \bm{$542 \pm 4$}\\
- & $(0.1,0.6,0.3)$ & $1423 \pm 6$ & \bm{$273 \pm 4$} & $583 \pm 5$\\
- & $(0.6,0.1,0.3)$ & $1388 \pm 6$ & $309 \pm 4$ & $559 \pm 5$\\
-\hline
- \multirow{3}{*}{\textbf{Map 5}} & $(0.33,0.33,0.33)$ & $1405 \pm 7$ & $467 \pm 4$& $559 \pm 4$\\
- & $(0.1,0.3,0.6)$ & $1431 \pm 9$ & $447 \pm 4$ & \bm{$541 \pm 4$}\\
- & $(0.1,0.6,0.3)$ & $1467 \pm 9$ & \bm{$411 \pm 5$} & $567 \pm 5$\\
- & $(0.6,0.1,0.3)$ & $1399 \pm 9$ & $469 \pm 4$ & $547 \pm 3$\\
-\hline
- \multirow{3}{*}{\textbf{Map 6}} & $(0.33,0.33,0.33)$ & $1575 \pm 7$ & $549 \pm 5$& $303 \pm 4$\\
- & $(0.1,0.3,0.6)$ & $1626 \pm 9$ & $540 \pm 6$ & \bm{$286 \pm 5$}\\
- & $(0.1,0.6,0.3)$ & $1703 \pm 11$ & \bm{$499 \pm 4$} & $316 \pm 7$\\
- & $(0.6,0.1,0.3)$ & $1571 \pm 7$ & $559 \pm 5$ & $294 \pm 4$\\
-\hline
- \multirow{3}{*}{\textbf{Map 7}} & $(0.33,0.33,0.33)$ & $1434 \pm 5$ & $599 \pm 6$& $284 \pm 6$\\
- & $(0.1,0.3,0.6)$ & $1475 \pm 10$ & $602 \pm 5$ & \bm{$243 \pm 6$}\\
- & $(0.1,0.6,0.3)$ & $1489 \pm 12$ & \bm{$549 \pm 3$} & $264 \pm 6$\\
- & $(0.6,0.1,0.3)$ & \bm{$1407 \pm 8$} & $618 \pm 5$ & $270 \pm 6$\\
-\hline
- \multirow{3}{*}{\textbf{Map 8}} & $(0.33,0.33,0.33)$ & $1761 \pm 9$ & $254 \pm 5$& $382 \pm 3$\\
- & $(0.1,0.3,0.6)$ & $1804 \pm 10$ & $269 \pm 4$ & \bm{$357 \pm 4$}\\
- & $(0.1,0.6,0.3)$ & $1826 \pm 10$ & \bm{$230 \pm 3$} & $392 \pm 7$\\
- & $(0.6,0.1,0.3)$ & \bm{$1732 \pm 9$} & $311 \pm 8$ & $379 \pm 6$\\
-\hline
- \multirow{3}{*}{\textbf{Map 9}} & $(0.33,0.33,0.33)$ & $2501 \pm 14$ & $926 \pm 6$& $574 \pm 9$\\
- & $(0.1,0.3,0.6)$ & $2503 \pm 10$ & $921 \pm 10$ & \bm{$524 \pm 8$}\\
- & $(0.1,0.6,0.3)$ & $2641 \pm 14$ & \bm{$833 \pm 5$} & $574 \pm 14$\\
- & $(0.6,0.1,0.3)$ & \bm{$2470 \pm 9$} & $956 \pm 5$ & $573 \pm 8$\\
-\hline
- \multirow{3}{*}{\textbf{Map 10}} & $(0.33,0.33,0.33)$ & $1430 \pm 8$ & $630 \pm 4$& $205 \pm 2$\\
- & $(0.1,0.3,0.6)$ & $1493 \pm 13$ & $615 \pm 4$ & $209 \pm 2$\\
- & $(0.1,0.6,0.3)$ & $1542 \pm 10$ & \bm{$554 \pm 4$} & $229 \pm 5$\\
- & $(0.6,0.1,0.3)$ & \bm{$1378 \pm 5$} & $663 \pm 6$ & $202 \pm 4$\\
-\hline
-\end{tabular}
-\caption{MO-MCTS results in MO-PTSP with different weight vectors. Significantly better results per map are in bold.}
-\label{tab:weights}
-\end{center}
-\end{table}
 

### 2014-05-31 paper/paperVersion1.tex
paper diff updated

--- 2014-05-28 paper/paper.tex
@@
 Application of such approaches could be real-time strategy games, where long term planning must be carried out in order to balance aspects such as attack units, defensive structures and resource gathering.
-This should allow smart balancing of heuristics.
+This should allow allow smart balancing of heuristics.
 
@@
 Secondly, the algorithm must be able to provide solutions across the multi-objective spectrum: by parametrizing the algorithm, it must be possible to prioritize one objective over the others and therefore converge to solutions according to these preferences.
-Therefore, the main contribution of this paper is the proposal of a multi-objective version of MCTS that is applicable to real-time domains and is able to provide different solutions across the multi-objective spectrum.
 
@@
 
-Each iteration of MCTS can be divided into several steps, as introduced by G.
-Chaslot et al.~\cite{Chaslot2007}: \textit{Tree selection}, \textit{Expansion}, \textit{Monte Carlo simulation} and \textit{Back-propagation} (all summarized in Figure~\ref{fig:mcts}).
+Each iteration of MCTS can be divided into several steps~\cite{Gelly2006}: \textit{Tree selection}, \textit{Expansion}, \textit{Monte Carlo simulation} and \textit{Back-propagation} (all summarized in Figure~\ref{fig:mcts}).
 When the algorithm starts, the tree is formed only by the root node, which holds the current state of the game.
@@
 In other words, it chooses between taking an action that leads to states with the best outcome found so far, and performing a move to go to less explored game states, respectively.
-In order to achieve this, Kocsis and Szepesv\'{a}ri~\cite{kocsis2006bandit} applied Upper Confidence Bound (UCB1, see Equation~\ref{eq:ucb1}) as a \textit{Tree Policy}.
+In order to achieve this, MCTS uses Upper Confidence Bound (UCB1, see Equation~\ref{eq:ucb1}) as a \textit{Tree Policy}.
 
@@
 Higher values of $C$ give added weight to the second term of the UCB1 Equation~\ref{eq:ucb1}, giving preference to those actions that have been explored less, at the expense of taking actions with the highest average reward $Q(s,a)$.
-A commonly used value for single-player games is $\sqrt{2}$, as it balances both facets of the search when the rewards are normalized between $0$ and $1$.
-The value of $C$ is application dependant, and it may vary from game to game.
+A commonly used value is $\sqrt{2}$, as it balances both facets of the search when the rewards are normalized between $0$ and $1$.
 It is worth noting that MCTS, when combined with UCB1, reaches asymptotically logarithmic regret on each node of the tree~\cite{coquelin2007bandit}.
@@
 MCTS has been employed extensively in real-time games.
-An example of this is the popular real-time game \textit{Ms.
+A clear example of this is the popular real-time game \textit{Ms.
 PacMan}.
@@
 
-MCTS has also been applied to single-player games, like SameGame~\cite{Schadd2008}, where the player's goal is to destroy contiguous tiles of the same colour, distributed in a rectangular grid.
-Another use of MCTS is in the popular puzzle Morpion Solitaire~\cite{Cazenave2007}, a connection game where the goal is to link nodes of a graph with straight lines that must contain at least five vertices.
+MCTS has also been applied to single-player games, like SameGame~\cite{Matsumoto2010}, where the player's goal is to destroy contiguous tiles of the same colour, distributed in a rectangular grid.
+Another use of MCTS is in the popular puzzle Morpion Solitaire~\cite{Edelkamp2010}, a connection game where the goal is to link nodes of a graph with straight lines that must contain at least five vertices.
 Finally, the PTSP has also been addressed by MCTS, both in the single-objective~\cite{Perez2013, Powley2012} and the multi-objective versions~\cite{Powley2013}.
@@
 	\includegraphics[scale=0.235]{img/moopDec.png}
-	\caption{Decision and objective spaces in a MOP with two variables ($x_1$ and $x_2$) and two objectives ($f_1$ and $f_2$).
+	\caption{Decision and objective spaces in a MOOP with two variables ($x_1$ and $x_2$) and two objectives ($f_1$ and $f_2$).
 In the objective space, yellow dots are non-optimal objective vectors, while blue dots form a non-dominated set.}
@@
 
-There are many different algorithms in the literature that are proposed to tackle multi-objective optimization problems~\cite{Deb2001}.
-A weighted-sum approach is one of the traditional methods in which the objectives are weighted according to user preference.
-The sum of the weighted objectives builds one objective function to be optimized.
+There are many different algorithms in the litrature which are proposed to tackle multi-objective optimization problems~\cite{Deb2001}.
+Weighted-sum approach is one of the traditional methods in which the objectives are weighted according to user preference.
+The sum of the weighted objectives builds one objective function which is to be optimized.
 The solution to this single-objective problem is one certain solution, ideally on a Pareto-front.
@@
 The goal of EMO algorithms is to find a set of optimal solutions with a good convergence to the Pareto-front as well as a good spread and diversity along the Pareto-front.
-One of the most well-known algorithms in the literature is the Non-dominated Sorting Evolutionary Algorithm II (NSGA-{II})~\cite{Deb2002}, the pseudocode of which is shown in Algorithm~\ref{alg:nsga2}.
-As in any evolutionary algorithm, NSGA-{II} evolves a set of $N$ individuals in a population denoted as $Pop$ with the difference that they are ranked according to a dominance criterion and a crowding distance measure, which are used to maintain a good diversity of solutions.
+One of the most well-known algorithms in the literature is the Non-dominated Sorting Evolutionary Algorithm II (NSGA-{II})~\cite{Deb2002}, pseudocode of which is shown in Algorithm~\ref{alg:nsga2}.
+As in any evolutionary algorithm, NSGA-{II} evolves a set of $N$ individuals in a population denoted as $Pop$ with the difference that they are ranked according to dominance criterion and a crowding distance measure which is used to maintain a good diversity of solutions.
 
@@
 \item Crowding distance: The function $\Call{crowdingDistanceAssignment}{F_i}$ assigns to each one of the individuals in $F_i$ a crowding distance value, which is the distance between the individual and its neighbours.
-The individuals with the smallest crowding distances are selected with a lower probability than the ones with larger values.
+The individuals with the smallest crowding distances are selected with a lower probability than the ones with the larger values.
 \item Elitism: The solutions in the first front $F_0$ always survive to the next population. 
@@
 
-The dynamics of an RL problem are usually captured by a Markov Decision Process (MDP) which is a tuple $(S, A, T, R)$, in which $S$ is the set of possible states in the problem (or game), and $s_0$ is the initial state.
+The dynamics of an RL problem are usually captured by a Markov Decision Process (MDP) which is a tuple $(S, A, T, R)$, where $S$ is the set of possible states in the problem (or game), and $s_0$ is the initial state.
 $A$ is the set of available actions the agent can make at any given time, and the transition model $T(s_i, a_i, s_{i+1})$ determines the probability of reaching the state $s_{i+1}$ when action $a_i$ is applied in state $s_i$.
@@
 Thus, MORL problems differ from RL in having more than one objective (here $m$) that must be maximized.
-If the objectives are independent or non-conflicting, scalarization approaches such as the weighted sum approach, could be suitable to tackle the problem.
+If the objectives are independent or non-conflicting, scalarization approaches such as weighted sum approach, could be suitable to tackle the problem.
 Essentially, this would mean using a conventional RL algorithm on a single objective where the global reward is obtained from a weighted-sum of the multiple rewards.
@@
 
-The proposed algorithm was employed successfully in two domains: the DST and the Grid Scheduling problem.
-In both cases, the results matched the state-of-the-art, albeit at the expense of a high computational cost.
-
-As mentioned before, one of the objectives of this paper is to propose an MO-MCTS algorithm that is suitable for real-time domains.
-Therefore, an approach different from Wang's is needed, in order to overcome the high computational cost involved in their approach.
+The proposed algorithm was employed successfully in two domains: the DST and the Grid Scheduling problem, matching state-of-the-art results in both cases, at the expense of a high computational cost.
+
+As mentioned before, the objective of this paper is to propose an MO-MCTS algorithm that is suitable for real-time domains.
+Although the work discussed here is influenced by Wang's approach, some modifications need to be done in order to overcome the high computational cost involved in their approach.
 
@@
 Here, if $\overline{r}$ is not dominated by the local Pareto front, it is added to the front and $\overline{r}$ is propagated to its parent.
-It may also happen that $\overline{r}$ dominates all or some solutions of the local front, in which case the new solution is included and the dominated solutions are removed from the Pareto front.
-If $\overline{r}$ is dominated by the node's Pareto front, the local Pareto front does not change and there is no need to maintain this propagation up the tree.
+In case $\overline{r}$ is dominated by the node's Pareto front, the local Pareto front does not change and there is no need to keep this propagation up the tree.
 Three observations can be made about the mechanism described here:
@@
 
-As shown before, it is straightforward to discover which actions lead to what points in the front given as a solution: the first gene in the best NSGA-II individual, a root's child in MO-MCTS.
-Hence, by identifying the point in the Pareto front that the algorithm should converge to, it is possible to execute the action that leads to that point.
+As shown before, it is straightforward to discover which actions lead to what points in the front given as a solution: first gene in the best NSGA-II individual, a root's child in MO-MCTS.
+Hence, by identifying the point in the Pareto front that the algorithm should converge to,
+it is possible to execute the action that leads to that point.
 
@@
 Note that, in the vanilla MCTS, there is no Pareto front obtained as a solution.
-Typically, in this case, rewards are calculated as a weighted sum of the objectives and a weight vector $W$.
-The action is then chosen following any of the mechanisms usually employed in the literature: the action taken more often from the root; the one that leads to the best reward found; the move with the highest expected reward; or the action that maximizes Equation~\ref{eq:ucb1} in the root.
+Typically, in this case, rewards are calculated as a weighted sum of the objectives and a weight vector $W$, and then the action is chosen following any of the mechanisms usually employed in the literature: the action taken more often from the root; the one that leads to the best reward found; the move with the highest expected reward; or the action that maximizes Equation~\ref{eq:ucb1} in the root.
 
@@
 The ship can perform four different moves: \textit{up}, \textit{down}, \textit{right}  and \textit{left}.
-If the action applied takes the ship off the grid or into the sea floor, the vessel's position will not change.
-There are two objectives in this game: the number of moves performed by the ship, which must be minimized, and the value of the treasure found, which should be maximized.
+If the action applied takes the ship off the grid or into the sea floor, the vessel's position does not change.
+There are two objectives in this game: the number of moves performed by the ship, that must be minimized, and the value of the treasure found, that should be maximized.
 As can be seen in Figure~\ref{fig:dstBase}, the most valuable treasures are at a greater distance from the initial position, so the objectives are in conflict.
@@
 
-\subsection{Heuristics for DST} \label{ssec:heurDST}
-
-The DST is a problem especially suited for the use of Transposition Tables (TT)~\cite{Childs2008} within MCTS.
-TT is a technique used to optimize tree search-based algorithms when two or more states can be found at different locations in the tree.
+\subsubsection{Transposition Tables in DST} \label{ssec:transTables}
+
+The DST is a problem specially suited for the use of Transposition Tables (TT)~\cite{Childs2008} within MCTS.
+TT is a technique used to optimize three search based algorithms when two or more states can be found at different locations in the tree.
 It consists of sharing information between these equivalent states in a centralized manner, in order to avoid managing these positions as completely different states.
@@
 TT tables are implemented in the MCTS algorithms tested in this benchmark by using hash tables that store a \textit{representative} node for each pair $(position, moves)$ found.
-The key of the hash map then needs to be obtained from three values: position coordinates $x$ and $y$ of the ship in the board, and number of moves, indicated by the depth of the node in the tree.
+The key of the hash map needs then to be obtained from three values: position coordinates $x$ and $y$ of the ship in the board, and number of moves, indicated by the depth of the node in the tree.
 Hence, transpositions can only happen at the same depth within the tree, a feature that has been successfully tried before in the literature~\cite{Kozelek2009}.
 
+\subsubsection{Heuristics for DST} \label{sssec:heurDST}
+
 As DST has two different objectives, number of moves and value of the treasure, the quality of a state can be assessed by two rewards, $\rho_p$ and $\rho_v$, for each objective respectively.
-$\rho_p$ is adjusted to be maximized using the maximum number of moves in the game, while $\rho_v$ is simply the value of the cell that holds the treasure.
-Therefore, the reward vector to maximize is defined as $\overline{r} = \{ \rho_p, \rho_v\}$.
+$\rho_p$ is adjusted to be maximized using the maximum number of moves in the game, while $\rho_v$ is just the value in the cell with the treasure.
 Equation~\ref{eq:dst} summarizes these rewards:
@@
 
-This needs to be accomplished while optimizing three different objectives:  1) \textbf{Time}: the player must collect all waypoints scattered around the maze in as few time steps as possible; 2) \textbf{Fuel}: the fuel consumption by the end of the game must be minimized; and 3) \textbf{Damage}: the ship should end the game with as little damage as possible.
+This needs to be done while optimizing three different objectives:  1) \textbf{Time}: the player must collect all waypoints scattered around the maze in as few time steps as possible; 2) \textbf{Fuel}: the fuel consumption by the end of the game must be minimized; and 3) \textbf{Damage}: the ship should end the game with as little damage as possible.
 
@@
 Regarding the third objective, the ship can suffer damage in two different ways: by colliding with obstacles and driving through lava.
-In the former case, the ship can collide with normal obstacles (that subtract $10$ units of damage) and especially damaging obstacles ($30$ units).
+In the former case, the ship can collide with normal obstacles (that subtract $10$ units of damage) and specially damaging obstacles ($30$ units).
 In the latter, lava lakes are abundant in MO-PTSP levels and, in contrast with normal surfaces, they deal one unit of damage for each time step the ship spends over this type of surface.
@@
 
-\subsection{Heuristics for MO-PTSP} \label{sssec:heurMOPTSP}
-
-All the algorithms tested in the MO-PTSP in this research employ macro-actions, a concept that can be used for coarsening the action space by applying the action chosen by the control algorithm in several consecutive cycles, instead of just in the next one.
+\subsubsection{Macro-actions for MO-PTSP} \label{sssec:macro}
+
+All algorithms tested in the MO-PTSP in this research employ macro-actions, a concept that can be used for coarsening the action space by applying the action chosen by the control algorithm in several consecutive cycles, instead of just in the next one.
 
@@
 
+\subsubsection{Heuristics for MO-PTSP} \label{sssec:heurMOPTSP}
+
 In order to evaluate a game state in MO-PTSP, three different measures or rewards are taken, $\rho_t, \rho_f, \rho_d$, one for each one of the objectives: time, fuel and damage, respectively.
@@
 Note that, as waypoints need to keep being visited, it is necessary to include a distance measure for this reward.
-Otherwise, an approach that prioritizes this objective would not minimize distance to waypoints at all (the ship could just stand still: no fuel consumed is optimal), and therefore would not complete the game.
-The value of $\alpha$ has been determined empirically, in order to provide a good balance between these two components, and is set to $0.66$.
-
-Finally, Equation~\ref{eq:dam} gives the method used to calculate the damage objective, $\rho_d$:
+Otherwise, an approach that prioritizes this objective would not minimize distance to waypoints at all (the ship could just stand still: no fuel consumed is optimal), and therefore could not complete the game.
+The value of $\alpha$ has been determined empirically, in order to provide a good balance between these two components, and it is set to $0.66$.
+
+Finally, Equation~\ref{eq:dam} gives the method to calculate the damage objective, $\rho_d$:
 
@@
 This is distinguished by the parameter $\gamma$, that can be seen as a threshold value for the ship's speed ($sp$).
-This differentiation is made in order to avoid low speeds in lava lakes, as this significantly increases the damage suffered.
+This differentiation is made in order to avoid low speeds in lava lakes, as this increase the damage suffered significantly.
 The values for these variables have also been determined empirically, and they are set to $\gamma = 0.8$, $\beta_{1} = 0.75$ and $\beta_{2} = 0.25$.
 
-The final reward vector to be maximized is therefore $\overline{r} = \{ \rho_t, \rho_f, \rho_d \}$.
-It is important to highlight again that these three rewards are the same for all the algorithms tested in the experiments.
+It is important to highlight again that these three rewards, $\rho_t$, $\rho_f$ and $\rho_d$, are the same for the three algorithms tested in the experiments.
 
@@
 This NSGA-II version evolves a population where the individuals are sequences of actions (macro-actions in the MO-PTSP case), obtaining the fitness from the state of the game after applying the sequence.
-The population sizes, determined empirically, were set to $20$ and $50$ individuals for DST and MO-PTSP respectively.
+The populations sizes, determined empirically, were set to $20$ and $50$ individuals for DST and MO-PTSP respectively.
 The value of $C$ in Equation~\ref{eq:ucb1} and~\ref{eq:moucb} is set to $\sqrt{2}$.
 
-All the algorithms have a limited number of evaluations before providing an action to perform.
+All algorithms have a limited number of evaluations before providing an action to perform.
 In order for these games to be real-time, the time budget allowed is close to $40$ milliseconds.
 With the objective of avoiding congestion peaks at the machine where the experiments are run, the average number of evaluations possible in $40$ ms is calculated and employed in the tests.
-This leads to $4500$ evaluations in the DST, and $500$ evaluations for MO-PTSP, using the same server where the PTSP and MO-PTSP competitions were run\footnote{Intel Core i5, 2.90GHz, 4GB of RAM.}.
+This leads to $4500$ evaluations in the DST, and $500$ evaluations for MO-PTSP, using the same server where the PTSP and MO-PTSP competitions were run\footnote{Intel Core i5, 230GHz, 4GB of RAM.}.
 
 \subsection{Results in DST} \label{ssec:resDST}
+
+As the optimal Pareto front of the DST is known, a measure of performance can be obtained by observing the percentage of times these solutions are found by the players.
+As the solution that the algorithms converge to depends on the weights vector employed during the search (see end of Section~\ref{sec:momcts}), the approach taken here is to provide different weight vectors $W$ and analyze them separately.
+
+The weight vector for DST has two dimensions.
+This vector is here referred to as $W = (w_p, w_v)$, where $w_p$ weights moves; and $w_v$ weights the treasure value ($w_v = 1 - w_p$).
+$w_p$ takes values between $0$ and $1$, with a precision of $0.01$, and $100$ experiments have been run for each pair $(w_p, w_v)$.
+Hence, the game has been played a total of $10000$ times, for each algorithm.
 
@@
 The results show clearly how approximating the optimal Pareto front allows for finding all possible solutions.
-Both the NSGA-II and MO-MCTS approaches are able to converge to any solution in the front given the appropriate weight vector.
-It is important to highlight that these two algorithms employed the Euclidean distance action selection (see Section~\ref{sec:momcts}).
+Both NSGA-II and MO-MCTS approaches are able to converge to any solution in the front given the appropriate weight vector.
+It is important to highlight that these two algorithms employed the euclidean distance action selection (see Section~\ref{sec:momcts}).
 Other experiments, not included in this paper, showed that weighted sum action selection provides similar results to MCTS (that is, convergence to the edges of the front).
-The crucial distinction to make here is that both algorithms, NSGA-II and MO-MCTS, allow for better action selection mechanisms, which are able to overcome this problem by approximating a global Pareto front.
+The crucial distinction to make here is that both algorithms, NSGA-II and MO-MCTS, allow for better action selection mechanisms, that are able to overcome this problem by approximating a global Pareto front.
 
@@
 The experiments are performed in the $10$ maps available within the framework~\footnote{Available at \url{www.ptsp-game.net}}.
-In this case, the optimal Pareto front is not known in advance, and normally differs from one map to another.
+In this case, the optimal Pareto front is not known in advance, and normally it will differ from one map to another.
 Hence, the mechanisms to compare the performance of the algorithms tested need to be different than the one used for DST.
@@
 First of all, $4$ different weight vectors are tested: $W_1 = (0.33,0.33,0.33)$, $W_2 = (0.1,0.3,0.6)$, $W_3 = (0.1,0.6,0.3)$ and $W_4 = (0.6,0.1,0.3)$, where each $w_i$ corresponds to the weight for an objective ($w_t$ for time, $w_f$ for fuel and $w_d$ for damage, in this order).
-$W_1$ treats all objectives as having the same weight, while the other three give more relevance to different objectives.
+$W_1$ treats all objectives with the same weight, while the other three give more relevance to different objectives.
 These vectors then provide a wide spectrum for weights in this benchmark.
@@
  \multirow{3}{*}{\textbf{Map 5}} & $(0.33,0.33,0.33)$ & $1405 \pm 7$ & $467 \pm 4$& $559 \pm 4$\\
- & $(0.1,0.3,0.6)$ & $1431 \pm 9$ & $447 \pm 4$ & \bm{$541 \pm 4$}\\
+ & $(0.1,0.3,0.6)$ & $1431 \pm 9$ & $447 \pm 4$ & $541 \pm 4$\\
  & $(0.1,0.6,0.3)$ & $1467 \pm 9$ & \bm{$411 \pm 5$} & $567 \pm 5$\\
@@
  \multirow{3}{*}{\textbf{Map 6}} & $(0.33,0.33,0.33)$ & $1575 \pm 7$ & $549 \pm 5$& $303 \pm 4$\\
- & $(0.1,0.3,0.6)$ & $1626 \pm 9$ & $540 \pm 6$ & \bm{$286 \pm 5$}\\
+ & $(0.1,0.3,0.6)$ & $1626 \pm 9$ & $540 \pm 6$ & $286 \pm 5$\\
  & $(0.1,0.6,0.3)$ & $1703 \pm 11$ & \bm{$499 \pm 4$} & $316 \pm 7$\\
@@
 
-Some exceptions can be explained by analysing the specific maps: in map 1, a fuel canister is always collected near the end, restoring the fuel level to its maximum.
+Some exceptions can be explained by analysing the map: in map 1, a fuel canister is always collected near the end, restoring the fuel level to its maximum.
 This leaves too few cycles to make a difference in the controller (note also that this is the map with the smallest fuel consumption).
-Also, in map 10, there is no difference in the damage objective, however map 10 is a map with no obstacles to damage the ship (thus only lava lakes deal damage).
+Also, in map 10, there is no difference in the damage objective: but map 10 is a map with no obstacles to damage the ship (thus only lava lakes deal damage).
 This also results in this map being the one with the lowest overall damage.
@@
 Now, it is time to compare these results with the other algorithms.
-In order to do this, the same number of runs is performed for NSGA-II and MCTS in the $10$ maps of the benchmark.
+In order to do this, the same number of experiments is run for NSGA-II and MCTS in the $10$ maps of the benchmark.
 
@@
 Extending this comparison to all maps, each pair of algorithms ends with a triplet $(D, \O, d)$, where $D$ is the number of maps where the first algorithm dominates the second, $\O$ is the amount of maps where no dominance can be established, and $d$ states the number of maps where the first is dominated by the second.
-Table~\ref{tab:resMOPTSP} summarizes these results for all the algorithms tested.
+Table~\ref{tab:resMOPTSP} summarizes these results for all algorithms tested.
 
@@
 \caption{Results in MO-PTSP: Each cell indicates the triplet $(D, \O, d)$, where $D$ is the number of maps where the row algorithm dominates the column one, $\O$ is the amount of maps where no dominance can be established, and $d$ states the number of maps where the row algorithm is dominated by the column one.
-All the algorithms followed the same route (order of waypoints and fuel canisters) in every map tested.}
+All algorithms followed the same route (order of waypoints and fuel canisters) in every map tested.}
 \label{tab:resMOPTSP}
@@
 One of the first things to notice is that MO-MCTS dominates MCTS and NSGA-II in most of the maps, and it is never dominated in any.
-In particular, the dominance of MO-MCTS over MCTS is outstanding, even dominating in all $10$ maps for two of the weight vectors.
-MO-MCTS also dominates NSGA-II in more maps than in those where there is no dominance, and  is never dominated by NSGA-II in any map.
+Particularly, the dominance of MO-MCTS over MCTS is outstanding, even dominating in all $10$ maps for two of the weight vectors.
+MO-MCTS also dominates NSGA-II in more maps than in those where there is no dominance, and neither is it ever dominated by NSGA-II in any map.
 
@@
 
-Nonetheless, the inclusion of \texttt{PurofMovio} in this comparison is not pointless: it is possible to assess the quality of the three algorithms tested here by comparing their performance relatively, against this high quality entry.
+Nonetheless, the inclusion of \texttt{PurofMovio} in this comparison is not pointless: it is possible to asses the quality of the three algorithms tested here by comparing their performance relatively, against this high quality entry.
 Attending to this criteria, it can be seen how MO-MCTS is the algorithm that is dominated less often by \texttt{PurofMovio}, producing similar results on an average of $4.75$ out of the $10$ maps, and being dominated in $5.25$ maps.
@@
 Specifically in the MO-PTSP, there are regions of the map where there are more obstacles or lava lakes, hence the ship is most likely to suffer higher damage there.
-Also, the route followed during the game affects the relative ideal speed between waypoints, or perhaps a fuel canister may be picked up, which will affect how the fuel objective will be managed.
+Also, the route followed during the game affects the relative ideal speed between waypoints, or perhaps a fuel canister is going to be picked up, which will affect how the fuel objective can be managed.
 In general, many real-time games go through different phases, with different objectives and priorities.
@@
 A way to provide different weights at different times in MO-PTSP is straightforward.
-Given the route of waypoints (and fuel canisters) being followed, one can divide it into \textit{segments}, where each segment starts and ends with a waypoint (or fuel canister).
+Given the route of waypoints (and fuel canisters) that is being followed, one can divide it into \textit{segments}, where each segment starts and ends with a waypoint (or fuel canister).
 Then, each segment can be assigned a particular weight vector $W$.
@@
 
-This section shows some initial results obtained when testing the third variant, using a stochastic hill climbing algorithm on each map.
+This section shows some initial results in order to test the third variant, by using a stochastic hill climbing algorithm on each map.
 The goal is to check if, by varying the weight vectors between segments, better solutions can be achieved.
@@
 
-Finally, the results described in this paper allow us to assume that multi-objective approaches can provide a high quality level of play in real-time games.
+Finally, the results described in this paper allow us to presume that multi-objective approaches can provide a high quality level of play in real-time games.
 Multi-objective problems pop up in different settings and by using MCTS we can balance not only between exploration and exploitation but between multiple objectives as well.

### 2014-06-04 paper/paper.tex
Sanaz changes for revision 1 in

@@
 However, there are some cases where it cannot be said that $\vec x \preceq \vec y$ or $\vec y \preceq \vec x$. 
-In this case, it is said that these solutions are non-dominated with respect to each other.
+In this case, it is said that these solutions are indifferent to each other.
 Solutions that are not dominated can be grouped in a \textit{non-dominated set}.
@@
 One of the most well-known algorithms in the literature is the Non-dominated Sorting Evolutionary Algorithm II (NSGA-{II})~\cite{Deb2002}, the pseudocode of which is shown in Algorithm~\ref{alg:nsga2}.
+NSGA-II is known to deliver solutions with very good diversity and convergence to the Pareto-front for 2-objective problems.
+However, for problems with more than 3 objectives, other approaches such as NSGA-III~\cite{Jain2013}, SMS-EMOA~\cite{Beume2007} and MOEA/D~\cite{Zhang2007a} are shown to outperform NSGA-II.
 As in any evolutionary algorithm, NSGA-{II} evolves a set of $N$ individuals in a population denoted as $Pop$ with the difference that they are ranked according to a dominance criterion and a crowding distance measure, which are used to maintain a good diversity of solutions.
@@
 
-The three main pillars of the NSGA-2 algorithm are:
+The three main pillars of the NSGA-II algorithm are:
 \begin{itemize}
@@
 
-\begin{table}[!t]
-\begin{center}
-%\begin{tabular}{|>{\centering\arraybackslash}m{0.65cm}|c|c|c|c|}
-\begin{tabular}{|p{0.8cm}|p{2.2cm}|m{1.13cm}|m{1.13cm}|m{1.13cm}|}
-\hline
-\textbf{Map}  & \textbf{$W:(w_t, w_f, w_d)$} & \textbf{Time} &  \textbf{Fuel} & \textbf{Damage} \\ 
-\hline
- \multirow{3}{*}{\textbf{Map 1}} & $(0.33,0.33,0.33)$ & $1654 \pm 7$ & $131 \pm 2$& $846 \pm 13$\\
- & $(0.1,0.3,0.6)$ & $1657 \pm 8$ & $130 \pm 2$ & \bm{$773 \pm 11$}\\
- & $(0.1,0.6,0.3)$ & $1681 \pm 11$ & $131 \pm 2$ & $837 \pm 15$\\
- & $(0.6,0.1,0.3)$ & $1649 \pm 8$ & $132 \pm 2$ & $833 \pm 13$\\
-\hline
- \multirow{3}{*}{\textbf{Map 2}} & $(0.33,0.33,0.33)$ & $1409 \pm 7$ & $235 \pm 4$& $364 \pm 3$\\
- & $(0.1,0.3,0.6)$ & $1402 \pm 6$ & $236 \pm 5$ & \bm{$354 \pm 2$}\\
- & $(0.1,0.6,0.3)$ & $1416 \pm 8$ & \bm{$219 \pm 4$} & $360 \pm 3$\\
- & $(0.6,0.1,0.3)$ & $1396 \pm 8$ & $245 \pm 5$ & $361 \pm 2$\\
-\hline
- \multirow{3}{*}{\textbf{Map 3}} & $(0.33,0.33,0.33)$ & $1373 \pm 6$ & $221 \pm 3$& $301 \pm 7$\\
- & $(0.1,0.3,0.6)$ & $1378 \pm 5$ & $211 \pm 3$ & \bm{$268 \pm 5$}\\
- & $(0.1,0.6,0.3)$ & $1385 \pm 6$ & \bm{$203 \pm 4$} & $291 \pm 7$\\
- & $(0.6,0.1,0.3)$ & \bm{$1363 \pm 4$} & $229 \pm 4$ & $285 \pm 7$\\
-\hline
- \multirow{3}{*}{\textbf{Map 4}} & $(0.33,0.33,0.33)$ & $1383 \pm 6$ & $291 \pm 5$& $565 \pm 5$\\
- & $(0.1,0.3,0.6)$ & $1385 \pm 7$ & $304 \pm 4$ & \bm{$542 \pm 4$}\\
- & $(0.1,0.6,0.3)$ & $1423 \pm 6$ & \bm{$273 \pm 4$} & $583 \pm 5$\\
- & $(0.6,0.1,0.3)$ & $1388 \pm 6$ & $309 \pm 4$ & $559 \pm 5$\\
-\hline
- \multirow{3}{*}{\textbf{Map 5}} & $(0.33,0.33,0.33)$ & $1405 \pm 7$ & $467 \pm 4$& $559 \pm 4$\\
- & $(0.1,0.3,0.6)$ & $1431 \pm 9$ & $447 \pm 4$ & $541 \pm 4$\\
- & $(0.1,0.6,0.3)$ & $1467 \pm 9$ & \bm{$411 \pm 5$} & $567 \pm 5$\\
- & $(0.6,0.1,0.3)$ & $1399 \pm 9$ & $469 \pm 4$ & $547 \pm 3$\\
-\hline
- \multirow{3}{*}{\textbf{Map 6}} & $(0.33,0.33,0.33)$ & $1575 \pm 7$ & $549 \pm 5$& $303 \pm 4$\\
- & $(0.1,0.3,0.6)$ & $1626 \pm 9$ & $540 \pm 6$ & $286 \pm 5$\\
- & $(0.1,0.6,0.3)$ & $1703 \pm 11$ & \bm{$499 \pm 4$} & $316 \pm 7$\\
- & $(0.6,0.1,0.3)$ & $1571 \pm 7$ & $559 \pm 5$ & $294 \pm 4$\\
-\hline
- \multirow{3}{*}{\textbf{Map 7}} & $(0.33,0.33,0.33)$ & $1434 \pm 5$ & $599 \pm 6$& $284 \pm 6$\\
- & $(0.1,0.3,0.6)$ & $1475 \pm 10$ & $602 \pm 5$ & \bm{$243 \pm 6$}\\
- & $(0.1,0.6,0.3)$ & $1489 \pm 12$ & \bm{$549 \pm 3$} & $264 \pm 6$\\
- & $(0.6,0.1,0.3)$ & \bm{$1407 \pm 8$} & $618 \pm 5$ & $270 \pm 6$\\
-\hline
- \multirow{3}{*}{\textbf{Map 8}} & $(0.33,0.33,0.33)$ & $1761 \pm 9$ & $254 \pm 5$& $382 \pm 3$\\
- & $(0.1,0.3,0.6)$ & $1804 \pm 10$ & $269 \pm 4$ & \bm{$357 \pm 4$}\\
- & $(0.1,0.6,0.3)$ & $1826 \pm 10$ & \bm{$230 \pm 3$} & $392 \pm 7$\\
- & $(0.6,0.1,0.3)$ & \bm{$1732 \pm 9$} & $311 \pm 8$ & $379 \pm 6$\\
-\hline
- \multirow{3}{*}{\textbf{Map 9}} & $(0.33,0.33,0.33)$ & $2501 \pm 14$ & $926 \pm 6$& $574 \pm 9$\\
- & $(0.1,0.3,0.6)$ & $2503 \pm 10$ & $921 \pm 10$ & \bm{$524 \pm 8$}\\
- & $(0.1,0.6,0.3)$ & $2641 \pm 14$ & \bm{$833 \pm 5$} & $574 \pm 14$\\
- & $(0.6,0.1,0.3)$ & \bm{$2470 \pm 9$} & $956 \pm 5$ & $573 \pm 8$\\
-\hline
- \multirow{3}{*}{\textbf{Map 10}} & $(0.33,0.33,0.33)$ & $1430 \pm 8$ & $630 \pm 4$& $205 \pm 2$\\
- & $(0.1,0.3,0.6)$ & $1493 \pm 13$ & $615 \pm 4$ & $209 \pm 2$\\
- & $(0.1,0.6,0.3)$ & $1542 \pm 10$ & \bm{$554 \pm 4$} & $229 \pm 5$\\
- & $(0.6,0.1,0.3)$ & \bm{$1378 \pm 5$} & $663 \pm 6$ & $202 \pm 4$\\
-\hline
-\end{tabular}
-\caption{MO-MCTS results in MO-PTSP with different weight vectors. Values in bold indicate the best results on each map, when they are separated at least by $2$ standard errors from the others.}
-\label{tab:weights}
-\end{center}
-\end{table}
-
 The same three algorithms have been tested in the MO-PTSP domain.
@@
 It can be seen that the highest weight in $W$ leads, in most of the cases, to the best solution in that objective in the map.
+
+\begin{table}[!t]
+\begin{center}
+%\begin{tabular}{|>{\centering\arraybackslash}m{0.65cm}|c|c|c|c|}
+\begin{tabular}{|p{0.8cm}|p{2.2cm}|m{1.13cm}|m{1.13cm}|m{1.13cm}|}
+\hline
+\textbf{Map}  & \textbf{$W:(w_t, w_f, w_d)$} & \textbf{Time} &  \textbf{Fuel} & \textbf{Damage} \\ 
+\hline
+ \multirow{3}{*}{\textbf{Map 1}} & $(0.33,0.33,0.33)$ & $1654 \pm 7$ & $131 \pm 2$& $846 \pm 13$\\
+ & $(0.1,0.3,0.6)$ & $1657 \pm 8$ & $130 \pm 2$ & \bm{$773 \pm 11$}\\
+ & $(0.1,0.6,0.3)$ & $1681 \pm 11$ & $131 \pm 2$ & $837 \pm 15$\\
+ & $(0.6,0.1,0.3)$ & $1649 \pm 8$ & $132 \pm 2$ & $833 \pm 13$\\
+\hline
+ \multirow{3}{*}{\textbf{Map 2}} & $(0.33,0.33,0.33)$ & $1409 \pm 7$ & $235 \pm 4$& $364 \pm 3$\\
+ & $(0.1,0.3,0.6)$ & $1402 \pm 6$ & $236 \pm 5$ & \bm{$354 \pm 2$}\\
+ & $(0.1,0.6,0.3)$ & $1416 \pm 8$ & \bm{$219 \pm 4$} & $360 \pm 3$\\
+ & $(0.6,0.1,0.3)$ & $1396 \pm 8$ & $245 \pm 5$ & $361 \pm 2$\\
+\hline
+ \multirow{3}{*}{\textbf{Map 3}} & $(0.33,0.33,0.33)$ & $1373 \pm 6$ & $221 \pm 3$& $301 \pm 7$\\
+ & $(0.1,0.3,0.6)$ & $1378 \pm 5$ & $211 \pm 3$ & \bm{$268 \pm 5$}\\
+ & $(0.1,0.6,0.3)$ & $1385 \pm 6$ & \bm{$203 \pm 4$} & $291 \pm 7$\\
+ & $(0.6,0.1,0.3)$ & \bm{$1363 \pm 4$} & $229 \pm 4$ & $285 \pm 7$\\
+\hline
+ \multirow{3}{*}{\textbf{Map 4}} & $(0.33,0.33,0.33)$ & $1383 \pm 6$ & $291 \pm 5$& $565 \pm 5$\\
+ & $(0.1,0.3,0.6)$ & $1385 \pm 7$ & $304 \pm 4$ & \bm{$542 \pm 4$}\\
+ & $(0.1,0.6,0.3)$ & $1423 \pm 6$ & \bm{$273 \pm 4$} & $583 \pm 5$\\
+ & $(0.6,0.1,0.3)$ & $1388 \pm 6$ & $309 \pm 4$ & $559 \pm 5$\\
+\hline
+ \multirow{3}{*}{\textbf{Map 5}} & $(0.33,0.33,0.33)$ & $1405 \pm 7$ & $467 \pm 4$& $559 \pm 4$\\
+ & $(0.1,0.3,0.6)$ & $1431 \pm 9$ & $447 \pm 4$ & $541 \pm 4$\\
+ & $(0.1,0.6,0.3)$ & $1467 \pm 9$ & \bm{$411 \pm 5$} & $567 \pm 5$\\
+ & $(0.6,0.1,0.3)$ & $1399 \pm 9$ & $469 \pm 4$ & $547 \pm 3$\\
+\hline
+ \multirow{3}{*}{\textbf{Map 6}} & $(0.33,0.33,0.33)$ & $1575 \pm 7$ & $549 \pm 5$& $303 \pm 4$\\
+ & $(0.1,0.3,0.6)$ & $1626 \pm 9$ & $540 \pm 6$ & $286 \pm 5$\\
+ & $(0.1,0.6,0.3)$ & $1703 \pm 11$ & \bm{$499 \pm 4$} & $316 \pm 7$\\
+ & $(0.6,0.1,0.3)$ & $1571 \pm 7$ & $559 \pm 5$ & $294 \pm 4$\\
+\hline
+ \multirow{3}{*}{\textbf{Map 7}} & $(0.33,0.33,0.33)$ & $1434 \pm 5$ & $599 \pm 6$& $284 \pm 6$\\
+ & $(0.1,0.3,0.6)$ & $1475 \pm 10$ & $602 \pm 5$ & \bm{$243 \pm 6$}\\
+ & $(0.1,0.6,0.3)$ & $1489 \pm 12$ & \bm{$549 \pm 3$} & $264 \pm 6$\\
+ & $(0.6,0.1,0.3)$ & \bm{$1407 \pm 8$} & $618 \pm 5$ & $270 \pm 6$\\
+\hline
+ \multirow{3}{*}{\textbf{Map 8}} & $(0.33,0.33,0.33)$ & $1761 \pm 9$ & $254 \pm 5$& $382 \pm 3$\\
+ & $(0.1,0.3,0.6)$ & $1804 \pm 10$ & $269 \pm 4$ & \bm{$357 \pm 4$}\\
+ & $(0.1,0.6,0.3)$ & $1826 \pm 10$ & \bm{$230 \pm 3$} & $392 \pm 7$\\
+ & $(0.6,0.1,0.3)$ & \bm{$1732 \pm 9$} & $311 \pm 8$ & $379 \pm 6$\\
+\hline
+ \multirow{3}{*}{\textbf{Map 9}} & $(0.33,0.33,0.33)$ & $2501 \pm 14$ & $926 \pm 6$& $574 \pm 9$\\
+ & $(0.1,0.3,0.6)$ & $2503 \pm 10$ & $921 \pm 10$ & \bm{$524 \pm 8$}\\
+ & $(0.1,0.6,0.3)$ & $2641 \pm 14$ & \bm{$833 \pm 5$} & $574 \pm 14$\\
+ & $(0.6,0.1,0.3)$ & \bm{$2470 \pm 9$} & $956 \pm 5$ & $573 \pm 8$\\
+\hline
+ \multirow{3}{*}{\textbf{Map 10}} & $(0.33,0.33,0.33)$ & $1430 \pm 8$ & $630 \pm 4$& $205 \pm 2$\\
+ & $(0.1,0.3,0.6)$ & $1493 \pm 13$ & $615 \pm 4$ & $209 \pm 2$\\
+ & $(0.1,0.6,0.3)$ & $1542 \pm 10$ & \bm{$554 \pm 4$} & $229 \pm 5$\\
+ & $(0.6,0.1,0.3)$ & \bm{$1378 \pm 5$} & $663 \pm 6$ & $202 \pm 4$\\
+\hline
+\end{tabular}
+\caption{MO-PTSP results with different weight vectors. Values in bold indicate the best results on each map, when they are separated at least by $2$ standard errors from the others.}
+\label{tab:weights}
+\end{center}
+\end{table}
 

### 2014-06-06 paper/paper.tex
approximating

@@
 In the vanilla algorithm, each node would use this vector $\overline{r}$ to update its own accumulated reward vector $\overline{R}$.
-In the algorithm proposed here, each node in the MO-MCTS algorithm also keeps a local Pareto front ($P$), updated at each iteration with the reward vector $\overline{r}$ obtained at the end of the simulation.
+In the algorithm proposed here, each node in the MO-MCTS algorithm also keeps a local Pareto front approximation ($P$), updated at each iteration with the reward vector $\overline{r}$ obtained at the end of the simulation.
 Algorithm~\ref{alg:updPareto} describes how the node statistics are updated in MO-MCTS.
@@
 
-Here, if $\overline{r}$ is not dominated by the local Pareto front, it is added to the front and $\overline{r}$ is propagated to its parent.
-It may also happen that $\overline{r}$ dominates all or some solutions of the local front, in which case the new solution is included and the dominated solutions are removed from the Pareto front.
-If $\overline{r}$ is dominated by the node's Pareto front, the local Pareto front does not change and there is no need to maintain this propagation up the tree.
+Here, if $\overline{r}$ is not dominated by the local Pareto front approximation, it is added to the front and $\overline{r}$ is propagated to its parent.
+It may also happen that $\overline{r}$ dominates all or some solutions of the local front, in which case the new solution is included and the dominated solutions are removed from the Pareto front approximation.
+If $\overline{r}$ is dominated by the node's Pareto front approximation, it does not change and there is no need to maintain this propagation up the tree.
 Three observations can be made about the mechanism described here:
@@
 \item By construction, if a reward $\overline{r}$ is dominated by the local front of a node, it is a given that it will be dominated by the nodes above in the tree, so there is no need to update the fronts of the upper nodes, producing little impact on the computational cost of the algorithm.
-\item It is easy to infer, from the last point, that the Pareto front of a node cannot be worse than the front of its children (in other words, the front of a child will never dominate that of its parent).
+\item It is easy to infer, from the last point, that the Pareto front approximation of a node cannot be worse than the front of its children (in other words, the front of a child will never dominate that of its parent).
 Therefore, the root of the tree contains the best non-dominated front ever found during the search.
@@
 This information can be used, when all iterations have been performed, to select the move to perform next.
-If weights for the different objectives are provided, these weights can be used to select the desired solution in the Pareto front of the root node, and hence select the action that leads to that point.
+If weights for the different objectives are provided, these weights can be used to select the desired solution in the Pareto front approximation of the root node, and hence select the action that leads to that point.
 Secondly, the root's Pareto front can be used to measure the global quality of the search using the hypervolume calculation.
 
-Finally, the information stored at each node regarding the local Pareto front can be used to substitute $Q(s,a)$ in the UCB1 equation.
+Finally, the information stored at each node regarding the local Pareto front approximation can be used to substitute $Q(s,a)$ in the UCB1 equation.
 The quality of a given pair $(s,a)$ can be obtained by measuring the HV of the Pareto front stored in the node reached from state $s$ after applying action $a$.
@@
 As shown before, it is straightforward to discover which actions lead to what points in the front given as a solution: the first gene in the best NSGA-II individual, a root's child in MO-MCTS.
-Hence, by identifying the point in the Pareto front that the algorithm should converge to, it is possible to execute the action that leads to that point.
+Hence, by identifying the point in the Pareto front approximation that the algorithm should converge to, it is possible to execute the action that leads to that point.
 
@@
 \item Weighted Sum: the action chosen is the one that maximizes the weighted sum of the reward vector multiplied by $W$, for each point in the front.
-\item Euclidean distance: the euclidean distance from each point in the Pareto front (normalized in $[0,1]$) to the vector $W$ is calculated.
+\item Euclidean distance: the euclidean distance from each point in the Pareto front approximation (normalized in $[0,1]$) to the vector $W$ is calculated.
 The action to choose would be the one that leads to the point in the Pareto front with the shortest distance to $W$.
@@
 
-These results suggest that the weights effectively work on making the algorithm converge to different points in the Pareto front discovered by the algorithm.
+These results suggest that the weights effectively work on making the algorithm converge to different points in the Pareto front approximated by the algorithm.
 Now, it is time to compare these results with the other algorithms.

### 2014-06-06 paper/paper.tex
dealing with table 1

@@
 Table~\ref{tab:weights} shows the results of executing the MO-MCTS controller during $30$ games in each one of the $10$ maps of the MO-PTSP, for every weight vector.
+This table includes the averages obtained and the p-value for the results of the objective with the highest weight, compared with the result from the default weight vector $W_1 = (0.33,0.33,0.33)$ in the same map.
 It can be seen that the highest weight in $W$ leads, in most of the cases, to the best solution in that objective in the map.
+
+%\begin{table}[!t]
+%\begin{center}
+%%\begin{tabular}{|>{\centering\arraybackslash}m{0.65cm}|c|c|c|c|}
+%\begin{tabular}{|p{0.8cm}|p{2.2cm}|m{1.13cm}|m{1.13cm}|m{1.13cm}|}
+%\hline
+%\textbf{Map}  & \textbf{$W:(w_t, w_f, w_d)$} & \textbf{Time} &  \textbf{Fuel} & \textbf{Damage} \\ 
+%\hline
+% \multirow{3}{*}{\textbf{Map 1}} & $(0.33,0.33,0.33)$ & $1654 \pm 7$ & $131 \pm 2$& $846 \pm 13$\\
+% & $(0.1,0.3,0.6)$ & $1657 \pm 8$ & $130 \pm 2$ & \bm{$773 \pm 11$}\\
+% & $(0.1,0.6,0.3)$ & $1681 \pm 11$ & $131 \pm 2$ & $837 \pm 15$\\
+% & $(0.6,0.1,0.3)$ & $1649 \pm 8$ & $132 \pm 2$ & $833 \pm 13$\\
+%\hline
+% \multirow{3}{*}{\textbf{Map 2}} & $(0.33,0.33,0.33)$ & $1409 \pm 7$ & $235 \pm 4$& $364 \pm 3$\\
+% & $(0.1,0.3,0.6)$ & $1402 \pm 6$ & $236 \pm 5$ & \bm{$354 \pm 2$}\\
+% & $(0.1,0.6,0.3)$ & $1416 \pm 8$ & \bm{$219 \pm 4$} & $360 \pm 3$\\
+% & $(0.6,0.1,0.3)$ & $1396 \pm 8$ & $245 \pm 5$ & $361 \pm 2$\\
+%\hline
+% \multirow{3}{*}{\textbf{Map 3}} & $(0.33,0.33,0.33)$ & $1373 \pm 6$ & $221 \pm 3$& $301 \pm 7$\\
+% & $(0.1,0.3,0.6)$ & $1378 \pm 5$ & $211 \pm 3$ & \bm{$268 \pm 5$}\\
+% & $(0.1,0.6,0.3)$ & $1385 \pm 6$ & \bm{$203 \pm 4$} & $291 \pm 7$\\
+% & $(0.6,0.1,0.3)$ & \bm{$1363 \pm 4$} & $229 \pm 4$ & $285 \pm 7$\\
+%\hline
+% \multirow{3}{*}{\textbf{Map 4}} & $(0.33,0.33,0.33)$ & $1383 \pm 6$ & $291 \pm 5$& $565 \pm 5$\\
+% & $(0.1,0.3,0.6)$ & $1385 \pm 7$ & $304 \pm 4$ & \bm{$542 \pm 4$}\\
+% & $(0.1,0.6,0.3)$ & $1423 \pm 6$ & \bm{$273 \pm 4$} & $583 \pm 5$\\
+% & $(0.6,0.1,0.3)$ & $1388 \pm 6$ & $309 \pm 4$ & $559 \pm 5$\\
+%\hline
+% \multirow{3}{*}{\textbf{Map 5}} & $(0.33,0.33,0.33)$ & $1405 \pm 7$ & $467 \pm 4$& $559 \pm 4$\\
+% & $(0.1,0.3,0.6)$ & $1431 \pm 9$ & $447 \pm 4$ & $541 \pm 4$\\
+% & $(0.1,0.6,0.3)$ & $1467 \pm 9$ & \bm{$411 \pm 5$} & $567 \pm 5$\\
+% & $(0.6,0.1,0.3)$ & $1399 \pm 9$ & $469 \pm 4$ & $547 \pm 3$\\
+%\hline
+% \multirow{3}{*}{\textbf{Map 6}} & $(0.33,0.33,0.33)$ & $1575 \pm 7$ & $549 \pm 5$& $303 \pm 4$\\
+% & $(0.1,0.3,0.6)$ & $1626 \pm 9$ & $540 \pm 6$ & $286 \pm 5$\\
+% & $(0.1,0.6,0.3)$ & $1703 \pm 11$ & \bm{$499 \pm 4$} & $316 \pm 7$\\
+% & $(0.6,0.1,0.3)$ & $1571 \pm 7$ & $559 \pm 5$ & $294 \pm 4$\\
+%\hline
+% \multirow{3}{*}{\textbf{Map 7}} & $(0.33,0.33,0.33)$ & $1434 \pm 5$ & $599 \pm 6$& $284 \pm 6$\\
+% & $(0.1,0.3,0.6)$ & $1475 \pm 10$ & $602 \pm 5$ & \bm{$243 \pm 6$}\\
+% & $(0.1,0.6,0.3)$ & $1489 \pm 12$ & \bm{$549 \pm 3$} & $264 \pm 6$\\
+% & $(0.6,0.1,0.3)$ & \bm{$1407 \pm 8$} & $618 \pm 5$ & $270 \pm 6$\\
+%\hline
+% \multirow{3}{*}{\textbf{Map 8}} & $(0.33,0.33,0.33)$ & $1761 \pm 9$ & $254 \pm 5$& $382 \pm 3$\\
+% & $(0.1,0.3,0.6)$ & $1804 \pm 10$ & $269 \pm 4$ & \bm{$357 \pm 4$}\\
+% & $(0.1,0.6,0.3)$ & $1826 \pm 10$ & \bm{$230 \pm 3$} & $392 \pm 7$\\
+% & $(0.6,0.1,0.3)$ & \bm{$1732 \pm 9$} & $311 \pm 8$ & $379 \pm 6$\\
+%\hline
+% \multirow{3}{*}{\textbf{Map 9}} & $(0.33,0.33,0.33)$ & $2501 \pm 14$ & $926 \pm 6$& $574 \pm 9$\\
+% & $(0.1,0.3,0.6)$ & $2503 \pm 10$ & $921 \pm 10$ & \bm{$524 \pm 8$}\\
+% & $(0.1,0.6,0.3)$ & $2641 \pm 14$ & \bm{$833 \pm 5$} & $574 \pm 14$\\
+% & $(0.6,0.1,0.3)$ & \bm{$2470 \pm 9$} & $956 \pm 5$ & $573 \pm 8$\\
+%\hline
+% \multirow{3}{*}{\textbf{Map 10}} & $(0.33,0.33,0.33)$ & $1430 \pm 8$ & $630 \pm 4$& $205 \pm 2$\\
+% & $(0.1,0.3,0.6)$ & $1493 \pm 13$ & $615 \pm 4$ & $209 \pm 2$\\
+% & $(0.1,0.6,0.3)$ & $1542 \pm 10$ & \bm{$554 \pm 4$} & $229 \pm 5$\\
+% & $(0.6,0.1,0.3)$ & \bm{$1378 \pm 5$} & $663 \pm 6$ & $202 \pm 4$\\
+%\hline
+%\end{tabular}
+%\caption{MO-PTSP results with different weight vectors. Values in bold indicate the best results on each map, when they are separated at least by $2$ standard errors from the others.}
+%\label{tab:weights}
+%\end{center}
+%\end{table}
 
@@
 %\begin{tabular}{|>{\centering\arraybackslash}m{0.65cm}|c|c|c|c|}
-\begin{tabular}{|p{0.8cm}|p{2.2cm}|m{1.13cm}|m{1.13cm}|m{1.13cm}|}
-\hline
-\textbf{Map}  & \textbf{$W:(w_t, w_f, w_d)$} & \textbf{Time} &  \textbf{Fuel} & \textbf{Damage} \\ 
-\hline
- \multirow{3}{*}{\textbf{Map 1}} & $(0.33,0.33,0.33)$ & $1654 \pm 7$ & $131 \pm 2$& $846 \pm 13$\\
- & $(0.1,0.3,0.6)$ & $1657 \pm 8$ & $130 \pm 2$ & \bm{$773 \pm 11$}\\
- & $(0.1,0.6,0.3)$ & $1681 \pm 11$ & $131 \pm 2$ & $837 \pm 15$\\
- & $(0.6,0.1,0.3)$ & $1649 \pm 8$ & $132 \pm 2$ & $833 \pm 13$\\
-\hline
- \multirow{3}{*}{\textbf{Map 2}} & $(0.33,0.33,0.33)$ & $1409 \pm 7$ & $235 \pm 4$& $364 \pm 3$\\
- & $(0.1,0.3,0.6)$ & $1402 \pm 6$ & $236 \pm 5$ & \bm{$354 \pm 2$}\\
- & $(0.1,0.6,0.3)$ & $1416 \pm 8$ & \bm{$219 \pm 4$} & $360 \pm 3$\\
- & $(0.6,0.1,0.3)$ & $1396 \pm 8$ & $245 \pm 5$ & $361 \pm 2$\\
-\hline
- \multirow{3}{*}{\textbf{Map 3}} & $(0.33,0.33,0.33)$ & $1373 \pm 6$ & $221 \pm 3$& $301 \pm 7$\\
- & $(0.1,0.3,0.6)$ & $1378 \pm 5$ & $211 \pm 3$ & \bm{$268 \pm 5$}\\
- & $(0.1,0.6,0.3)$ & $1385 \pm 6$ & \bm{$203 \pm 4$} & $291 \pm 7$\\
- & $(0.6,0.1,0.3)$ & \bm{$1363 \pm 4$} & $229 \pm 4$ & $285 \pm 7$\\
-\hline
- \multirow{3}{*}{\textbf{Map 4}} & $(0.33,0.33,0.33)$ & $1383 \pm 6$ & $291 \pm 5$& $565 \pm 5$\\
- & $(0.1,0.3,0.6)$ & $1385 \pm 7$ & $304 \pm 4$ & \bm{$542 \pm 4$}\\
- & $(0.1,0.6,0.3)$ & $1423 \pm 6$ & \bm{$273 \pm 4$} & $583 \pm 5$\\
- & $(0.6,0.1,0.3)$ & $1388 \pm 6$ & $309 \pm 4$ & $559 \pm 5$\\
-\hline
- \multirow{3}{*}{\textbf{Map 5}} & $(0.33,0.33,0.33)$ & $1405 \pm 7$ & $467 \pm 4$& $559 \pm 4$\\
- & $(0.1,0.3,0.6)$ & $1431 \pm 9$ & $447 \pm 4$ & $541 \pm 4$\\
- & $(0.1,0.6,0.3)$ & $1467 \pm 9$ & \bm{$411 \pm 5$} & $567 \pm 5$\\
- & $(0.6,0.1,0.3)$ & $1399 \pm 9$ & $469 \pm 4$ & $547 \pm 3$\\
-\hline
- \multirow{3}{*}{\textbf{Map 6}} & $(0.33,0.33,0.33)$ & $1575 \pm 7$ & $549 \pm 5$& $303 \pm 4$\\
- & $(0.1,0.3,0.6)$ & $1626 \pm 9$ & $540 \pm 6$ & $286 \pm 5$\\
- & $(0.1,0.6,0.3)$ & $1703 \pm 11$ & \bm{$499 \pm 4$} & $316 \pm 7$\\
- & $(0.6,0.1,0.3)$ & $1571 \pm 7$ & $559 \pm 5$ & $294 \pm 4$\\
-\hline
- \multirow{3}{*}{\textbf{Map 7}} & $(0.33,0.33,0.33)$ & $1434 \pm 5$ & $599 \pm 6$& $284 \pm 6$\\
- & $(0.1,0.3,0.6)$ & $1475 \pm 10$ & $602 \pm 5$ & \bm{$243 \pm 6$}\\
- & $(0.1,0.6,0.3)$ & $1489 \pm 12$ & \bm{$549 \pm 3$} & $264 \pm 6$\\
- & $(0.6,0.1,0.3)$ & \bm{$1407 \pm 8$} & $618 \pm 5$ & $270 \pm 6$\\
-\hline
- \multirow{3}{*}{\textbf{Map 8}} & $(0.33,0.33,0.33)$ & $1761 \pm 9$ & $254 \pm 5$& $382 \pm 3$\\
- & $(0.1,0.3,0.6)$ & $1804 \pm 10$ & $269 \pm 4$ & \bm{$357 \pm 4$}\\
- & $(0.1,0.6,0.3)$ & $1826 \pm 10$ & \bm{$230 \pm 3$} & $392 \pm 7$\\
- & $(0.6,0.1,0.3)$ & \bm{$1732 \pm 9$} & $311 \pm 8$ & $379 \pm 6$\\
-\hline
- \multirow{3}{*}{\textbf{Map 9}} & $(0.33,0.33,0.33)$ & $2501 \pm 14$ & $926 \pm 6$& $574 \pm 9$\\
- & $(0.1,0.3,0.6)$ & $2503 \pm 10$ & $921 \pm 10$ & \bm{$524 \pm 8$}\\
- & $(0.1,0.6,0.3)$ & $2641 \pm 14$ & \bm{$833 \pm 5$} & $574 \pm 14$\\
- & $(0.6,0.1,0.3)$ & \bm{$2470 \pm 9$} & $956 \pm 5$ & $573 \pm 8$\\
-\hline
- \multirow{3}{*}{\textbf{Map 10}} & $(0.33,0.33,0.33)$ & $1430 \pm 8$ & $630 \pm 4$& $205 \pm 2$\\
- & $(0.1,0.3,0.6)$ & $1493 \pm 13$ & $615 \pm 4$ & $209 \pm 2$\\
- & $(0.1,0.6,0.3)$ & $1542 \pm 10$ & \bm{$554 \pm 4$} & $229 \pm 5$\\
- & $(0.6,0.1,0.3)$ & \bm{$1378 \pm 5$} & $663 \pm 6$ & $202 \pm 4$\\
+\bgroup
+%\def\arraystretch{0.9}%  1 is the default, change whatever you need
+\renewcommand\tabcolsep{5pt}
+\begin{tabular}{|p{0.9cm}|p{0.5cm}|m{1.9cm}|m{1.75cm}|m{1.75cm}|}
+\hline
+\textbf{Map}  & \textbf{$W$} & \textbf{Time} &  \textbf{Fuel} & \textbf{Damage} \\ 
+\hline
+ \multirow{3}{*}{\textbf{Map 1}} & $W_1$ & $1654$ & $131$& $846$\\
+ & $W_2$ & $1657$ & $130$ & \textbf{773} \bm{$(<0.01)$}\\
+ & $W_3$ & $1681$ & $131$ $(1.0)$ & $837$\\
+ & $W_4$ & $1649$ $(0.63)$ & $132$ & $833$\\
+\hline
+ \multirow{3}{*}{\textbf{Map 2}} & $W_1$ & $1409$ & $235$& $364$\\
+ & $W_2$ & $1402$ & $236$ & \textbf{354} \bm{$(<0.01)$}\\
+ & $W_3$ & $1416$ & \textbf{219} \bm{$(<0.01)$} & $360$\\
+ & $W_4$ & $1396$ $(0.22)$ & $245$ & $361$\\
+\hline
+ \multirow{3}{*}{\textbf{Map 3}} & $W_1$ & $1373$ & $221$& $301$\\
+ & $W_2$ & $1378$ & $211$ & \textbf{268} \bm{$(<0.01)$}\\
+ & $W_3$ & $1385$ & \textbf{203} \bm{$(<0.01)$} & $291$\\
+ & $W_4$ & $1363$ $(0.17)$ & $229$ & $285$\\
+\hline
+ \multirow{3}{*}{\textbf{Map 4}} & $W_1$ & $1383$ & $291$& $565$\\
+ & $W_2$ & $1385$ & $304$ & \textbf{542} \bm{$(<0.01)$}\\
+ & $W_3$ & $1423 $ & \textbf{273} \bm{$(<0.01)$} & $583$\\
+ & $W_4$ & $1388$ $(0.55)$ & $309$ & $559$\\
+\hline
+ \multirow{3}{*}{\textbf{Map 5}} & $W_1$ & $1405$ & $467$& $559$\\
+ & $W_2$ & $1431$ & $447$ & \textbf{541} \bm{$(<0.01)$}\\
+ & $W_3$ & $1467$ & \textbf{411} \bm{$(<0.01)$} & $567$\\
+ & $W_4$ & $1399$ $(0.60)$ & $469$ & $547$\\
+\hline
+ \multirow{3}{*}{\textbf{Map 6}} & $W_1$ & $1575$ & $549$& $303$\\
+ & $W_2$ & $1626$ & $540$ & \textbf{286} \bm{$(0.01)$}\\
+ & $W_3$ & $1703$ & \textbf{499} \bm{$(<0.01)$} & $316$\\
+ & $W_4$ & $1571$ $(0.68)$ & $559$ & $294$\\
+\hline
+ \multirow{3}{*}{\textbf{Map 7}} & $W_1$ & $1434$ & $599$& $284$\\
+ & $W_2$ & $1475$ & $602$ & \textbf{243} \bm{$(0.01)$}\\
+ & $W_3$ & $1489$ & \textbf{549} \bm{$(<0.01)$} & $264$\\
+ & $W_4$ & \textbf{1407} \bm{$(<0.01)$} & $618$ & $270$\\
+\hline
+ \multirow{3}{*}{\textbf{Map 8}} & $W_1$ & $1761 $ & $254$& $382$\\
+ & $W_2$ & $1804$ & $269$ & \textbf{357} \bm{$(<0.01)$}\\
+ & $W_3$ & $1826$ & 230 \bm{$(<0.01)$} & $392$\\
+ & $W_4$ & \textbf{1732} \bm{$(0.02)$} & $311$ & $379$\\
+\hline
+ \multirow{3}{*}{\textbf{Map 9}} & $W_1$ & $2501$ & $926$& $574$\\
+ & $W_2$ & $2503$ & $921$ & \textbf{524} \bm{$(<0.01)$}\\
+ & $W_3$ & $2641$ & \textbf{833} \bm{$(<0.01)$} & $574$\\
+ & $W_4$ & $2470$ $(0.06)$ & $956$ & $573$\\
+\hline
+ \multirow{3}{*}{\textbf{Map 10}} & $W_1$ & $1430$ & $630$& $205$\\
+ & $W_2$ & $1493$ & $615$ & $209$ $(0.16)$\\
+ & $W_3$ & $1542$ & \textbf{554} \bm{$(<0.01)$} & $229$\\
+ & $W_4$ & \textbf{1378} \bm{$(<0.01)$} & $663$ & $202$\\
 \hline
 \end{tabular}
-\caption{MO-PTSP results with different weight vectors. Values in bold indicate the best results on each map, when they are separated at least by $2$ standard errors from the others.}
+\egroup
+\caption{MO-PTSP averages with different weight vectors. Includes p-values for the objective with the highest weight. Results in bold are statistically significant (p-value $<0.05$).}
 \label{tab:weights}
@@
 \end{table}
+
+%\begin{table}[!t]
+%\begin{center}
+%%\begin{tabular}{|>{\centering\arraybackslash}m{0.65cm}|c|c|c|c|}
+%\bgroup
+%%\def\arraystretch{0.9}%  1 is the default, change whatever you need
+%\renewcommand\tabcolsep{5pt}
+%\begin{tabular}{|p{0.9cm}|p{0.5cm}|m{1.9cm}|m{1.75cm}|m{1.75cm}|}
+%\hline
+%\textbf{Map}  & \textbf{$W$} & \textbf{Time} &  \textbf{Fuel} & \textbf{Damage} \\ 
+%\hline
+% \multirow{3}{*}{\textbf{Map 1}} & $W_1$ & $1654$ & $131$& $846$\\
+% & $W_2$ & $1657$ & $130$ & \textbf{773} \bm{$(<0.01)$}\\
+% & $W_3$ & $1681$ & \textbf{131} $(0.72)$ & $837$\\
+% & $W_4$ & \textbf{1649} $(0.63)$ & $132$ & $833$\\
+%\hline
+% \multirow{3}{*}{\textbf{Map 2}} & $W_1$ & $1409$ & $235$& $364$\\
+% & $W_2$ & $1402$ & $236$ & $354$ $(0.1)$\\
+% & $W_3$ & $1416$ & \textbf{219} \bm{$(0.01)$} & $360$\\
+% & $W_4$ & \textbf{1396} $(0.55)$ & $245$ & $361$\\
+%\hline
+% \multirow{3}{*}{\textbf{Map 3}} & $W_1$ & $1373$ & $221$& $301$\\
+% & $W_2$ & $1378$ & $211$ & \textbf{268} $(0.053)$\\
+% & $W_3$ & $1385$ & \textbf{203} $(0.11)$ & $291$\\
+% & $W_4$ & \textbf{1363} $(0.17)$ & $229$ & $285$\\
+%\hline
+% \multirow{3}{*}{\textbf{Map 4}} & $W_1$ & \textbf{1383} & $291$& $565$\\
+% & $W_2$ & $1385$ & $304$ & \textbf{542} \bm{$(0.01)$}\\
+% & $W_3$ & $1423 $ & \textbf{273} \bm{$(0.01)$} & $583$\\
+% & $W_4$ & $1388$ $(0.74)$ & $309$ & $559$\\
+%\hline
+% \multirow{3}{*}{\textbf{Map 5}} & $W_1$ & $1405$ & $467$& $559$\\
+% & $W_2$ & $1431$ & $447$ & \textbf{541} $(0.23)$\\
+% & $W_3$ & $1467$ & \textbf{411} \bm{$(<0.01)$} & $567$\\
+% & $W_4$ & \textbf{1399} $(0.60)$ & $469$ & $547$\\
+%\hline
+% \multirow{3}{*}{\textbf{Map 6}} & $W_1$ & $1575$ & $549$& $303$\\
+% & $W_2$ & $1626$ & $540$ & \textbf{286} $(0.21)$\\
+% & $W_3$ & $1703$ & \textbf{499} \bm{$(<0.01)$} & $316$\\
+% & $W_4$ & \textbf{1571} $(0.68)$ & $559$ & $294$\\
+%\hline
+% \multirow{3}{*}{\textbf{Map 7}} & $W_1$ & $1434$ & $599$& $284$\\
+% & $W_2$ & $1475$ & $602$ & \textbf{243} \bm{$(0.01)$}\\
+% & $W_3$ & $1489$ & \textbf{549} \bm{$(<0.01)$} & $264$\\
+% & $W_4$ & \textbf{1407} \bm{$(0.01)$} & $618$ & $270$\\
+%\hline
+% \multirow{3}{*}{\textbf{Map 8}} & $W_1$ & $1761 $ & $254$& $382$\\
+% & $W_2$ & $1804$ & $269$ & 357 \bm{$(<0.01)$}\\
+% & $W_3$ & $1826$ & 230 \bm{$(<0.01)$} & $392$\\
+% & $W_4$ & \textbf{1732} \bm{$(0.02)$} & $311$ & $379$\\
+%\hline
+% \multirow{3}{*}{\textbf{Map 9}} & $W_1$ & $2501$ & $926$& $574$\\
+% & $W_2$ & $2503$ & $921$ & \textbf{524} \bm{$(<0.01)$}\\
+% & $W_3$ & $2641$ & \textbf{833} \bm{$(<0.01)$} & $574$\\
+% & $W_4$ & \textbf{2470} $(0.06)$ & $956$ & $573$\\
+%\hline
+% \multirow{3}{*}{\textbf{Map 10}} & $W_1$ & $1430$ & $630$& $205$\\
+% & $W_2$ & $1493$ & $615$ & \textbf{209} $(0.16)$\\
+% & $W_3$ & $1542$ & \textbf{554} \bm{$(0.00)$} & $229$\\
+% & $W_4$ & \textbf{1378} \bm{$(<0.01)$} & $663$ & $202$\\
+%\hline
+%\end{tabular}
+%\egroup
+%\caption{MO-PTSP results with different weight vectors. Values in bold indicate the best results on each map, when they are separated at least by $2$ standard errors from the others.}
+%\label{tab:weights}
+%\end{center}
+%\end{table}
 

### 2014-06-06 paper/spyros/stats.py

@@ -1,0 +1,117 @@
+import numpy as np
+import scipy as sp
+import scipy.stats as st
+import pandas as pd
+import itertools
+
+
+def mean_confidence_interval(data, confidence=0.95):
+	a = 1.0*np.array(data)
+	n = len(a)
+	m, se = np.mean(a), scipy.stats.sem(a)
+	h = se * st.t.ppf((1+confidence)/2., n-1)
+	return m, m-h, m+h
+
+
+
+def ttest_ind(m0, m1,se_0, se_1, n,  axis=0, equal_var=True):
+ 	#print type(m0)
+	m0 = np.array(m0,dtype = "f")
+	m1 = np.array(m1,dtype = "f")
+	se_0 = np.array(se_0,dtype = "f")
+	se_1 = np.array(se_1,dtype = "f")
+	v1 = se_0*np.sqrt(n)
+	v2 = se_1*np.sqrt(n)
+	v1 = v1*v1
+	v2 = v2*v2
+	n1 = n
+	n2 = n
+
+	#print v1, v2
+
+	if (equal_var):
+		df = n1 + n2 - 2
+		svar = ((n1 - 1) * v1 + (n2 - 1) * v2) / float(df)
+		denom = np.sqrt(svar * (1.0 / n1 + 1.0 / n2))
+	else:
+		vn1 = v1 / n1
+		vn2 = v2 / n2
+		df = ((vn1 + vn2)**2) / ((vn1**2) / (n1 - 1) + (vn2**2) / (n2 - 1))
+
+        # If df is undefined, variances are zero (assumes n1 > 0 & n2 > 0).
+        # Hence it doesn't matter what df is as long as it's not NaN.
+        df = np.where(np.isnan(df), 1, df)
+        denom = np.sqrt(vn1 + vn2)
+
+	d = m0 - m1
+	##print "====================="
+	#print d, denom
+	t = np.divide(d, denom)
+	##print m0,m1, se_0, se_1
+	##print "--->t", t, df
+	t, prob = st.stats._ttest_finish(df, t)
+	##print t, prob
+	##print "===================="
+	return t, prob
+
+def mean_confidence_interval_nodata(m, se, n, confidence = 0.95):
+	z = st.t.ppf((1+confidence)/2., n-1)
+	h = se * z
+	#print st.t._ppf((1+confidence)/2., n-1)
+	#print (-z,z)
+	#print st.norm.interval(0.95, loc=0, scale=1)
+	return m, h
+
+def printtestline(prefix, comb, mean, se, tt):
[…]

### 2014-06-06 paper/paper.tex
adding ttest

@@
 Table~\ref{tab:weights} shows the results of executing the MO-MCTS controller during $30$ games in each one of the $10$ maps of the MO-PTSP, for every weight vector.
-This table includes the averages obtained and the p-value for the results of the objective with the highest weight, compared with the result from the default weight vector $W_1 = (0.33,0.33,0.33)$ in the same map.
+This table includes the averages obtained and the independent t-test p-value for the results of the objective with the highest weight, compared with the result from the default weight vector $W_1 = (0.33,0.33,0.33)$ in the same map.
 It can be seen that the highest weight in $W$ leads, in most of the cases, to the best solution in that objective in the map.

### 2014-06-09 paper/paper.tex
final version for tables 1 and 3. Letter for reviewers and diff pdf updated

--- 2014-06-06 paper/paper.tex
@@
 Table~\ref{tab:weights} shows the results of executing the MO-MCTS controller during $30$ games in each one of the $10$ maps of the MO-PTSP, for every weight vector.
+This table includes the averages and standard errors obtained.
+Results in bold are those with an independent t-test p-value smaller than $0.01$, comparing the results of the objective with the highest weight with the result from the default weight vector $W_1 = (0.33,0.33,0.33)$ in the same map.
 It can be seen that the highest weight in $W$ leads, in most of the cases, to the best solution in that objective in the map.
@@
 %\begin{tabular}{|>{\centering\arraybackslash}m{0.65cm}|c|c|c|c|}
-\begin{tabular}{|p{0.8cm}|p{2.2cm}|m{1.13cm}|m{1.13cm}|m{1.13cm}|}
+\begin{tabular}{|p{0.8cm}|p{2.2cm}|m{1.2cm}|m{1.13cm}|m{1.13cm}|}
 \hline
@@
 \hline
- \multirow{3}{*}{\textbf{Map 1}} & $(0.33,0.33,0.33)$ & $1654 \pm 7$ & $131 \pm 2$& $846 \pm 13$\\
- & $(0.1,0.3,0.6)$ & $1657 \pm 8$ & $130 \pm 2$ & \bm{$773 \pm 11$}\\
- & $(0.1,0.6,0.3)$ & $1681 \pm 11$ & $131 \pm 2$ & $837 \pm 15$\\
- & $(0.6,0.1,0.3)$ & $1649 \pm 8$ & $132 \pm 2$ & $833 \pm 13$\\
-\hline
- \multirow{3}{*}{\textbf{Map 2}} & $(0.33,0.33,0.33)$ & $1409 \pm 7$ & $235 \pm 4$& $364 \pm 3$\\
- & $(0.1,0.3,0.6)$ & $1402 \pm 6$ & $236 \pm 5$ & \bm{$354 \pm 2$}\\
- & $(0.1,0.6,0.3)$ & $1416 \pm 8$ & \bm{$219 \pm 4$} & $360 \pm 3$\\
- & $(0.6,0.1,0.3)$ & $1396 \pm 8$ & $245 \pm 5$ & $361 \pm 2$\\
-\hline
- \multirow{3}{*}{\textbf{Map 3}} & $(0.33,0.33,0.33)$ & $1373 \pm 6$ & $221 \pm 3$& $301 \pm 7$\\
- & $(0.1,0.3,0.6)$ & $1378 \pm 5$ & $211 \pm 3$ & \bm{$268 \pm 5$}\\
- & $(0.1,0.6,0.3)$ & $1385 \pm 6$ & \bm{$203 \pm 4$} & $291 \pm 7$\\
- & $(0.6,0.1,0.3)$ & \bm{$1363 \pm 4$} & $229 \pm 4$ & $285 \pm 7$\\
-\hline
- \multirow{3}{*}{\textbf{Map 4}} & $(0.33,0.33,0.33)$ & $1383 \pm 6$ & $291 \pm 5$& $565 \pm 5$\\
- & $(0.1,0.3,0.6)$ & $1385 \pm 7$ & $304 \pm 4$ & \bm{$542 \pm 4$}\\
- & $(0.1,0.6,0.3)$ & $1423 \pm 6$ & \bm{$273 \pm 4$} & $583 \pm 5$\\
- & $(0.6,0.1,0.3)$ & $1388 \pm 6$ & $309 \pm 4$ & $559 \pm 5$\\
-\hline
- \multirow{3}{*}{\textbf{Map 5}} & $(0.33,0.33,0.33)$ & $1405 \pm 7$ & $467 \pm 4$& $559 \pm 4$\\
- & $(0.1,0.3,0.6)$ & $1431 \pm 9$ & $447 \pm 4$ & $541 \pm 4$\\
- & $(0.1,0.6,0.3)$ & $1467 \pm 9$ & \bm{$411 \pm 5$} & $567 \pm 5$\\
- & $(0.6,0.1,0.3)$ & $1399 \pm 9$ & $469 \pm 4$ & $547 \pm 3$\\
-\hline
- \multirow{3}{*}{\textbf{Map 6}} & $(0.33,0.33,0.33)$ & $1575 \pm 7$ & $549 \pm 5$& $303 \pm 4$\\
- & $(0.1,0.3,0.6)$ & $1626 \pm 9$ & $540 \pm 6$ & $286 \pm 5$\\
- & $(0.1,0.6,0.3)$ & $1703 \pm 11$ & \bm{$499 \pm 4$} & $316 \pm 7$\\
- & $(0.6,0.1,0.3)$ & $1571 \pm 7$ & $559 \pm 5$ & $294 \pm 4$\\
-\hline
- \multirow{3}{*}{\textbf{Map 7}} & $(0.33,0.33,0.33)$ & $1434 \pm 5$ & $599 \pm 6$& $284 \pm 6$\\
- & $(0.1,0.3,0.6)$ & $1475 \pm 10$ & $602 \pm 5$ & \bm{$243 \pm 6$}\\
- & $(0.1,0.6,0.3)$ & $1489 \pm 12$ & \bm{$549 \pm 3$} & $264 \pm 6$\\
- & $(0.6,0.1,0.3)$ & \bm{$1407 \pm 8$} & $618 \pm 5$ & $270 \pm 6$\\
-\hline
- \multirow{3}{*}{\textbf{Map 8}} & $(0.33,0.33,0.33)$ & $1761 \pm 9$ & $254 \pm 5$& $382 \pm 3$\\
- & $(0.1,0.3,0.6)$ & $1804 \pm 10$ & $269 \pm 4$ & \bm{$357 \pm 4$}\\
- & $(0.1,0.6,0.3)$ & $1826 \pm 10$ & \bm{$230 \pm 3$} & $392 \pm 7$\\
- & $(0.6,0.1,0.3)$ & \bm{$1732 \pm 9$} & $311 \pm 8$ & $379 \pm 6$\\
-\hline
- \multirow{3}{*}{\textbf{Map 9}} & $(0.33,0.33,0.33)$ & $2501 \pm 14$ & $926 \pm 6$& $574 \pm 9$\\
- & $(0.1,0.3,0.6)$ & $2503 \pm 10$ & $921 \pm 10$ & \bm{$524 \pm 8$}\\
- & $(0.1,0.6,0.3)$ & $2641 \pm 14$ & \bm{$833 \pm 5$} & $574 \pm 14$\\
- & $(0.6,0.1,0.3)$ & \bm{$2470 \pm 9$} & $956 \pm 5$ & $573 \pm 8$\\
-\hline
- \multirow{3}{*}{\textbf{Map 10}} & $(0.33,0.33,0.33)$ & $1430 \pm 8$ & $630 \pm 4$& $205 \pm 2$\\
- & $(0.1,0.3,0.6)$ & $1493 \pm 13$ & $615 \pm 4$ & $209 \pm 2$\\
- & $(0.1,0.6,0.3)$ & $1542 \pm 10$ & \bm{$554 \pm 4$} & $229 \pm 5$\\
- & $(0.6,0.1,0.3)$ & \bm{$1378 \pm 5$} & $663 \pm 6$ & $202 \pm 4$\\
+ \multirow{3}{*}{\textbf{Map 1}} & $(0.33,0.33,0.33)$ & $1654$ $(7)$ & $131$ $(2)$& $846$ $(13)$\\
+ & $(0.1,0.3,0.6)$ & $1657$ $(8)$ & $130$ $(2)$ & \bm{$773$ $(11)$}\\
+ & $(0.1,0.6,0.3)$ & $1681$ $(11)$ & $131$ $(2)$ & $837$ $(15)$\\
+ & $(0.6,0.1,0.3)$ & $1649$ $(8)$ & $132$ $(2)$ & $833$ $(13)$\\
+\hline
+ \multirow{3}{*}{\textbf{Map 2}} & $(0.33,0.33,0.33)$ & $1409$ $(7)$ & $235$ $(4)$& $364$ $(3)$\\
+ & $(0.1,0.3,0.6)$ & $1402$ $(6)$ & $236$ $(5)$ & \bm{$354$ $(2)$}\\
+ & $(0.1,0.6,0.3)$ & $1416$ $(8)$ & \bm{$219$ $(4)$} & $360$ $(3)$\\
+ & $(0.6,0.1,0.3)$ & $1396$ $(8)$ & $245$ $(5)$ & $361$ $(2)$\\
+\hline
+ \multirow{3}{*}{\textbf{Map 3}} & $(0.33,0.33,0.33)$ & $1373$ $(6)$ & $221$ $(3)$& $301$ $(7)$\\
+ & $(0.1,0.3,0.6)$ & $1378$ $(5)$ & $211$ $(3)$ & \bm{$268$ $(5)$}\\
+ & $(0.1,0.6,0.3)$ & $1385$ $(6)$ & \bm{$203$ $(4)$} & $291$ $(7)$\\
+ & $(0.6,0.1,0.3)$ & $1363$ $(4)$ & $229$ $(4)$ & $285$ $(7)$\\
+\hline
+ \multirow{3}{*}{\textbf{Map 4}} & $(0.33,0.33,0.33)$ & $1383$ $(6)$ & $291$ $(5)$& $565$ $(5)$\\
+ & $(0.1,0.3,0.6)$ & $1385$ $(7)$ & $304$ $(4)$ & \bm{$542$ $(4)$}\\
+ & $(0.1,0.6,0.3)$ & $1423$ $(6)$ & \bm{$273$ $(4)$} & $583$ $(5)$\\
+ & $(0.6,0.1,0.3)$ & $1388$ $(6)$ & $309$ $(4)$ & $559$ $(5)$\\
+\hline
+ \multirow{3}{*}{\textbf{Map 5}} & $(0.33,0.33,0.33)$ & $1405$ $(7)$ & $467$ $(4)$& $559$ $(4)$\\
+ & $(0.1,0.3,0.6)$ & $1431$ $(9)$ & $447$ $(4)$ & \bm{$541$ $(4)$}\\
+ & $(0.1,0.6,0.3)$ & $1467$ $(9)$ & \bm{$411$ $(5)$} & $567$ $(5)$\\
+ & $(0.6,0.1,0.3)$ & $1399$ $(9)$ & $469$ $(4)$ & $547$ $(3)$\\
+\hline
+ \multirow{3}{*}{\textbf{Map 6}} & $(0.33,0.33,0.33)$ & $1575$ $(7)$ & $549$ $(5)$& $303$ $(4)$\\
+ & $(0.1,0.3,0.6)$ & $1626$ $(9)$ & $540$ $(6)$ & $286$ $(5)$\\
+ & $(0.1,0.6,0.3)$ & $1703$ $(11)$ & \bm{$499$ $(4)$} & $316$ $(7)$\\
+ & $(0.6,0.1,0.3)$ & $1571$ $(7)$ & $559$ $(5)$ & $294$ $(4)$\\
+\hline
+ \multirow{3}{*}{\textbf{Map 7}} & $(0.33,0.33,0.33)$ & $1434$ $(5)$ & $599$ $(6)$& $284$ $(6)$\\
+ & $(0.1,0.3,0.6)$ & $1475$ $(10)$ & $602$ $(5)$ & $243$ $(6)$\\
+ & $(0.1,0.6,0.3)$ & $1489$ $(12)$ & \bm{$549$ $(3)$} & $264$ $(6)$\\
+ & $(0.6,0.1,0.3)$ & \bm{$1407$ $(8)$} & $618$ $(5)$ & $270$ $(6)$\\
+\hline
+ \multirow{3}{*}{\textbf{Map 8}} & $(0.33,0.33,0.33)$ & $1761$ $(9)$ & $254$ $(5)$& $382$ $(3)$\\
+ & $(0.1,0.3,0.6)$ & $1804$ $(10)$ & $269$ $(4)$ & \bm{$357$ $(4)$}\\
+ & $(0.1,0.6,0.3)$ & $1826$ $(10)$ & $230$ $(3)$ & $392$ $(7)$\\
+ & $(0.6,0.1,0.3)$ & $1732$ $(9)$ & $311$ $(8)$ & $379$ $(6)$\\
+\hline
+ \multirow{3}{*}{\textbf{Map 9}} & $(0.33,0.33,0.33)$ & $2501$ $(14)$ & $926$ $(6)$& $574$ $(9)$\\
+ & $(0.1,0.3,0.6)$ & $2503$ $(10)$ & $921$ $(10)$ & \bm{$524$ $(8)$}\\
+ & $(0.1,0.6,0.3)$ & $2641$ $(14)$ & \bm{$833$ $(5)$} & $574$ $(14)$\\
+ & $(0.6,0.1,0.3)$ & $2470$ $(9)$ & $956$ $(5)$ & $573$ $(8)$\\
+\hline
+ \multirow{3}{*}{\textbf{Map 10}} & $(0.33,0.33,0.33)$ & $1430$ $(8)$ & $630$ $(4)$& $205$ $(2)$\\
+ & $(0.1,0.3,0.6)$ & $1493$ $(13)$ & $615$ $(4)$ & $209$ $(2)$\\
+ & $(0.1,0.6,0.3)$ & $1542$ $(10)$ & \bm{$554$ $(4)$} & $229$ $(5)$\\
+ & $(0.6,0.1,0.3)$ & \bm{$1378$ $(5)$} & $663$ $(6)$ & $202$ $(4)$\\
 \hline
 \end{tabular}
-\caption{MO-PTSP results with different weight vectors. Values in bold indicate the best results on each map, when they are separated at least by $2$ standard errors from the others.}
+\caption{MO-PTSP averages (plus standard error) with different weight vectors. Results in bold obtained an independent t-test p-value $<0.01$.}
 \label{tab:weights}
@@
 \end{table}
+
+%\begin{table}[!t]
+%\begin{center}
+%%\begin{tabular}{|>{\centering\arraybackslash}m{0.65cm}|c|c|c|c|}
+%\bgroup
+%%\def\arraystretch{0.9}%  1 is the default, change whatever you need
+%\renewcommand\tabcolsep{5pt}
+%\begin{tabular}{|p{0.9cm}|p{0.5cm}|m{1.9cm}|m{1.75cm}|m{1.75cm}|}
+%\hline
+%\textbf{Map}  & \textbf{$W$} & \textbf{Time} &  \textbf{Fuel} & \textbf{Damage} \\ 
+%\hline
+% \multirow{3}{*}{\textbf{Map 1}} & $W_1$ & $1654$ & $131$& $846$\\
+% & $W_2$ & $1657$ & $130$ & \textbf{773} \bm{$(<0.01)$}\\
+% & $W_3$ & $1681$ & $131$ $(1.0)$ & $837$\\
+% & $W_4$ & $1649$ $(0.63)$ & $132$ & $833$\\
+%\hline
+% \multirow{3}{*}{\textbf{Map 2}} & $W_1$ & $1409$ & $235$& $364$\\
+% & $W_2$ & $1402$ & $236$ & \textbf{354} \bm{$(<0.01)$}\\
+% & $W_3$ & $1416$ & \textbf{219} \bm{$(<0.01)$} & $360$\\
+% & $W_4$ & $1396$ $(0.22)$ & $245$ & $361$\\
+%\hline
+% \multirow{3}{*}{\textbf{Map 3}} & $W_1$ & $1373$ & $221$& $301$\\
+% & $W_2$ & $1378$ & $211$ & \textbf{268} \bm{$(<0.01)$}\\
+% & $W_3$ & $1385$ & \textbf{203} \bm{$(<0.01)$} & $291$\\
+% & $W_4$ & $1363$ $(0.17)$ & $229$ & $285$\\
+%\hline
+% \multirow{3}{*}{\textbf{Map 4}} & $W_1$ & $1383$ & $291$& $565$\\
+% & $W_2$ & $1385$ & $304$ & \textbf{542} \bm{$(<0.01)$}\\
+% & $W_3$ & $1423 $ & \textbf{273} \bm{$(<0.01)$} & $583$\\
+% & $W_4$ & $1388$ $(0.55)$ & $309$ & $559$\\
+%\hline
+% \multirow{3}{*}{\textbf{Map 5}} & $W_1$ & $1405$ & $467$& $559$\\
+% & $W_2$ & $1431$ & $447$ & \textbf{541} \bm{$(<0.01)$}\\
+% & $W_3$ & $1467$ & \textbf{411} \bm{$(<0.01)$} & $567$\\
+% & $W_4$ & $1399$ $(0.60)$ & $469$ & $547$\\
+%\hline
+% \multirow{3}{*}{\textbf{Map 6}} & $W_1$ & $1575$ & $549$& $303$\\
+% & $W_2$ & $1626$ & $540$ & \textbf{286} \bm{$(0.01)$}\\
+% & $W_3$ & $1703$ & \textbf{499} \bm{$(<0.01)$} & $316$\\
+% & $W_4$ & $1571$ $(0.68)$ & $559$ & $294$\\
+%\hline
+% \multirow{3}{*}{\textbf{Map 7}} & $W_1$ & $1434$ & $599$& $284$\\
+% & $W_2$ & $1475$ & $602$ & \textbf{243} \bm{$(0.01)$}\\
+% & $W_3$ & $1489$ & \textbf{549} \bm{$(<0.01)$} & $264$\\
+% & $W_4$ & \textbf{1407} \bm{$(<0.01)$} & $618$ & $270$\\
+%\hline
+% \multirow{3}{*}{\textbf{Map 8}} & $W_1$ & $1761 $ & $254$& $382$\\
+% & $W_2$ & $1804$ & $269$ & \textbf{357} \bm{$(<0.01)$}\\
+% & $W_3$ & $1826$ & 230 \bm{$(<0.01)$} & $392$\\
+% & $W_4$ & \textbf{1732} \bm{$(0.02)$} & $311$ & $379$\\
+%\hline
+% \multirow{3}{*}{\textbf{Map 9}} & $W_1$ & $2501$ & $926$& $574$\\
+% & $W_2$ & $2503$ & $921$ & \textbf{524} \bm{$(<0.01)$}\\
+% & $W_3$ & $2641$ & \textbf{833} \bm{$(<0.01)$} & $574$\\
+% & $W_4$ & $2470$ $(0.06)$ & $956$ & $573$\\
+%\hline
+% \multirow{3}{*}{\textbf{Map 10}} & $W_1$ & $1430$ & $630$& $205$\\
+% & $W_2$ & $1493$ & $615$ & $209$ $(0.16)$\\
+% & $W_3$ & $1542$ & \textbf{554} \bm{$(<0.01)$} & $229$\\
+% & $W_4$ & \textbf{1378} \bm{$(<0.01)$} & $663$ & $202$\\
+%\hline
+%\end{tabular}
+%\egroup
+%\caption{MO-PTSP averages with different weight vectors. Includes p-values for the objective with the highest weight. Results in bold are statistically significant (p-value $<0.05$).}
+%\label{tab:weights}
+%\end{center}
+%\end{table}
+
+%\begin{table}[!t]
+%\begin{center}
+%%\begin{tabular}{|>{\centering\arraybackslash}m{0.65cm}|c|c|c|c|}
+%\bgroup
+%%\def\arraystretch{0.9}%  1 is the default, change whatever you need
+%\renewcommand\tabcolsep{5pt}
+%\begin{tabular}{|p{0.9cm}|p{0.5cm}|m{1.9cm}|m{1.75cm}|m{1.75cm}|}
+%\hline
+%\textbf{Map}  & \textbf{$W$} & \textbf{Time} &  \textbf{Fuel} & \textbf{Damage} \\ 
+%\hline
+% \multirow{3}{*}{\textbf{Map 1}} & $W_1$ & $1654$ & $131$& $846$\\
+% & $W_2$ & $1657$ & $130$ & \textbf{773} \bm{$(<0.01)$}\\
+% & $W_3$ & $1681$ & \textbf{131} $(0.72)$ & $837$\\
+% & $W_4$ & \textbf{1649} $(0.63)$ & $132$ & $833$\\
+%\hline
+% \multirow{3}{*}{\textbf{Map 2}} & $W_1$ & $1409$ & $235$& $364$\\
+% & $W_2$ & $1402$ & $236$ & $354$ $(0.1)$\\
+% & $W_3$ & $1416$ & \textbf{219} \bm{$(0.01)$} & $360$\\
+% & $W_4$ & \textbf{1396} $(0.55)$ & $245$ & $361$\\
+%\hline
+% \multirow{3}{*}{\textbf{Map 3}} & $W_1$ & $1373$ & $221$& $301$\\
+% & $W_2$ & $1378$ & $211$ & \textbf{268} $(0.053)$\\
+% & $W_3$ & $1385$ & \textbf{203} $(0.11)$ & $291$\\
+% & $W_4$ & \textbf{1363} $(0.17)$ & $229$ & $285$\\
+%\hline
+% \multirow{3}{*}{\textbf{Map 4}} & $W_1$ & \textbf{1383} & $291$& $565$\\
+% & $W_2$ & $1385$ & $304$ & \textbf{542} \bm{$(0.01)$}\\
+% & $W_3$ & $1423 $ & \textbf{273} \bm{$(0.01)$} & $583$\\
+% & $W_4$ & $1388$ $(0.74)$ & $309$ & $559$\\
+%\hline
+% \multirow{3}{*}{\textbf{Map 5}} & $W_1$ & $1405$ & $467$& $559$\\
+% & $W_2$ & $1431$ & $447$ & \textbf{541} $(0.23)$\\
+% & $W_3$ & $1467$ & \textbf{411} \bm{$(<0.01)$} & $567$\\
+% & $W_4$ & \textbf{1399} $(0.60)$ & $469$ & $547$\\
+%\hline
+% \multirow{3}{*}{\textbf{Map 6}} & $W_1$ & $1575$ & $549$& $303$\\
+% & $W_2$ & $1626$ & $540$ & \textbf{286} $(0.21)$\\
+% & $W_3$ & $1703$ & \textbf{499} \bm{$(<0.01)$} & $316$\\
+% & $W_4$ & \textbf{1571} $(0.68)$ & $559$ & $294$\\
+%\hline
+% \multirow{3}{*}{\textbf{Map 7}} & $W_1$ & $1434$ & $599$& $284$\\
+% & $W_2$ & $1475$ & $602$ & \textbf{243} \bm{$(0.01)$}\\
+% & $W_3$ & $1489$ & \textbf{549} \bm{$(<0.01)$} & $264$\\
+% & $W_4$ & \textbf{1407} \bm{$(0.01)$} & $618$ & $270$\\
+%\hline
+% \multirow{3}{*}{\textbf{Map 8}} & $W_1$ & $1761 $ & $254$& $382$\\
+% & $W_2$ & $1804$ & $269$ & 357 \bm{$(<0.01)$}\\
+% & $W_3$ & $1826$ & 230 \bm{$(<0.01)$} & $392$\\
+% & $W_4$ & \textbf{1732} \bm{$(0.02)$} & $311$ & $379$\\
+%\hline
+% \multirow{3}{*}{\textbf{Map 9}} & $W_1$ & $2501$ & $926$& $574$\\
+% & $W_2$ & $2503$ & $921$ & \textbf{524} \bm{$(<0.01)$}\\
+% & $W_3$ & $2641$ & \textbf{833} \bm{$(<0.01)$} & $574$\\
+% & $W_4$ & \textbf{2470} $(0.06)$ & $956$ & $573$\\
+%\hline
+% \multirow{3}{*}{\textbf{Map 10}} & $W_1$ & $1430$ & $630$& $205$\\
+% & $W_2$ & $1493$ & $615$ & \textbf{209} $(0.16)$\\
+% & $W_3$ & $1542$ & \textbf{554} \bm{$(0.00)$} & $229$\\
+% & $W_4$ & \textbf{1378} \bm{$(<0.01)$} & $663$ & $202$\\
+%\hline
+%\end{tabular}
+%\egroup
+%\caption{MO-PTSP results with different weight vectors. Values in bold indicate the best results on each map, when they are separated at least by $2$ standard errors from the others.}
+%\label{tab:weights}
+%\end{center}
+%\end{table}
 
@@
 \begin{center}
-\begin{tabular}{|p{0.8cm}|p{1.95cm}|m{1.13cm}|m{1.13cm}|m{1.13cm}|m{0.1cm}|}
+\begin{tabular}{|p{0.8cm}|p{1.95cm}|m{1.2cm}|m{1.05cm}|m{1.05cm}|m{0.1cm}|}
 \hline
@@
 \hline
- \multirow{3}{*}{\textbf{Map 1}} & $11111111111111$ & $1654 \pm 7$ & $131 \pm 2$& $846 \pm 13$ & $\preceq$ \\
- & $22222222222222$ & $1657 \pm 8$ & $130 \pm 2$ & $773 \pm 11$ & $\preceq$ \\
- & $33333333333333$ & $1681 \pm 11$ & $131 \pm 2$ & $837 \pm 15$ & $\preceq$ \\
- & $32312212331112$ & $1619 \pm 12$ & $130 \pm 2$ & $744 \pm 15$ & \\
-\hline
- \multirow{3}{*}{\textbf{Map 2}} & $11111111111111$ & $1409 \pm 7$ & $235 \pm 4$& $364 \pm 3$ & $\preceq$ \\
- & $22222222222222$ & $1402 \pm 6$ & $236 \pm 5$ & $354 \pm 2$ & $\preceq$ \\
- & $33333333333333$ & $1416 \pm 8$ & $219 \pm 4$ & $360 \pm 3$ & $\preceq$ \\
- & $23131312323213$ & $1390 \pm 10$ & $210 \pm 3$ & $353 \pm 3$ & \\
-\hline
- \multirow{3}{*}{\textbf{Map 3}} & $11111111111111$ & $1373 \pm 6$ & $221 \pm 3$& $301 \pm 7$ & $\preceq$ \\
- & $22222222222222$ & $1378 \pm 5$ & $211 \pm 3$ & $268 \pm 5$ & $\preceq$ \\
- & $33333333333333$ & $1385 \pm 6$ & $203 \pm 4$ & $291 \pm 7$ & $\O$ \\
- & $11122222112231$ & $1358 \pm 9$ & $219 \pm 7$ & $263 \pm 12$ & \\
-\hline
- \multirow{3}{*}{\textbf{Map 4}} & $11111111111111$ & $1383 \pm 6$ & $291 \pm 5$& $565 \pm 5$ & $\preceq$ \\
- & $22222222222222$ & $1385 \pm 7$ & $304 \pm 4$ & $542 \pm 4$ & $\preceq$ \\
- & $33333333333333$ & $1423 \pm 6$ & $273 \pm 4$ & $583 \pm 5$ &  $\O$ \\
- & $11121131212112$ & $1360 \pm 4$ & $282 \pm 5$ & $540 \pm 4$ & \\
-\hline
- \multirow{3}{*}{\textbf{Map 5}} & $11111111111111$ & $1405 \pm 7$ & $467 \pm 4$& $559 \pm 4$ & $\preceq$ \\
- & $22222222222222$ & $1431 \pm 9$ & $447 \pm 4$ &  $541 \pm 4$ & $\preceq$ \\
- & $33333333333333$ & $1467 \pm 9$ & $411 \pm 5$ & $567 \pm 5$ & $\O$ \\
- & $21311213111211$ & $1397 \pm 11$ & $448 \pm 10$ & $535 \pm 5$ & \\
-\hline
- \multirow{3}{*}{\textbf{Map 6}} & $11111111111111$ & $1575 \pm 7$ & $549 \pm 5$& $303 \pm 4$ & $\preceq$ \\
- & $22222222222222$ & $1626 \pm 9$ & $540 \pm 6$ & $286 \pm 5$ & $\preceq$ \\
- & $33333333333333$ & $1703 \pm 11$ & $499 \pm 4$ & $316 \pm 7$ & $\O$ \\
- & $31121312111111$ & $1570 \pm 16$ & $535 \pm 10$ & $266 \pm 6$ & \\
-\hline
- \multirow{3}{*}{\textbf{Map 7}} & $11111111111111$ & $1434 \pm 5$ & $599 \pm 6$& $284 \pm 6$ & $\preceq$ \\
- & $22222222222222$ & $1475 \pm 10$ & $602 \pm 5$ & $243 \pm 6$ & $\preceq$ \\
- & $33333333333333$ & $1489 \pm 12$ & $549 \pm 3$ & $264 \pm 6$ & $\O$ \\
- & $11332211321332$ & $1401 \pm 12$ & $563 \pm 4$ & $230 \pm 12$ & \\
-\hline
- \multirow{3}{*}{\textbf{Map 8}} & $11111111111111$ & $1761 \pm 9$ & $254 \pm 5$& $382 \pm 3$ & $\preceq$ \\
- & $22222222222222$ & $1804 \pm 10$ & $269 \pm 4$ & $357 \pm 4$ & $\preceq$ \\
- & $33333333333333$ & $1826 \pm 10$ & $230 \pm 3$ & $392 \pm 7$ & $\O$ \\
- & $23221131313323$ & $1747 \pm 11$ & $247 \pm 9$ & $363 \pm 9$ & \\
-\hline
- \multirow{3}{*}{\textbf{Map 9}} & $11111111111111$ & $2501 \pm 14$ & $926 \pm 6$& $574 \pm 9$ & $\preceq$ \\
- & $22222222222222$ & $2503 \pm 10$ & $921 \pm 10$ & $524 \pm 8$ & $\preceq$ \\
- & $33333333333333$ & $2641 \pm 14$ & $833 \pm 5$ & $574 \pm 14$ & $\O$ \\
- & $21132331333223$ & $2463 \pm 19$ & $891 \pm 8$ & $523 \pm 9$ & \\
-\hline
- \multirow{3}{*}{\textbf{Map 10}} & $11111111111111$ & $1430 \pm 8$ & $630 \pm 4$& $205 \pm 2$ & $\preceq$ \\
- & $22222222222222$ & $1493 \pm 13$ & $615 \pm 4$ & $209 \pm 2$ & $\preceq$ \\
- & $33333333333333$ & $1542 \pm 10$ & $554 \pm 4$ & $229 \pm 5$ & $\O$ \\
- & $11311111322231$ & $1418 \pm 9$ & $623 \pm 9$ & $197 \pm 2$ & \\
+ \multirow{3}{*}{\textbf{Map 1}} & $11111111111111$ & $1654$ $(7)$ & $131$ $(2)$& $846$ $(13)$ & $\preceq$ \\
+ & $22222222222222$ & $1657$ $(8)$ & $130$ $(2)$ & $773$ $(11)$ & $\preceq$ \\
+ & $33333333333333$ & $1681$ $(11)$ & $131$ $(2)$ & $837$ $(15)$ & $\preceq$ \\
+ & $32312212331112$ & $1619$ $(12)$ & $130$ $(2)$ & $744$ $(15)$ & \\
+\hline
+ \multirow{3}{*}{\textbf{Map 2}} & $11111111111111$ & $1409$ $(7)$ & $235$ $(4)$& $364$ $(3)$ & $\preceq$ \\
+ & $22222222222222$ & $1402$ $(6)$ & $236$ $(5)$ & $354$ $(2)$ & $\preceq$ \\
+ & $33333333333333$ & $1416$ $(8)$ & $219$ $(4)$ & $360$ $(3)$ & $\preceq$ \\
+ & $23131312323213$ & $1390$ $(10)$ & $210$ $(3)$ & $353$ $(3)$ & \\
+\hline
+ \multirow{3}{*}{\textbf{Map 3}} & $11111111111111$ & $1373$ $(6)$ & $221$ $(3)$& $301$ $(7)$ & $\preceq$ \\
+ & $22222222222222$ & $1378$ $(5)$ & $211$ $(3)$ & $268$ $(5)$ & $\preceq$ \\
+ & $33333333333333$ & $1385$ $(6)$ & $203$ $(4)$ & $291$ $(7)$ & $\O$ \\
+ & $11122222112231$ & $1358$ $(9)$ & $219$ $(7)$ & $263$ $(12)$ & \\
+\hline
+ \multirow{3}{*}{\textbf{Map 4}} & $11111111111111$ & $1383$ $(6)$ & $291$ $(5)$& $565$ $(5)$ & $\preceq$ \\
+ & $22222222222222$ & $1385$ $(7)$ & $304$ $(4)$ & $542$ $(4)$ & $\preceq$ \\
+ & $33333333333333$ & $1423$ $(6)$ & $273$ $(4)$ & $583$ $(5)$ &  $\O$ \\
+ & $11121131212112$ & $1360$ $(4)$ & $282$ $(5)$ & $540$ $(4)$ & \\
+\hline
+ \multirow{3}{*}{\textbf{Map 5}} & $11111111111111$ & $1405$ $(7)$ & $467$ $(4)$& $559$ $(4)$ & $\preceq$ \\
+ & $22222222222222$ & $1431$ $(9)$ & $447$ $(4)$ &  $541$ $(4)$ & $\preceq$ \\
+ & $33333333333333$ & $1467$ $(9)$ & $411$ $(5)$ & $567$ $(5)$ & $\O$ \\
+ & $21311213111211$ & $1397$ $(11)$ & $448$ $(10)$ & $535$ $(5)$ & \\
+\hline
+ \multirow{3}{*}{\textbf{Map 6}} & $11111111111111$ & $1575$ $(7)$ & $549$ $(5)$& $303$ $(4)$ & $\preceq$ \\
+ & $22222222222222$ & $1626$ $(9)$ & $540$ $(6)$ & $286$ $(5)$ & $\preceq$ \\
+ & $33333333333333$ & $1703$ $(11)$ & $499$ $(4)$ & $316$ $(7)$ & $\O$ \\
+ & $31121312111111$ & $1570$ $(16)$ & $535$ $(10)$ & $266$ $(6)$ & \\
+\hline
+ \multirow{3}{*}{\textbf{Map 7}} & $11111111111111$ & $1434$ $(5)$ & $599$ $(6)$& $284$ $(6)$ & $\preceq$ \\
+ & $22222222222222$ & $1475$ $(10)$ & $602$ $(5)$ & $243$ $(6)$ & $\preceq$ \\
+ & $33333333333333$ & $1489$ $(12)$ & $549$ $(3)$ & $264$ $(6)$ & $\O$ \\
+ & $11332211321332$ & $1401$ $(12)$ & $563$ $(4)$ & $230$ $(12)$ & \\
+\hline
+ \multirow{3}{*}{\textbf{Map 8}} & $11111111111111$ & $1761$ $(9)$ & $254$ $(5)$& $382$ $(3)$ & $\preceq$ \\
+ & $22222222222222$ & $1804$ $(10)$ & $269$ $(4)$ & $357$ $(4)$ & $\preceq$ \\
+ & $33333333333333$ & $1826$ $(10)$ & $230$ $(3)$ & $392$ $(7)$ & $\O$ \\
+ & $23221131313323$ & $1747$ $(11)$ & $247$ $(9)$ & $363$ $(9)$ & \\
+\hline
+ \multirow{3}{*}{\textbf{Map 9}} & $11111111111111$ & $2501$ $(14)$ & $926$ $(6)$& $574$ $(9)$ & $\preceq$ \\
+ & $22222222222222$ & $2503$ $(10)$ & $921$ $(10)$ & $524$ $(8)$ & $\preceq$ \\
+ & $33333333333333$ & $2641$ $(14)$ & $833$ $(5)$ & $574$ $(14)$ & $\O$ \\
+ & $21132331333223$ & $2463$ $(19)$ & $891$ $(8)$ & $523$ $(9)$ & \\
+\hline
+ \multirow{3}{*}{\textbf{Map 10}} & $11111111111111$ & $1430$ $(8)$ & $630$ $(4)$& $205$ $(2)$ & $\preceq$ \\
+ & $22222222222222$ & $1493$ $(13)$ & $615$ $(4)$ & $209$ $(2)$ & $\preceq$ \\
+ & $33333333333333$ & $1542$ $(10)$ & $554$ $(4)$ & $229$ $(5)$ & $\O$ \\
+ & $11311111322231$ & $1418$ $(9)$ & $623$ $(9)$ & $197$ $(2)$ & \\
 \hline

### 2014-06-11 paper/paper.tex
revision 1 submitted

@@
 
-Two main goals can be identified in this paper: first, the proposed algorithm must be applicable to real-time domains (those where the next move to make must be decided within a small time budget) and it should obtain better or at least the same performance than the other state-of-the-art algorithms.
+Two main goals can be identified in this paper: first, the proposed algorithm must be applicable to real-time domains (those where the next move to make must be decided within a small time budget) and it should obtain better or at least the same performance than the other algorithms.
 It is important to highlight that all three algorithms tested employ the same heuristic functions to evaluate the features of a given game state.
@@
 \item Elitism: The individuals from the first front $F_0$ are always ranked first, according to the dominance criterion and their crowding distance. 
-This ensures that the best solutions always survive to the next generation.
 \end{itemize}
@@
 Finally, in the comparison between NSGA-II and MO-MCTS, the latter algorithm obtains higher percentages for each one of the points in the front.
-This result suggests that, with a limited number of iterations/evaluations, the proposed algorithm is able to explore the search space more efficiently than a rolling horizon version of a state-of-the-art NSGA-II.
+This result suggests that, with a limited number of iterations/evaluations, the proposed algorithm is able to explore the search space more efficiently than a rolling horizon version of NSGA-II.
 

### 2014-07-29 paper/paper.tex
preparing camera ready. bios to be updated before submitting final files
bio updated

@@
 	\begin{center}
-	\includegraphics[scale=0.4,natwidth=708,natheight=484]{img/mcts.png}
+	\includegraphics[scale=0.4,natwidth=708,natheight=484]{figures/perez1.png}
 	\caption{MCTS algorithm steps.}
@@
 	\begin{center}
-	\includegraphics[scale=0.235]{img/moopDec2.png}
+	\includegraphics[scale=0.235]{figures/perez2.png}
 	\caption{Decision and objective spaces in a MOP with two variables ($x_1$ and $x_2$) and two objectives ($f_1$ and $f_2$).
@@
 \begin{center}
-\includegraphics[width=0.75\columnwidth]{img/hv}
+\includegraphics[width=0.75\columnwidth]{figures/perez3}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=0.65\columnwidth]{img/wangHV}
+\includegraphics[width=0.65\columnwidth]{figures/perez4}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=0.75\columnwidth]{img/dst}
+\includegraphics[width=0.75\columnwidth]{figures/perez5}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=1.0\columnwidth]{img/optDST.png}
+\includegraphics[width=1.0\columnwidth]{figures/perez6}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=0.9\columnwidth]{img/dstTransTables}
+\includegraphics[width=0.9\columnwidth]{figures/perez7}
 \end{center}
@@
 	\begin{center}
-	\includegraphics[width=0.8\columnwidth]{img/moptspmap}
+	\includegraphics[width=0.8\columnwidth]{figures/perez8}
 	\caption{Sample MO-PTSP map.}
@@
 	\centering
-	\includegraphics[width=2.1\columnwidth]{img/dstAll}
+	\includegraphics[width=2.1\columnwidth]{figures/perez9}
 	\caption{Results in DST: percentages of each optima found during $100$ games played with different weight vectors.
@@
 is currently pursuing a Ph.D.
-in Artificial Intelligence applied to games at the University of Essex, Colchester.
-He has published in the domain of Game AI, participated in several Game AI competitions and organized the Physical Travelling Salesman Problem competitions, held in IEEE conferences.
+in Artificial Intelligence applied to games at the University of Essex (UK).
+He has published in the domain of Game AI, with research interests on Reinforcement Learning and Evolutionary Computation.
+He has organized several Game AI competitions, as the Physical Travelling Salesman Problem and the General Video Game AI competitions, both held in IEEE conferences.
 He also has programming experience in the videogames industry with titles published for game consoles and PC.
@@
 \begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{./photos/spyro.jpg}}]{Spyridon Samothrakis}
- is currently pursuing a PhD in Computational Intelligence and Games at the University of Essex.
-His interests include game theory, computational neuroscience, evolutionary algorithms and consciousness.
+ is currently a Senior Research Officer at the University of Essex.
+He holds a B.Sc.
+from the University of Sheffield (Computer Science), an M.Sc.
+from the University of Sussex (Intelligent Systems) and a Ph.D (Computer Science) from the University of Essex.
+His interests include game theory, machine learning, evolutionary algorithms and consciousness.
 \end{IEEEbiography}

### 2014-08-15 lpaper/lunarcont.tex
Final Commit before Demo

@@
 
-In general, the optimisation problem can be defined as: Given a function $f:\bm{X}\rightarrow Y $, where $X \in \mathcal{R}^n$, with $n$ being the length of vector $\bm{X}$.
+In general, the optimisation problem can be defined as: Given a function $f:\bm{X}\rightarrow \bm{Y} $, where $\bm{X} \in \mathcal{R}^n$, with $n$ being the length of vector $\bm{X}$.
 The domain $\bm{X}$ is usually called the search space.
@@
 Thus, for child $C_l(\nu)$ ,   $\bm{x^{C_l}_{\mathit{max,k}}} =\bm{x^{C}_{\mathit{min,k}}} +  \left(\bm{x^{C}_{\mathit{min,k}}} - \bm{x^{C}_{\mathit{max,k}}}\right)/2 $.
-For the other child,  $\bm{x^{C_r}_{\mathit{min,k}}} = \bm{x^{C}_{\mathit{min,k}}} +  \left(\bm{x^{C}_{\mathit{min,k}}} - \bm{x^{C}_{\mathit{max,k}}}\right)/2 $.
+For the other child,  $\bm{x^{C_r}_{\mathit{min,k}}} =  \bm{x^{C_l}_{\mathit{max,k}}}$.
 In other words, the father node is split into two children nodes, choosing which dimension to split uniform random (or according to some other criterion).
@@
 
-If a dissimilarity metric is defined between different values of $\bm{X}$ as ${\lVert \bm{x}_1 - \bm{x}_2) \rVert}^\alpha$, one can set $\upsilon_1 =\sqrt{|A|}/2)^\alpha$ and $\rho = 2^{-\alpha/|A|}$.
-In our experiments $a = 2$.
+If a dissimilarity metric is defined between different values of $\bm{X}$ as ${\lVert \bm{x}_1 - \bm{x}_2 \rVert}^\alpha$, one can set $\upsilon_1 =\sqrt{|A|}/2)^\alpha$ and $\rho = 2^{-\alpha/|A|}$.
+In our experiments $\alpha = 2$.
 If $N(\nu) = 0$, $B(\nu) = U(\nu) = \infty$, i.e.
@@
 \begin{itemize}
-\item $S$ is a set of states, $s \in S$, where $s_j^t$ is state $j$ at time $t$ and $s'_j$ is the state j at time $t+1$.
+\item $S$ is a set of states, $s \in S$, with $s'$ being the next state in time. 
 \item $A$ is a set of actions, each action named $a_j$.
@@
 The goal of an agent in an MDP environment is to maximise its long term value.
-Note that, since we are dealing with continuous actions and states here, sets $A$ and $S$ are metric (and infinitive).
+Note that, since we are dealing with continuous actions and states here, sets $A$ and $S$ are metric (and not countable).
 Also note that we are NOT dealing with continuous time MDPs in this paper, although there is a notion of time.
@@
   \includegraphics[width=0.45\textwidth]{graphics/gamma-crop.pdf}
-  \caption{Examples of how rewards are affected by $\gamma$.}
+  \caption{Examples of how rewards are affected by $\gamma$. Split probability reflects the probability of T-HOO splitting a node into two, while action depth reflects look-ahead depth}
   \label{fig:gamma}
@@
 
-When learning how to act using a generative model, one can make a third distinction as to the type of sets represented by $A$ and $S$.
+When learning how to act using a generative model (i.e.
+an internal simulator of feature events conditioned on actions), one can make a third distinction as to the type of sets represented by $A$ and $S$.
 In this paper the focus is on an $A$ and $S$ that come from a metric space.
@@
 There are examples of closed loop rolling horizon papers, but these are beyond the scope of this article~\cite{xiang2012differential,samothrakis2010planning}.
-
-\subsection{Notes on planning using a Generative Model}
-
-The Markov Decision Process simply describes an environment alongside the rewards that said environment provides to an agent.
-A problem central to Artificial Intelligence is what actions an agent must take to maximise its long term reward.
-The most basic assumption one can make is that the agent knows nothing about the environment.
-In this case the agent gets thrown into the environment and interacts (``trains'') for some time.
-After the training is complete, the agent should be able to act intelligently (presumably with some success).
-
-Another approach one can take is to assume that the agent has a copy of the MDP in its ``mind'' and thus is able to plan accordingly.
-This is usually termed dynamic programming~\cite{sutton1998reinforcement}, however it cannot be used in the case of continuous states or actions without the use of a function approximator (FA) or some kind of discretization.
-Another approach, and one which has made a significant impact recently, arguably presented for the first time by Kearns~\cite{citeulike:1584916} and popularised mostly as Upper Confidence Bounds for Trees (UCT)~\cite{kocsis2006bandit}, is to use a generative model.
-
-A generative model is nothing more than a simulator of future events.
-This is less of a requirement than knowing the full model (which entails knowing the full function $T$), but still requires the agent to have some knowledge of the world.
-This category of algorithms, which are broadly termed ``Rolling Horizon'', has the following steps.
-From the current state, a number of simulations of possible futures is run by executing a set of action sequences (all of this in the ``mind'' of the agent).
-The simulations last for a certain number of steps (hence the ``rolling horizon'') or continue until an end state is encountered (i.e.
-when the MDP has ``absorbing'' states that cannot be acted from).
-
-Once a satisfactory plan/policy has been estimated (or the agent runs out of time, iterations etc.) it proceeds by executing the first action of the plan, thus moving to the next state.
-Once there, the agent performs the same sequence of ``mental'' actions, (i.e.
-plans and executes the first action), until some predefined stopping condition.
-In this paper we will concern ourselves \textit{only} with this method of planning.
-What changes from method to method is how the plan is formed.
-The main version of the algorithm used can be seen in Algorithm \ref{algo:ooop}.
-
-In standard UCT, a procedure similar to temporal difference learning is followed (or, more closely, Monte Carlo methods), where one learns the value of each state.
-An agent then forms a policy by choosing a value greedily.
-In our case, because our state space has continuous states and actions, this is impossible without the use of a function approximator (or an infinite discretization of space), which is not attempted here.
-What is attempted though is to learn the policy function directly without reference to state, while planning continuously from the current state.
-So while the agent forms open-loop plans at each clock tick, in reality it is executing a closed-loop policy.
-While the strategy is not provably optimal, it can work well in practice~\cite{weinstein2012bandit}.
-This will be seen by our experimental results (see Section \ref{sec:Experiments}).
 
@@
 		\While{$!StoppingCondition$}
-		    \State $a \gets \mathsc{FindOpenLoopPlan}(depth, T',S')[0]$ \Comment{Execute first action(s) of the policy}
-		    \State $s \sim T(s'|s,a)$ \Comment{Stop only at an absorbing state}
+		    \State $a \gets \mathsc{FindOpenLoopPlan}(depth, T',S')[0]$ \Comment{Execute first action(s) of the policy - [0] gets the first action}
+		    \State $s \sim T(s'|s,a)$ \Comment{Stop only at an absorbing state, sampling ($\sim$) from the environment}
 		\EndWhile
@@
 \subsection{Lunar Lander}
-In this section we give a brief overview of the game of Lunar Lander, a popular arcade game.
+In this section we give a brief overview of a problem based upon {\itshape Lunar Lander}, a popular arcade game.
+The goal in {\itshape Lunar Lander} is to land a spaceship on a flat plateau on an otherwise jagged landscape.
+In order to do this, the player must use a supply of fuel to apply thrust to a spaceship, which has inertia.
+The ship is able to rotate at no cost to fuel usage, but must contend with rotational inertia.
 
@@
 Anything that passes from the left edge of the playing field moves to the right instantaneously, and vice versa.
-
-The other feature of note is the jagged landscape, a simplistic representation of the noisy, crater-filled lunar landscape the game represents.
-This landscape is constructed as a series of line segments, with each vertex of the jagged landscape being distributed equally horizontally, and randomly vertically.
-% unclear?
+The landscape is constructed as a series of line segments, with each vertex being distributed equally horizontally, and randomly vertically.
 
@@
 
-The spaceship within this game world is modelled as a circular mass with some basic physical properties, such as a position within the game world $\bm {s}$, a velocity $\bm {v}$, an orientation $\bm{d}$, an angular velocity $\omega$ and a radius of a bounding sphere $r$ used for collision detection with the landscape.
+The spaceship within this game world is modelled as a circular mass with some physical properties, such as a position within the game world $\bm {s}$, a velocity $\bm {v}$, an orientation $\bm{d}$, an angular velocity $\omega$ and a radius of a bounding sphere $r$ used for collision detection with the landscape.
 
-This collision detection method is based on the lowest point of the spaceship's bounding circle coming into intersection with the highest point of the landscape for the line perpendicular between the ship and the bottom of the playing field, as can be seen in Figure \ref{fig:diagram_landscapepoint}.
+This collision detection method is based on the spaceship's bounding circle overlapping the corresponding point on the landscape on the same vertical axis, as can be seen in Figure \ref{fig:diagram_landscapepoint}.
 
-% this needs a diagram for clarity: reuse the old one? reusing it temporarily
 \begin{figure}[hbtp]
 \centering
-\includegraphics[scale=0.4]{graphics/landscapepoint}
+\includegraphics[scale=0.35]{graphics/landscapepoint}
 \caption{The point closest to the spaceship on the landscape lies in between two of the defined vertices of the landscape, and requires interpolation to calculate the y co-ordinate from the x co-ordinate of the ship.}
@@
 
-As the landscape is stored as a series of line segments, interpolation must be used in the event the ship is not perfectly aligned with one of the landscape axes.
-The point that lies on the line segment that the ship is being checked against is calculated through simple linear interpolation based on which two vertices the ship is horizontally closest to.
 For the ship's centre $\bm {s}$, the left nearest landscape vertex $\bm {p}^{l}$ and the right nearest landscape vertex $\bm {p}^{r}$, the point of collision $\bm {p}^{c}$ against the landscape is calculated as
-
-% does this constitute self-plaigarism???
 
@@
 The ship colliding with the landscape constitutes the end of the {\itshape Lunar Lander} game.
-The conditions surrounding this collision, including speedand orientation of the ship, constitute whether the nature of the collision is a success or a failure.
+The conditions surrounding this collision, including speed and orientation of the ship, constitute whether the nature of the collision is a success or a failure.
 In this experimental setup we accept a velocity of $5$ and an angle of $< 0.1$ radians as successful landing criteria.
@@
 The controllers have to identify a vector of real numbers $\bm{A}$, with each element of the set $a_i \in [0,1]$.
-Each even element of the vector represents the linear thrust, whereas each odd element represents angular velocity.
-Each linear thrust vector element is transformed by applying as follows:
+Each even element of the vector represents the linear thrust, whereas each odd element represents angular velocity, actions that the agent can take.
+Each linear thrust vector element is transformed by applying the following:
 
 \begin{equation}
-\bm{a_I} = 5(\bm{a_I} - 0.5)
-\bm{a_T} = 300\bm{a_T} 
+\bm{a_I} \gets 5(\bm{a_I} - 0.5)
+\bm{a_T} \gets 300\bm{a_T} 
 \end{equation}
@@
 \begin{align}
-f &= \twopartdef{-\sum{arccos(\bm{dl}\cdot\bm{d}/\norm{\bm{dl}}\norm{\bm{d}}})}{\bm{np_x} < \bm{s_x}}{-\sum{arccos(\bm{dr}\cdot\bm{d}/\norm{\bm{dr}}\norm{\bm{d}}})} {otherwise}\notag\\
-l &= -\sum{arccos(\bm{du}\cdot\bm{d}/\norm{\bm{du}}\norm{\bm{d}}}) - \abs{\bm {v_y} - V_{y}^{max}}\notag\\
+f &= \twopartdef{-\sum{arccos(\bm{dl}\cdot\bm{d}/(\norm{\bm{dl}}\norm{\bm{d}}}))}{\bm{np_x} < \bm{s_x}}{-\sum{arccos(\bm{dr}\cdot\bm{d}/(\norm{\bm{dr}}\norm{\bm{d}}}))} {otherwise}\notag\\
+l &= -\sum{arccos(\bm{du}\cdot\bm{d}/(\norm{\bm{du}}\norm{\bm{d}}})) - \abs{\bm {v_y} - V_{y}^{max}}\notag\\
 h &= \twopartdef{ \abs{-(\norm{\bm {v}} - C) - \abs{\bm {s}_{x} - \bm {np}_{x}}} - f}{\norm{\bm {s} - \bm {np}} < D}
@@
                 \includegraphics[width=1.0\textwidth]{./graphics/map2.pdf}
-                \caption{Map A,  where the ship is lower then the pad.}
+                \caption{Map A,  where the ship is lower than the pad.}
                 \label{fig:map2}
@@
 each algorithm at each step plans for 50 steps ahead.
-We set the noise levels to $0.1 (\mathcal{U}(0,1)-0.5)*2 l$ uniform noise for the double integrator experiment and $10 (\mathcal{U}(0,1)-0.5)*2 l$  for the Inverted Pendulum.
+We set the noise levels to $0.1 (\mathcal{U}(-0.5,0.5)) l$ uniform noise for the double integrator experiment and $10 (\mathcal{U}(-0.5,0.5)) l$  for the Inverted Pendulum (different noise levels for different problems come from the fact that the problems have unequal features).
 The variable $l$ is the noise level which defines how noisy the environment is.
 Initially, we performed two experiments, with a variable number of iterations and a fixed noise level of one.
+In all experiments the default CMA implementation is used~\footnote{\url{https://www.lri.fr/~hansen/cmaes_java.tar.gz}}.
 
@@
 Unfortunately, it seems that different algorithms require different heuristics to work properly in the case that absorbing states are not met during rollouts.
+Definite conclusions are hard to draw about the quality and/or the suitability of each method.
+There are a multitude of hyperparameters to optimise and each method responds better to different heuristics.
+On top of that, implementations of each method might have different memory/speed requirements on different platforms - which makes direct comparison even harder.
+If anything, such intricacies require more specific set of experiments and a conscious effort to understand which method behaves better under what conditions.
+Overall however we think it might be possible to incorporate simplifying assumptions into T-HOLOP priors, thus making it easier to search harder distributions.
+In all cases however, one might want to include sensible defaults in all algorithms, and use them without too much hassle; CMA-ES is probably as close to that as one can imagine for continuous settings, so it might be worth considering it as a ``first option'' in most settings (including the one we use here).
 In future research we aim to improve MCTS-like algorithms for continuous settings by including a learning module, possibly in the form of actor-critic algorithms.
-This should automate learning the heuristics (after some ad-hoc played games) and help guide the rollouts towards more sensible solutions, even in the case where heuristics do not favour T-HOLOP.
+This should automate learning heuristics (after some training games) and help guide the rollouts towards more sensible solutions, even in the case where heuristics do not favour T-HOLOP.
 
@@
 \bibliography{lunarcontbib}
-\end{document}
+\end{document}=

### 2014-08-15 code/src/lunarcode/LunarRun.java

@@ -22,8 +22,8 @@ public class LunarRun {
     public static boolean m_verbose = false;
 
     public static Game game;
-    //public static LunarLanderEvoCont cont;
-    public static LunarLanderHOOCont cont;
+    public static LunarLanderEvoCont cont;
+    //public static LunarLanderHOOCont cont;
 
     public static void main(String[] args) {
         runGame();
@@ -33,8 +33,8 @@ public class LunarRun {
 
     public static void prepareGame() {
         game = new LunarGame();
-        //cont = new LunarLanderEvoCont();
-        cont = new LunarLanderHOOCont();
+        cont = new LunarLanderEvoCont();
+       // cont = new LunarLanderHOOCont();
     }
 
     public static void runGame() {

### 2015-05-14 paper/paper.tex
new figures, after viva

@@
 
+\begin{IEEEkeywords}
+Monte Carlo Tree Search, Multi-Objective Optimization, Real-Time Games.
+\end{IEEEkeywords}
+
 % For peer review papers, you can put extra information on the cover
@@
 \begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{photos/sanaz.jpg}}]
-{Sanaz Mostaghim}
-is a professor of computer science at the Otto von Guericke University Magdeburg, Germany.
-Her current research interests are in the field of evolutionary computation, particularly in the areas of evolutionary multi-objective optimization, swarm intelligence and their applications in science and industry.
-She is an active IEEE member and is serving as an associate editor for IEEE Transactions on Evolutionary Computation and IEEE Transactions on Cybernetics.
+{Sanaz Mostaghim} is a full professor of computer science at the Otto von Guericke University Magdeburg, Germany.
+She holds a PhD degree (2004) in electrical engineering and computer science from the University of Paderborn, Germany.
+She worked as a postdoctoral fellow at ETH Zurich in Switzerland and as a lecturer at Karlsruhe Institute of technology (KIT), Germany.
+Her research interests are in the area of swarm intelligence, evolutionary multi-objective optimisation and their applications in robotics, science and industry.
+She is an active member in several German and international societies such as IEEE and GI.
+She serves as an associate editor for IEEE Transactions on evolutionary computation and IEEE Transactions on cybernetics.
 \end{IEEEbiography}

### 2015-06-22 paper/paper.tex
paper with the correct images

@@
 \begin{center}
-\includegraphics[width=0.75\columnwidth]{figures/perez3}
+\includegraphics[width=0.75\columnwidth]{figures/perez3b}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=1.0\columnwidth]{figures/perez6}
+\includegraphics[width=1.0\columnwidth]{figures/perez6b}
 \end{center}
@@
 \begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{photos/perez.jpg}}]
-{Diego Perez}
-is currently pursuing a Ph.D.
-in Artificial Intelligence applied to games at the University of Essex (UK).
+{Diego Perez} is a Senior Research Officer at the University of Essex (UK).
+He holds a PhD degree (2015) in Computer Science and Electronic Engineering at the University of Essex.
 He has published in the domain of Game AI, with research interests on Reinforcement Learning and Evolutionary Computation.
 He has organized several Game AI competitions, as the Physical Travelling Salesman Problem and the General Video Game AI competitions, both held in IEEE conferences.
-He also has programming experience in the videogames industry with titles published for game consoles and PC.
+He has programming experience in the videogames industry with titles published for game consoles and PC.
 \end{IEEEbiography}
@@
 {Sanaz Mostaghim} is a full professor of computer science at the Otto von Guericke University Magdeburg, Germany.
-She holds a PhD degree (2004) in electrical engineering and computer science from the University of Paderborn, Germany.
+She holds a PhD degree (2004) in Electrical Engineering and Computer Science from the University of Paderborn, Germany.
 She worked as a postdoctoral fellow at ETH Zurich in Switzerland and as a lecturer at Karlsruhe Institute of technology (KIT), Germany.
@@
 \begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{photos/lucas.png}}]{Simon Lucas}
-(SMIEEE) is a professor of Computer Science at the University of Essex (UK) where he leads the Game Intelligence Group.
-His main research interests are games, evolutionary computation, and machine learning, and he has published widely in these fields with over 160 peer-reviewed papers.
+(SMIEEE) is a professor of Computer Science at the University of Essex (UK) where he is the Head of School and leads the Game Intelligence Group.
+He holds a PhD degree (1991) in Electronics and Computer Science from the University of Southampton.
+His main research interests are games, evolutionary computation, and machine learning, and he has published widely in these fields with over 180 peer-reviewed papers.
 He is the inventor of the scanning n-tuple classifier, and is the founding Editor-in-Chief of the IEEE Transactions on Computational Intelligence and AI in Games.
